### 2019-04-30 icdm/submission/paper.tex
initial import

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Paper Title*\\
{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
should not be used}
\thanks{Identify applicable funding agency here. If none, delete this.}
}

\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address}
\and
\IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address}
\and
\IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address}
\and
\IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address}
\and
\IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address}
}

\maketitle

\begin{abstract}
This document is a model and instructions for \LaTeX.
This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
or Math in Paper Title or Abstract.
\end{abstract}

\begin{IEEEkeywords}
component, formatting, style, styling, insert
\end{IEEEkeywords}

\section{Introduction}
This document is a model and instructions for \LaTeX.
Please observe the conference page limits.

\section{Ease of Use}

\subsection{Maintaining the Integrity of the Specifications}

The IEEEtran class file is used to format your paper and style the text. All margins, 
column widths, line spaces, and text fonts are prescribed; please do not 
alter them. You may note peculiarities. For example, the head margin
measures proportionately more than is customary. This measurement 
and others are deliberate, using specifications that anticipate your paper 
as one part of the entire proceedings, and not as an independent document. 
Please do not revise any of the current designations.

\section{Prepare Your Paper Before Styling}
Before you begin to format your paper, first write and save the content as a 
separate text file. Complete all content and organizational editing before 
formatting. Please note sections \ref{AA}--\ref{SCM} below for more information on 
proofreading, spelling and grammar.

Keep your text and graphic files separate until after the text has been 
formatted and styled. Do not number text heads---{\LaTeX} will do that 
for you.

\subsection{Abbreviations and Acronyms}\label{AA}
Define abbreviations and acronyms the first time they are used in the text, 
even after they have been defined in the abstract. Abbreviations such as 
IEEE, SI, MKS, CGS, ac, dc, and rms do not have to be defined. Do not use 
abbreviations in the title or heads unless they are unavoidable.

\subsection{Units}
\begin{itemize}
\item Use either SI (MKS) or CGS as primary units. (SI units are encouraged.) English units may be used as secondary units (in parentheses). An exception would be the use of English units as identifiers in trade, such as ``3.5-inch disk drive''.
\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds. This often leads to confusion because equations do not balance dimensionally. If you must use mixed units, clearly state the units for each quantity that you use in an equation.
\item Do not mix complete spellings and abbreviations of units: ``Wb/m\textsuperscript{2}'' or ``webers per square meter'', not ``webers/m\textsuperscript{2}''. Spell out units when they appear in text: ``. . . a few henries'', not ``. . . a few H''.
\item Use a zero before decimal points: ``0.25'', not ``.25''. Use ``cm\textsuperscript{3}'', not ``cc''.)
\end{itemize}

\subsection{Equations}
Number equations consecutively. To make your 
equations more compact, you may use the solidus (~/~), the exp function, or 
appropriate exponents. Italicize Roman symbols for quantities and variables, 
but not Greek symbols. Use a long dash rather than a hyphen for a minus 
sign. Punctuate equations with commas or periods when they are part of a 
sentence, as in:
\begin{equation}
a+b=\gamma\label{eq}
\end{equation}

Be sure that the 
symbols in your equation have been defined before or immediately following 
the equation. Use ``\eqref{eq}'', not ``Eq.~\eqref{eq}'' or ``equation \eqref{eq}'', except at 
the beginning of a sentence: ``Equation \eqref{eq} is . . .''

\subsection{\LaTeX-Specific Advice}

Please use ``soft'' (e.g., \verb|\eqref{Eq}|) cross references instead
of ``hard'' references (e.g., \verb|(1)|). That will make it possible
to combine sections, add equations, or change the order of figures or
citations without having to go through the file line by line.

Please don't use the \verb|{eqnarray}| equation environment. Use
\verb|{align}| or \verb|{IEEEeqnarray}| instead. The \verb|{eqnarray}|
environment leaves unsightly spaces around relation symbols.

Please note that the \verb|{subequations}| environment in {\LaTeX}
will increment the main equation counter even when there are no
equation numbers displayed. If you forget that, you might write an
article in which the equation numbers skip from (17) to (20), causing
the copy editors to wonder if you've discovered a new method of
counting.

{\BibTeX} does not work by magic. It doesn't get the bibliographic
data from thin air but from .bib files. If you use {\BibTeX} to produce a
bibliography you must send the .bib files.

{\LaTeX} can't read your mind. If you assign the same label to a
subsubsection and a table, you might find that Table I has been cross
referenced as Table IV-B3.

{\LaTeX} does not have precognitive abilities. If you put a
\verb|\label| command before the command that updates the counter it's
supposed to be using, the label will pick up the last counter to be
cross referenced instead. In particular, a \verb|\label| command
should not go before the caption of a figure or a table.

Do not use \verb|\nonumber| inside the \verb|{array}| environment. It
will not stop equation numbers inside \verb|{array}| (there won't be
any anyway) and it might stop a wanted equation number in the
surrounding equation.

\subsection{Some Common Mistakes}\label{SCM}
\begin{itemize}
\item The word ``data'' is plural, not singular.
\item The subscript for the permeability of vacuum $\mu_{0}$, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ``o''.
\item In American English, commas, semicolons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation. When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks. A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this). (A parenthetical sentence is punctuated within the parentheses.)
\item A graph within a graph is an ``inset'', not an ``insert''. The word alternatively is preferred to the word ``alternately'' (unless you really mean something that alternates).
\item Do not use the word ``essentially'' to mean ``approximately'' or ``effectively''.
\item In your paper title, if the words ``that uses'' can accurately replace the word ``using'', capitalize the ``u''; if not, keep using lower-cased.
\item Be aware of the different meanings of the homophones ``affect'' and ``effect'', ``complement'' and ``compliment'', ``discreet'' and ``discrete'', ``principal'' and ``principle''.
\item Do not confuse ``imply'' and ``infer''.
\item The prefix ``non'' is not a word; it should be joined to the word it modifies, usually without a hyphen.
\item There is no period after the ``et'' in the Latin abbreviation ``et al.''.
\item The abbreviation ``i.e.'' means ``that is'', and the abbreviation ``e.g.'' means ``for example''.
\end{itemize}
An excellent style manual for science writers is \cite{b7}.

\subsection{Authors and Affiliations}
\textbf{The class file is designed for, but not limited to, six authors.} A 
minimum of one author is required for all conference articles. Author names 
should be listed starting from left to right and then moving down to the 
next line. This is the author sequence that will be used in future citations 
and by indexing services. Names should not be listed in columns nor group by 
affiliation. Please keep your affiliations as succinct as possible (for 
example, do not differentiate among departments of the same organization).

\subsection{Identify the Headings}
Headings, or heads, are organizational devices that guide the reader through 
your paper. There are two types: component heads and text heads.

Component heads identify the different components of your paper and are not 
topically subordinate to each other. Examples include Acknowledgments and 
References and, for these, the correct style to use is ``Heading 5''. Use 
``figure caption'' for your Figure captions, and ``table head'' for your 
table title. Run-in heads, such as ``Abstract'', will require you to apply a 
style (in this case, italic) in addition to the style provided by the drop 
down menu to differentiate the head from the text.

Text heads organize the topics on a relational, hierarchical basis. For 
example, the paper title is the primary text head because all subsequent 
material relates and elaborates on this one topic. If there are two or more 
sub-topics, the next level head (uppercase Roman numerals) should be used 
and, conversely, if there are not at least two sub-topics, then no subheads 
should be introduced.

\subsection{Figures and Tables}
\paragraph{Positioning Figures and Tables} Place figures and tables at the top and 
bottom of columns. Avoid placing them in the middle of columns. Large 
figures and tables may span across both columns. Figure captions should be 
below the figures; table heads should appear above the tables. Insert 
figures and tables after they are cited in the text. Use the abbreviation 
``Fig.~\ref{fig}'', even at the beginning of a sentence.

\begin{table}[htbp]
\caption{Table Type Styles}
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Table}&\multicolumn{3}{|c|}{\textbf{Table Column Head}} \\
\cline{2-4} 
\textbf{Head} & \textbf{\textit{Table column subhead}}& \textbf{\textit{Subhead}}& \textbf{\textit{Subhead}} \\
\hline
copy& More table copy$^{\mathrm{a}}$& &  \\
\hline
\multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
\end{tabular}
\label{tab1}
\end{center}
\end{table}

Figure Labels: Use 8 point Times New Roman for Figure labels. Use words 
rather than symbols or abbreviations when writing Figure axis labels to 
avoid confusing the reader. As an example, write the quantity 
``Magnetization'', or ``Magnetization, M'', not just ``M''. If including 
units in the label, present them within parentheses. Do not label axes only 
with units. In the example, write ``Magnetization (A/m)'' or ``Magnetization 
\{A[m(1)]\}'', not just ``A/m''. Do not label axes with a ratio of 
quantities and units. For example, write ``Temperature (K)'', not 
``Temperature/K''.

\section*{Acknowledgment}

The preferred spelling of the word ``acknowledgment'' in America is without 
an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B. 
G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor 
acknowledgments in the unnumbered footnote on the first page.

\section*{References}

Please number citations consecutively within brackets \cite{b1}. The 
sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference 
number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at 
the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''

Number footnotes separately in superscripts. Place the actual footnote at 
the bottom of the column in which it was cited. Do not put footnotes in the 
abstract or reference list. Use letters for table footnotes.

Unless there are six authors or more give all authors' names; do not use 
``et al.''. Papers that have not been published, even if they have been 
submitted for publication, should be cited as ``unpublished'' \cite{b4}. Papers 
that have been accepted for publication should be cited as ``in press'' \cite{b5}. 
Capitalize only the first word in a paper title, except for proper nouns and 
element symbols.

For papers published in translation journals, please give the English 
citation first, followed by the original foreign-language citation \cite{b6}.

\bibliographystyle{IEEEtran}
\bibliography{refs}

% use bibfile instead (2 lines above)
% \begin{thebibliography}{00}
% \bibitem{b1} G. Eason, B. Noble, and I. N. Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil. Trans. Roy. Soc. London, vol. A247, pp. 529--551, April 1955.
% \bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
% \bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
% \bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
% \bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
% \bibitem{b6} Y. Yorozu, M. Hirano, K. Oka, and Y. Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl. J. Magn. Japan, vol. 2, pp. 740--741, August 1987 [Digests 9th Annual Conf. Magnetics Japan, p. 301, 1982].
% \bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
% \end{thebibliography}
\vspace{12pt}
\color{red}
IEEE conference templates contain guidance text for composing and formatting conference papers. Please ensure that all template text is removed from your conference paper prior to submission to the conference. Failure to remove the template text from your paper may result in your paper not being published.

\end{document}

### 2019-05-01 icdm/submission/paper.tex
added images and text from bachelor thesis

@@
     T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+
+% Personal Packages
+\usepackage{csquotes}
 \begin{document}
@@
 
-\section{Introduction}
-This document is a model and instructions for \LaTeX.
-Please observe the conference page limits.
+\section{Introduction}\label{chap:introduction}
 
-\section{Ease of Use}
+Representing words as vectors, i.e word embeddings (WE) is a fundamental aspect of Natural Language Processing (NLP).
+There are two ways to create such WE, either arbitrarily or with the purpose of capturing the semantics of the words, i.e.
+vector representations of words that are syntactically or semantically similar will be near to each other.
+By capturing semantic or syntactic meaning WE have shown to facilitate a lot of subsequent NLP tasks, such as entity recognition, machine translation or sentence classification.
+The first attempt to create WE with neural networks was made by Bengio et al.
+\cite{bengio} but more recently Mikolov et al.
+\cite{mikolov} introduced a software package called w2vec that uses a simpler network and with this approach Mikolov et al.
+produced state of the art results.
+One of the proposed algorithms in this software package is the Skip-Gram Model (SGM).
+The SGM is an algorithm, that trains a neural network, on the task of predicting the neighboring words in a sentence.
+The weights of this network are then used as WE.
 
-\subsection{Maintaining the Integrity of the Specifications}
+The SGM gained a lot of attention, as it achieved very good results for a very simplistic model.
+As a consequence, a lot of effort went into optimizing it.
+Most of this effort was trying to improve the throughput of the model, i.e the number of words that are processed per second.
+The SGM uses Stochastic Gradient Descent as an optimization algorithm and is therefore inherently sequential.
+To remedy this problem Mikolov et al.
+used Hogwild \cite{hogwild}, where different threads can access a shared model and update it.
+As this is not an optimal solution Yi et al.
+\cite{intel} tried to optimize it, by using a mini-batch like approach and converting vector to vector multiplications into matrix to matrix multiplications.
+This yielded two consequences: First the model is updated less frequently leading to less overwriting and offering the possibility to parallelize more.
+Secondly, it transformed level-1 BLAS operation into level-3 BLAS operations, and the algorithm could therefore effectively use computational resources.
+Another attempt at optimizing the throughput was made by Seulki and Youngmin \cite{gpu}.
+Their goal was to parallelize the algorithm on GPU's.
+They, therefore, chose to parallelize the update of the dimensions of each word representation.
+Both of these approaches and most of the literature are focused on improving the throughput of the model, but not the convergence time.
+Therefore one could ask if the convergence time of the SGM can be optimized while at the same time maintaining its accuracy.
+This work proposes an approach that uses advanced optimizers and input shuffling to optimize the convergence time of the SGM.
+In combination both of these techniques allowed us to decrease the convergence time of the SGM.
+If combined with an optimized throughput these result could lead to an overall decrease in runtime.
 
-The IEEEtran class file is used to format your paper and style the text. All margins, 
-column widths, line spaces, and text fonts are prescribed; please do not 
-alter them. You may note peculiarities. For example, the head margin
-measures proportionately more than is customary. This measurement 
-and others are deliberate, using specifications that anticipate your paper 
-as one part of the entire proceedings, and not as an independent document. 
-Please do not revise any of the current designations.
+This work is structured as follows: subsection \ref{chap:background}  describes the SGM and its optimizations.
+Furthermore, we will give an extended explanation of the gradient descent optimizers used in this work.
+The reader is introduced to our Implementation in subsection \ref{chap:implementation}.
+Results are presented in  subsection \ref{chap:results} , where we will also describe used datasets, the measure applied to compare the quality of word embeddings, and finally the empirical results.
+The last part will focus on the discussion of our results, and possible future work in subsection \ref{chap:discussion} followed by a conclusion in subsection \ref{chap:conclusion}.
 
-\section{Prepare Your Paper Before Styling}
-Before you begin to format your paper, first write and save the content as a 
-separate text file. Complete all content and organizational editing before 
-formatting. Please note sections \ref{AA}--\ref{SCM} below for more information on 
-proofreading, spelling and grammar.
+\section{Background}\label{chap:background}
+This subsection gives an overview of the SGM and related work on its optimization. Furthermore, the used optimizers in our experiments are explained.
+\subsection{The Skip-Gram Model}
+The SGM is a very simple model used to learn WE.
+The idea is to train the model, a neural network, on a fake task and then use some of the weights of the neural network as embeddings.
+To understand this fake task center and context words must be defined.
+The center word is any given word in a sentence, from which we want to learn the WE.
+The context words of this specific center word are the words left and right in a given window $m$ in the sentence.
+See Figure \ref{fig:window_ex} as an example, where the context words are highlighted.
+An important thing to notice, is that instead of having a fixed window size $m$, each word will randomly have a window size between $1$ and $m$.
+The idea behind having different window sizes is that words that are further away in the sentence, have less semantic correlation to the center word.
 
-Keep your text and graphic files separate until after the text has been 
-formatted and styled. Do not number text heads---{\LaTeX} will do that 
-for you.
+\begin{figure}[h]
+\centering
+\includegraphics[scale=0.3]{images/window_ex}
+\caption{Example of a center word and it's context words}
+\label{fig:window_ex}
+\end{figure}
+%TODO think about example matrix
+With these definitions the fake task can be defined: given a pair of center and context words, which will be our training samples, the goal of the network is to predict the probability of each word to appear in the context of the center word.
+One may ask how such a network is built.
+The network consists of two matrices, a projection and an output layer.
+Both of these matrices store word embeddings.
+The first one, the projection layer, will store the WE for one specific word in one specific row.
+It will have dimension $T\times d$, with $T$ being the size of our vocabulary and $d$ the dimension of our WE.
+The second layer, the output layer, will store one WE in each column.
+The idea behind these 2 layers is that the input layer will represent our words as center word and the output layer as context words.
+Since the task of the network being to predict the probability of each word appearing in the context of a given center word, the following probability should be maximized:\\
+\begin{equation} \label{eq:basicSkip}
+\prod_{t=1}^T \prod_{-m<j<m} p(w_{t+j}|w_t)
+\end{equation}
+Where T is the number of words in the corpus data, $w_t$ the $t^{th}$ word in the corpus data and $m$ is the context window.
+This means that the $m$ nearest words to $w$ are considered as context words.
+Equation \ref{eq:basicSkip} can be transformed into sums by using log probabilities:
+\begin{equation}
+\sum _{t=1}^T \sum_{-m<j<m} log( p(w_{t+j}|w_t) )
+\end{equation}
+where the parameters are the same as in Equation \ref{eq:basicSkip}.
 
-\subsection{Abbreviations and Acronyms}\label{AA}
-Define abbreviations and acronyms the first time they are used in the text, 
-even after they have been defined in the abstract. Abbreviations such as 
-IEEE, SI, MKS, CGS, ac, dc, and rms do not have to be defined. Do not use 
-abbreviations in the title or heads unless they are unavoidable.
+The basic Skip-Gram Model uses a classical Softmax to calculate the conditional probability $p(w_{t+j}|w_t)$:
+\begin{equation} \label{eq:softmax}
+p(w_{t+j}|w_t)= \frac{exp( \tilde{v}_{w_{t+j}}^Tv_{w_t})}{\sum_{w=1}^v exp(\tilde{v}_w^Tv_{ w_t})}
+\end{equation}
 
-\subsection{Units}
+Here $\tilde{v}_{w_t}$ and $ v_{w_t}$ are the vector representations, stored respectively in the projection Layer (first Matrix) and in the output layer (second Matrix).
+There is a problem with the classical softmax.
+As a matter of fact, it is unsuitable to compute the softmax.
+For the computation of $\sum_{w=1}^v exp(\tilde{v_w}^T w_t)$, the denominator in Equation \ref{eq:softmax}, one has to go over the whole corpus data.
+As very big data sets are needed to train the model, this is not a solution.
+Mikolov et al.
+\cite{mikolov2} proposed different solutions.
+The first solution is to use a Hierarchical softmax introduced by Morin and Bengio \cite{hsoftmax}.
+In this model, the probability distribution of the output nodes is saved in a binary tree which gives one a logarithmic computation time for each of these probabilities and makes it suitable to compute the softmax.
+Another possibility is the use of negative sampling which is discussed in the next subsection.
+
+\subsection{Negative Sampling}
+A second alternative solution to the use of a classical softmax is negative sampling.
+Negative Sampling is a simplification of Noise Contrastive Estimation (NCE) which was introduced by Gutmann and Hyv{\"a}rinen \cite{nce-original}, and first applied to NLP by Mnih and Teh \cite{mnih}.
+This subsection will shortly describe NCE and then how Mikolov et al.
+\cite{mikolov2} simplified it to create the technique called negative sampling.
+\\ The idea behind NCE is to distinguish data from noise.
+It does so by reducing the problem to a logistic regression task and does it by maximizing the log probability.
+The SGM is only interested in good word representation, hence the probability of the word is not meaningful as long as the quality of the word representations remains high.
+Mikolov et al.
+\cite{mikolov2} simplified NCE and called it Negative Sampling, which will be explained next.\\
+The main goal of negative sampling is to only update the output nodes of certain words.
+This will obviously save an enormous amount of computation time.
+The idea is that given a pair $(c,w) \in D$, where $c$ is a word in the context window of $w$ we select $K$ random words $k_i$ from the corpus data $D$.
+We assume those words do not appear in the context of $w$.
+We denote the score that the $(c,w)$ wasn't drawn at random the following way: $p(y=1|c,w)$, and if $(k,w) $ is chosen at random this way: $p(y=0|k,w)$.
+Now we use logistic regression to update the weights of the $k$ selected context words and $c$.
+By doing so we will only have to update $k+1$ nodes.
+
+Let's look at how we construct our objective function for a given word $w$ and one of its context words $c$:
+
+\begin{align*}
+p(c|w) &= p(y=1|c,w) + \prod_{k\in K} p(y=0|k,c)
+\\&= p(y=1|c,w) + \prod_{k\in K} 1- p(y=1|k,c)
+\\&= log((p(y=1|c,w)) + \sum_{k\in K} log(1- p(y=1|k,c))
+\\&= log(\frac{1}{1+e^{-v_c \tilde{v_w }}}) + \sum_{k\in K} log(1-\frac{1}{1+e^{-v_c \tilde{v_k}}})
+\\&= log(\frac{1}{1+e^{-v_c \tilde{v_w } }}) + \sum_{k\in K} log(\frac{1}{1+e^{v_c \tilde{v_k} }})
+\\&= log(\sigma(v_c \tilde{v_w } ) + \sum_{k\in K} \sigma(log(-v_c \tilde{v_k} )) &&\text{where, $\sigma(x) = \frac{1}{1+e^{-x}}$}
+\end{align*}\label{eq:obj_neg_samples}
+
+Where $v_c$ and $\tilde{v_w }$, can be interpreted as before in Equation \ref{eq:softmax}.
+The goal is to maximize this objective function.
+Another way to look at this function, apart from logistic regression, and understand why it's working so well, is to assume that two vectors are similar if their dot product is high.
+Therefore, we maximize the dot product of similar words $(w,c)$ ($c$ appears in the context of $w$) and minimize it for dissimilar words $(w,k_i)$ (those were selected randomly).
+We see that to compute our objective function we will only have to compute the sum over the number of negative samples $K$, which is very small in practice (2-20).
+To put things in perspective lets imagine our data set consists of 100000 words, we set $K=2$.
+Assume each output neuron has a weight vector $v$ with $|v| = 300$.
+In consequence, when updating our weights we would only update $0.2*10^{-2}$ of the 300 million weights in the output layer.
+
+One question remains: how do we choose our random words? Mikolov et al. \cite{mikolov2} used the following unigram distribution:
+
+\begin{equation} \label{eq:unigram}
+P(w)=\frac{f(w)^{\frac{3}{4}}}{\sum_{t=0}^{T} f(w_t)^{\frac{3}{4}}}
+\end{equation}
+where $f(w)$ is the frequency of $w_t$.
+The value of $\frac{3}{4}$ is set empirically.
+Raising the unigram distribution to power of $\frac{3}{4}$ makes it less likely for a word to be drawn if it often appears in the dataset in comparison to the basic unigram distribution.
+See figure \ref{fig:frequency_ex} for an example.
+\begin{figure}[ht]
+\centering
+\includegraphics[scale=0.45]{images/frequency_ex}
+\caption{Probabililty of a word, of the text8 dataset (sampled), to be chosen at random according to its frequency and the power to which the unigram distribution is raised}
+\label{fig:frequency_ex}
+\end{figure}
+It's quite easily observable that this approach will outperform the classical softmax in computation time, as it only compute the sum over $K$ output nodes.
+Now the question arises if the accuracy is good enough but according to Mikolov et al.
+\cite{mikolov2} the negative sampling method "is an extremely simple training method that learns accurate representations".
+As a matter of fact, Mikolov et al.
+\cite{mikolov2} reported a 6\% accuracy improvement in comparison to a Hierarchical Softmax model.
+Therefore, it is a good solution to the problem raised by the classical softmax.
+We now have enough background knowledge about the SGM to look at how it can be optimized. The next subsection introduces earlier approaches to optimize the SGM.
+
+\section{Related Work}
+Due to the popularity of the SGM, a lot of research went into optimizing it.
+This research can be divided into two categories, optimization of the throughput and the optimization of the algorithm's accuracy.
+The latter was achieved by allowing words to have multiple meanings, also called context-sensitive word embedings.
+For our work, the optimization of the throughput is of big interest while the semantic optimization is aimed at giving the reader a more holistic comprehension of the possible research directions.
+This subsection will first give an overview of the optimization of the throughput and then present one paper that focused on context-sensitive word embeddings.
+
+\subsection{Optimization of the throughput}
+
+In the original model, the optimization is done with Stochastic Gradient Descent (SGD), which is a sequential algorithm.
+This process does not favor parallelization.
+To deal with this specific problem Mikolov et al.\cite{mikolov2} used a Hogwild tree proposed by Recht et al.\cite{hogwild}.
+The approach is to allow multiple threads to access a shared memory, in this case, the single model.
+In the original SGM, the threads are constructed as follows: at the beginning of training, the dataset is split into $N$ evenly sized chunks and each of these chunks will be processed by a one thread.
+The threads run parallel and have access to the shared memory.
+Therefore, overwriting errors are bound to happen.
+According to Recht et al.\cite{hogwild} the overwriting errors won't lead to a significant accuracy loss if the data is sparse enough.
+But in the case of NLP, the problem seems to be a bit more significant, and especially for word embedding, as many words share the same context words.
+There were several attempts at solving this issue, and we are going to cover a few of them in the following subsubsections.
+
+\paragraph{Parallelization by the use of caching}
+This idea was proposed by Vuurens et al.
+\cite{efficient}.
+The architecture used here is the basic skip gram model with a hierarchical softmax.
+The general idea is to cache the most frequently used nodes of the binary tree used to memorize the probability distribution and update them on the shared single model after a certain number of seen words (the paper used the number 10).
+The paper produced interesting results as they managed to decrease execution time by increasing the number of cores used for the calculation.
+This is very powerful because in the original implementation the execution time regressed after using more than 8 cores.
+It seems to indicate that too much overwriting was happening, as the number of concurrent threads surpasses a certain threshold.
+This can be seen in Figure \ref{fig:efficient}, where c31 is the model proposed by Vuurens et al.\cite{efficient}.
+The model did not suffer any accuracy loss in comparison to the original SGM model.
+\begin{figure}[ht]
+\centering
+\includegraphics[scale=0.3]{images/cachingEfficiency.png}
+\caption{Comparasion of the execution time in relation to the number of used cores \cite{efficient}}
+\label{fig:efficient}
+\end{figure}
+This work proposes a very good way to parallelize the SGM, as in particular, it allows using more cores during the computation.
+As this approached focused on the Hierarchical softmax, in contrast to our work which used negative sampling, the next subsection covers optimizations of the Skip-Gram Model with negative sampling(SGNS).
+
+\paragraph{Parallization in shared and Distributed Memory}
+The first parallelized solution which was proposed by Ji et al.
+\cite{intel}, is to try to reduce the cost of our vector multiplication.
+The main idea in this paper is to convert the level 1-BLAS (Basic linear subprogram) vector to vector operations to a level 3-BLAS matrix multiplication operation, which are efficiently implemented into hardware and consequently faster.
+This is achieved, by using the same negative samples for each context word of a given word $w$.
+Instead of using for each context word a vector to vector multiplication we can transform this into a matrix multiplication, under the assumption that we will not lose accuracy by sharing the same negative samples.
+The matrix multiplication can be represented in the following way.
+\[
+\begin{bmatrix}
+w \\
+w_{n_1} \\
+\vdots \\
+w_{n_k}\\
+\end{bmatrix}
+*
+\begin{bmatrix}
+w_{c_1}\\
+\vdots\\
+w_{c_{2m}}\\
+\end{bmatrix}
+\]
+
+where $w$ is our given word, $w_{n_1}...w_{n_k}$ are the shared negative samples, with $k \in [5,20]$, and $w_{c_1}...w_{c_2m}$ are the words inside of the context window $m$ of $w$, with $m \in [10,20]$, also called a batch of input context words.
+After each batch the model updates the weights of the used vectors.
+This model achieves a 3.6 fold increase in throughput, by only losing 1\% of accuracy.
+An aspect that is not as useful to us is that the experiments were done on CPU, as modern GPU's are often used in many machine learning libraries, as with CUDA for example, there still need work to be done to optimize it with GPU's.
+This was done by Seulki and Youngmin \cite{gpu}, which is described in the next subsection.
+
+\paragraph{Accelleration of word2vec by Using GPU's}
+Seulki and Youngmin \cite{gpu} focused on getting a better throughput on the SGM when using GPU's.
+As the SGM is a sequential algorithm, is is not easy to parallelize it, especially if one wants to parallelize the training of individual training samples.
+As the algorithm goes sequentially over a sentence, the samples next to each other, in order of execution, will almost every time have the same input word.
+Consequently, it's very hard to parallelize at this level.
+To solve this problem, Seulki and Youngmin \cite{gpu} proposed the idea to parallelize the update of each dimension of the word embedding, since those are completely independent of each other.
+They achieved this by mapping each dimension to a CUDA thread while mapping each sentence to a CUDA block.
+As each CUDA block runs independently, the training of the sentences is parallelized, and the fact that sentences have different length is of no problem.
+If the execution time of the GPU kernel is greater than time used to read the sentences, it could be a smart choice to use multiple GPU's.
+According to Seulki and Youngmin \cite{gpu}, if multiple GPU's are used, there is a need for synchronizing the model, which will hinder run time performance.
+They achieved their best results with 2 concurrent GPU'S.
+The accomplished results were very good as they managed a 20x speedup compared to a single threaded CPU execution, which is a 3x increase in comparison to the original C code from Mikolov et al.
+\cite{mikolov2}, with no loss in accuracy.
+The problem with this and all the above optimization is that the code is not easily available.
+Therefore, we need an optimized implementation of the SGM that is easily available.
+This is provided by Gensim \cite{gensim}, which will be outlined in the next subsection.
+%%%%%%%%%%%
+\paragraph{Gensim}\label{ssec:gensim}
+Gensim \cite{gensim} is a pure Python library that holds a state of the art implementation of the SGM.
+Gensim is written in Cython, which first allowed Gensim to have the same runtime as the original C code.
+Furthermore, it made use of BLAS's and precomputed sigmoid tables, while also trying to parallelize the training of different sentences.
+This finally yielded in a 4x speedup in run time.
+Gensim is an important tool as it allows us, as a python library, to compare our data rather easily.
+It was also used in related work \cite{intel} and is therefore of value, as it allows to us compare our work in a simpler way.
+This concludes our overview of the optimizations of the throughput of the SGM.
+In the next subsection, we give a quick outlook of what has been done in the field of context-sensitive word embeddings.
+
+\subsubsection{Context sensitive word embedding}
+A word has different meanings dependent on its context.
+This is a problem that is not addressed by the SGM.
+Some new models, that have taken this issue into consideration, were proposed.
+A lot of work has been done in this direction, Liu et al.\cite{topicalWE} and Bartunov et al.\cite{breaking} for example, but the one reporting the best results is Liu et al.
+\cite{contextWithTensor}.
+The concept of this approach is to change the way we compute the objective function and variables we use in our conditional probability.
+The idea is to look if a word given a certain context word matches to a topic.
+\textit{Bank} would match to \textit{finance} given the context word \textit{money}.
+\textit{Bank} would also match to \textit{nature} if \textit{river} was the given context word.
+But \textit{Bank} would not match to \textit{nature} with the context word \textit{money}.
+Now one could ask how to achieve such a context sensitive word embedding?
+First, we have to introduce new variables, therefore let's look at the objective function used:
+\begin{equation}
+J(\Omega) = \sum_{(w,t,c)\in D} \sum_{(w,\tilde{t},\tilde{c} \in{\tilde{D}})} max(0,1- g(w,t,c) + g(w,\tilde{t},\tilde{c})) \lambda||\Omega||_{2}^2
+\end{equation}
+
+This approach uses the same negative sample technique as described in the previous subsections, $D$ is the corpus data and $\tilde{D}$ is the set of negative samples and $\lambda$is the hyper parameter used for the standard $L_2$ standardization.
+What is interesting here is the function $g(w,c,t)$, where $w$ is a word, $c$ the context word, and $t$ the context in which the word appears.
+$g$ is defined as follows:
+\begin{equation}
+g(w,c,t) = u^T \sigma(w^TM^{[1:k]}t+V_c^T(w \oplus t) + b_c)
+\end{equation}
+where, $u, V_c, b_c$ are standard parameters for a neural network.
+$\oplus$ is the vector concatenation.
+The most important parameter is $M^{[1:k]}$, which is a tensor layer, the tensor layer is used because of its ability to model multiple interactions in the data, as this will be useful for multiple contexts.
+They used SGD for the optimization of this objective function.
+They achieved interesting results as shown in \ref{fig:multipleContext}.\\
+\begin{figure}[ht]
+\centering
+\includegraphics[scale=0.7]{images/multipleContext.png}
+\caption{"Nearest neighbor words model and Skip-
+Gram. The first line in each block is the results of Skip-Gram;
+and the rest lines are the results of our model" \cite{contextWithTensor}}
+\label{fig:multipleContext}
+\end{figure}
+
+This will conclude our overview of the related work. We will now give the reader an outline of the different Gradient Descent Optimizer used in our experiments.
+
+\section{Results}\label{chap:results}
+
+%Describe the experimental setup, the used datasets/parameters and the experimental results achieved
+
+This subsection gives an overview of the used datasets, the used metric to evaluate our models, the configuration of our model and finally, the experimental results achieved.
+
+\subsection{Dataset}\label{sec:dataset}
+In this implementation we only used the text8 \footnote{http://mattmahoney.net/dc/enwik8.zip} dataset.
+We chose this dataset for two reasons.
+First of all, it's a very small dataset, that allowed us to do a lot of computations.
+Secondly, this dataset was used in related work \cite{intel} hence giving us a very good benchmark.
+The text8 dataset consists of 1702 lines of 1000 words, with a vocabulary of roughly 63000 words.
+Conveniently, there is no punctuation in the dataset.
+Therefore, we had to choose between building arbitrary sentences and keeping the dataset as it is.
+We chose the first option because it gives us a faster computation time, and did not show any significant loss in quality empirically, as shown in Table \ref{table:with_20}.
+We chose a length of sentences of 20.
+Furthermore, we applied a technique called subsampling to reduce the data set size, which is explained in the following subsection.
+
+\begin{table}[]
+\centering
+\begin{tabular}{|l|l|l|l|l|}
+\hline
+Length of Sentences & 10 & 20 & 40 & 1 Document \\ \hline
+Training Time for one batch &8min & 10min & 11min & 18min \\ \hline
+Convergence Time &4 epochs & 3 epochs & 3 epochs & 3 epochs \\ \hline
+Word Similarity& 0.65 & 0.66 & 0.66 & 0.66 \\ \hline
+\end{tabular}
+\caption{Training and Convergence time according to choice of the length of sentences in text8 dataset}
+\label{table:with_20}
+\end{table}
+
+\subsubsection{Subsampling}
+Subsampling is a technique introduced by Mikolov et al.
+\cite{mikolov} to reduce the dataset size while at the same time increasing the quality of the dataset, i.e getting better word embeddings with it.
+The idea behind subsampling is the removal of very frequent words such as: "the, as, it, of".
+These words do not give an intrinsic value to the words that appear in its context.
+Therefore, the goal of subsampling is to delete such words from the dataset.
+This will decrease the computation time, as it will reduce the number of training samples, and should, in theory, increase the accuracy of the model.
+The increase in accuracy can also be explained by the fact that words that would not have appeared in the context of each other, may now do because words between have been deleted.
+To choose which word to delete, Mikolov et al. \cite{mikolov2} chose the following equation to compute the deletion of a word $w$ in the data set:
+\begin{equation} \label{eq:sampling}
+P(w) = 1- \sqrt{{\frac{t}{f(w)}}}
+\end{equation}
+
+where $f(w)$ is the frequency of w, and $t$ is a threshold set empirically.
+As Equation \ref{eq:sampling} is a probability, subsampling is not a deterministic procedure, words that may have been deleted with a threshold of $10^-2$ may stay in the dataset with a lower threshold, as shown in Table \ref{table:treshold_examples}.
+Mikolov et al.
+recommend a value between $0$ and $10^{-5}$, depending on the size of the dataset.
+We experimented with different values and $10^{-4}$ seemed the most suited.
+We did this by looking at a random set of sentences and judging the results manually.
+An example of the first sentence with different sampling thresholds can be found in Table \ref{table:treshold_examples}.
+The table shows the first 20 words of our dataset, without the words that were subsampled according to a threshold sample.
+Stats about subsampling can be found in Table \ref{table:treshold}.
+\begin{table}[h]
+\centering
+\begin{tabular}{|l|l|l|l|l|l|}
+\hline
+Sampling Treshhold & 0 & $ 10^{-1}$&$ 10^{-2}$& $10^{-3} $ &$10^{-4} $ \\ \hline
+Number of words in Dataset & 16 mio & 15mio & 11 mio & 8mio & 4 mio \\ \hline
+\end{tabular}
+\caption{Size of the preprocessed text8 dataset according to sampling treshold}
+\label{table:treshold}
+\end{table}
+
+\begin{table}[]
+\begin{tabular}{|l|l|}
+\hline
+Sampling Treshold & First sentence of Dataset \\ \hline
+$10^{-1}$ & \begin{tabular}[c]{@{}l@{}}Anarchism originated as a term of abuse first used against early\\ working class radicals including the diggers of the english\end{tabular} \\ \hline
+$10^{-2}$ & \begin{tabular}[c]{@{}l@{}}Anarchism originated as a term of abuse first used against early \\ working class radicals including diggers of english\end{tabular} \\ \hline
+$10^{-3}$ & \begin{tabular}[c]{@{}l@{}}Anarchism originated a term abuse first used against early \\ working class radicals including diggers the english\end{tabular} \\ \hline
+$10^{-4}$ & \begin{tabular}[c]{@{}l@{}}Anarchism originated abuse used against working class radicals\\ diggers english\end{tabular} \\ \hline
+$10^{-5}$ & against radicals diggers \\ \hline
+\end{tabular}
+\caption{Example of a sentence with different sampling tresholds}\label{table:treshold_examples}
+\end{table}
+
+\subsubsection{Min count}
+We also deleted every word that did not appear more than 5 times in our dataset.
+We got this technique from Gensim \cite{gensim}, that introduced this parameter into their training.
+This is a good technique because of three reasons: First certain words of our data sets do not appear in a common lexicon (twigleg, melqu), or come from a foreign language (Volksvereinigung), or are names and acronyms.
+Secondly, each document often has spelling mistakes, those (as long as the same spelling mistake does not appear too often, what should be avoided in practice) would be deleted by sampling too, as the words do not have any meaning.
+Lastly, a word that only appears one time in our dataset will be very dependent on its original initialization.
+This is the case because it will only be updated with its context pairs once, which is only a dozen of times in practice and then won't be updated anymore.
+For all the above reasons, we applied this technique.
+Similar to subsampling, it should in theory improve the quality of the word embeddings and will decrease the computation time.
+
+\subsection{Evaluating word embedings}
+Evaluating word embedding is not an easy task, such as evaluating the accuracy of a common classifier.
+We cannot split our data set into train and test set.
+As the task that the network is learning is of no interest to us.
+Therefore, we need to verify that our embedding is of quality with other techniques.
+To define quality we first need to define a measure of similarity between two vectors.
+This requires knowledge of the Cosine distance, which is introduced in the following subsection.
+\subsubsection{Cosine distance}
+The cosine similarity, this is not the cosine distance, of vectors $v$ and $w$ is the cosine of the angle between the two vectors It can be calculated by taking the dot product of $v$ and $w$ and dividing it by the magnitude of $v$ and $w$ multiplied with each other.
+We get:
+\begin{equation}
+cos\_sim(v,w) = \frac{v \cdot w}{|v| |w|}
+\end{equation}
+The cosine of 0\textdegree ~is 1, it's 0 for two vectors that are orthogonal to each other and vectors that point in the opposite direction will have a cosine of their angle of -1.
+This is not a good distance measure as -1 is smaller than 0, and therefore two vectors pointing away from each other would be closer than two orthogonal vectors, but by subtracting 1 from the cosine of the angle we can create a good distance measure between the two vectors.
+This distance does not take into account any order of magnitude.
+Hence, for our tasks, two vectors will be considered equal if they are of different magnitude but point in the same direction.
+Apart from measuring the quality of word embedding well, this technique has another advantage.
+By normalizing the vectors the calculation of the cosine angle becomes the dot product of the two vectors.
+Which can be computed very fast on modern GPU's.
+Knowing that we have a measure to compute the similarity of two vectors let us introduce a way to rate the quality of our embeddings.
+
+\subsubsection{Word similarity and wordsim353}
+To measure the quality of our word embedding we will need a dataset to compare our results too.
+We chose wordsim353\footnote{http://www.cs.technion.ac.il/~gabr/resources/data/wordsim353/wordsim353.zip} for this task, as it's the most used in the related literature.
+The data set consists of 353 pairs of words rated by humans on their similarity.
+The similarity score is in the range of 1 and 10, with 10 being the best score.
+An example for two of such pairs can be found in Figure \ref{fig:ws353_ex}.
+We will rank our embeddings on the Pearson correlation coefficient between the cosine distance and the scores attributed by humans, as this is a common procedure.
+
+\begin{table}[]
+\centering
+\begin{tabular}{|l|l|l|}
+\hline
+Word1 & Word2 & Score \\ \hline
+\textquote{FBI} & \textquote{Investigation} & \textquote{8.31} \\ \hline
+\textquote{Mars} & \textquote{scientist} & \textquote{5.63} \\ \hline
+\end{tabular}
+\caption{Example of pairs and their rating in wordsim353}
+\label{fig:ws353_ex}
+\end{table}
+
+\subsection{Configuration of the network}
+The skip gram model has a lot of possible parameters, that can be tuned.
+We experimented with different models and finally decided for one that we tried to optimize.
+This subsection will give a short overview of each parameter, where we will explain the process in which we chose the value of the given parameter.
+The explanation of the parameters will be structured as follows:
+\texttt{Parameter} - Description and tuning - \textit{Value}
 \begin{itemize}
-\item Use either SI (MKS) or CGS as primary units.
-(SI units are encouraged.) English units may be used as secondary units (in parentheses).
-An exception would be the use of English units as identifiers in trade, such as ``3.5-inch disk drive''.
-\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds.
-This often leads to confusion because equations do not balance dimensionally.
-If you must use mixed units, clearly state the units for each quantity that you use in an equation.
-\item Do not mix complete spellings and abbreviations of units: ``Wb/m\textsuperscript{2}'' or ``webers per square meter'', not ``webers/m\textsuperscript{2}''.
-Spell out units when they appear in text: ``.
-.
-.
-a few henries'', not ``.
-.
-.
-a few H''.
-\item Use a zero before decimal points: ``0.25'', not ``.25''. Use ``cm\textsuperscript{3}'', not ``cc''.)
+\item \texttt{Negative Samples:} Here we have to find a trade-off between, setting the parameter too high which will result in increased accuracy but a longer computation time.
+For smaller data sets a higher number of negative samples is often needed.
+In their original paper Mikolov et al.
+\cite{mikolov2} recommend a value of 5-20.
+We tested a few values in the range of 5 to 15, as 10 yielded state of the art results we chose this value.
+- \textit{10}
+\item \texttt{Context Window:} The bigger the window the more training examples the network will have, but if the window is too big the semantic meaning of the window will be erased.
+Mikolov et al.
+\cite{mikolov} proposed a setting between 2-10, as all our sentences are of size 20 we chose 5.
+- \textit{5}
+\item\texttt{ Dimension of the embedding}: Here the choice is less obvious, as the dimension needs to be high enough to capture the meaning, but cannot be too high as this leads to a decrease in performance as shown by Yin and Shen \cite{dimension_size}.
+We, therefore, used Gensim to find the best embedding possible.
+- \textit{100}
+\item \texttt{Batch size}: As described in subsection \ref{ssec:b_SGM}, there is a tradeoff to find between quality and training time.
+We first used a batch size of 5000, but then decide after non conclusive results (see \ref{ssec:bs_lf}) that 2000 would be better - \textit{2000}
+\item \texttt{Alpha}: learning rate, this hyperparameter was tuned in every optimizer therefore only the range will be indicated - \textit{(1e-5,1)}
 \end{itemize}
 
-\subsection{Equations}
-Number equations consecutively. To make your 
-equations more compact, you may use the solidus (~/~), the exp function, or 
-appropriate exponents. Italicize Roman symbols for quantities and variables, 
-but not Greek symbols. Use a long dash rather than a hyphen for a minus 
-sign. Punctuate equations with commas or periods when they are part of a 
-sentence, as in:
-\begin{equation}
-a+b=\gamma\label{eq}
-\end{equation}
+\iffalse
+\begin{table}[h]
+\centering
+\begin{tabular}{|l|l|}
+\hline
+Embedding Size & Word Similarity on Gensim \\ \hline
+50 & 0.65 \\ \hline
+100 & 0.67 \\ \hline
+200 & 0.65 \\ \hline
+300 & 0.63 \\ \hline
+\end{tabular}
+\caption{Word similarity in relation to the size of the embedding}
+\label{table:gensim_emb_size}
+\end{table}
+\fi
+\subsection{Input Shuffling}
+We used input shuffling as a technique to optimize the skip gram model.
+We will first describe input shuffling in a general way and then explain why we suppose that input shuffling could work well on the skip gram model.
+\\
+Let $X = {x_1...x_n}$ be our input data set. Input Shuffling describes the process of taking a random permutation of the dataset as input at each epoch.
+The idea behind this technique is to present our optimizer with different loss surfaces so that it's able to find the best optimum.
+Therefore, it's easier for the neural network to escape a local minimum.
+As for example if a network had converged to a local minimum after one epoch it could not escape it as all the parameters are the same.
+But if we change the shape of the loss function, by input shuffling, then there would be a greater probability for the network to escape the local minimum.
+\\
+There are two reasons why we think that input shuffling is particularly well suited for the skip gram model.
+The first one has to do with the fact that when we read our words sequentially that words that only appear very early will not benefit from the context words being already updated from others.
+The second idea is that we used the special batch technique described in Section \ref{ssec:b_SGM}.
+When using this technique and not using shuffling we will always have words that appear next to each other in a batch and will, therefore, update the same words at the same time.
+We, therefore, lose some quality.
+But if we would use input shuffling instead, then the words in one batch would likely not be similar and therefore only taking the average of a small part of pairs with the same words will be less likely.
 
-Be sure that the 
-symbols in your equation have been defined before or immediately following 
-the equation. Use ``\eqref{eq}'', not ``Eq.~\eqref{eq}'' or ``equation \eqref{eq}'', except at 
-the beginning of a sentence: ``Equation \eqref{eq} is . . .''
+\subsection{Convergence time}
+To optimize convergence time we have to define it first.
+Therefore, we used the already available implementation of Gensim \cite{gensim}.
+Gensim is an open source software that proposes an implementation of the SGNS in Python.
+It is also written in Cython, therefore it has a fast computation time, but can be used inside a python implementation.
+Together with the knowledge from Ji et al.\cite{intel} that a score of $0.66$ in the task of word similarity, with the text8 dataset, is the state of the art, we tested Gensim (more on this process in subsection \ref{chap:discussion} and found out that it took 4 epochs to converge.
+Therefore, we defined the following criteria for convergence: \\
+$\rho - \rho_{prev} < 0.009 \vee \rho = 0.66$ \\
+where $\rho$ is the Pearson coefficient on the wordsim353 task.
+We also stopped computation, if it took more then 20 epochs to converge.
 
-\subsection{\LaTeX-Specific Advice}
+\subsection{Results by optimizer}
+We ran multiple experiments for each optimizer.
+This subsection will only give an overview of the achieved results.
+Each subsection will give an explanation over the achieved result with a specific optimizer.
 
-Please use ``soft'' (e.g., \verb|\eqref{Eq}|) cross references instead
-of ``hard'' references (e.g., \verb|(1)|). That will make it possible
-to combine sections, add equations, or change the order of figures or
-citations without having to go through the file line by line.
+\subsubsection{SGD}
+The first challenge for each optimizer was to find a correct learning rate.
+As SGD is the optimizer used in Gensim \cite{gensim} we first tried the same learning rate as Gensim \cite{gensim} and then performed a random search to find a better one.
+As expected a bell curve shape resulted, a learning rate that is too high leads to diversion and a learning rate that is too low leads to a training time that is too slow.
+The best value that we found for the learning rate is $0.0075$.
+With this setting SGD converged in 11 epochs.
+The second experiment was to add input shuffling.
+As seen in Figure \ref{fig:results_sgd}, for almost every learning rate the convergence time decreased.
+Our model, with the best setting, now converges in only 7 epochs.
+Another interesting fact to point out from Figure \ref{fig:results_sgd} is that with input shuffling we achieved better results with higher learning rates.
+As for learning rates of $0.01$ and $0.025$ we did converge in 11 epochs with input shuffling but did not converge in 20 epochs without it.
 
-Please don't use the \verb|{eqnarray}| equation environment. Use
-\verb|{align}| or \verb|{IEEEeqnarray}| instead. The \verb|{eqnarray}|
-environment leaves unsightly spaces around relation symbols.
-
-Please note that the \verb|{subequations}| environment in {\LaTeX}
-will increment the main equation counter even when there are no
-equation numbers displayed. If you forget that, you might write an
-article in which the equation numbers skip from (17) to (20), causing
-the copy editors to wonder if you've discovered a new method of
-counting.
-
-{\BibTeX} does not work by magic. It doesn't get the bibliographic
-data from thin air but from .bib files. If you use {\BibTeX} to produce a
-bibliography you must send the .bib files.
-
-{\LaTeX} can't read your mind. If you assign the same label to a
-subsubsection and a table, you might find that Table I has been cross
-referenced as Table IV-B3.
-
-{\LaTeX} does not have precognitive abilities. If you put a
-\verb|\label| command before the command that updates the counter it's
-supposed to be using, the label will pick up the last counter to be
-cross referenced instead. In particular, a \verb|\label| command
-should not go before the caption of a figure or a table.
-
-Do not use \verb|\nonumber| inside the \verb|{array}| environment. It
-will not stop equation numbers inside \verb|{array}| (there won't be
-any anyway) and it might stop a wanted equation number in the
-surrounding equation.
-
-\subsection{Some Common Mistakes}\label{SCM}
-\begin{itemize}
-\item The word ``data'' is plural, not singular.
-\item The subscript for the permeability of vacuum $\mu_{0}$, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ``o''.
-\item In American English, commas, semicolons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation.
-When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks.
-A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this).
-(A parenthetical sentence is punctuated within the parentheses.)
-\item A graph within a graph is an ``inset'', not an ``insert''. The word alternatively is preferred to the word ``alternately'' (unless you really mean something that alternates).
-\item Do not use the word ``essentially'' to mean ``approximately'' or ``effectively''.
-\item In your paper title, if the words ``that uses'' can accurately replace the word ``using'', capitalize the ``u''; if not, keep using lower-cased.
-\item Be aware of the different meanings of the homophones ``affect'' and ``effect'', ``complement'' and ``compliment'', ``discreet'' and ``discrete'', ``principal'' and ``principle''.
-\item Do not confuse ``imply'' and ``infer''.
-\item The prefix ``non'' is not a word; it should be joined to the word it modifies, usually without a hyphen.
-\item There is no period after the ``et'' in the Latin abbreviation ``et al.''.
-\item The abbreviation ``i.e.'' means ``that is'', and the abbreviation ``e.g.'' means ``for example''.
-\end{itemize}
-An excellent style manual for science writers is \cite{b7}.
-
-\subsection{Authors and Affiliations}
-\textbf{The class file is designed for, but not limited to, six authors.} A 
-minimum of one author is required for all conference articles. Author names 
-should be listed starting from left to right and then moving down to the 
-next line. This is the author sequence that will be used in future citations 
-and by indexing services. Names should not be listed in columns nor group by 
-affiliation. Please keep your affiliations as succinct as possible (for 
-example, do not differentiate among departments of the same organization).
-
-\subsection{Identify the Headings}
-Headings, or heads, are organizational devices that guide the reader through 
-your paper. There are two types: component heads and text heads.
-
-Component heads identify the different components of your paper and are not 
-topically subordinate to each other. Examples include Acknowledgments and 
-References and, for these, the correct style to use is ``Heading 5''. Use 
-``figure caption'' for your Figure captions, and ``table head'' for your 
-table title. Run-in heads, such as ``Abstract'', will require you to apply a 
-style (in this case, italic) in addition to the style provided by the drop 
-down menu to differentiate the head from the text.
-
-Text heads organize the topics on a relational, hierarchical basis. For 
-example, the paper title is the primary text head because all subsequent 
-material relates and elaborates on this one topic. If there are two or more 
-sub-topics, the next level head (uppercase Roman numerals) should be used 
-and, conversely, if there are not at least two sub-topics, then no subheads 
-should be introduced.
-
-\subsection{Figures and Tables}
-\paragraph{Positioning Figures and Tables} Place figures and tables at the top and 
-bottom of columns. Avoid placing them in the middle of columns. Large 
-figures and tables may span across both columns. Figure captions should be 
-below the figures; table heads should appear above the tables. Insert 
-figures and tables after they are cited in the text. Use the abbreviation 
-``Fig.~\ref{fig}'', even at the beginning of a sentence.
-
-\begin{table}[htbp]
-\caption{Table Type Styles}
-\begin{center}
-\begin{tabular}{|c|c|c|c|}
+\begin{figure}[h]
+\centering
+\includegraphics[scale=0.45]{images/results_sgd_shuffle}
+\caption{Training time Stochastic Gradient Descent with input Shuffling}
+\label{fig:results_sgd}
+\end{figure}
+\subsubsection{Momentum and Nesterov}
+Momentum and NAG \cite{nag} both have an additional hyperparameter $\gamma$, that, as described in Section \ref{optimizers}, defines the percentage of the previous gradient that will be added to the current gradients.
+We set $\gamma = 0.9$ as this is the typical value and did not alter it during our experiments.
+Momentum and Nesterov alone respectively only slightly decrease or increase the convergence time.
+Momentum optimally converges in 9 epochs and Nesterov in 13.
+If we combine these optimizers with input shuffling, interestingly the same phenomena as with plain SGD appear.
+The convergence time gets better, 8 epochs for Momentum and 3 epochs for NAG.
+The phenomena that a higher learning rat yields better results also happens with both of the optimizers.
+As Momentum does not converge in 20 epochs with a learning rate of 0.002 but does in 8 with input shuffling.
+\begin{figure}[h]
+\centering
+\begin{minipage}{.5\textwidth}
+\centering
+\includegraphics[scale=0.3]{images/results_mom_shuffle}
+\caption{Training time Momentum with input Shuffling}
+\label{fig:results_mom}
+\label{fig:test1}
+\end{minipage}%
+\begin{minipage}{.5\textwidth}
+\centering
+\includegraphics[scale=0.3]{images/results_nag_shuffle}
+\caption{Training time Nesterov with input Shuffling}
+\label{fig:results_nag}
+\end{minipage}
+\end{figure}
+\subsubsection{Adagrad}
+Adagrad \cite{adagrad} is a very interesting tool for learning word embeddings as it decreases the learning rate for very frequent occurring features, and vice versa (this is explained in detail in Section \ref{ssec:adagrad}).
+Because words that appear very frequently often do not have a semantic gain that is as important as words that appear less frequently to their context words, it's good to have a lower learning rate.
+So, in theory, Adagrad is particularly well suited for our task.
+This was confirmed empirically as our model converged in 4 epochs.
+When combined with shuffling Adagrad only took 3 epochs to converge.
+This shows the tendency of the skip gram model to converge faster with input shuffling and the big impact of having different learning rate for each feature.
+Here it's interesting to notice that a higher learning rate combined with input shuffling did not yield better results than without shuffling.
+Both of our best results happened with a learning rate of $0.1$, as shown in Figure \ref{fig:results_adagrad_shuffle}.
+\begin{figure}[h]
+\centering
+\includegraphics[scale=0.45]{images/results_adagrad_shuffle}
+\caption{Training time Adagrad with input Shuffling}
+\label{fig:results_adagrad_shuffle}
+\end{figure}
+\subsubsection{Adadelta}
+In theory Adadelta \cite{adadelta} should outperform Adagrad as it's an extension of the former.
+Because it didn't have any learning rate to tune, we only did 2 experiments, with and without input shuffling.
+We are aware of the fact that there are additional hyperparameters but did, for simplicity reason, and because their effect is not as high as the learning rate in other optimizers, decide not to tune it.
+We left it to its default value $\rho = 0.9$.
+This parameter defines the percentage taken when calculating the exponentially decaying average of past gradients, as explained in \ref{ssec:adadelta} Adadelta did not manage to achieve a word similarity of 0.66.
+It only converged to a similarity of 0.59.
+It did this in 20 epochs without input shuffling and in 3 with input shuffling, as can be seen in Table \ref{table:results_adadelta}
+\begin{table}[h]
+\centering
+\begin{tabular}{|l|l|l|}
 \hline
-\textbf{Table}&\multicolumn{3}{|c|}{\textbf{Table Column Head}} \\
-\cline{2-4} 
-\textbf{Head} & \textbf{\textit{Table column subhead}}& \textbf{\textit{Subhead}}& \textbf{\textit{Subhead}} \\
-\hline
-copy& More table copy$^{\mathrm{a}}$& &  \\
-\hline
-\multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
+Adadelta Model & Convergence Time & Word similarity \\ \hline
+Without Shuffling & 20 & 0.59 \\ \hline
+With Shuffling & 3 & 0.59 \\ \hline
 \end{tabular}
-\label{tab1}
-\end{center}
+\caption{Convergence Time and Quality with Adadelta}
+\label{table:results_adadelta}
 \end{table}
 
-Figure Labels: Use 8 point Times New Roman for Figure labels. Use words 
-rather than symbols or abbreviations when writing Figure axis labels to 
-avoid confusing the reader. As an example, write the quantity 
-``Magnetization'', or ``Magnetization, M'', not just ``M''. If including 
-units in the label, present them within parentheses. Do not label axes only 
-with units. In the example, write ``Magnetization (A/m)'' or ``Magnetization 
-\{A[m(1)]\}'', not just ``A/m''. Do not label axes with a ratio of 
-quantities and units. For example, write ``Temperature (K)'', not 
-``Temperature/K''.
+\subsubsection{Adam}
+Adam is the most advanced of all the optimizers used in our experiments and did yield the best results as seen in Figure \ref{fig:results_adam_shuffle}.
+Adam converged in 3 epochs without shuffling and 2 with.
+This is the best result that we got with any optimizer.
+It's also interesting to note that as same as with Adagrad it did not react to input shuffling the same way as SGD did.
+In fact, it worked in the opposite direction, as we achieved our best result with input shuffling while having a lower learning rate $0.001$ then we used to achieve the best result without input shuffling $0.05$.
+\begin{figure}[h]
+    \centering
+            \includegraphics[scale=0.45]{images/results_adam_shuffle} 
+    \caption{Training time Adam with input Shuffling}
+    \label{fig:results_adam_shuffle}
+\end{figure}
 
-\section*{Acknowledgment}
+\section{Discussion}\label{chap:discussion}
 
-The preferred spelling of the word ``acknowledgment'' in America is without 
-an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B. 
-G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor 
-acknowledgments in the unnumbered footnote on the first page.
+%Discuss the results. What is the outcome of your experimetns?
+This subsection shortly discusses the empirical results and then extensively compares the findings of this work to the existing literature while trying to explain some differences.
+It is followed by a subsection describing the limitations and possible extensions of this work.
+This subsection is concluded by the problems we encountered during the implementation so others can avoid them in the future.
 
-\section*{References}
+\subsection{Our work}
+This subsection discusses our findings, first we will analyze how input shuffling could influence the learning rate and conclude with the discussion of unexpected results.
 
-Please number citations consecutively within brackets \cite{b1}. The 
-sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference 
-number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at 
-the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''
+\subsubsection{Shuffling and learning rate with SGD}
+When the model used SGD as an optimizer it was able to use higher learning rates when input shuffling was used, as shown in figures \ref{fig:results_sgd}, \ref{fig:results_mom} and \ref{fig:results_nag} Therefore arises the questions why these phenomena happen, especially as it did not happen with advanced optimizers.
+One possibility is the use of the batched version of the SGNS.
+In consequence, when the input is not shuffled the same word will often appear in one batch, hence the optimizer is presented with an average value of the gradient, which can be imprecise.
+This is counter-attacked by advanced optimizers as they have adaptive learning rates, i.e a frequent appearing feature will have a low learning rate.
 
-Number footnotes separately in superscripts. Place the actual footnote at 
-the bottom of the column in which it was cited. Do not put footnotes in the 
-abstract or reference list. Use letters for table footnotes.
+\subsubsection{Large differences with NAG and SGD when using shuffling}
+SGD and NAG both have very different values when using shuffling in comparison to unshuffled input, as shown in figures \ref{fig:results_sgd} and \ref{fig:results_nag}, plain SGD.
+We do not only attribute those results to input shuffling but partially also to a good random initialization guess.
+Due to a lack of time these results were not replicate more than once.
 
-Unless there are six authors or more give all authors' names; do not use 
-``et al.''. Papers that have not been published, even if they have been 
-submitted for publication, should be cited as ``unpublished'' \cite{b4}. Papers 
-that have been accepted for publication should be cited as ``in press'' \cite{b5}. 
-Capitalize only the first word in a paper title, except for proper nouns and 
-element symbols.
+\subsection{Comparison to Gensim}
+This subsection will compare our finding extensively to Gensim \cite{gensim}.
+As explained in Section \ref{ssec:gensim}, Gensim is optimized to have a very high throughput, this made it possible to achieve a lot of computations.
+Furthermore, Gensim provides access to the loss and the resulting word embeddings, which facilitated the comparison process.
 
-For papers published in translation journals, please give the English 
-citation first, followed by the original foreign-language citation \cite{b6}.
+\iffalse
+As Mikolov et al.
+published the original paper which introduced the SGNS, it is, of course, relevant to compare ourselves to their work.
+The first important thing to take under account is that Mikolov et al.
+only trained their model on a very large google news dataset incorporating more than 3 billion words.
+This makes the comparison of our work more difficult.
+But we will make some assumptions, as they could be of value if true.
+In their original paper, Mikolov et al.
+reported results from computations that took 1 and 3 epochs.
+We accord these good results, which clearly outperformed our SGD model and Gensim, to the very large dataset and furthermore, as a matter of fact, their results are better with 3 than with 1 epoch.
+We do not have any information about the convergence time or criterion.
+Hence it would be interesting to use their dataset for comparison.
+\\
+One thing we can compare is the quality of our word embeddings.
+Mikolov et al.
+did not report any results on their model with the text8 dataset, but they, therefore, published their code.
+Which was then used by Ji et al \cite{intel} and they tested the model on the text8 dataset.
+They reported a word similarity of 0.63 on the word similarity task.
+This is obviously outperformed by all our models.
+We did not find any explanation on why those results differ as much.
+\\
+The final assumption is that an advanced optimizer could maybe outperform SGD in terms of quality on a large data set. This will be discussed in further work.
+
+\fi
+
+\subsubsection{Configuration of Gensim}
+The training with Gensim has a lot of possible parameters an extended list can be found in the appendix.
+This subsection will only describe the parameters we changed from the default setting.
+The description of each parameter has the following form: \\
+\texttt{name} (type) -- \textit{Description} -- Value\\
+Parameters:
+\begin{itemize}
+\item \texttt{sentences} (iterable of iterables) -- \textit{Dataset} -- text8 document splitted into sentences of 20 words
+\item \texttt{size}(int) – \textit{Dimensionality of the word vectors } -- 100
+\item \texttt{window} (int) --\textit {Maximum distance between the current and predicted word within a sentence }-- 5
+\item \texttt{min\_count }(int) --\textit{ Ignores all words with total frequency lower than this }-- 5
+\item \texttt{workers} (int) -- \textit{ Number of threads used to train the model} -- 4
+\item \texttt{sg} ({0, 1}) --\textit{ Training algorithm: 1 for skip--gram}--1
+\item \texttt{negative} (int) --\textit{Number of negative samples}-- 10
+\item \texttt{ ns\_exponent} (float) --\textit{ Exponent in the unigram distribution, when choosing random samples, as shown in Equation \ref{eq:unigram} }-- 0.75
+\item \texttt{alpha} (float) --\textit{ The initial learning rate. }-- 0.025
+\item \texttt{min\_alpha} (float) --\textit{ Learning rate will linearly drop to min\_alpha as training progresses }--0.0001
+\item \texttt{sample} (float) --\textit{ Treshold for subsampling as described in Equation \ref{eq:sampling}} -- 1e--4
+\item \texttt{iter} (int) --\textit{ Number of iterations (epochs) over the corpus. }-- 10
+\
+\item \texttt{compute\_loss} (bool) --\textit{If True, loss is stored at the end of each batch}-- True
+\item \texttt{callbacks} (iterable of CallbackAny2Vec) --\textit{ Set of functions executed in order to follow the loss and the progress of the model in word similarity }-- see Appendix
+\end{itemize}
+
+\subsubsection{Gensim vs. SGD}
+As stated earlier, we are not going to compare this work to Gensim in run time.
+Gensim is heavily optimized and written in cython\footnote{https://rare-technologies.com/word2vec-in-python-part-two-optimizing/}, which is 23x faster than plain Numpy.
+Since we used PyTorch the difference is not that big, but still remains.
+As shown in Figure \ref{fig:gensim_vs_adam} the convergence time was not the same between our implementation and Gensim.
+There are different possible reasons why this could be the case:\\ First, our batched approach could hinder performance in terms of convergence since our loss function is not exactly the same.
+When a word appears more than once in our batch, the gradient will be an average over the gradients of each pair alone, as it is done by Gensim.\\ Secondly, a difference to our implementation is the fact that Gensim checks whether negative samples are not equal to the context word.
+If that is the case Gensim selects a new random sample.
+Therefore, the learning of the input and output context is optimized.
+\\Finally, another possibility is the decay of the learning rate used by Gensim.
+In fact, decaying the learning rate has been proven in a lot of work to decrease the convergence time.
+Gensim linearly decreases the learning rate, as we did not use this technique, the decay of the learning rate could help explain the noted differences.
+\\ The first hypothesis, the fact that we used a batched approach, may be confirmed by the fact that when combined with input shuffling SGD does perform closer to Gensim, going from 11 to 7 epochs to converge, as input shuffling reduces the number of co-occurrence of the same word in a batch.
+Now the question arises if the 3 epochs, that Gensim is better, can be explained by the selection of better negative samples and the learning rate decay?
+
+\subsubsection{Gensim vs. Adam}
+The Adam optimizer did outperform the Gensim application in quality of word embeddings (only slightly: 0.01 correlation coefficient better) and convergence time.
+Adam converged in 2 epochs while Gensim in 4.
+To confirm the achieved results we ran each computation 40 times.
+The results can be seen in Figure \ref{fig:gensim_vs_adam}.
+
+\begin{figure}[h]
+\centering
+\includegraphics[scale=0.45]{images/gensim_vs_adam}
+\caption{Convergence time of SGD and Adam compared to Gensim}
+\label{fig:gensim_vs_adam}
+\end{figure}\subsection{Challenges faced}
+During our implementation we did encounter some problems, this subsection has the purpose of preventing the scientific community from making the same mistakes.
+\subsubsection{Using the wrong embeddings}
+To start with, we did not choose the same initialization value as Gensim in our word embeddings.
+We initialized the projection layer with a normal distribution between (-1,1), as opposed to Gensim which initializes all the weights to 0.
+The reason for this, is that at the beginning of our experiments we did not achieve good results with the same initialization as Gensim.
+Retrospectively we accord this to a learning rate that was too low.
+After a few simulations, we saw that we did not perform as good as Gensim.
+And as we changed the initialization of the projection layer back to 0, and already had adjusted the learning rate, we achieved good results.
+Therefore, this is a recommendation to future work to not set the initialization to (-1,1).
+
+\subsubsection{Batch size and loss function adjustements}\label{ssec:bs_lf}
+During our experiments, we faced a moment where we needed to us a very high learning rate (50) to achieve good results.
+As this is the complete opposite of what is standard we made a few changes to remedy the problem.\\
+First, we needed to find a good batch size. During the above-explained raise of the learning rate, we used a batch size of 5000. We then decided to take a batch size of 2000.\\
+Secondly, at the same time, we did experiment with different loss functions.
+As our batched model does not have exactly the same loss function as introduced by Mikolov et al.
+\cite{mikolov}, we needed to find one that suited our goals the best.
+At the beginning of our experiment, we took the average of all the scores stored in our final matrix, as explained in \ref{ssec:b_SGM}.
+But then chose to take the sum as the training did not seem optimal.\\
+With these two changes, we did increase convergence time and word similarity, while at the same time having a usual learning rate. \\
+Know the question arises why this happened?
+Is this because of the shape of the loss function better suits our advanced optimizers?
+Or because the loss function is closer to its original form?
+Our hypothesis looks the following way:
+When taking the average, our loss function is represented as follows:
+\begin{equation}
+\frac{-1}{b }* \sum_{x\in X} loss(x)
+\end{equation}
+where $X$ is our batch and $b = |X|$.
+When taking the partial derivative of each parameter, the constant $ \frac{-1}{b }$ will remain the same.
+Therefore, it's the same as taking the sum as the loss function and multiplying each gradient with $\frac{-1}{b }$.
+As our batch size was really high, i.e 2000 and 5000, this could highly influence the learning rate.
+This is also the conclusion of Goyal et al.
+\cite{fb}, that worked with very high batch sizes, i.e 8000: "When the minibatch size is multiplied by k, multiply the learning rate by k".
+
+\subsection{Future Work}
+This work shows that the convergence time of the SGNS could be improved by using input shuffling and advanced optimizers.
+As with every work, there still exists possible extensions.
+First and foremost an aspect of our implementation that can be prejudicial is that we only extensively tested our model with one small dataset.
+The fact that we only used one dataset as well that it's a small one is problematic.
+By using a very small dataset we do not use the model in the condition it is most needed for, as the dataset used in practice usually consists of more than 1 billion words.
+There is a small argument that can be made for machine translation as the use of small parallel corpus is not unusual in this field.
+But the main issue with using only one data set it that it has been shown that some optimizers perform better with specific shapes of loss functions.
+To make a compelling argument it's, therefore, necessary to show that our model with the use of Input shuffling and Adam as its optimizer also outperform Gensim with other data sets, so that the claim can hold consistently, it needs to be confirmed with other datasets as well.
+Furthermore, our implementation did not outperform Gensim in run time, as this was not the goal of our work.
+Therefore, one could improve an already existing, optimized version, with input shuffling and advanced optimizers and should achieve a better run time than Gensim.
+\section{Conclusion}\label{chap:conclusion}
+
+This work provides an overview of the Skip Gram Model with negative Sampling (SGNS) and the numerous successful attempts of optimizing the throughput of the model.
+As this is the case, no effort went into optimizng the convergence time of the SGNS, therefore this work focused on this point.
+We decided to use advanced optimizers and input shuffling as optimizing techniques.
+After giving a short overview over Gradient Descent algorithms this work proposes a slighlty altered version of the SGNS, where the idea is to compute the loss over the sum of a high number of training samples, i.e 2000,  instead of computing it for each individually.
+We did this as it allowed us to compute more models and analyze the convergence time faster.
+We used the text8 dataset and used  word similarity as a quality measure for the word embeddings (WE).
+We used the State of the art implementation Gensim to compare our self.
+We did achieve a better convergence time than gensim with Adam as an optimizer and the use of input shuffling.
+Gensim convereged in 4 epochs to a word similarity of 0.66 and our model only took 2 epochs to achieve the same quality.
+Those results still need to be confirmed with more datasets.
+Finally, if this work would be combined with an optimized throughput it  could improve the state of the art run time of the SGNS.
 

### 2019-05-01 icdm/submission/includes/background.tex
Splitted papaer into multiple files

--- 2019-04-30 icdm/submission/paper.tex
@@
-\documentclass[conference]{IEEEtran}
-\IEEEoverridecommandlockouts
-% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
-\usepackage{cite}
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{algorithmic}
-\usepackage{graphicx}
-\usepackage{textcomp}
-\usepackage{xcolor}
-\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
-    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
-\begin{document}
+\section{Background}\label{chap:background}
+This subsection gives an overview of the SGM and related work on its optimization. Furthermore, the used optimizers in our experiments are explained.
+\subsection{The Skip-Gram Model}
+The SGM is a very simple model used to learn WE.
+The idea is to train the model, a neural network, on a fake task and then use some of the weights of the neural network as embeddings.
+To understand this fake task center and context words must be defined.
+The center word is any given word in a sentence, from which we want to learn the WE.
+The context words of this specific center word are the words left and right in a given window $m$ in the sentence.
+See Figure \ref{fig:window_ex} as an example, where the context words are highlighted.
+An important thing to notice, is that instead of having a fixed window size $m$, each word will randomly have a window size between $1$ and $m$.
+The idea behind having different window sizes is that words that are further away in the sentence, have less semantic correlation to the center word.
 
-\title{Paper Title*\\
-{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
-should not be used}
-\thanks{Identify applicable funding agency here. If none, delete this.}
-}
+\begin{figure}[h]
+\centering
+\includegraphics[scale=0.3]{images/window_ex}
+\caption{Example of a center word and it's context words}
+\label{fig:window_ex}
+\end{figure}
+%TODO think about example matrix
+With these definitions the fake task can be defined: given a pair of center and context words, which will be our training samples, the goal of the network is to predict the probability of each word to appear in the context of the center word.
+One may ask how such a network is built.
+The network consists of two matrices, a projection and an output layer.
+Both of these matrices store word embeddings.
+The first one, the projection layer, will store the WE for one specific word in one specific row.
+It will have dimension $T\times d$, with $T$ being the size of our vocabulary and $d$ the dimension of our WE.
+The second layer, the output layer, will store one WE in each column.
+The idea behind these 2 layers is that the input layer will represent our words as center word and the output layer as context words.
+Since the task of the network being to predict the probability of each word appearing in the context of a given center word, the following probability should be maximized:\\
+\begin{equation} \label{eq:basicSkip}
+\prod_{t=1}^T \prod_{-m<j<m} p(w_{t+j}|w_t)
+\end{equation}
+Where T is the number of words in the corpus data, $w_t$ the $t^{th}$ word in the corpus data and $m$ is the context window.
+This means that the $m$ nearest words to $w$ are considered as context words.
+Equation \ref{eq:basicSkip} can be transformed into sums by using log probabilities:
+\begin{equation}
+\sum _{t=1}^T \sum_{-m<j<m} log( p(w_{t+j}|w_t) )
+\end{equation}
+where the parameters are the same as in Equation \ref{eq:basicSkip}.
 
-\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-\and
-\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-\and
-\IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-\and
-\IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-\and
-\IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-\and
-\IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-}
-
-\maketitle
-
-\begin{abstract}
-This document is a model and instructions for \LaTeX.
-This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
-or Math in Paper Title or Abstract.
-\end{abstract}
-
-\begin{IEEEkeywords}
-component, formatting, style, styling, insert
-\end{IEEEkeywords}
-
-\section{Introduction}
-This document is a model and instructions for \LaTeX.
-Please observe the conference page limits.
-
-\section{Ease of Use}
-
-\subsection{Maintaining the Integrity of the Specifications}
-
-The IEEEtran class file is used to format your paper and style the text. All margins, 
-column widths, line spaces, and text fonts are prescribed; please do not 
-alter them. You may note peculiarities. For example, the head margin
-measures proportionately more than is customary. This measurement 
-and others are deliberate, using specifications that anticipate your paper 
-as one part of the entire proceedings, and not as an independent document. 
-Please do not revise any of the current designations.
-
-\section{Prepare Your Paper Before Styling}
-Before you begin to format your paper, first write and save the content as a 
-separate text file. Complete all content and organizational editing before 
-formatting. Please note sections \ref{AA}--\ref{SCM} below for more information on 
-proofreading, spelling and grammar.
-
-Keep your text and graphic files separate until after the text has been 
-formatted and styled. Do not number text heads---{\LaTeX} will do that 
-for you.
-
-\subsection{Abbreviations and Acronyms}\label{AA}
-Define abbreviations and acronyms the first time they are used in the text, 
-even after they have been defined in the abstract. Abbreviations such as 
-IEEE, SI, MKS, CGS, ac, dc, and rms do not have to be defined. Do not use 
-abbreviations in the title or heads unless they are unavoidable.
-
-\subsection{Units}
-\begin{itemize}
-\item Use either SI (MKS) or CGS as primary units.
-(SI units are encouraged.) English units may be used as secondary units (in parentheses).
-An exception would be the use of English units as identifiers in trade, such as ``3.5-inch disk drive''.
-\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds.
-This often leads to confusion because equations do not balance dimensionally.
-If you must use mixed units, clearly state the units for each quantity that you use in an equation.
-\item Do not mix complete spellings and abbreviations of units: ``Wb/m\textsuperscript{2}'' or ``webers per square meter'', not ``webers/m\textsuperscript{2}''.
-Spell out units when they appear in text: ``.
-.
-.
-a few henries'', not ``.
-.
-.
-a few H''.
-\item Use a zero before decimal points: ``0.25'', not ``.25''. Use ``cm\textsuperscript{3}'', not ``cc''.)
-\end{itemize}
-
-\subsection{Equations}
-Number equations consecutively. To make your 
-equations more compact, you may use the solidus (~/~), the exp function, or 
-appropriate exponents. Italicize Roman symbols for quantities and variables, 
-but not Greek symbols. Use a long dash rather than a hyphen for a minus 
-sign. Punctuate equations with commas or periods when they are part of a 
-sentence, as in:
-\begin{equation}
-a+b=\gamma\label{eq}
+The basic Skip-Gram Model uses a classical Softmax to calculate the conditional probability $p(w_{t+j}|w_t)$:
+\begin{equation} \label{eq:softmax}
+p(w_{t+j}|w_t)= \frac{exp( \tilde{v}_{w_{t+j}}^Tv_{w_t})}{\sum_{w=1}^v exp(\tilde{v}_w^Tv_{ w_t})}
 \end{equation}
 
-Be sure that the 
-symbols in your equation have been defined before or immediately following 
-the equation. Use ``\eqref{eq}'', not ``Eq.~\eqref{eq}'' or ``equation \eqref{eq}'', except at 
-the beginning of a sentence: ``Equation \eqref{eq} is . . .''
+Here $\tilde{v}_{w_t}$ and $ v_{w_t}$ are the vector representations, stored respectively in the projection Layer (first Matrix) and in the output layer (second Matrix).
+There is a problem with the classical softmax.
+As a matter of fact, it is unsuitable to compute the softmax.
+For the computation of $\sum_{w=1}^v exp(\tilde{v_w}^T w_t)$, the denominator in Equation \ref{eq:softmax}, one has to go over the whole corpus data.
+As very big data sets are needed to train the model, this is not a solution.
+Mikolov et al.
+\cite{mikolov2} proposed different solutions.
+The first solution is to use a Hierarchical softmax introduced by Morin and Bengio \cite{hsoftmax}.
+In this model, the probability distribution of the output nodes is saved in a binary tree which gives one a logarithmic computation time for each of these probabilities and makes it suitable to compute the softmax.
+Another possibility is the use of negative sampling which is discussed in the next subsection.
 
-\subsection{\LaTeX-Specific Advice}
+\subsection{Negative Sampling}
+A second alternative solution to the use of a classical softmax is negative sampling.
+Negative Sampling is a simplification of Noise Contrastive Estimation (NCE) which was introduced by Gutmann and Hyv{\"a}rinen \cite{nce-original}, and first applied to NLP by Mnih and Teh \cite{mnih}.
+This subsection will shortly describe NCE and then how Mikolov et al.
+\cite{mikolov2} simplified it to create the technique called negative sampling.
+\\ The idea behind NCE is to distinguish data from noise.
+It does so by reducing the problem to a logistic regression task and does it by maximizing the log probability.
+The SGM is only interested in good word representation, hence the probability of the word is not meaningful as long as the quality of the word representations remains high.
+Mikolov et al.
+\cite{mikolov2} simplified NCE and called it Negative Sampling, which will be explained next.\\
+The main goal of negative sampling is to only update the output nodes of certain words.
+This will obviously save an enormous amount of computation time.
+The idea is that given a pair $(c,w) \in D$, where $c$ is a word in the context window of $w$ we select $K$ random words $k_i$ from the corpus data $D$.
+We assume those words do not appear in the context of $w$.
+We denote the score that the $(c,w)$ wasn't drawn at random the following way: $p(y=1|c,w)$, and if $(k,w) $ is chosen at random this way: $p(y=0|k,w)$.
+Now we use logistic regression to update the weights of the $k$ selected context words and $c$.
+By doing so we will only have to update $k+1$ nodes.
 
-Please use ``soft'' (e.g., \verb|\eqref{Eq}|) cross references instead
-of ``hard'' references (e.g., \verb|(1)|). That will make it possible
-to combine sections, add equations, or change the order of figures or
-citations without having to go through the file line by line.
+Let's look at how we construct our objective function for a given word $w$ and one of its context words $c$:
 
-Please don't use the \verb|{eqnarray}| equation environment. Use
-\verb|{align}| or \verb|{IEEEeqnarray}| instead. The \verb|{eqnarray}|
-environment leaves unsightly spaces around relation symbols.
+\begin{align*}
+p(c|w) &= p(y=1|c,w) + \prod_{k\in K} p(y=0|k,c)
+\\&= p(y=1|c,w) + \prod_{k\in K} 1- p(y=1|k,c)
+\\&= log((p(y=1|c,w)) + \sum_{k\in K} log(1- p(y=1|k,c))
+\\&= log(\frac{1}{1+e^{-v_c \tilde{v_w }}}) + \sum_{k\in K} log(1-\frac{1}{1+e^{-v_c \tilde{v_k}}})
+\\&= log(\frac{1}{1+e^{-v_c \tilde{v_w } }}) + \sum_{k\in K} log(\frac{1}{1+e^{v_c \tilde{v_k} }})
+\\&= log(\sigma(v_c \tilde{v_w } ) + \sum_{k\in K} \sigma(log(-v_c \tilde{v_k} )) &&\text{where, $\sigma(x) = \frac{1}{1+e^{-x}}$}
+\end{align*}\label{eq:obj_neg_samples}
 
-Please note that the \verb|{subequations}| environment in {\LaTeX}
-will increment the main equation counter even when there are no
-equation numbers displayed. If you forget that, you might write an
-article in which the equation numbers skip from (17) to (20), causing
-the copy editors to wonder if you've discovered a new method of
-counting.
+Where $v_c$ and $\tilde{v_w }$, can be interpreted as before in Equation \ref{eq:softmax}.
+The goal is to maximize this objective function.
+Another way to look at this function, apart from logistic regression, and understand why it's working so well, is to assume that two vectors are similar if their dot product is high.
+Therefore, we maximize the dot product of similar words $(w,c)$ ($c$ appears in the context of $w$) and minimize it for dissimilar words $(w,k_i)$ (those were selected randomly).
+We see that to compute our objective function we will only have to compute the sum over the number of negative samples $K$, which is very small in practice (2-20).
+To put things in perspective lets imagine our data set consists of 100000 words, we set $K=2$.
+Assume each output neuron has a weight vector $v$ with $|v| = 300$.
+In consequence, when updating our weights we would only update $0.2*10^{-2}$ of the 300 million weights in the output layer.
 
-{\BibTeX} does not work by magic. It doesn't get the bibliographic
-data from thin air but from .bib files. If you use {\BibTeX} to produce a
-bibliography you must send the .bib files.
+One question remains: how do we choose our random words? Mikolov et al. \cite{mikolov2} used the following unigram distribution:
 
-{\LaTeX} can't read your mind. If you assign the same label to a
-subsubsection and a table, you might find that Table I has been cross
-referenced as Table IV-B3.
-
-{\LaTeX} does not have precognitive abilities. If you put a
-\verb|\label| command before the command that updates the counter it's
-supposed to be using, the label will pick up the last counter to be
-cross referenced instead. In particular, a \verb|\label| command
-should not go before the caption of a figure or a table.
-
-Do not use \verb|\nonumber| inside the \verb|{array}| environment. It
-will not stop equation numbers inside \verb|{array}| (there won't be
-any anyway) and it might stop a wanted equation number in the
-surrounding equation.
-
-\subsection{Some Common Mistakes}\label{SCM}
-\begin{itemize}
-\item The word ``data'' is plural, not singular.
-\item The subscript for the permeability of vacuum $\mu_{0}$, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ``o''.
-\item In American English, commas, semicolons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation.
-When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks.
-A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this).
-(A parenthetical sentence is punctuated within the parentheses.)
-\item A graph within a graph is an ``inset'', not an ``insert''. The word alternatively is preferred to the word ``alternately'' (unless you really mean something that alternates).
-\item Do not use the word ``essentially'' to mean ``approximately'' or ``effectively''.
-\item In your paper title, if the words ``that uses'' can accurately replace the word ``using'', capitalize the ``u''; if not, keep using lower-cased.
-\item Be aware of the different meanings of the homophones ``affect'' and ``effect'', ``complement'' and ``compliment'', ``discreet'' and ``discrete'', ``principal'' and ``principle''.
-\item Do not confuse ``imply'' and ``infer''.
-\item The prefix ``non'' is not a word; it should be joined to the word it modifies, usually without a hyphen.
-\item There is no period after the ``et'' in the Latin abbreviation ``et al.''.
-\item The abbreviation ``i.e.'' means ``that is'', and the abbreviation ``e.g.'' means ``for example''.
-\end{itemize}
-An excellent style manual for science writers is \cite{b7}.
-
-\subsection{Authors and Affiliations}
-\textbf{The class file is designed for, but not limited to, six authors.} A 
-minimum of one author is required for all conference articles. Author names 
-should be listed starting from left to right and then moving down to the 
-next line. This is the author sequence that will be used in future citations 
-and by indexing services. Names should not be listed in columns nor group by 
-affiliation. Please keep your affiliations as succinct as possible (for 
-example, do not differentiate among departments of the same organization).
-
-\subsection{Identify the Headings}
-Headings, or heads, are organizational devices that guide the reader through 
-your paper. There are two types: component heads and text heads.
-
-Component heads identify the different components of your paper and are not 
-topically subordinate to each other. Examples include Acknowledgments and 
-References and, for these, the correct style to use is ``Heading 5''. Use 
-``figure caption'' for your Figure captions, and ``table head'' for your 
-table title. Run-in heads, such as ``Abstract'', will require you to apply a 
-style (in this case, italic) in addition to the style provided by the drop 
-down menu to differentiate the head from the text.
-
-Text heads organize the topics on a relational, hierarchical basis. For 
-example, the paper title is the primary text head because all subsequent 
-material relates and elaborates on this one topic. If there are two or more 
-sub-topics, the next level head (uppercase Roman numerals) should be used 
-and, conversely, if there are not at least two sub-topics, then no subheads 
-should be introduced.
-
-\subsection{Figures and Tables}
-\paragraph{Positioning Figures and Tables} Place figures and tables at the top and 
-bottom of columns. Avoid placing them in the middle of columns. Large 
-figures and tables may span across both columns. Figure captions should be 
-below the figures; table heads should appear above the tables. Insert 
-figures and tables after they are cited in the text. Use the abbreviation 
-``Fig.~\ref{fig}'', even at the beginning of a sentence.
-
-\begin{table}[htbp]
-\caption{Table Type Styles}
-\begin{center}
-\begin{tabular}{|c|c|c|c|}
-\hline
-\textbf{Table}&\multicolumn{3}{|c|}{\textbf{Table Column Head}} \\
-\cline{2-4} 
-\textbf{Head} & \textbf{\textit{Table column subhead}}& \textbf{\textit{Subhead}}& \textbf{\textit{Subhead}} \\
-\hline
-copy& More table copy$^{\mathrm{a}}$& &  \\
-\hline
-\multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
-\end{tabular}
-\label{tab1}
-\end{center}
-\end{table}
-
-Figure Labels: Use 8 point Times New Roman for Figure labels. Use words 
-rather than symbols or abbreviations when writing Figure axis labels to 
-avoid confusing the reader. As an example, write the quantity 
-``Magnetization'', or ``Magnetization, M'', not just ``M''. If including 
-units in the label, present them within parentheses. Do not label axes only 
-with units. In the example, write ``Magnetization (A/m)'' or ``Magnetization 
-\{A[m(1)]\}'', not just ``A/m''. Do not label axes with a ratio of 
-quantities and units. For example, write ``Temperature (K)'', not 
-``Temperature/K''.
-
-\section*{Acknowledgment}
-
-The preferred spelling of the word ``acknowledgment'' in America is without 
-an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B. 
-G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor 
-acknowledgments in the unnumbered footnote on the first page.
-
-\section*{References}
-
-Please number citations consecutively within brackets \cite{b1}. The 
-sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference 
-number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at 
-the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''
-
-Number footnotes separately in superscripts. Place the actual footnote at 
-the bottom of the column in which it was cited. Do not put footnotes in the 
-abstract or reference list. Use letters for table footnotes.
-
-Unless there are six authors or more give all authors' names; do not use 
-``et al.''. Papers that have not been published, even if they have been 
-submitted for publication, should be cited as ``unpublished'' \cite{b4}. Papers 
-that have been accepted for publication should be cited as ``in press'' \cite{b5}. 
-Capitalize only the first word in a paper title, except for proper nouns and 
-element symbols.
-
-For papers published in translation journals, please give the English 
-citation first, followed by the original foreign-language citation \cite{b6}.
-
-\bibliographystyle{IEEEtran}
-\bibliography{refs}
-
-% use bibfile instead (2 lines above)
-% \begin{thebibliography}{00}
-% \bibitem{b1} G.
-Eason, B.
-Noble, and I.
-N.
-Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil.
-Trans.
-Roy.
-Soc.
-London, vol.
-A247, pp.
-529--551, April 1955.
-% \bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
-% \bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
-% \bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
-% \bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
-% \bibitem{b6} Y.
-Yorozu, M.
-Hirano, K.
-Oka, and Y.
-Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl.
-J.
-Magn.
-Japan, vol.
-2, pp.
-740--741, August 1987 [Digests 9th Annual Conf.
-Magnetics Japan, p.
-301, 1982].
-% \bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
-% \end{thebibliography}
-\vspace{12pt}
-\color{red}
-IEEE conference templates contain guidance text for composing and formatting conference papers.
-Please ensure that all template text is removed from your conference paper prior to submission to the conference.
-Failure to remove the template text from your paper may result in your paper not being published.
-
-\end{document}
+\begin{equation} \label{eq:unigram}
+P(w)=\frac{f(w)^{\frac{3}{4}}}{\sum_{t=0}^{T} f(w_t)^{\frac{3}{4}}}
+\end{equation}
+where $f(w)$ is the frequency of $w_t$.
+The value of $\frac{3}{4}$ is set empirically.
+Raising the unigram distribution to power of $\frac{3}{4}$ makes it less likely for a word to be drawn if it often appears in the dataset in comparison to the basic unigram distribution.
+See figure \ref{fig:frequency_ex} for an example.
+\begin{figure}[ht]
+\centering
+\includegraphics[scale=0.45]{images/frequency_ex}
+\caption{Probabililty of a word, of the text8 dataset (sampled), to be chosen at random according to its frequency and the power to which the unigram distribution is raised}
+\label{fig:frequency_ex}
+\end{figure}
+It's quite easily observable that this approach will outperform the classical softmax in computation time, as it only compute the sum over $K$ output nodes.
+Now the question arises if the accuracy is good enough but according to Mikolov et al.
+\cite{mikolov2} the negative sampling method "is an extremely simple training method that learns accurate representations".
+As a matter of fact, Mikolov et al.
+\cite{mikolov2} reported a 6\% accuracy improvement in comparison to a Hierarchical Softmax model.
+Therefore, it is a good solution to the problem raised by the classical softmax.
+We now have enough background knowledge about the SGM to look at how it can be optimized. The next subsection introduces earlier approaches to optimize the SGM.

### 2019-05-01 icdm/submission/includes/conclusion.tex
Splitted papaer into multiple files

--- 2019-04-30 icdm/submission/paper.tex
@@
-\documentclass[conference]{IEEEtran}
-\IEEEoverridecommandlockouts
-% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
-\usepackage{cite}
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{algorithmic}
-\usepackage{graphicx}
-\usepackage{textcomp}
-\usepackage{xcolor}
-\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
-    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
-\begin{document}
+\section{Conclusion}\label{chap:conclusion}
 
-\title{Paper Title*\\
-{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
-should not be used}
-\thanks{Identify applicable funding agency here. If none, delete this.}
-}
-
-\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-\and
-\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-\and
-\IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-\and
-\IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-\and
-\IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-\and
-\IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-}
-
-\maketitle
-
-\begin{abstract}
-This document is a model and instructions for \LaTeX.
-This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
-or Math in Paper Title or Abstract.
-\end{abstract}
-
-\begin{IEEEkeywords}
-component, formatting, style, styling, insert
-\end{IEEEkeywords}
-
-\section{Introduction}
-This document is a model and instructions for \LaTeX.
-Please observe the conference page limits.
-
-\section{Ease of Use}
-
-\subsection{Maintaining the Integrity of the Specifications}
-
-The IEEEtran class file is used to format your paper and style the text. All margins, 
-column widths, line spaces, and text fonts are prescribed; please do not 
-alter them. You may note peculiarities. For example, the head margin
-measures proportionately more than is customary. This measurement 
-and others are deliberate, using specifications that anticipate your paper 
-as one part of the entire proceedings, and not as an independent document. 
-Please do not revise any of the current designations.
-
-\section{Prepare Your Paper Before Styling}
-Before you begin to format your paper, first write and save the content as a 
-separate text file. Complete all content and organizational editing before 
-formatting. Please note sections \ref{AA}--\ref{SCM} below for more information on 
-proofreading, spelling and grammar.
-
-Keep your text and graphic files separate until after the text has been 
-formatted and styled. Do not number text heads---{\LaTeX} will do that 
-for you.
-
-\subsection{Abbreviations and Acronyms}\label{AA}
-Define abbreviations and acronyms the first time they are used in the text, 
-even after they have been defined in the abstract. Abbreviations such as 
-IEEE, SI, MKS, CGS, ac, dc, and rms do not have to be defined. Do not use 
-abbreviations in the title or heads unless they are unavoidable.
-
-\subsection{Units}
-\begin{itemize}
-\item Use either SI (MKS) or CGS as primary units.
-(SI units are encouraged.) English units may be used as secondary units (in parentheses).
-An exception would be the use of English units as identifiers in trade, such as ``3.5-inch disk drive''.
-\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds.
-This often leads to confusion because equations do not balance dimensionally.
-If you must use mixed units, clearly state the units for each quantity that you use in an equation.
-\item Do not mix complete spellings and abbreviations of units: ``Wb/m\textsuperscript{2}'' or ``webers per square meter'', not ``webers/m\textsuperscript{2}''.
-Spell out units when they appear in text: ``.
-.
-.
-a few henries'', not ``.
-.
-.
-a few H''.
-\item Use a zero before decimal points: ``0.25'', not ``.25''. Use ``cm\textsuperscript{3}'', not ``cc''.)
-\end{itemize}
-
-\subsection{Equations}
-Number equations consecutively. To make your 
-equations more compact, you may use the solidus (~/~), the exp function, or 
-appropriate exponents. Italicize Roman symbols for quantities and variables, 
-but not Greek symbols. Use a long dash rather than a hyphen for a minus 
-sign. Punctuate equations with commas or periods when they are part of a 
-sentence, as in:
-\begin{equation}
-a+b=\gamma\label{eq}
-\end{equation}
-
-Be sure that the 
-symbols in your equation have been defined before or immediately following 
-the equation. Use ``\eqref{eq}'', not ``Eq.~\eqref{eq}'' or ``equation \eqref{eq}'', except at 
-the beginning of a sentence: ``Equation \eqref{eq} is . . .''
-
-\subsection{\LaTeX-Specific Advice}
-
-Please use ``soft'' (e.g., \verb|\eqref{Eq}|) cross references instead
-of ``hard'' references (e.g., \verb|(1)|). That will make it possible
-to combine sections, add equations, or change the order of figures or
-citations without having to go through the file line by line.
-
-Please don't use the \verb|{eqnarray}| equation environment. Use
-\verb|{align}| or \verb|{IEEEeqnarray}| instead. The \verb|{eqnarray}|
-environment leaves unsightly spaces around relation symbols.
-
-Please note that the \verb|{subequations}| environment in {\LaTeX}
-will increment the main equation counter even when there are no
-equation numbers displayed. If you forget that, you might write an
-article in which the equation numbers skip from (17) to (20), causing
-the copy editors to wonder if you've discovered a new method of
-counting.
-
-{\BibTeX} does not work by magic. It doesn't get the bibliographic
-data from thin air but from .bib files. If you use {\BibTeX} to produce a
-bibliography you must send the .bib files.
-
-{\LaTeX} can't read your mind. If you assign the same label to a
-subsubsection and a table, you might find that Table I has been cross
-referenced as Table IV-B3.
-
-{\LaTeX} does not have precognitive abilities. If you put a
-\verb|\label| command before the command that updates the counter it's
-supposed to be using, the label will pick up the last counter to be
-cross referenced instead. In particular, a \verb|\label| command
-should not go before the caption of a figure or a table.
-
-Do not use \verb|\nonumber| inside the \verb|{array}| environment. It
-will not stop equation numbers inside \verb|{array}| (there won't be
-any anyway) and it might stop a wanted equation number in the
-surrounding equation.
-
-\subsection{Some Common Mistakes}\label{SCM}
-\begin{itemize}
-\item The word ``data'' is plural, not singular.
-\item The subscript for the permeability of vacuum $\mu_{0}$, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ``o''.
-\item In American English, commas, semicolons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation.
-When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks.
-A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this).
-(A parenthetical sentence is punctuated within the parentheses.)
-\item A graph within a graph is an ``inset'', not an ``insert''. The word alternatively is preferred to the word ``alternately'' (unless you really mean something that alternates).
-\item Do not use the word ``essentially'' to mean ``approximately'' or ``effectively''.
-\item In your paper title, if the words ``that uses'' can accurately replace the word ``using'', capitalize the ``u''; if not, keep using lower-cased.
-\item Be aware of the different meanings of the homophones ``affect'' and ``effect'', ``complement'' and ``compliment'', ``discreet'' and ``discrete'', ``principal'' and ``principle''.
-\item Do not confuse ``imply'' and ``infer''.
-\item The prefix ``non'' is not a word; it should be joined to the word it modifies, usually without a hyphen.
-\item There is no period after the ``et'' in the Latin abbreviation ``et al.''.
-\item The abbreviation ``i.e.'' means ``that is'', and the abbreviation ``e.g.'' means ``for example''.
-\end{itemize}
-An excellent style manual for science writers is \cite{b7}.
-
-\subsection{Authors and Affiliations}
-\textbf{The class file is designed for, but not limited to, six authors.} A 
-minimum of one author is required for all conference articles. Author names 
-should be listed starting from left to right and then moving down to the 
-next line. This is the author sequence that will be used in future citations 
-and by indexing services. Names should not be listed in columns nor group by 
-affiliation. Please keep your affiliations as succinct as possible (for 
-example, do not differentiate among departments of the same organization).
-
-\subsection{Identify the Headings}
-Headings, or heads, are organizational devices that guide the reader through 
-your paper. There are two types: component heads and text heads.
-
-Component heads identify the different components of your paper and are not 
-topically subordinate to each other. Examples include Acknowledgments and 
-References and, for these, the correct style to use is ``Heading 5''. Use 
-``figure caption'' for your Figure captions, and ``table head'' for your 
-table title. Run-in heads, such as ``Abstract'', will require you to apply a 
-style (in this case, italic) in addition to the style provided by the drop 
-down menu to differentiate the head from the text.
-
-Text heads organize the topics on a relational, hierarchical basis. For 
-example, the paper title is the primary text head because all subsequent 
-material relates and elaborates on this one topic. If there are two or more 
-sub-topics, the next level head (uppercase Roman numerals) should be used 
-and, conversely, if there are not at least two sub-topics, then no subheads 
-should be introduced.
-
-\subsection{Figures and Tables}
-\paragraph{Positioning Figures and Tables} Place figures and tables at the top and 
-bottom of columns. Avoid placing them in the middle of columns. Large 
-figures and tables may span across both columns. Figure captions should be 
-below the figures; table heads should appear above the tables. Insert 
-figures and tables after they are cited in the text. Use the abbreviation 
-``Fig.~\ref{fig}'', even at the beginning of a sentence.
-
-\begin{table}[htbp]
-\caption{Table Type Styles}
-\begin{center}
-\begin{tabular}{|c|c|c|c|}
-\hline
-\textbf{Table}&\multicolumn{3}{|c|}{\textbf{Table Column Head}} \\
-\cline{2-4} 
-\textbf{Head} & \textbf{\textit{Table column subhead}}& \textbf{\textit{Subhead}}& \textbf{\textit{Subhead}} \\
-\hline
-copy& More table copy$^{\mathrm{a}}$& &  \\
-\hline
-\multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
-\end{tabular}
-\label{tab1}
-\end{center}
-\end{table}
-
-Figure Labels: Use 8 point Times New Roman for Figure labels. Use words 
-rather than symbols or abbreviations when writing Figure axis labels to 
-avoid confusing the reader. As an example, write the quantity 
-``Magnetization'', or ``Magnetization, M'', not just ``M''. If including 
-units in the label, present them within parentheses. Do not label axes only 
-with units. In the example, write ``Magnetization (A/m)'' or ``Magnetization 
-\{A[m(1)]\}'', not just ``A/m''. Do not label axes with a ratio of 
-quantities and units. For example, write ``Temperature (K)'', not 
-``Temperature/K''.
-
-\section*{Acknowledgment}
-
-The preferred spelling of the word ``acknowledgment'' in America is without 
-an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B. 
-G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor 
-acknowledgments in the unnumbered footnote on the first page.
-
-\section*{References}
-
-Please number citations consecutively within brackets \cite{b1}. The 
-sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference 
-number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at 
-the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''
-
-Number footnotes separately in superscripts. Place the actual footnote at 
-the bottom of the column in which it was cited. Do not put footnotes in the 
-abstract or reference list. Use letters for table footnotes.
-
-Unless there are six authors or more give all authors' names; do not use 
-``et al.''. Papers that have not been published, even if they have been 
-submitted for publication, should be cited as ``unpublished'' \cite{b4}. Papers 
-that have been accepted for publication should be cited as ``in press'' \cite{b5}. 
-Capitalize only the first word in a paper title, except for proper nouns and 
-element symbols.
-
-For papers published in translation journals, please give the English 
-citation first, followed by the original foreign-language citation \cite{b6}.
-
-\bibliographystyle{IEEEtran}
-\bibliography{refs}
-
-% use bibfile instead (2 lines above)
-% \begin{thebibliography}{00}
-% \bibitem{b1} G.
-Eason, B.
-Noble, and I.
-N.
-Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil.
-Trans.
-Roy.
-Soc.
-London, vol.
-A247, pp.
-529--551, April 1955.
-% \bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
-% \bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
-% \bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
-% \bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
-% \bibitem{b6} Y.
-Yorozu, M.
-Hirano, K.
-Oka, and Y.
-Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl.
-J.
-Magn.
-Japan, vol.
-2, pp.
-740--741, August 1987 [Digests 9th Annual Conf.
-Magnetics Japan, p.
-301, 1982].
-% \bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
-% \end{thebibliography}
-\vspace{12pt}
-\color{red}
-IEEE conference templates contain guidance text for composing and formatting conference papers.
-Please ensure that all template text is removed from your conference paper prior to submission to the conference.
-Failure to remove the template text from your paper may result in your paper not being published.
-
-\end{document}
+This work provides an overview of the Skip Gram Model with negative Sampling (SGNS) and the numerous successful attempts of optimizing the throughput of the model.
+As this is the case, no effort went into optimizng the convergence time of the SGNS, therefore this work focused on this point.
+We decided to use advanced optimizers and input shuffling as optimizing techniques.
+After giving a short overview over Gradient Descent algorithms this work proposes a slighlty altered version of the SGNS, where the idea is to compute the loss over the sum of a high number of training samples, i.e 2000,  instead of computing it for each individually.
+We did this as it allowed us to compute more models and analyze the convergence time faster.
+We used the text8 dataset and used  word similarity as a quality measure for the word embeddings (WE).
+We used the State of the art implementation Gensim to compare our self.
+We did achieve a better convergence time than gensim with Adam as an optimizer and the use of input shuffling.
+Gensim convereged in 4 epochs to a word similarity of 0.66 and our model only took 2 epochs to achieve the same quality.
+Those results still need to be confirmed with more datasets.
+Finally, if this work would be combined with an optimized throughput it  could improve the state of the art run time of the SGNS.

### 2019-05-01 icdm/submission/includes/contribution.tex
Splitted papaer into multiple files

\section{Batched SkipGramModel} \label{ssec:b_SGM}
This section will give an overview of our implementation. First, it will give a broad overview of the implementation idea and process. Then it will go into detail, explaining the forward process of our model and the construction of our dataset, by the use of the \texttt{Dataset} class. The proposed implementation is a slighlty altered version of the original SGNS. The idea behind this altering, is to compute the loss for multiple words and context pairs at the same time. The exact process will be described in the following paragraph.
\subsection{Forwarding}
As this represents the challenging part of our Implementation the forwarding method is explained step by step. Each time step is illustrated to make the explanation clearer.\\
Let $X = {(v_1,c_1),(v_2,c_2),(v_3,c_3)}$, be the training batch, where $(v_i,c_i)$ is a training sample constituted by a word and on of its context words. \\
\textit{Input:}\\
The forwarding method will accept two vectors $v$ and $c$, and a matrix $A$ as an input. The first vector represents all the center words in a batch, the second one the context words. The Matrix represents the negative samples. The two vectors are of the same length, defined as $n$. The matrix must be of dimension $n \times k$, with $k$ being the number of negative samples per pair. This means the $i^{th}$ row will store the negative samples for the $i^{th}$ word context pair.\\
The input can be illustrated as follows: \\
$v = \begin{bmatrix}
v_1 & v_2 & v_3
\end{bmatrix}, c = \begin{bmatrix}
c1\\
c2\\
c3\end{bmatrix}$ and $A =
\begin{bmatrix}
k_{1,1} & k_{2,1} & k_{3,1}\\
k_{1,2} & k_{2,2} & k_{3,2}\\
k_{1,3} & k_{2,3} & k_{3,3}\\
\end{bmatrix}$\\

\textit{Concatenation of samples:}\\
The concatenation of the vector $c$ and the Matrix $A$ will result in a Matrix $\tilde{A}$, with:\\
$\tilde{A} = \begin{bmatrix}
c_1 & k_{1,1} & k_{2,1} & k_{3,1}\\
c_2 & k_{1,2} & k_{2,2} & k_{3,2}\\
c_3 & k_{1,3} & k_{2,3}& k_{3,3}\\
\end{bmatrix}$\\

\textit{Embeddings:}\\
Now it's necessary to access the word embeddings. For this purpose a Matrix $E_v$ of dimensionality $n \times d$ where $d$ is the dimension of our word embedding is created. $E_v$ stores the word embedding of the $i^{th}$ word from our input vector $v$ in it's $i^{th}$ row. The same is done for the other Matrix $\tilde{A}$. This will result in a $n \times k+1 \times d$ Array $E_c$. \\
$E_v = \begin{bmatrix}
\tilde{v_1}_1 & \ldots & \tilde{v_1}_d\\
\tilde{v_2}_1 & \ldots & \tilde{v_2}_d\\
\tilde{v_3}_1 & \ldots & \tilde{v_3}_d\\
\end{bmatrix}
$, where $\tilde{v_i} = \begin{bmatrix}
\tilde{v_i}_1 & \ldots & \tilde{v_i}_d \end{bmatrix}$ is the embedding of $v_i$. \\

$E_c = \begin{bmatrix}
\tilde{c_1 }& \tilde{k_{1,1}} & \tilde{k_{2,1}} \\
\tilde{c_2 }& \tilde{k_{1,2}}& \tilde{k_{2,2}} \\
\tilde{c_3 }&\tilde{ k_{1,3} }& \tilde{k_{2,3}}\\
\end{bmatrix}$,
where each entry of the matrix is a vector of dimension $d$\\

\textit{Batch multiplication and negation of samples:}\\
Now we need to compute the dot product of each word vector with its pair and the negative samples, exactly as done as in the original loss function of Mikolov et al. shown in Equation \ref{eq:obj_neg_samples}. For this task some definitions are necessary: let $A_j$ be the $j^th$ row of the matrix A, let $E_c(i,j)$ be the $d$ dimensional embedding of the word stored in $\tilde{A}(i,j)$. The dot product is computed with a so-called batch multiplication\footnote{Documentation: https://pytorch.org/docs/stable/torch.html\#torch.bmm} which will result in a matrix $S$ where $S(i,j) = E_c(i,j) \cdot A_j$. This will result in a $n\times k+1$ Matrix $S$. Now we only have to negate the last $k$ rows with minus one. The sum of each row represents the loss computed in Equation \ref{eq:obj_neg_samples}, for each word context pair.
Since computation time is too long with this approach the loss function is altered\\
$S = \begin{bmatrix}
\tilde{v_1} \cdot \tilde{c_1} & -\tilde{v_1} \cdot \tilde{k_{1,1}} & -\tilde{v_1} \cdot \tilde{k_{2,1}}& -\tilde{v_1} \cdot \tilde{k_{3,1}}\\
\tilde{v_2} \cdot \tilde{c_2} & -\tilde{v_2} \cdot \tilde{k_{1,2}} & -\tilde{v_2} \cdot \tilde{k_{2,2}} & -\tilde{v_2} \cdot \tilde{k_{3,2}}\\
\tilde{v_3} \cdot \tilde{c_3} &-\tilde{v_3} \cdot c_3 \tilde{k_{1,3}} & -\tilde{v_3} \cdot c_3 \tilde{k_{2,3}}&-\tilde{v_3} \cdot \tilde{k_{3,3}}\\
\end{bmatrix}$\\

\textit{Loss function:}\\
Summing the matrix $S$ and multiplying it with $-1$ (to make the problem a minimizing problem) results in the loss for our entire batch. As some words may appear more than once in the batch this will more be an average of it than as with Mikolov et al. \cite{mikolov2} the exact update per pair. \\
L = $- \sum_{(i,j) \in k \times n} S(i,j) $\\

This algorithm provides a way to compute the loss, to make the implementation successful, one still needs to access the context-pairs. The process of doing so is described in the next paragraph.

### 2019-05-01 icdm/submission/includes/discussion.tex
Splitted papaer into multiple files

--- 2019-04-30 icdm/submission/paper.tex
@@
-\documentclass[conference]{IEEEtran}
-\IEEEoverridecommandlockouts
-% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
-\usepackage{cite}
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{algorithmic}
-\usepackage{graphicx}
-\usepackage{textcomp}
-\usepackage{xcolor}
-\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
-    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
-\begin{document}
+\section{Discussion}\label{chap:discussion}
 
-\title{Paper Title*\\
-{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
-should not be used}
-\thanks{Identify applicable funding agency here. If none, delete this.}
-}
+%Discuss the results. What is the outcome of your experimetns?
+This subsection shortly discusses the empirical results and then extensively compares the findings of this work to the existing literature while trying to explain some differences.
+It is followed by a subsection describing the limitations and possible extensions of this work.
+This subsection is concluded by the problems we encountered during the implementation so others can avoid them in the future.
 
-\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-\and
-\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-\and
-\IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-\and
-\IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-\and
-\IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-\and
-\IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-}
+\subsection{Our work}
+This subsection discusses our findings, first we will analyze how input shuffling could influence the learning rate and conclude with the discussion of unexpected results.
 
-\maketitle
+\subsubsection{Shuffling and learning rate with SGD}
+When the model used SGD as an optimizer it was able to use higher learning rates when input shuffling was used, as shown in figures \ref{fig:results_sgd}, \ref{fig:results_mom} and \ref{fig:results_nag} Therefore arises the questions why these phenomena happen, especially as it did not happen with advanced optimizers.
+One possibility is the use of the batched version of the SGNS.
+In consequence, when the input is not shuffled the same word will often appear in one batch, hence the optimizer is presented with an average value of the gradient, which can be imprecise.
+This is counter-attacked by advanced optimizers as they have adaptive learning rates, i.e a frequent appearing feature will have a low learning rate.
 
-\begin{abstract}
-This document is a model and instructions for \LaTeX.
-This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
-or Math in Paper Title or Abstract.
-\end{abstract}
+\subsubsection{Large differences with NAG and SGD when using shuffling}
+SGD and NAG both have very different values when using shuffling in comparison to unshuffled input, as shown in figures \ref{fig:results_sgd} and \ref{fig:results_nag}, plain SGD.
+We do not only attribute those results to input shuffling but partially also to a good random initialization guess.
+Due to a lack of time these results were not replicate more than once.
 
-\begin{IEEEkeywords}
-component, formatting, style, styling, insert
-\end{IEEEkeywords}
+\subsection{Comparison to Gensim}
+This subsection will compare our finding extensively to Gensim \cite{gensim}.
+As explained in Section \ref{ssec:gensim}, Gensim is optimized to have a very high throughput, this made it possible to achieve a lot of computations.
+Furthermore, Gensim provides access to the loss and the resulting word embeddings, which facilitated the comparison process.
 
-\section{Introduction}
-This document is a model and instructions for \LaTeX.
-Please observe the conference page limits.
+\iffalse
+As Mikolov et al.
+published the original paper which introduced the SGNS, it is, of course, relevant to compare ourselves to their work.
+The first important thing to take under account is that Mikolov et al.
+only trained their model on a very large google news dataset incorporating more than 3 billion words.
+This makes the comparison of our work more difficult.
+But we will make some assumptions, as they could be of value if true.
+In their original paper, Mikolov et al.
+reported results from computations that took 1 and 3 epochs.
+We accord these good results, which clearly outperformed our SGD model and Gensim, to the very large dataset and furthermore, as a matter of fact, their results are better with 3 than with 1 epoch.
+We do not have any information about the convergence time or criterion.
+Hence it would be interesting to use their dataset for comparison.
+\\
+One thing we can compare is the quality of our word embeddings.
+Mikolov et al.
+did not report any results on their model with the text8 dataset, but they, therefore, published their code.
+Which was then used by Ji et al \cite{intel} and they tested the model on the text8 dataset.
+They reported a word similarity of 0.63 on the word similarity task.
+This is obviously outperformed by all our models.
+We did not find any explanation on why those results differ as much.
+\\
+The final assumption is that an advanced optimizer could maybe outperform SGD in terms of quality on a large data set. This will be discussed in further work.
 
-\section{Ease of Use}
+\fi
 
-\subsection{Maintaining the Integrity of the Specifications}
-
-The IEEEtran class file is used to format your paper and style the text. All margins, 
-column widths, line spaces, and text fonts are prescribed; please do not 
-alter them. You may note peculiarities. For example, the head margin
-measures proportionately more than is customary. This measurement 
-and others are deliberate, using specifications that anticipate your paper 
-as one part of the entire proceedings, and not as an independent document. 
-Please do not revise any of the current designations.
-
-\section{Prepare Your Paper Before Styling}
-Before you begin to format your paper, first write and save the content as a 
-separate text file. Complete all content and organizational editing before 
-formatting. Please note sections \ref{AA}--\ref{SCM} below for more information on 
-proofreading, spelling and grammar.
-
-Keep your text and graphic files separate until after the text has been 
-formatted and styled. Do not number text heads---{\LaTeX} will do that 
-for you.
-
-\subsection{Abbreviations and Acronyms}\label{AA}
-Define abbreviations and acronyms the first time they are used in the text, 
-even after they have been defined in the abstract. Abbreviations such as 
-IEEE, SI, MKS, CGS, ac, dc, and rms do not have to be defined. Do not use 
-abbreviations in the title or heads unless they are unavoidable.
-
-\subsection{Units}
+\subsubsection{Configuration of Gensim}
+The training with Gensim has a lot of possible parameters an extended list can be found in the appendix.
+This subsection will only describe the parameters we changed from the default setting.
+The description of each parameter has the following form: \\
+\texttt{name} (type) -- \textit{Description} -- Value\\
+Parameters:
 \begin{itemize}
-\item Use either SI (MKS) or CGS as primary units.
-(SI units are encouraged.) English units may be used as secondary units (in parentheses).
-An exception would be the use of English units as identifiers in trade, such as ``3.5-inch disk drive''.
-\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds.
-This often leads to confusion because equations do not balance dimensionally.
-If you must use mixed units, clearly state the units for each quantity that you use in an equation.
-\item Do not mix complete spellings and abbreviations of units: ``Wb/m\textsuperscript{2}'' or ``webers per square meter'', not ``webers/m\textsuperscript{2}''.
-Spell out units when they appear in text: ``.
-.
-.
-a few henries'', not ``.
-.
-.
-a few H''.
-\item Use a zero before decimal points: ``0.25'', not ``.25''. Use ``cm\textsuperscript{3}'', not ``cc''.)
+\item \texttt{sentences} (iterable of iterables) -- \textit{Dataset} -- text8 document splitted into sentences of 20 words
+\item \texttt{size}(int) – \textit{Dimensionality of the word vectors } -- 100
+\item \texttt{window} (int) --\textit {Maximum distance between the current and predicted word within a sentence }-- 5
+\item \texttt{min\_count }(int) --\textit{ Ignores all words with total frequency lower than this }-- 5
+\item \texttt{workers} (int) -- \textit{ Number of threads used to train the model} -- 4
+\item \texttt{sg} ({0, 1}) --\textit{ Training algorithm: 1 for skip--gram}--1
+\item \texttt{negative} (int) --\textit{Number of negative samples}-- 10
+\item \texttt{ ns\_exponent} (float) --\textit{ Exponent in the unigram distribution, when choosing random samples, as shown in Equation \ref{eq:unigram} }-- 0.75
+\item \texttt{alpha} (float) --\textit{ The initial learning rate. }-- 0.025
+\item \texttt{min\_alpha} (float) --\textit{ Learning rate will linearly drop to min\_alpha as training progresses }--0.0001
+\item \texttt{sample} (float) --\textit{ Treshold for subsampling as described in Equation \ref{eq:sampling}} -- 1e--4
+\item \texttt{iter} (int) --\textit{ Number of iterations (epochs) over the corpus. }-- 10
+\
+\item \texttt{compute\_loss} (bool) --\textit{If True, loss is stored at the end of each batch}-- True
+\item \texttt{callbacks} (iterable of CallbackAny2Vec) --\textit{ Set of functions executed in order to follow the loss and the progress of the model in word similarity }-- see Appendix
 \end{itemize}
 
-\subsection{Equations}
-Number equations consecutively. To make your 
-equations more compact, you may use the solidus (~/~), the exp function, or 
-appropriate exponents. Italicize Roman symbols for quantities and variables, 
-but not Greek symbols. Use a long dash rather than a hyphen for a minus 
-sign. Punctuate equations with commas or periods when they are part of a 
-sentence, as in:
+\subsubsection{Gensim vs. SGD}
+As stated earlier, we are not going to compare this work to Gensim in run time.
+Gensim is heavily optimized and written in cython\footnote{https://rare-technologies.com/word2vec-in-python-part-two-optimizing/}, which is 23x faster than plain Numpy.
+Since we used PyTorch the difference is not that big, but still remains.
+As shown in Figure \ref{fig:gensim_vs_adam} the convergence time was not the same between our implementation and Gensim.
+There are different possible reasons why this could be the case:\\ First, our batched approach could hinder performance in terms of convergence since our loss function is not exactly the same.
+When a word appears more than once in our batch, the gradient will be an average over the gradients of each pair alone, as it is done by Gensim.\\ Secondly, a difference to our implementation is the fact that Gensim checks whether negative samples are not equal to the context word.
+If that is the case Gensim selects a new random sample.
+Therefore, the learning of the input and output context is optimized.
+\\Finally, another possibility is the decay of the learning rate used by Gensim.
+In fact, decaying the learning rate has been proven in a lot of work to decrease the convergence time.
+Gensim linearly decreases the learning rate, as we did not use this technique, the decay of the learning rate could help explain the noted differences.
+\\ The first hypothesis, the fact that we used a batched approach, may be confirmed by the fact that when combined with input shuffling SGD does perform closer to Gensim, going from 11 to 7 epochs to converge, as input shuffling reduces the number of co-occurrence of the same word in a batch.
+Now the question arises if the 3 epochs, that Gensim is better, can be explained by the selection of better negative samples and the learning rate decay?
+
+\subsubsection{Gensim vs. Adam}
+The Adam optimizer did outperform the Gensim application in quality of word embeddings (only slightly: 0.01 correlation coefficient better) and convergence time.
+Adam converged in 2 epochs while Gensim in 4.
+To confirm the achieved results we ran each computation 40 times.
+The results can be seen in Figure \ref{fig:gensim_vs_adam}.
+
+\begin{figure}[h]
+\centering
+\includegraphics[scale=0.45]{images/gensim_vs_adam}
+\caption{Convergence time of SGD and Adam compared to Gensim}
+\label{fig:gensim_vs_adam}
+\end{figure}\subsection{Challenges faced}
+During our implementation we did encounter some problems, this subsection has the purpose of preventing the scientific community from making the same mistakes.
+\subsubsection{Using the wrong embeddings}
+To start with, we did not choose the same initialization value as Gensim in our word embeddings.
+We initialized the projection layer with a normal distribution between (-1,1), as opposed to Gensim which initializes all the weights to 0.
+The reason for this, is that at the beginning of our experiments we did not achieve good results with the same initialization as Gensim.
+Retrospectively we accord this to a learning rate that was too low.
+After a few simulations, we saw that we did not perform as good as Gensim.
+And as we changed the initialization of the projection layer back to 0, and already had adjusted the learning rate, we achieved good results.
+Therefore, this is a recommendation to future work to not set the initialization to (-1,1).
+
+\subsubsection{Batch size and loss function adjustements}\label{ssec:bs_lf}
+During our experiments, we faced a moment where we needed to us a very high learning rate (50) to achieve good results.
+As this is the complete opposite of what is standard we made a few changes to remedy the problem.\\
+First, we needed to find a good batch size. During the above-explained raise of the learning rate, we used a batch size of 5000. We then decided to take a batch size of 2000.\\
+Secondly, at the same time, we did experiment with different loss functions.
+As our batched model does not have exactly the same loss function as introduced by Mikolov et al.
+\cite{mikolov}, we needed to find one that suited our goals the best.
+At the beginning of our experiment, we took the average of all the scores stored in our final matrix, as explained in \ref{ssec:b_SGM}.
+But then chose to take the sum as the training did not seem optimal.\\
+With these two changes, we did increase convergence time and word similarity, while at the same time having a usual learning rate. \\
+Know the question arises why this happened?
+Is this because of the shape of the loss function better suits our advanced optimizers?
+Or because the loss function is closer to its original form?
+Our hypothesis looks the following way:
+When taking the average, our loss function is represented as follows:
 \begin{equation}
-a+b=\gamma\label{eq}
+\frac{-1}{b }* \sum_{x\in X} loss(x)
 \end{equation}
+where $X$ is our batch and $b = |X|$.
+When taking the partial derivative of each parameter, the constant $ \frac{-1}{b }$ will remain the same.
+Therefore, it's the same as taking the sum as the loss function and multiplying each gradient with $\frac{-1}{b }$.
+As our batch size was really high, i.e 2000 and 5000, this could highly influence the learning rate.
+This is also the conclusion of Goyal et al.
+\cite{fb}, that worked with very high batch sizes, i.e 8000: "When the minibatch size is multiplied by k, multiply the learning rate by k".
 
-Be sure that the 
-symbols in your equation have been defined before or immediately following 
-the equation. Use ``\eqref{eq}'', not ``Eq.~\eqref{eq}'' or ``equation \eqref{eq}'', except at 
-the beginning of a sentence: ``Equation \eqref{eq} is . . .''
-
-\subsection{\LaTeX-Specific Advice}
-
-Please use ``soft'' (e.g., \verb|\eqref{Eq}|) cross references instead
-of ``hard'' references (e.g., \verb|(1)|). That will make it possible
-to combine sections, add equations, or change the order of figures or
-citations without having to go through the file line by line.
-
-Please don't use the \verb|{eqnarray}| equation environment. Use
-\verb|{align}| or \verb|{IEEEeqnarray}| instead. The \verb|{eqnarray}|
-environment leaves unsightly spaces around relation symbols.
-
-Please note that the \verb|{subequations}| environment in {\LaTeX}
-will increment the main equation counter even when there are no
-equation numbers displayed. If you forget that, you might write an
-article in which the equation numbers skip from (17) to (20), causing
-the copy editors to wonder if you've discovered a new method of
-counting.
-
-{\BibTeX} does not work by magic. It doesn't get the bibliographic
-data from thin air but from .bib files. If you use {\BibTeX} to produce a
-bibliography you must send the .bib files.
-
-{\LaTeX} can't read your mind. If you assign the same label to a
-subsubsection and a table, you might find that Table I has been cross
-referenced as Table IV-B3.
-
-{\LaTeX} does not have precognitive abilities. If you put a
-\verb|\label| command before the command that updates the counter it's
-supposed to be using, the label will pick up the last counter to be
-cross referenced instead. In particular, a \verb|\label| command
-should not go before the caption of a figure or a table.
-
-Do not use \verb|\nonumber| inside the \verb|{array}| environment. It
-will not stop equation numbers inside \verb|{array}| (there won't be
-any anyway) and it might stop a wanted equation number in the
-surrounding equation.
-
-\subsection{Some Common Mistakes}\label{SCM}
-\begin{itemize}
-\item The word ``data'' is plural, not singular.
-\item The subscript for the permeability of vacuum $\mu_{0}$, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ``o''.
-\item In American English, commas, semicolons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation.
-When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks.
-A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this).
-(A parenthetical sentence is punctuated within the parentheses.)
-\item A graph within a graph is an ``inset'', not an ``insert''. The word alternatively is preferred to the word ``alternately'' (unless you really mean something that alternates).
-\item Do not use the word ``essentially'' to mean ``approximately'' or ``effectively''.
-\item In your paper title, if the words ``that uses'' can accurately replace the word ``using'', capitalize the ``u''; if not, keep using lower-cased.
-\item Be aware of the different meanings of the homophones ``affect'' and ``effect'', ``complement'' and ``compliment'', ``discreet'' and ``discrete'', ``principal'' and ``principle''.
-\item Do not confuse ``imply'' and ``infer''.
-\item The prefix ``non'' is not a word; it should be joined to the word it modifies, usually without a hyphen.
-\item There is no period after the ``et'' in the Latin abbreviation ``et al.''.
-\item The abbreviation ``i.e.'' means ``that is'', and the abbreviation ``e.g.'' means ``for example''.
-\end{itemize}
-An excellent style manual for science writers is \cite{b7}.
-
-\subsection{Authors and Affiliations}
-\textbf{The class file is designed for, but not limited to, six authors.} A 
-minimum of one author is required for all conference articles. Author names 
-should be listed starting from left to right and then moving down to the 
-next line. This is the author sequence that will be used in future citations 
-and by indexing services. Names should not be listed in columns nor group by 
-affiliation. Please keep your affiliations as succinct as possible (for 
-example, do not differentiate among departments of the same organization).
-
-\subsection{Identify the Headings}
-Headings, or heads, are organizational devices that guide the reader through 
-your paper. There are two types: component heads and text heads.
-
-Component heads identify the different components of your paper and are not 
-topically subordinate to each other. Examples include Acknowledgments and 
-References and, for these, the correct style to use is ``Heading 5''. Use 
-``figure caption'' for your Figure captions, and ``table head'' for your 
-table title. Run-in heads, such as ``Abstract'', will require you to apply a 
-style (in this case, italic) in addition to the style provided by the drop 
-down menu to differentiate the head from the text.
-
-Text heads organize the topics on a relational, hierarchical basis. For 
-example, the paper title is the primary text head because all subsequent 
-material relates and elaborates on this one topic. If there are two or more 
-sub-topics, the next level head (uppercase Roman numerals) should be used 
-and, conversely, if there are not at least two sub-topics, then no subheads 
-should be introduced.
-
-\subsection{Figures and Tables}
-\paragraph{Positioning Figures and Tables} Place figures and tables at the top and 
-bottom of columns. Avoid placing them in the middle of columns. Large 
-figures and tables may span across both columns. Figure captions should be 
-below the figures; table heads should appear above the tables. Insert 
-figures and tables after they are cited in the text. Use the abbreviation 
-``Fig.~\ref{fig}'', even at the beginning of a sentence.
-
-\begin{table}[htbp]
-\caption{Table Type Styles}
-\begin{center}
-\begin{tabular}{|c|c|c|c|}
-\hline
-\textbf{Table}&\multicolumn{3}{|c|}{\textbf{Table Column Head}} \\
-\cline{2-4} 
-\textbf{Head} & \textbf{\textit{Table column subhead}}& \textbf{\textit{Subhead}}& \textbf{\textit{Subhead}} \\
-\hline
-copy& More table copy$^{\mathrm{a}}$& &  \\
-\hline
-\multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
-\end{tabular}
-\label{tab1}
-\end{center}
-\end{table}
-
-Figure Labels: Use 8 point Times New Roman for Figure labels. Use words 
-rather than symbols or abbreviations when writing Figure axis labels to 
-avoid confusing the reader. As an example, write the quantity 
-``Magnetization'', or ``Magnetization, M'', not just ``M''. If including 
-units in the label, present them within parentheses. Do not label axes only 
-with units. In the example, write ``Magnetization (A/m)'' or ``Magnetization 
-\{A[m(1)]\}'', not just ``A/m''. Do not label axes with a ratio of 
-quantities and units. For example, write ``Temperature (K)'', not 
-``Temperature/K''.
-
-\section*{Acknowledgment}
-
-The preferred spelling of the word ``acknowledgment'' in America is without 
-an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B. 
-G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor 
-acknowledgments in the unnumbered footnote on the first page.
-
-\section*{References}
-
-Please number citations consecutively within brackets \cite{b1}. The 
-sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference 
-number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at 
-the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''
-
-Number footnotes separately in superscripts. Place the actual footnote at 
-the bottom of the column in which it was cited. Do not put footnotes in the 
-abstract or reference list. Use letters for table footnotes.
-
-Unless there are six authors or more give all authors' names; do not use 
-``et al.''. Papers that have not been published, even if they have been 
-submitted for publication, should be cited as ``unpublished'' \cite{b4}. Papers 
-that have been accepted for publication should be cited as ``in press'' \cite{b5}. 
-Capitalize only the first word in a paper title, except for proper nouns and 
-element symbols.
-
-For papers published in translation journals, please give the English 
-citation first, followed by the original foreign-language citation \cite{b6}.
-
-\bibliographystyle{IEEEtran}
-\bibliography{refs}
-
-% use bibfile instead (2 lines above)
-% \begin{thebibliography}{00}
-% \bibitem{b1} G.
-Eason, B.
-Noble, and I.
-N.
-Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil.
-Trans.
-Roy.
-Soc.
-London, vol.
-A247, pp.
-529--551, April 1955.
-% \bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
-% \bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
-% \bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
-% \bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
-% \bibitem{b6} Y.
-Yorozu, M.
-Hirano, K.
-Oka, and Y.
-Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl.
-J.
-Magn.
-Japan, vol.
-2, pp.
-740--741, August 1987 [Digests 9th Annual Conf.
-Magnetics Japan, p.
-301, 1982].
-% \bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
-% \end{thebibliography}
-\vspace{12pt}
-\color{red}
-IEEE conference templates contain guidance text for composing and formatting conference papers.
-Please ensure that all template text is removed from your conference paper prior to submission to the conference.
-Failure to remove the template text from your paper may result in your paper not being published.
-
-\end{document}
+\subsection{Future Work}
+This work shows that the convergence time of the SGNS could be improved by using input shuffling and advanced optimizers.
+As with every work, there still exists possible extensions.
+First and foremost an aspect of our implementation that can be prejudicial is that we only extensively tested our model with one small dataset.
+The fact that we only used one dataset as well that it's a small one is problematic.
+By using a very small dataset we do not use the model in the condition it is most needed for, as the dataset used in practice usually consists of more than 1 billion words.
+There is a small argument that can be made for machine translation as the use of small parallel corpus is not unusual in this field.
+But the main issue with using only one data set it that it has been shown that some optimizers perform better with specific shapes of loss functions.
+To make a compelling argument it's, therefore, necessary to show that our model with the use of Input shuffling and Adam as its optimizer also outperform Gensim with other data sets, so that the claim can hold consistently, it needs to be confirmed with other datasets as well.
+Furthermore, our implementation did not outperform Gensim in run time, as this was not the goal of our work.
+Therefore, one could improve an already existing, optimized version, with input shuffling and advanced optimizers and should achieve a better run time than Gensim.

### 2019-05-01 icdm/submission/includes/introduction.tex
Splitted papaer into multiple files

--- 2019-04-30 icdm/submission/paper.tex
@@
-\documentclass[conference]{IEEEtran}
-\IEEEoverridecommandlockouts
-% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
-\usepackage{cite}
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{algorithmic}
-\usepackage{graphicx}
-\usepackage{textcomp}
-\usepackage{xcolor}
-\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
-    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
-\begin{document}
+\section{Introduction}\label{chap:introduction}
 
-\title{Paper Title*\\
-{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
-should not be used}
-\thanks{Identify applicable funding agency here. If none, delete this.}
-}
+Representing words as vectors, i.e word embeddings (WE) is a fundamental aspect of Natural Language Processing (NLP).
+There are two ways to create such WE, either arbitrarily or with the purpose of capturing the semantics of the words, i.e.
+vector representations of words that are syntactically or semantically similar will be near to each other.
+By capturing semantic or syntactic meaning WE have shown to facilitate a lot of subsequent NLP tasks, such as entity recognition, machine translation or sentence classification.
+The first attempt to create WE with neural networks was made by Bengio et al.
+\cite{bengio} but more recently Mikolov et al.
+\cite{mikolov} introduced a software package called w2vec that uses a simpler network and with this approach Mikolov et al.
+produced state of the art results.
+One of the proposed algorithms in this software package is the Skip-Gram Model (SGM).
+The SGM is an algorithm, that trains a neural network, on the task of predicting the neighboring words in a sentence.
+The weights of this network are then used as WE.
 
-\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-\and
-\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-\and
-\IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-\and
-\IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-\and
-\IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-\and
-\IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-}
+The SGM gained a lot of attention, as it achieved very good results for a very simplistic model.
+As a consequence, a lot of effort went into optimizing it.
+Most of this effort was trying to improve the throughput of the model, i.e the number of words that are processed per second.
+The SGM uses Stochastic Gradient Descent as an optimization algorithm and is therefore inherently sequential.
+To remedy this problem Mikolov et al.
+used Hogwild \cite{hogwild}, where different threads can access a shared model and update it.
+As this is not an optimal solution Yi et al.
+\cite{intel} tried to optimize it, by using a mini-batch like approach and converting vector to vector multiplications into matrix to matrix multiplications.
+This yielded two consequences: First the model is updated less frequently leading to less overwriting and offering the possibility to parallelize more.
+Secondly, it transformed level-1 BLAS operation into level-3 BLAS operations, and the algorithm could therefore effectively use computational resources.
+Another attempt at optimizing the throughput was made by Seulki and Youngmin \cite{gpu}.
+Their goal was to parallelize the algorithm on GPU's.
+They, therefore, chose to parallelize the update of the dimensions of each word representation.
+Both of these approaches and most of the literature are focused on improving the throughput of the model, but not the convergence time.
+Therefore one could ask if the convergence time of the SGM can be optimized while at the same time maintaining its accuracy.
+This work proposes an approach that uses advanced optimizers and input shuffling to optimize the convergence time of the SGM.
+In combination both of these techniques allowed us to decrease the convergence time of the SGM.
+If combined with an optimized throughput these result could lead to an overall decrease in runtime.
 
-\maketitle
-
-\begin{abstract}
-This document is a model and instructions for \LaTeX.
-This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
-or Math in Paper Title or Abstract.
-\end{abstract}
-
-\begin{IEEEkeywords}
-component, formatting, style, styling, insert
-\end{IEEEkeywords}
-
-\section{Introduction}
-This document is a model and instructions for \LaTeX.
-Please observe the conference page limits.
-
-\section{Ease of Use}
-
-\subsection{Maintaining the Integrity of the Specifications}
-
-The IEEEtran class file is used to format your paper and style the text. All margins, 
-column widths, line spaces, and text fonts are prescribed; please do not 
-alter them. You may note peculiarities. For example, the head margin
-measures proportionately more than is customary. This measurement 
-and others are deliberate, using specifications that anticipate your paper 
-as one part of the entire proceedings, and not as an independent document. 
-Please do not revise any of the current designations.
-
-\section{Prepare Your Paper Before Styling}
-Before you begin to format your paper, first write and save the content as a 
-separate text file. Complete all content and organizational editing before 
-formatting. Please note sections \ref{AA}--\ref{SCM} below for more information on 
-proofreading, spelling and grammar.
-
-Keep your text and graphic files separate until after the text has been 
-formatted and styled. Do not number text heads---{\LaTeX} will do that 
-for you.
-
-\subsection{Abbreviations and Acronyms}\label{AA}
-Define abbreviations and acronyms the first time they are used in the text, 
-even after they have been defined in the abstract. Abbreviations such as 
-IEEE, SI, MKS, CGS, ac, dc, and rms do not have to be defined. Do not use 
-abbreviations in the title or heads unless they are unavoidable.
-
-\subsection{Units}
-\begin{itemize}
-\item Use either SI (MKS) or CGS as primary units.
-(SI units are encouraged.) English units may be used as secondary units (in parentheses).
-An exception would be the use of English units as identifiers in trade, such as ``3.5-inch disk drive''.
-\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds.
-This often leads to confusion because equations do not balance dimensionally.
-If you must use mixed units, clearly state the units for each quantity that you use in an equation.
-\item Do not mix complete spellings and abbreviations of units: ``Wb/m\textsuperscript{2}'' or ``webers per square meter'', not ``webers/m\textsuperscript{2}''.
-Spell out units when they appear in text: ``.
-.
-.
-a few henries'', not ``.
-.
-.
-a few H''.
-\item Use a zero before decimal points: ``0.25'', not ``.25''. Use ``cm\textsuperscript{3}'', not ``cc''.)
-\end{itemize}
-
-\subsection{Equations}
-Number equations consecutively. To make your 
-equations more compact, you may use the solidus (~/~), the exp function, or 
-appropriate exponents. Italicize Roman symbols for quantities and variables, 
-but not Greek symbols. Use a long dash rather than a hyphen for a minus 
-sign. Punctuate equations with commas or periods when they are part of a 
-sentence, as in:
-\begin{equation}
-a+b=\gamma\label{eq}
-\end{equation}
-
-Be sure that the 
-symbols in your equation have been defined before or immediately following 
-the equation. Use ``\eqref{eq}'', not ``Eq.~\eqref{eq}'' or ``equation \eqref{eq}'', except at 
-the beginning of a sentence: ``Equation \eqref{eq} is . . .''
-
-\subsection{\LaTeX-Specific Advice}
-
-Please use ``soft'' (e.g., \verb|\eqref{Eq}|) cross references instead
-of ``hard'' references (e.g., \verb|(1)|). That will make it possible
-to combine sections, add equations, or change the order of figures or
-citations without having to go through the file line by line.
-
-Please don't use the \verb|{eqnarray}| equation environment. Use
-\verb|{align}| or \verb|{IEEEeqnarray}| instead. The \verb|{eqnarray}|
-environment leaves unsightly spaces around relation symbols.
-
-Please note that the \verb|{subequations}| environment in {\LaTeX}
-will increment the main equation counter even when there are no
-equation numbers displayed. If you forget that, you might write an
-article in which the equation numbers skip from (17) to (20), causing
-the copy editors to wonder if you've discovered a new method of
-counting.
-
-{\BibTeX} does not work by magic. It doesn't get the bibliographic
-data from thin air but from .bib files. If you use {\BibTeX} to produce a
-bibliography you must send the .bib files.
-
-{\LaTeX} can't read your mind. If you assign the same label to a
-subsubsection and a table, you might find that Table I has been cross
-referenced as Table IV-B3.
-
-{\LaTeX} does not have precognitive abilities. If you put a
-\verb|\label| command before the command that updates the counter it's
-supposed to be using, the label will pick up the last counter to be
-cross referenced instead. In particular, a \verb|\label| command
-should not go before the caption of a figure or a table.
-
-Do not use \verb|\nonumber| inside the \verb|{array}| environment. It
-will not stop equation numbers inside \verb|{array}| (there won't be
-any anyway) and it might stop a wanted equation number in the
-surrounding equation.
-
-\subsection{Some Common Mistakes}\label{SCM}
-\begin{itemize}
-\item The word ``data'' is plural, not singular.
-\item The subscript for the permeability of vacuum $\mu_{0}$, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ``o''.
-\item In American English, commas, semicolons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation.
-When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks.
-A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this).
-(A parenthetical sentence is punctuated within the parentheses.)
-\item A graph within a graph is an ``inset'', not an ``insert''. The word alternatively is preferred to the word ``alternately'' (unless you really mean something that alternates).
-\item Do not use the word ``essentially'' to mean ``approximately'' or ``effectively''.
-\item In your paper title, if the words ``that uses'' can accurately replace the word ``using'', capitalize the ``u''; if not, keep using lower-cased.
-\item Be aware of the different meanings of the homophones ``affect'' and ``effect'', ``complement'' and ``compliment'', ``discreet'' and ``discrete'', ``principal'' and ``principle''.
-\item Do not confuse ``imply'' and ``infer''.
-\item The prefix ``non'' is not a word; it should be joined to the word it modifies, usually without a hyphen.
-\item There is no period after the ``et'' in the Latin abbreviation ``et al.''.
-\item The abbreviation ``i.e.'' means ``that is'', and the abbreviation ``e.g.'' means ``for example''.
-\end{itemize}
-An excellent style manual for science writers is \cite{b7}.
-
-\subsection{Authors and Affiliations}
-\textbf{The class file is designed for, but not limited to, six authors.} A 
-minimum of one author is required for all conference articles. Author names 
-should be listed starting from left to right and then moving down to the 
-next line. This is the author sequence that will be used in future citations 
-and by indexing services. Names should not be listed in columns nor group by 
-affiliation. Please keep your affiliations as succinct as possible (for 
-example, do not differentiate among departments of the same organization).
-
-\subsection{Identify the Headings}
-Headings, or heads, are organizational devices that guide the reader through 
-your paper. There are two types: component heads and text heads.
-
-Component heads identify the different components of your paper and are not 
-topically subordinate to each other. Examples include Acknowledgments and 
-References and, for these, the correct style to use is ``Heading 5''. Use 
-``figure caption'' for your Figure captions, and ``table head'' for your 
-table title. Run-in heads, such as ``Abstract'', will require you to apply a 
-style (in this case, italic) in addition to the style provided by the drop 
-down menu to differentiate the head from the text.
-
-Text heads organize the topics on a relational, hierarchical basis. For 
-example, the paper title is the primary text head because all subsequent 
-material relates and elaborates on this one topic. If there are two or more 
-sub-topics, the next level head (uppercase Roman numerals) should be used 
-and, conversely, if there are not at least two sub-topics, then no subheads 
-should be introduced.
-
-\subsection{Figures and Tables}
-\paragraph{Positioning Figures and Tables} Place figures and tables at the top and 
-bottom of columns. Avoid placing them in the middle of columns. Large 
-figures and tables may span across both columns. Figure captions should be 
-below the figures; table heads should appear above the tables. Insert 
-figures and tables after they are cited in the text. Use the abbreviation 
-``Fig.~\ref{fig}'', even at the beginning of a sentence.
-
-\begin{table}[htbp]
-\caption{Table Type Styles}
-\begin{center}
-\begin{tabular}{|c|c|c|c|}
-\hline
-\textbf{Table}&\multicolumn{3}{|c|}{\textbf{Table Column Head}} \\
-\cline{2-4} 
-\textbf{Head} & \textbf{\textit{Table column subhead}}& \textbf{\textit{Subhead}}& \textbf{\textit{Subhead}} \\
-\hline
-copy& More table copy$^{\mathrm{a}}$& &  \\
-\hline
-\multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
-\end{tabular}
-\label{tab1}
-\end{center}
-\end{table}
-
-Figure Labels: Use 8 point Times New Roman for Figure labels. Use words 
-rather than symbols or abbreviations when writing Figure axis labels to 
-avoid confusing the reader. As an example, write the quantity 
-``Magnetization'', or ``Magnetization, M'', not just ``M''. If including 
-units in the label, present them within parentheses. Do not label axes only 
-with units. In the example, write ``Magnetization (A/m)'' or ``Magnetization 
-\{A[m(1)]\}'', not just ``A/m''. Do not label axes with a ratio of 
-quantities and units. For example, write ``Temperature (K)'', not 
-``Temperature/K''.
-
-\section*{Acknowledgment}
-
-The preferred spelling of the word ``acknowledgment'' in America is without 
-an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B. 
-G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor 
-acknowledgments in the unnumbered footnote on the first page.
-
-\section*{References}
-
-Please number citations consecutively within brackets \cite{b1}. The 
-sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference 
-number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at 
-the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''
-
-Number footnotes separately in superscripts. Place the actual footnote at 
-the bottom of the column in which it was cited. Do not put footnotes in the 
-abstract or reference list. Use letters for table footnotes.
-
-Unless there are six authors or more give all authors' names; do not use 
-``et al.''. Papers that have not been published, even if they have been 
-submitted for publication, should be cited as ``unpublished'' \cite{b4}. Papers 
-that have been accepted for publication should be cited as ``in press'' \cite{b5}. 
-Capitalize only the first word in a paper title, except for proper nouns and 
-element symbols.
-
-For papers published in translation journals, please give the English 
-citation first, followed by the original foreign-language citation \cite{b6}.
-
-\bibliographystyle{IEEEtran}
-\bibliography{refs}
-
-% use bibfile instead (2 lines above)
-% \begin{thebibliography}{00}
-% \bibitem{b1} G.
-Eason, B.
-Noble, and I.
-N.
-Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil.
-Trans.
-Roy.
-Soc.
-London, vol.
-A247, pp.
-529--551, April 1955.
-% \bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
-% \bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
-% \bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
-% \bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
-% \bibitem{b6} Y.
-Yorozu, M.
-Hirano, K.
-Oka, and Y.
-Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl.
-J.
-Magn.
-Japan, vol.
-2, pp.
-740--741, August 1987 [Digests 9th Annual Conf.
-Magnetics Japan, p.
-301, 1982].
-% \bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
-% \end{thebibliography}
-\vspace{12pt}
-\color{red}
-IEEE conference templates contain guidance text for composing and formatting conference papers.
-Please ensure that all template text is removed from your conference paper prior to submission to the conference.
-Failure to remove the template text from your paper may result in your paper not being published.
-
-\end{document}
+This work is structured as follows: subsection \ref{chap:background}  describes the SGM and its optimizations.
+Furthermore, we will give an extended explanation of the gradient descent optimizers used in this work.
+The reader is introduced to our Implementation in subsection \ref{chap:implementation}.
+Results are presented in  subsection \ref{chap:results} , where we will also describe used datasets, the measure applied to compare the quality of word embeddings, and finally the empirical results.
+The last part will focus on the discussion of our results, and possible future work in subsection \ref{chap:discussion} followed by a conclusion in subsection \ref{chap:conclusion}.

### 2019-05-01 icdm/submission/includes/related_work.tex
Splitted papaer into multiple files

--- 2019-04-30 icdm/submission/paper.tex
@@
-\documentclass[conference]{IEEEtran}
-\IEEEoverridecommandlockouts
-% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
-\usepackage{cite}
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{algorithmic}
-\usepackage{graphicx}
-\usepackage{textcomp}
-\usepackage{xcolor}
-\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
-    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
-\begin{document}
+\section{Related Work}
+Due to the popularity of the SGM, a lot of research went into optimizing it.
+This research can be divided into two categories, optimization of the throughput and the optimization of the algorithm's accuracy.
+The latter was achieved by allowing words to have multiple meanings, also called context-sensitive word embedings.
+For our work, the optimization of the throughput is of big interest while the semantic optimization is aimed at giving the reader a more holistic comprehension of the possible research directions.
+This subsection will first give an overview of the optimization of the throughput and then present one paper that focused on context-sensitive word embeddings.
 
-\title{Paper Title*\\
-{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
-should not be used}
-\thanks{Identify applicable funding agency here. If none, delete this.}
-}
+\subsection{Optimization of the throughput}
 
-\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-\and
-\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-\and
-\IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-\and
-\IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-\and
-\IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-\and
-\IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-}
+In the original model, the optimization is done with Stochastic Gradient Descent (SGD), which is a sequential algorithm.
+This process does not favor parallelization.
+To deal with this specific problem Mikolov et al.\cite{mikolov2} used a Hogwild tree proposed by Recht et al.\cite{hogwild}.
+The approach is to allow multiple threads to access a shared memory, in this case, the single model.
+In the original SGM, the threads are constructed as follows: at the beginning of training, the dataset is split into $N$ evenly sized chunks and each of these chunks will be processed by a one thread.
+The threads run parallel and have access to the shared memory.
+Therefore, overwriting errors are bound to happen.
+According to Recht et al.\cite{hogwild} the overwriting errors won't lead to a significant accuracy loss if the data is sparse enough.
+But in the case of NLP, the problem seems to be a bit more significant, and especially for word embedding, as many words share the same context words.
+There were several attempts at solving this issue, and we are going to cover a few of them in the following subsubsections.
 
-\maketitle
+\paragraph{Parallelization by the use of caching}
+This idea was proposed by Vuurens et al.
+\cite{efficient}.
+The architecture used here is the basic skip gram model with a hierarchical softmax.
+The general idea is to cache the most frequently used nodes of the binary tree used to memorize the probability distribution and update them on the shared single model after a certain number of seen words (the paper used the number 10).
+The paper produced interesting results as they managed to decrease execution time by increasing the number of cores used for the calculation.
+This is very powerful because in the original implementation the execution time regressed after using more than 8 cores.
+It seems to indicate that too much overwriting was happening, as the number of concurrent threads surpasses a certain threshold.
+This can be seen in Figure \ref{fig:efficient}, where c31 is the model proposed by Vuurens et al.\cite{efficient}.
+The model did not suffer any accuracy loss in comparison to the original SGM model.
+\begin{figure}[ht]
+\centering
+\includegraphics[scale=0.3]{images/cachingEfficiency.png}
+\caption{Comparasion of the execution time in relation to the number of used cores \cite{efficient}}
+\label{fig:efficient}
+\end{figure}
+This work proposes a very good way to parallelize the SGM, as in particular, it allows using more cores during the computation.
+As this approached focused on the Hierarchical softmax, in contrast to our work which used negative sampling, the next subsection covers optimizations of the Skip-Gram Model with negative sampling(SGNS).
 
-\begin{abstract}
-This document is a model and instructions for \LaTeX.
-This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
-or Math in Paper Title or Abstract.
-\end{abstract}
+\paragraph{Parallization in shared and Distributed Memory}
+The first parallelized solution which was proposed by Ji et al.
+\cite{intel}, is to try to reduce the cost of our vector multiplication.
+The main idea in this paper is to convert the level 1-BLAS (Basic linear subprogram) vector to vector operations to a level 3-BLAS matrix multiplication operation, which are efficiently implemented into hardware and consequently faster.
+This is achieved, by using the same negative samples for each context word of a given word $w$.
+Instead of using for each context word a vector to vector multiplication we can transform this into a matrix multiplication, under the assumption that we will not lose accuracy by sharing the same negative samples.
+The matrix multiplication can be represented in the following way.
+\[
+\begin{bmatrix}
+w \\
+w_{n_1} \\
+\vdots \\
+w_{n_k}\\
+\end{bmatrix}
+*
+\begin{bmatrix}
+w_{c_1}\\
+\vdots\\
+w_{c_{2m}}\\
+\end{bmatrix}
+\]
 
-\begin{IEEEkeywords}
-component, formatting, style, styling, insert
-\end{IEEEkeywords}
+where $w$ is our given word, $w_{n_1}...w_{n_k}$ are the shared negative samples, with $k \in [5,20]$, and $w_{c_1}...w_{c_2m}$ are the words inside of the context window $m$ of $w$, with $m \in [10,20]$, also called a batch of input context words.
+After each batch the model updates the weights of the used vectors.
+This model achieves a 3.6 fold increase in throughput, by only losing 1\% of accuracy.
+An aspect that is not as useful to us is that the experiments were done on CPU, as modern GPU's are often used in many machine learning libraries, as with CUDA for example, there still need work to be done to optimize it with GPU's.
+This was done by Seulki and Youngmin \cite{gpu}, which is described in the next subsection.
 
-\section{Introduction}
-This document is a model and instructions for \LaTeX.
-Please observe the conference page limits.
+\paragraph{Accelleration of word2vec by Using GPU's}
+Seulki and Youngmin \cite{gpu} focused on getting a better throughput on the SGM when using GPU's.
+As the SGM is a sequential algorithm, is is not easy to parallelize it, especially if one wants to parallelize the training of individual training samples.
+As the algorithm goes sequentially over a sentence, the samples next to each other, in order of execution, will almost every time have the same input word.
+Consequently, it's very hard to parallelize at this level.
+To solve this problem, Seulki and Youngmin \cite{gpu} proposed the idea to parallelize the update of each dimension of the word embedding, since those are completely independent of each other.
+They achieved this by mapping each dimension to a CUDA thread while mapping each sentence to a CUDA block.
+As each CUDA block runs independently, the training of the sentences is parallelized, and the fact that sentences have different length is of no problem.
+If the execution time of the GPU kernel is greater than time used to read the sentences, it could be a smart choice to use multiple GPU's.
+According to Seulki and Youngmin \cite{gpu}, if multiple GPU's are used, there is a need for synchronizing the model, which will hinder run time performance.
+They achieved their best results with 2 concurrent GPU'S.
+The accomplished results were very good as they managed a 20x speedup compared to a single threaded CPU execution, which is a 3x increase in comparison to the original C code from Mikolov et al.
+\cite{mikolov2}, with no loss in accuracy.
+The problem with this and all the above optimization is that the code is not easily available.
+Therefore, we need an optimized implementation of the SGM that is easily available.
+This is provided by Gensim \cite{gensim}, which will be outlined in the next subsection.
+%%%%%%%%%%%
+\paragraph{Gensim}\label{ssec:gensim}
+Gensim \cite{gensim} is a pure Python library that holds a state of the art implementation of the SGM.
+Gensim is written in Cython, which first allowed Gensim to have the same runtime as the original C code.
+Furthermore, it made use of BLAS's and precomputed sigmoid tables, while also trying to parallelize the training of different sentences.
+This finally yielded in a 4x speedup in run time.
+Gensim is an important tool as it allows us, as a python library, to compare our data rather easily.
+It was also used in related work \cite{intel} and is therefore of value, as it allows to us compare our work in a simpler way.
+This concludes our overview of the optimizations of the throughput of the SGM.
+In the next subsection, we give a quick outlook of what has been done in the field of context-sensitive word embeddings.
 
-\section{Ease of Use}
-
-\subsection{Maintaining the Integrity of the Specifications}
-
-The IEEEtran class file is used to format your paper and style the text. All margins, 
-column widths, line spaces, and text fonts are prescribed; please do not 
-alter them. You may note peculiarities. For example, the head margin
-measures proportionately more than is customary. This measurement 
-and others are deliberate, using specifications that anticipate your paper 
-as one part of the entire proceedings, and not as an independent document. 
-Please do not revise any of the current designations.
-
-\section{Prepare Your Paper Before Styling}
-Before you begin to format your paper, first write and save the content as a 
-separate text file. Complete all content and organizational editing before 
-formatting. Please note sections \ref{AA}--\ref{SCM} below for more information on 
-proofreading, spelling and grammar.
-
-Keep your text and graphic files separate until after the text has been 
-formatted and styled. Do not number text heads---{\LaTeX} will do that 
-for you.
-
-\subsection{Abbreviations and Acronyms}\label{AA}
-Define abbreviations and acronyms the first time they are used in the text, 
-even after they have been defined in the abstract. Abbreviations such as 
-IEEE, SI, MKS, CGS, ac, dc, and rms do not have to be defined. Do not use 
-abbreviations in the title or heads unless they are unavoidable.
-
-\subsection{Units}
-\begin{itemize}
-\item Use either SI (MKS) or CGS as primary units.
-(SI units are encouraged.) English units may be used as secondary units (in parentheses).
-An exception would be the use of English units as identifiers in trade, such as ``3.5-inch disk drive''.
-\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds.
-This often leads to confusion because equations do not balance dimensionally.
-If you must use mixed units, clearly state the units for each quantity that you use in an equation.
-\item Do not mix complete spellings and abbreviations of units: ``Wb/m\textsuperscript{2}'' or ``webers per square meter'', not ``webers/m\textsuperscript{2}''.
-Spell out units when they appear in text: ``.
-.
-.
-a few henries'', not ``.
-.
-.
-a few H''.
-\item Use a zero before decimal points: ``0.25'', not ``.25''. Use ``cm\textsuperscript{3}'', not ``cc''.)
-\end{itemize}
-
-\subsection{Equations}
-Number equations consecutively. To make your 
-equations more compact, you may use the solidus (~/~), the exp function, or 
-appropriate exponents. Italicize Roman symbols for quantities and variables, 
-but not Greek symbols. Use a long dash rather than a hyphen for a minus 
-sign. Punctuate equations with commas or periods when they are part of a 
-sentence, as in:
+\subsubsection{Context sensitive word embedding}
+A word has different meanings dependent on its context.
+This is a problem that is not addressed by the SGM.
+Some new models, that have taken this issue into consideration, were proposed.
+A lot of work has been done in this direction, Liu et al.\cite{topicalWE} and Bartunov et al.\cite{breaking} for example, but the one reporting the best results is Liu et al.
+\cite{contextWithTensor}.
+The concept of this approach is to change the way we compute the objective function and variables we use in our conditional probability.
+The idea is to look if a word given a certain context word matches to a topic.
+\textit{Bank} would match to \textit{finance} given the context word \textit{money}.
+\textit{Bank} would also match to \textit{nature} if \textit{river} was the given context word.
+But \textit{Bank} would not match to \textit{nature} with the context word \textit{money}.
+Now one could ask how to achieve such a context sensitive word embedding?
+First, we have to introduce new variables, therefore let's look at the objective function used:
 \begin{equation}
-a+b=\gamma\label{eq}
+J(\Omega) = \sum_{(w,t,c)\in D} \sum_{(w,\tilde{t},\tilde{c} \in{\tilde{D}})} max(0,1- g(w,t,c) + g(w,\tilde{t},\tilde{c})) \lambda||\Omega||_{2}^2
 \end{equation}
 
-Be sure that the 
-symbols in your equation have been defined before or immediately following 
-the equation. Use ``\eqref{eq}'', not ``Eq.~\eqref{eq}'' or ``equation \eqref{eq}'', except at 
-the beginning of a sentence: ``Equation \eqref{eq} is . . .''
+This approach uses the same negative sample technique as described in the previous subsections, $D$ is the corpus data and $\tilde{D}$ is the set of negative samples and $\lambda$is the hyper parameter used for the standard $L_2$ standardization.
+What is interesting here is the function $g(w,c,t)$, where $w$ is a word, $c$ the context word, and $t$ the context in which the word appears.
+$g$ is defined as follows:
+\begin{equation}
+g(w,c,t) = u^T \sigma(w^TM^{[1:k]}t+V_c^T(w \oplus t) + b_c)
+\end{equation}
+where, $u, V_c, b_c$ are standard parameters for a neural network.
+$\oplus$ is the vector concatenation.
+The most important parameter is $M^{[1:k]}$, which is a tensor layer, the tensor layer is used because of its ability to model multiple interactions in the data, as this will be useful for multiple contexts.
+They used SGD for the optimization of this objective function.
+They achieved interesting results as shown in \ref{fig:multipleContext}.\\
+\begin{figure}[ht]
+\centering
+\includegraphics[scale=0.7]{images/multipleContext.png}
+\caption{"Nearest neighbor words model and Skip-
+Gram. The first line in each block is the results of Skip-Gram;
+and the rest lines are the results of our model" \cite{contextWithTensor}}
+\label{fig:multipleContext}
+\end{figure}
 
-\subsection{\LaTeX-Specific Advice}
-
-Please use ``soft'' (e.g., \verb|\eqref{Eq}|) cross references instead
-of ``hard'' references (e.g., \verb|(1)|). That will make it possible
-to combine sections, add equations, or change the order of figures or
-citations without having to go through the file line by line.
-
-Please don't use the \verb|{eqnarray}| equation environment. Use
-\verb|{align}| or \verb|{IEEEeqnarray}| instead. The \verb|{eqnarray}|
-environment leaves unsightly spaces around relation symbols.
-
-Please note that the \verb|{subequations}| environment in {\LaTeX}
-will increment the main equation counter even when there are no
-equation numbers displayed. If you forget that, you might write an
-article in which the equation numbers skip from (17) to (20), causing
-the copy editors to wonder if you've discovered a new method of
-counting.
-
-{\BibTeX} does not work by magic. It doesn't get the bibliographic
-data from thin air but from .bib files. If you use {\BibTeX} to produce a
-bibliography you must send the .bib files.
-
-{\LaTeX} can't read your mind. If you assign the same label to a
-subsubsection and a table, you might find that Table I has been cross
-referenced as Table IV-B3.
-
-{\LaTeX} does not have precognitive abilities. If you put a
-\verb|\label| command before the command that updates the counter it's
-supposed to be using, the label will pick up the last counter to be
-cross referenced instead. In particular, a \verb|\label| command
-should not go before the caption of a figure or a table.
-
-Do not use \verb|\nonumber| inside the \verb|{array}| environment. It
-will not stop equation numbers inside \verb|{array}| (there won't be
-any anyway) and it might stop a wanted equation number in the
-surrounding equation.
-
-\subsection{Some Common Mistakes}\label{SCM}
-\begin{itemize}
-\item The word ``data'' is plural, not singular.
-\item The subscript for the permeability of vacuum $\mu_{0}$, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ``o''.
-\item In American English, commas, semicolons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation.
-When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks.
-A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this).
-(A parenthetical sentence is punctuated within the parentheses.)
-\item A graph within a graph is an ``inset'', not an ``insert''. The word alternatively is preferred to the word ``alternately'' (unless you really mean something that alternates).
-\item Do not use the word ``essentially'' to mean ``approximately'' or ``effectively''.
-\item In your paper title, if the words ``that uses'' can accurately replace the word ``using'', capitalize the ``u''; if not, keep using lower-cased.
-\item Be aware of the different meanings of the homophones ``affect'' and ``effect'', ``complement'' and ``compliment'', ``discreet'' and ``discrete'', ``principal'' and ``principle''.
-\item Do not confuse ``imply'' and ``infer''.
-\item The prefix ``non'' is not a word; it should be joined to the word it modifies, usually without a hyphen.
-\item There is no period after the ``et'' in the Latin abbreviation ``et al.''.
-\item The abbreviation ``i.e.'' means ``that is'', and the abbreviation ``e.g.'' means ``for example''.
-\end{itemize}
-An excellent style manual for science writers is \cite{b7}.
-
-\subsection{Authors and Affiliations}
-\textbf{The class file is designed for, but not limited to, six authors.} A 
-minimum of one author is required for all conference articles. Author names 
-should be listed starting from left to right and then moving down to the 
-next line. This is the author sequence that will be used in future citations 
-and by indexing services. Names should not be listed in columns nor group by 
-affiliation. Please keep your affiliations as succinct as possible (for 
-example, do not differentiate among departments of the same organization).
-
-\subsection{Identify the Headings}
-Headings, or heads, are organizational devices that guide the reader through 
-your paper. There are two types: component heads and text heads.
-
-Component heads identify the different components of your paper and are not 
-topically subordinate to each other. Examples include Acknowledgments and 
-References and, for these, the correct style to use is ``Heading 5''. Use 
-``figure caption'' for your Figure captions, and ``table head'' for your 
-table title. Run-in heads, such as ``Abstract'', will require you to apply a 
-style (in this case, italic) in addition to the style provided by the drop 
-down menu to differentiate the head from the text.
-
-Text heads organize the topics on a relational, hierarchical basis. For 
-example, the paper title is the primary text head because all subsequent 
-material relates and elaborates on this one topic. If there are two or more 
-sub-topics, the next level head (uppercase Roman numerals) should be used 
-and, conversely, if there are not at least two sub-topics, then no subheads 
-should be introduced.
-
-\subsection{Figures and Tables}
-\paragraph{Positioning Figures and Tables} Place figures and tables at the top and 
-bottom of columns. Avoid placing them in the middle of columns. Large 
-figures and tables may span across both columns. Figure captions should be 
-below the figures; table heads should appear above the tables. Insert 
-figures and tables after they are cited in the text. Use the abbreviation 
-``Fig.~\ref{fig}'', even at the beginning of a sentence.
-
-\begin{table}[htbp]
-\caption{Table Type Styles}
-\begin{center}
-\begin{tabular}{|c|c|c|c|}
-\hline
-\textbf{Table}&\multicolumn{3}{|c|}{\textbf{Table Column Head}} \\
-\cline{2-4} 
-\textbf{Head} & \textbf{\textit{Table column subhead}}& \textbf{\textit{Subhead}}& \textbf{\textit{Subhead}} \\
-\hline
-copy& More table copy$^{\mathrm{a}}$& &  \\
-\hline
-\multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
-\end{tabular}
-\label{tab1}
-\end{center}
-\end{table}
-
-Figure Labels: Use 8 point Times New Roman for Figure labels. Use words 
-rather than symbols or abbreviations when writing Figure axis labels to 
-avoid confusing the reader. As an example, write the quantity 
-``Magnetization'', or ``Magnetization, M'', not just ``M''. If including 
-units in the label, present them within parentheses. Do not label axes only 
-with units. In the example, write ``Magnetization (A/m)'' or ``Magnetization 
-\{A[m(1)]\}'', not just ``A/m''. Do not label axes with a ratio of 
-quantities and units. For example, write ``Temperature (K)'', not 
-``Temperature/K''.
-
-\section*{Acknowledgment}
-
-The preferred spelling of the word ``acknowledgment'' in America is without 
-an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B. 
-G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor 
-acknowledgments in the unnumbered footnote on the first page.
-
-\section*{References}
-
-Please number citations consecutively within brackets \cite{b1}. The 
-sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference 
-number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at 
-the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''
-
-Number footnotes separately in superscripts. Place the actual footnote at 
-the bottom of the column in which it was cited. Do not put footnotes in the 
-abstract or reference list. Use letters for table footnotes.
-
-Unless there are six authors or more give all authors' names; do not use 
-``et al.''. Papers that have not been published, even if they have been 
-submitted for publication, should be cited as ``unpublished'' \cite{b4}. Papers 
-that have been accepted for publication should be cited as ``in press'' \cite{b5}. 
-Capitalize only the first word in a paper title, except for proper nouns and 
-element symbols.
-
-For papers published in translation journals, please give the English 
-citation first, followed by the original foreign-language citation \cite{b6}.
-
-\bibliographystyle{IEEEtran}
-\bibliography{refs}
-
-% use bibfile instead (2 lines above)
-% \begin{thebibliography}{00}
-% \bibitem{b1} G.
-Eason, B.
-Noble, and I.
-N.
-Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil.
-Trans.
-Roy.
-Soc.
-London, vol.
-A247, pp.
-529--551, April 1955.
-% \bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
-% \bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
-% \bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
-% \bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
-% \bibitem{b6} Y.
-Yorozu, M.
-Hirano, K.
-Oka, and Y.
-Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl.
-J.
-Magn.
-Japan, vol.
-2, pp.
-740--741, August 1987 [Digests 9th Annual Conf.
-Magnetics Japan, p.
-301, 1982].
-% \bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
-% \end{thebibliography}
-\vspace{12pt}
-\color{red}
-IEEE conference templates contain guidance text for composing and formatting conference papers.
-Please ensure that all template text is removed from your conference paper prior to submission to the conference.
-Failure to remove the template text from your paper may result in your paper not being published.
-
-\end{document}
+This will conclude our overview of the related work. We will now give the reader an outline of the different Gradient Descent Optimizer used in our experiments.

### 2019-05-01 icdm/submission/includes/results.tex
Splitted papaer into multiple files

--- 2019-04-30 icdm/submission/paper.tex
@@
-\documentclass[conference]{IEEEtran}
-\IEEEoverridecommandlockouts
-% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
-\usepackage{cite}
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{algorithmic}
-\usepackage{graphicx}
-\usepackage{textcomp}
-\usepackage{xcolor}
-\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
-    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
-\begin{document}
+\section{Results}\label{chap:results}
 
-\title{Paper Title*\\
-{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
-should not be used}
-\thanks{Identify applicable funding agency here. If none, delete this.}
-}
+%Describe the experimental setup, the used datasets/parameters and the experimental results achieved
 
-\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-\and
-\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-\and
-\IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-\and
-\IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-\and
-\IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-\and
-\IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-}
+This subsection gives an overview of the used datasets, the used metric to evaluate our models, the configuration of our model and finally, the experimental results achieved.
 
-\maketitle
+\subsection{Dataset}\label{sec:dataset}
+In this implementation we only used the text8 \footnote{http://mattmahoney.net/dc/enwik8.zip} dataset.
+We chose this dataset for two reasons.
+First of all, it's a very small dataset, that allowed us to do a lot of computations.
+Secondly, this dataset was used in related work \cite{intel} hence giving us a very good benchmark.
+The text8 dataset consists of 1702 lines of 1000 words, with a vocabulary of roughly 63000 words.
+Conveniently, there is no punctuation in the dataset.
+Therefore, we had to choose between building arbitrary sentences and keeping the dataset as it is.
+We chose the first option because it gives us a faster computation time, and did not show any significant loss in quality empirically, as shown in Table \ref{table:with_20}.
+We chose a length of sentences of 20.
+Furthermore, we applied a technique called subsampling to reduce the data set size, which is explained in the following subsection.
 
-\begin{abstract}
-This document is a model and instructions for \LaTeX.
-This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
-or Math in Paper Title or Abstract.
-\end{abstract}
+\begin{table}[]
+\centering
+\begin{tabular}{|l|l|l|l|l|}
+\hline
+Length of Sentences & 10 & 20 & 40 & 1 Document \\ \hline
+Training Time for one batch &8min & 10min & 11min & 18min \\ \hline
+Convergence Time &4 epochs & 3 epochs & 3 epochs & 3 epochs \\ \hline
+Word Similarity& 0.65 & 0.66 & 0.66 & 0.66 \\ \hline
+\end{tabular}
+\caption{Training and Convergence time according to choice of the length of sentences in text8 dataset}
+\label{table:with_20}
+\end{table}
 
-\begin{IEEEkeywords}
-component, formatting, style, styling, insert
-\end{IEEEkeywords}
+\subsubsection{Subsampling}
+Subsampling is a technique introduced by Mikolov et al.
+\cite{mikolov} to reduce the dataset size while at the same time increasing the quality of the dataset, i.e getting better word embeddings with it.
+The idea behind subsampling is the removal of very frequent words such as: "the, as, it, of".
+These words do not give an intrinsic value to the words that appear in its context.
+Therefore, the goal of subsampling is to delete such words from the dataset.
+This will decrease the computation time, as it will reduce the number of training samples, and should, in theory, increase the accuracy of the model.
+The increase in accuracy can also be explained by the fact that words that would not have appeared in the context of each other, may now do because words between have been deleted.
+To choose which word to delete, Mikolov et al. \cite{mikolov2} chose the following equation to compute the deletion of a word $w$ in the data set:
+\begin{equation} \label{eq:sampling}
+P(w) = 1- \sqrt{{\frac{t}{f(w)}}}
+\end{equation}
 
-\section{Introduction}
-This document is a model and instructions for \LaTeX.
-Please observe the conference page limits.
+where $f(w)$ is the frequency of w, and $t$ is a threshold set empirically.
+As Equation \ref{eq:sampling} is a probability, subsampling is not a deterministic procedure, words that may have been deleted with a threshold of $10^-2$ may stay in the dataset with a lower threshold, as shown in Table \ref{table:treshold_examples}.
+Mikolov et al.
+recommend a value between $0$ and $10^{-5}$, depending on the size of the dataset.
+We experimented with different values and $10^{-4}$ seemed the most suited.
+We did this by looking at a random set of sentences and judging the results manually.
+An example of the first sentence with different sampling thresholds can be found in Table \ref{table:treshold_examples}.
+The table shows the first 20 words of our dataset, without the words that were subsampled according to a threshold sample.
+Stats about subsampling can be found in Table \ref{table:treshold}.
+\begin{table}[h]
+\centering
+\begin{tabular}{|l|l|l|l|l|l|}
+\hline
+Sampling Treshhold & 0 & $ 10^{-1}$&$ 10^{-2}$& $10^{-3} $ &$10^{-4} $ \\ \hline
+Number of words in Dataset & 16 mio & 15mio & 11 mio & 8mio & 4 mio \\ \hline
+\end{tabular}
+\caption{Size of the preprocessed text8 dataset according to sampling treshold}
+\label{table:treshold}
+\end{table}
 
-\section{Ease of Use}
+\begin{table}[]
+\begin{tabular}{|l|l|}
+\hline
+Sampling Treshold & First sentence of Dataset \\ \hline
+$10^{-1}$ & \begin{tabular}[c]{@{}l@{}}Anarchism originated as a term of abuse first used against early\\ working class radicals including the diggers of the english\end{tabular} \\ \hline
+$10^{-2}$ & \begin{tabular}[c]{@{}l@{}}Anarchism originated as a term of abuse first used against early \\ working class radicals including diggers of english\end{tabular} \\ \hline
+$10^{-3}$ & \begin{tabular}[c]{@{}l@{}}Anarchism originated a term abuse first used against early \\ working class radicals including diggers the english\end{tabular} \\ \hline
+$10^{-4}$ & \begin{tabular}[c]{@{}l@{}}Anarchism originated abuse used against working class radicals\\ diggers english\end{tabular} \\ \hline
+$10^{-5}$ & against radicals diggers \\ \hline
+\end{tabular}
+\caption{Example of a sentence with different sampling tresholds}\label{table:treshold_examples}
+\end{table}
 
-\subsection{Maintaining the Integrity of the Specifications}
+\subsubsection{Min count}
+We also deleted every word that did not appear more than 5 times in our dataset.
+We got this technique from Gensim \cite{gensim}, that introduced this parameter into their training.
+This is a good technique because of three reasons: First certain words of our data sets do not appear in a common lexicon (twigleg, melqu), or come from a foreign language (Volksvereinigung), or are names and acronyms.
+Secondly, each document often has spelling mistakes, those (as long as the same spelling mistake does not appear too often, what should be avoided in practice) would be deleted by sampling too, as the words do not have any meaning.
+Lastly, a word that only appears one time in our dataset will be very dependent on its original initialization.
+This is the case because it will only be updated with its context pairs once, which is only a dozen of times in practice and then won't be updated anymore.
+For all the above reasons, we applied this technique.
+Similar to subsampling, it should in theory improve the quality of the word embeddings and will decrease the computation time.
 
-The IEEEtran class file is used to format your paper and style the text. All margins, 
-column widths, line spaces, and text fonts are prescribed; please do not 
-alter them. You may note peculiarities. For example, the head margin
-measures proportionately more than is customary. This measurement 
-and others are deliberate, using specifications that anticipate your paper 
-as one part of the entire proceedings, and not as an independent document. 
-Please do not revise any of the current designations.
+\subsection{Evaluating word embedings}
+Evaluating word embedding is not an easy task, such as evaluating the accuracy of a common classifier.
+We cannot split our data set into train and test set.
+As the task that the network is learning is of no interest to us.
+Therefore, we need to verify that our embedding is of quality with other techniques.
+To define quality we first need to define a measure of similarity between two vectors.
+This requires knowledge of the Cosine distance, which is introduced in the following subsection.
+\subsubsection{Cosine distance}
+The cosine similarity, this is not the cosine distance, of vectors $v$ and $w$ is the cosine of the angle between the two vectors It can be calculated by taking the dot product of $v$ and $w$ and dividing it by the magnitude of $v$ and $w$ multiplied with each other.
+We get:
+\begin{equation}
+cos\_sim(v,w) = \frac{v \cdot w}{|v| |w|}
+\end{equation}
+The cosine of 0\textdegree ~is 1, it's 0 for two vectors that are orthogonal to each other and vectors that point in the opposite direction will have a cosine of their angle of -1.
+This is not a good distance measure as -1 is smaller than 0, and therefore two vectors pointing away from each other would be closer than two orthogonal vectors, but by subtracting 1 from the cosine of the angle we can create a good distance measure between the two vectors.
+This distance does not take into account any order of magnitude.
+Hence, for our tasks, two vectors will be considered equal if they are of different magnitude but point in the same direction.
+Apart from measuring the quality of word embedding well, this technique has another advantage.
+By normalizing the vectors the calculation of the cosine angle becomes the dot product of the two vectors.
+Which can be computed very fast on modern GPU's.
+Knowing that we have a measure to compute the similarity of two vectors let us introduce a way to rate the quality of our embeddings.
 
-\section{Prepare Your Paper Before Styling}
-Before you begin to format your paper, first write and save the content as a 
-separate text file. Complete all content and organizational editing before 
-formatting. Please note sections \ref{AA}--\ref{SCM} below for more information on 
-proofreading, spelling and grammar.
+\subsubsection{Word similarity and wordsim353}
+To measure the quality of our word embedding we will need a dataset to compare our results too.
+We chose wordsim353\footnote{http://www.cs.technion.ac.il/~gabr/resources/data/wordsim353/wordsim353.zip} for this task, as it's the most used in the related literature.
+The data set consists of 353 pairs of words rated by humans on their similarity.
+The similarity score is in the range of 1 and 10, with 10 being the best score.
+An example for two of such pairs can be found in Figure \ref{fig:ws353_ex}.
+We will rank our embeddings on the Pearson correlation coefficient between the cosine distance and the scores attributed by humans, as this is a common procedure.
 
-Keep your text and graphic files separate until after the text has been 
-formatted and styled. Do not number text heads---{\LaTeX} will do that 
-for you.
+\begin{table}[]
+\centering
+\begin{tabular}{|l|l|l|}
+\hline
+Word1 & Word2 & Score \\ \hline
+\textquote{FBI} & \textquote{Investigation} & \textquote{8.31} \\ \hline
+\textquote{Mars} & \textquote{scientist} & \textquote{5.63} \\ \hline
+\end{tabular}
+\caption{Example of pairs and their rating in wordsim353}
+\label{fig:ws353_ex}
+\end{table}
 
-\subsection{Abbreviations and Acronyms}\label{AA}
-Define abbreviations and acronyms the first time they are used in the text, 
-even after they have been defined in the abstract. Abbreviations such as 
-IEEE, SI, MKS, CGS, ac, dc, and rms do not have to be defined. Do not use 
-abbreviations in the title or heads unless they are unavoidable.
-
-\subsection{Units}
+\subsection{Configuration of the network}
+The skip gram model has a lot of possible parameters, that can be tuned.
+We experimented with different models and finally decided for one that we tried to optimize.
+This subsection will give a short overview of each parameter, where we will explain the process in which we chose the value of the given parameter.
+The explanation of the parameters will be structured as follows:
+\texttt{Parameter} - Description and tuning - \textit{Value}
 \begin{itemize}
-\item Use either SI (MKS) or CGS as primary units.
-(SI units are encouraged.) English units may be used as secondary units (in parentheses).
-An exception would be the use of English units as identifiers in trade, such as ``3.5-inch disk drive''.
-\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds.
-This often leads to confusion because equations do not balance dimensionally.
-If you must use mixed units, clearly state the units for each quantity that you use in an equation.
-\item Do not mix complete spellings and abbreviations of units: ``Wb/m\textsuperscript{2}'' or ``webers per square meter'', not ``webers/m\textsuperscript{2}''.
-Spell out units when they appear in text: ``.
-.
-.
-a few henries'', not ``.
-.
-.
-a few H''.
-\item Use a zero before decimal points: ``0.25'', not ``.25''. Use ``cm\textsuperscript{3}'', not ``cc''.)
+\item \texttt{Negative Samples:} Here we have to find a trade-off between, setting the parameter too high which will result in increased accuracy but a longer computation time.
+For smaller data sets a higher number of negative samples is often needed.
+In their original paper Mikolov et al.
+\cite{mikolov2} recommend a value of 5-20.
+We tested a few values in the range of 5 to 15, as 10 yielded state of the art results we chose this value.
+- \textit{10}
+\item \texttt{Context Window:} The bigger the window the more training examples the network will have, but if the window is too big the semantic meaning of the window will be erased.
+Mikolov et al.
+\cite{mikolov} proposed a setting between 2-10, as all our sentences are of size 20 we chose 5.
+- \textit{5}
+\item\texttt{ Dimension of the embedding}: Here the choice is less obvious, as the dimension needs to be high enough to capture the meaning, but cannot be too high as this leads to a decrease in performance as shown by Yin and Shen \cite{dimension_size}.
+We, therefore, used Gensim to find the best embedding possible.
+- \textit{100}
+\item \texttt{Batch size}: As described in subsection \ref{ssec:b_SGM}, there is a tradeoff to find between quality and training time.
+We first used a batch size of 5000, but then decide after non conclusive results (see \ref{ssec:bs_lf}) that 2000 would be better - \textit{2000}
+\item \texttt{Alpha}: learning rate, this hyperparameter was tuned in every optimizer therefore only the range will be indicated - \textit{(1e-5,1)}
 \end{itemize}
 
-\subsection{Equations}
-Number equations consecutively. To make your 
-equations more compact, you may use the solidus (~/~), the exp function, or 
-appropriate exponents. Italicize Roman symbols for quantities and variables, 
-but not Greek symbols. Use a long dash rather than a hyphen for a minus 
-sign. Punctuate equations with commas or periods when they are part of a 
-sentence, as in:
-\begin{equation}
-a+b=\gamma\label{eq}
-\end{equation}
+\iffalse
+\begin{table}[h]
+\centering
+\begin{tabular}{|l|l|}
+\hline
+Embedding Size & Word Similarity on Gensim \\ \hline
+50 & 0.65 \\ \hline
+100 & 0.67 \\ \hline
+200 & 0.65 \\ \hline
+300 & 0.63 \\ \hline
+\end{tabular}
+\caption{Word similarity in relation to the size of the embedding}
+\label{table:gensim_emb_size}
+\end{table}
+\fi
+\subsection{Input Shuffling}
+We used input shuffling as a technique to optimize the skip gram model.
+We will first describe input shuffling in a general way and then explain why we suppose that input shuffling could work well on the skip gram model.
+\\
+Let $X = {x_1...x_n}$ be our input data set. Input Shuffling describes the process of taking a random permutation of the dataset as input at each epoch.
+The idea behind this technique is to present our optimizer with different loss surfaces so that it's able to find the best optimum.
+Therefore, it's easier for the neural network to escape a local minimum.
+As for example if a network had converged to a local minimum after one epoch it could not escape it as all the parameters are the same.
+But if we change the shape of the loss function, by input shuffling, then there would be a greater probability for the network to escape the local minimum.
+\\
+There are two reasons why we think that input shuffling is particularly well suited for the skip gram model.
+The first one has to do with the fact that when we read our words sequentially that words that only appear very early will not benefit from the context words being already updated from others.
+The second idea is that we used the special batch technique described in Section \ref{ssec:b_SGM}.
+When using this technique and not using shuffling we will always have words that appear next to each other in a batch and will, therefore, update the same words at the same time.
+We, therefore, lose some quality.
+But if we would use input shuffling instead, then the words in one batch would likely not be similar and therefore only taking the average of a small part of pairs with the same words will be less likely.
 
-Be sure that the 
-symbols in your equation have been defined before or immediately following 
-the equation. Use ``\eqref{eq}'', not ``Eq.~\eqref{eq}'' or ``equation \eqref{eq}'', except at 
-the beginning of a sentence: ``Equation \eqref{eq} is . . .''
+\subsection{Convergence time}
+To optimize convergence time we have to define it first.
+Therefore, we used the already available implementation of Gensim \cite{gensim}.
+Gensim is an open source software that proposes an implementation of the SGNS in Python.
+It is also written in Cython, therefore it has a fast computation time, but can be used inside a python implementation.
+Together with the knowledge from Ji et al.\cite{intel} that a score of $0.66$ in the task of word similarity, with the text8 dataset, is the state of the art, we tested Gensim (more on this process in subsection \ref{chap:discussion} and found out that it took 4 epochs to converge.
+Therefore, we defined the following criteria for convergence: \\
+$\rho - \rho_{prev} < 0.009 \vee \rho = 0.66$ \\
+where $\rho$ is the Pearson coefficient on the wordsim353 task.
+We also stopped computation, if it took more then 20 epochs to converge.
 
-\subsection{\LaTeX-Specific Advice}
+\subsection{Results by optimizer}
+We ran multiple experiments for each optimizer.
+This subsection will only give an overview of the achieved results.
+Each subsection will give an explanation over the achieved result with a specific optimizer.
 
-Please use ``soft'' (e.g., \verb|\eqref{Eq}|) cross references instead
-of ``hard'' references (e.g., \verb|(1)|). That will make it possible
-to combine sections, add equations, or change the order of figures or
-citations without having to go through the file line by line.
+\subsubsection{SGD}
+The first challenge for each optimizer was to find a correct learning rate.
+As SGD is the optimizer used in Gensim \cite{gensim} we first tried the same learning rate as Gensim \cite{gensim} and then performed a random search to find a better one.
+As expected a bell curve shape resulted, a learning rate that is too high leads to diversion and a learning rate that is too low leads to a training time that is too slow.
+The best value that we found for the learning rate is $0.0075$.
+With this setting SGD converged in 11 epochs.
+The second experiment was to add input shuffling.
+As seen in Figure \ref{fig:results_sgd}, for almost every learning rate the convergence time decreased.
+Our model, with the best setting, now converges in only 7 epochs.
+Another interesting fact to point out from Figure \ref{fig:results_sgd} is that with input shuffling we achieved better results with higher learning rates.
+As for learning rates of $0.01$ and $0.025$ we did converge in 11 epochs with input shuffling but did not converge in 20 epochs without it.
 
-Please don't use the \verb|{eqnarray}| equation environment. Use
-\verb|{align}| or \verb|{IEEEeqnarray}| instead. The \verb|{eqnarray}|
-environment leaves unsightly spaces around relation symbols.
-
-Please note that the \verb|{subequations}| environment in {\LaTeX}
-will increment the main equation counter even when there are no
-equation numbers displayed. If you forget that, you might write an
-article in which the equation numbers skip from (17) to (20), causing
-the copy editors to wonder if you've discovered a new method of
-counting.
-
-{\BibTeX} does not work by magic. It doesn't get the bibliographic
-data from thin air but from .bib files. If you use {\BibTeX} to produce a
-bibliography you must send the .bib files.
-
-{\LaTeX} can't read your mind. If you assign the same label to a
-subsubsection and a table, you might find that Table I has been cross
-referenced as Table IV-B3.
-
-{\LaTeX} does not have precognitive abilities. If you put a
-\verb|\label| command before the command that updates the counter it's
-supposed to be using, the label will pick up the last counter to be
-cross referenced instead. In particular, a \verb|\label| command
-should not go before the caption of a figure or a table.
-
-Do not use \verb|\nonumber| inside the \verb|{array}| environment. It
-will not stop equation numbers inside \verb|{array}| (there won't be
-any anyway) and it might stop a wanted equation number in the
-surrounding equation.
-
-\subsection{Some Common Mistakes}\label{SCM}
-\begin{itemize}
-\item The word ``data'' is plural, not singular.
-\item The subscript for the permeability of vacuum $\mu_{0}$, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ``o''.
-\item In American English, commas, semicolons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation.
-When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks.
-A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this).
-(A parenthetical sentence is punctuated within the parentheses.)
-\item A graph within a graph is an ``inset'', not an ``insert''. The word alternatively is preferred to the word ``alternately'' (unless you really mean something that alternates).
-\item Do not use the word ``essentially'' to mean ``approximately'' or ``effectively''.
-\item In your paper title, if the words ``that uses'' can accurately replace the word ``using'', capitalize the ``u''; if not, keep using lower-cased.
-\item Be aware of the different meanings of the homophones ``affect'' and ``effect'', ``complement'' and ``compliment'', ``discreet'' and ``discrete'', ``principal'' and ``principle''.
-\item Do not confuse ``imply'' and ``infer''.
-\item The prefix ``non'' is not a word; it should be joined to the word it modifies, usually without a hyphen.
-\item There is no period after the ``et'' in the Latin abbreviation ``et al.''.
-\item The abbreviation ``i.e.'' means ``that is'', and the abbreviation ``e.g.'' means ``for example''.
-\end{itemize}
-An excellent style manual for science writers is \cite{b7}.
-
-\subsection{Authors and Affiliations}
-\textbf{The class file is designed for, but not limited to, six authors.} A 
-minimum of one author is required for all conference articles. Author names 
-should be listed starting from left to right and then moving down to the 
-next line. This is the author sequence that will be used in future citations 
-and by indexing services. Names should not be listed in columns nor group by 
-affiliation. Please keep your affiliations as succinct as possible (for 
-example, do not differentiate among departments of the same organization).
-
-\subsection{Identify the Headings}
-Headings, or heads, are organizational devices that guide the reader through 
-your paper. There are two types: component heads and text heads.
-
-Component heads identify the different components of your paper and are not 
-topically subordinate to each other. Examples include Acknowledgments and 
-References and, for these, the correct style to use is ``Heading 5''. Use 
-``figure caption'' for your Figure captions, and ``table head'' for your 
-table title. Run-in heads, such as ``Abstract'', will require you to apply a 
-style (in this case, italic) in addition to the style provided by the drop 
-down menu to differentiate the head from the text.
-
-Text heads organize the topics on a relational, hierarchical basis. For 
-example, the paper title is the primary text head because all subsequent 
-material relates and elaborates on this one topic. If there are two or more 
-sub-topics, the next level head (uppercase Roman numerals) should be used 
-and, conversely, if there are not at least two sub-topics, then no subheads 
-should be introduced.
-
-\subsection{Figures and Tables}
-\paragraph{Positioning Figures and Tables} Place figures and tables at the top and 
-bottom of columns. Avoid placing them in the middle of columns. Large 
-figures and tables may span across both columns. Figure captions should be 
-below the figures; table heads should appear above the tables. Insert 
-figures and tables after they are cited in the text. Use the abbreviation 
-``Fig.~\ref{fig}'', even at the beginning of a sentence.
-
-\begin{table}[htbp]
-\caption{Table Type Styles}
-\begin{center}
-\begin{tabular}{|c|c|c|c|}
+\begin{figure}[h]
+\centering
+\includegraphics[scale=0.45]{images/results_sgd_shuffle}
+\caption{Training time Stochastic Gradient Descent with input Shuffling}
+\label{fig:results_sgd}
+\end{figure}
+\subsubsection{Momentum and Nesterov}
+Momentum and NAG \cite{nag} both have an additional hyperparameter $\gamma$, that, as described in Section \ref{optimizers}, defines the percentage of the previous gradient that will be added to the current gradients.
+We set $\gamma = 0.9$ as this is the typical value and did not alter it during our experiments.
+Momentum and Nesterov alone respectively only slightly decrease or increase the convergence time.
+Momentum optimally converges in 9 epochs and Nesterov in 13.
+If we combine these optimizers with input shuffling, interestingly the same phenomena as with plain SGD appear.
+The convergence time gets better, 8 epochs for Momentum and 3 epochs for NAG.
+The phenomena that a higher learning rat yields better results also happens with both of the optimizers.
+As Momentum does not converge in 20 epochs with a learning rate of 0.002 but does in 8 with input shuffling.
+\begin{figure}[h]
+\centering
+\begin{minipage}{.5\textwidth}
+\centering
+\includegraphics[scale=0.3]{images/results_mom_shuffle}
+\caption{Training time Momentum with input Shuffling}
+\label{fig:results_mom}
+\label{fig:test1}
+\end{minipage}%
+\begin{minipage}{.5\textwidth}
+\centering
+\includegraphics[scale=0.3]{images/results_nag_shuffle}
+\caption{Training time Nesterov with input Shuffling}
+\label{fig:results_nag}
+\end{minipage}
+\end{figure}
+\subsubsection{Adagrad}
+Adagrad \cite{adagrad} is a very interesting tool for learning word embeddings as it decreases the learning rate for very frequent occurring features, and vice versa (this is explained in detail in Section \ref{ssec:adagrad}).
+Because words that appear very frequently often do not have a semantic gain that is as important as words that appear less frequently to their context words, it's good to have a lower learning rate.
+So, in theory, Adagrad is particularly well suited for our task.
+This was confirmed empirically as our model converged in 4 epochs.
+When combined with shuffling Adagrad only took 3 epochs to converge.
+This shows the tendency of the skip gram model to converge faster with input shuffling and the big impact of having different learning rate for each feature.
+Here it's interesting to notice that a higher learning rate combined with input shuffling did not yield better results than without shuffling.
+Both of our best results happened with a learning rate of $0.1$, as shown in Figure \ref{fig:results_adagrad_shuffle}.
+\begin{figure}[h]
+\centering
+\includegraphics[scale=0.45]{images/results_adagrad_shuffle}
+\caption{Training time Adagrad with input Shuffling}
+\label{fig:results_adagrad_shuffle}
+\end{figure}
+\subsubsection{Adadelta}
+In theory Adadelta \cite{adadelta} should outperform Adagrad as it's an extension of the former.
+Because it didn't have any learning rate to tune, we only did 2 experiments, with and without input shuffling.
+We are aware of the fact that there are additional hyperparameters but did, for simplicity reason, and because their effect is not as high as the learning rate in other optimizers, decide not to tune it.
+We left it to its default value $\rho = 0.9$.
+This parameter defines the percentage taken when calculating the exponentially decaying average of past gradients, as explained in \ref{ssec:adadelta} Adadelta did not manage to achieve a word similarity of 0.66.
+It only converged to a similarity of 0.59.
+It did this in 20 epochs without input shuffling and in 3 with input shuffling, as can be seen in Table \ref{table:results_adadelta}
+\begin{table}[h]
+\centering
+\begin{tabular}{|l|l|l|}
 \hline
-\textbf{Table}&\multicolumn{3}{|c|}{\textbf{Table Column Head}} \\
-\cline{2-4} 
-\textbf{Head} & \textbf{\textit{Table column subhead}}& \textbf{\textit{Subhead}}& \textbf{\textit{Subhead}} \\
-\hline
-copy& More table copy$^{\mathrm{a}}$& &  \\
-\hline
-\multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
+Adadelta Model & Convergence Time & Word similarity \\ \hline
+Without Shuffling & 20 & 0.59 \\ \hline
+With Shuffling & 3 & 0.59 \\ \hline
 \end{tabular}
-\label{tab1}
-\end{center}
+\caption{Convergence Time and Quality with Adadelta}
+\label{table:results_adadelta}
 \end{table}
 
-Figure Labels: Use 8 point Times New Roman for Figure labels. Use words 
-rather than symbols or abbreviations when writing Figure axis labels to 
-avoid confusing the reader. As an example, write the quantity 
-``Magnetization'', or ``Magnetization, M'', not just ``M''. If including 
-units in the label, present them within parentheses. Do not label axes only 
-with units. In the example, write ``Magnetization (A/m)'' or ``Magnetization 
-\{A[m(1)]\}'', not just ``A/m''. Do not label axes with a ratio of 
-quantities and units. For example, write ``Temperature (K)'', not 
-``Temperature/K''.
-
-\section*{Acknowledgment}
-
-The preferred spelling of the word ``acknowledgment'' in America is without 
-an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B. 
-G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor 
-acknowledgments in the unnumbered footnote on the first page.
-
-\section*{References}
-
-Please number citations consecutively within brackets \cite{b1}. The 
-sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference 
-number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at 
-the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''
-
-Number footnotes separately in superscripts. Place the actual footnote at 
-the bottom of the column in which it was cited. Do not put footnotes in the 
-abstract or reference list. Use letters for table footnotes.
-
-Unless there are six authors or more give all authors' names; do not use 
-``et al.''. Papers that have not been published, even if they have been 
-submitted for publication, should be cited as ``unpublished'' \cite{b4}. Papers 
-that have been accepted for publication should be cited as ``in press'' \cite{b5}. 
-Capitalize only the first word in a paper title, except for proper nouns and 
-element symbols.
-
-For papers published in translation journals, please give the English 
-citation first, followed by the original foreign-language citation \cite{b6}.
-
-\bibliographystyle{IEEEtran}
-\bibliography{refs}
-
-% use bibfile instead (2 lines above)
-% \begin{thebibliography}{00}
-% \bibitem{b1} G.
-Eason, B.
-Noble, and I.
-N.
-Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil.
-Trans.
-Roy.
-Soc.
-London, vol.
-A247, pp.
-529--551, April 1955.
-% \bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
-% \bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
-% \bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
-% \bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
-% \bibitem{b6} Y.
-Yorozu, M.
-Hirano, K.
-Oka, and Y.
-Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl.
-J.
-Magn.
-Japan, vol.
-2, pp.
-740--741, August 1987 [Digests 9th Annual Conf.
-Magnetics Japan, p.
-301, 1982].
-% \bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
-% \end{thebibliography}
-\vspace{12pt}
-\color{red}
-IEEE conference templates contain guidance text for composing and formatting conference papers.
-Please ensure that all template text is removed from your conference paper prior to submission to the conference.
-Failure to remove the template text from your paper may result in your paper not being published.
-
-\end{document}
+\subsubsection{Adam}
+Adam is the most advanced of all the optimizers used in our experiments and did yield the best results as seen in Figure \ref{fig:results_adam_shuffle}.
+Adam converged in 3 epochs without shuffling and 2 with.
+This is the best result that we got with any optimizer.
+It's also interesting to note that as same as with Adagrad it did not react to input shuffling the same way as SGD did.
+In fact, it worked in the opposite direction, as we achieved our best result with input shuffling while having a lower learning rate $0.001$ then we used to achieve the best result without input shuffling $0.05$.
+\begin{figure}[h]
+    \centering
+            \includegraphics[scale=0.45]{images/results_adam_shuffle} 
+    \caption{Training time Adam with input Shuffling}
+    \label{fig:results_adam_shuffle}
+\end{figure}

### 2019-05-01 icdm/submission/paper.tex
Splitted papaer into multiple files

--- 2019-04-30 icdm/submission/paper.tex
@@
 
-\section{Introduction}
-This document is a model and instructions for \LaTeX.
-Please observe the conference page limits.
-
-\section{Ease of Use}
-
-\subsection{Maintaining the Integrity of the Specifications}
-
-The IEEEtran class file is used to format your paper and style the text. All margins, 
-column widths, line spaces, and text fonts are prescribed; please do not 
-alter them. You may note peculiarities. For example, the head margin
-measures proportionately more than is customary. This measurement 
-and others are deliberate, using specifications that anticipate your paper 
-as one part of the entire proceedings, and not as an independent document. 
-Please do not revise any of the current designations.
-
-\section{Prepare Your Paper Before Styling}
-Before you begin to format your paper, first write and save the content as a 
-separate text file. Complete all content and organizational editing before 
-formatting. Please note sections \ref{AA}--\ref{SCM} below for more information on 
-proofreading, spelling and grammar.
-
-Keep your text and graphic files separate until after the text has been 
-formatted and styled. Do not number text heads---{\LaTeX} will do that 
-for you.
-
-\subsection{Abbreviations and Acronyms}\label{AA}
-Define abbreviations and acronyms the first time they are used in the text, 
-even after they have been defined in the abstract. Abbreviations such as 
-IEEE, SI, MKS, CGS, ac, dc, and rms do not have to be defined. Do not use 
-abbreviations in the title or heads unless they are unavoidable.
-
-\subsection{Units}
-\begin{itemize}
-\item Use either SI (MKS) or CGS as primary units.
-(SI units are encouraged.) English units may be used as secondary units (in parentheses).
-An exception would be the use of English units as identifiers in trade, such as ``3.5-inch disk drive''.
-\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds.
-This often leads to confusion because equations do not balance dimensionally.
-If you must use mixed units, clearly state the units for each quantity that you use in an equation.
-\item Do not mix complete spellings and abbreviations of units: ``Wb/m\textsuperscript{2}'' or ``webers per square meter'', not ``webers/m\textsuperscript{2}''.
-Spell out units when they appear in text: ``.
-.
-.
-a few henries'', not ``.
-.
-.
-a few H''.
-\item Use a zero before decimal points: ``0.25'', not ``.25''. Use ``cm\textsuperscript{3}'', not ``cc''.)
-\end{itemize}
-
-\subsection{Equations}
-Number equations consecutively. To make your 
-equations more compact, you may use the solidus (~/~), the exp function, or 
-appropriate exponents. Italicize Roman symbols for quantities and variables, 
-but not Greek symbols. Use a long dash rather than a hyphen for a minus 
-sign. Punctuate equations with commas or periods when they are part of a 
-sentence, as in:
-\begin{equation}
-a+b=\gamma\label{eq}
-\end{equation}
-
-Be sure that the 
-symbols in your equation have been defined before or immediately following 
-the equation. Use ``\eqref{eq}'', not ``Eq.~\eqref{eq}'' or ``equation \eqref{eq}'', except at 
-the beginning of a sentence: ``Equation \eqref{eq} is . . .''
-
-\subsection{\LaTeX-Specific Advice}
-
-Please use ``soft'' (e.g., \verb|\eqref{Eq}|) cross references instead
-of ``hard'' references (e.g., \verb|(1)|). That will make it possible
-to combine sections, add equations, or change the order of figures or
-citations without having to go through the file line by line.
-
-Please don't use the \verb|{eqnarray}| equation environment. Use
-\verb|{align}| or \verb|{IEEEeqnarray}| instead. The \verb|{eqnarray}|
-environment leaves unsightly spaces around relation symbols.
-
-Please note that the \verb|{subequations}| environment in {\LaTeX}
-will increment the main equation counter even when there are no
-equation numbers displayed. If you forget that, you might write an
-article in which the equation numbers skip from (17) to (20), causing
-the copy editors to wonder if you've discovered a new method of
-counting.
-
-{\BibTeX} does not work by magic. It doesn't get the bibliographic
-data from thin air but from .bib files. If you use {\BibTeX} to produce a
-bibliography you must send the .bib files.
-
-{\LaTeX} can't read your mind. If you assign the same label to a
-subsubsection and a table, you might find that Table I has been cross
-referenced as Table IV-B3.
-
-{\LaTeX} does not have precognitive abilities. If you put a
-\verb|\label| command before the command that updates the counter it's
-supposed to be using, the label will pick up the last counter to be
-cross referenced instead. In particular, a \verb|\label| command
-should not go before the caption of a figure or a table.
-
-Do not use \verb|\nonumber| inside the \verb|{array}| environment. It
-will not stop equation numbers inside \verb|{array}| (there won't be
-any anyway) and it might stop a wanted equation number in the
-surrounding equation.
-
-\subsection{Some Common Mistakes}\label{SCM}
-\begin{itemize}
-\item The word ``data'' is plural, not singular.
-\item The subscript for the permeability of vacuum $\mu_{0}$, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ``o''.
-\item In American English, commas, semicolons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation.
-When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks.
-A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this).
-(A parenthetical sentence is punctuated within the parentheses.)
-\item A graph within a graph is an ``inset'', not an ``insert''. The word alternatively is preferred to the word ``alternately'' (unless you really mean something that alternates).
-\item Do not use the word ``essentially'' to mean ``approximately'' or ``effectively''.
-\item In your paper title, if the words ``that uses'' can accurately replace the word ``using'', capitalize the ``u''; if not, keep using lower-cased.
-\item Be aware of the different meanings of the homophones ``affect'' and ``effect'', ``complement'' and ``compliment'', ``discreet'' and ``discrete'', ``principal'' and ``principle''.
-\item Do not confuse ``imply'' and ``infer''.
-\item The prefix ``non'' is not a word; it should be joined to the word it modifies, usually without a hyphen.
-\item There is no period after the ``et'' in the Latin abbreviation ``et al.''.
-\item The abbreviation ``i.e.'' means ``that is'', and the abbreviation ``e.g.'' means ``for example''.
-\end{itemize}
-An excellent style manual for science writers is \cite{b7}.
-
-\subsection{Authors and Affiliations}
-\textbf{The class file is designed for, but not limited to, six authors.} A 
-minimum of one author is required for all conference articles. Author names 
-should be listed starting from left to right and then moving down to the 
-next line. This is the author sequence that will be used in future citations 
-and by indexing services. Names should not be listed in columns nor group by 
-affiliation. Please keep your affiliations as succinct as possible (for 
-example, do not differentiate among departments of the same organization).
-
-\subsection{Identify the Headings}
-Headings, or heads, are organizational devices that guide the reader through 
-your paper. There are two types: component heads and text heads.
-
-Component heads identify the different components of your paper and are not 
-topically subordinate to each other. Examples include Acknowledgments and 
-References and, for these, the correct style to use is ``Heading 5''. Use 
-``figure caption'' for your Figure captions, and ``table head'' for your 
-table title. Run-in heads, such as ``Abstract'', will require you to apply a 
-style (in this case, italic) in addition to the style provided by the drop 
-down menu to differentiate the head from the text.
-
-Text heads organize the topics on a relational, hierarchical basis. For 
-example, the paper title is the primary text head because all subsequent 
-material relates and elaborates on this one topic. If there are two or more 
-sub-topics, the next level head (uppercase Roman numerals) should be used 
-and, conversely, if there are not at least two sub-topics, then no subheads 
-should be introduced.
-
-\subsection{Figures and Tables}
-\paragraph{Positioning Figures and Tables} Place figures and tables at the top and 
-bottom of columns. Avoid placing them in the middle of columns. Large 
-figures and tables may span across both columns. Figure captions should be 
-below the figures; table heads should appear above the tables. Insert 
-figures and tables after they are cited in the text. Use the abbreviation 
-``Fig.~\ref{fig}'', even at the beginning of a sentence.
-
-\begin{table}[htbp]
-\caption{Table Type Styles}
-\begin{center}
-\begin{tabular}{|c|c|c|c|}
-\hline
-\textbf{Table}&\multicolumn{3}{|c|}{\textbf{Table Column Head}} \\
-\cline{2-4} 
-\textbf{Head} & \textbf{\textit{Table column subhead}}& \textbf{\textit{Subhead}}& \textbf{\textit{Subhead}} \\
-\hline
-copy& More table copy$^{\mathrm{a}}$& &  \\
-\hline
-\multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
-\end{tabular}
-\label{tab1}
-\end{center}
-\end{table}
-
-Figure Labels: Use 8 point Times New Roman for Figure labels. Use words 
-rather than symbols or abbreviations when writing Figure axis labels to 
-avoid confusing the reader. As an example, write the quantity 
-``Magnetization'', or ``Magnetization, M'', not just ``M''. If including 
-units in the label, present them within parentheses. Do not label axes only 
-with units. In the example, write ``Magnetization (A/m)'' or ``Magnetization 
-\{A[m(1)]\}'', not just ``A/m''. Do not label axes with a ratio of 
-quantities and units. For example, write ``Temperature (K)'', not 
-``Temperature/K''.
-
-\section*{Acknowledgment}
-
-The preferred spelling of the word ``acknowledgment'' in America is without 
-an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B. 
-G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor 
-acknowledgments in the unnumbered footnote on the first page.
-
-\section*{References}
-
-Please number citations consecutively within brackets \cite{b1}. The 
-sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference 
-number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at 
-the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''
-
-Number footnotes separately in superscripts. Place the actual footnote at 
-the bottom of the column in which it was cited. Do not put footnotes in the 
-abstract or reference list. Use letters for table footnotes.
-
-Unless there are six authors or more give all authors' names; do not use 
-``et al.''. Papers that have not been published, even if they have been 
-submitted for publication, should be cited as ``unpublished'' \cite{b4}. Papers 
-that have been accepted for publication should be cited as ``in press'' \cite{b5}. 
-Capitalize only the first word in a paper title, except for proper nouns and 
-element symbols.
-
-For papers published in translation journals, please give the English 
-citation first, followed by the original foreign-language citation \cite{b6}.
+\input{includes/introduction}
+\input{includes/background}
+\input{includes/related_work}
+\input{includes/contribution}
+\input{includes/results}
+\input{includes/conclusion}
 

### 2019-05-01 icdm/submission/paper.tex
removed unused bibliography and guidance text

@@
 
-% use bibfile instead (2 lines above)
-% \begin{thebibliography}{00}
-% \bibitem{b1} G.
-Eason, B.
-Noble, and I.
-N.
-Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil.
-Trans.
-Roy.
-Soc.
-London, vol.
-A247, pp.
-529--551, April 1955.
-% \bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
-% \bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
-% \bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
-% \bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
-% \bibitem{b6} Y.
-Yorozu, M.
-Hirano, K.
-Oka, and Y.
-Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl.
-J.
-Magn.
-Japan, vol.
-2, pp.
-740--741, August 1987 [Digests 9th Annual Conf.
-Magnetics Japan, p.
-301, 1982].
-% \bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
-% \end{thebibliography}
-\vspace{12pt}
-\color{red}
-IEEE conference templates contain guidance text for composing and formatting conference papers.
-Please ensure that all template text is removed from your conference paper prior to submission to the conference.
-Failure to remove the template text from your paper may result in your paper not being published.
-
 \end{document}

### 2019-05-01 icdm/submission/includes/background.tex
added example table and rescaled images

@@
 \section{Background}\label{chap:background}
+
+\begin{table*}[tb]\centering
+    \caption{Sample Caption}
+    %\scriptsize
+    \begin{tabular}{l r r r r r r r}%
+        \toprule
+        \textbf{App} & \textbf{NSGA-II} & \textbf{Random} & \textbf{Heuristic} & \textbf{Simple {ga}} & \textbf{MOSA} & \textbf{MIO} \\%
+        \midrule%
+        moneywallet        & 0.20 & 0.40 & 0.40 & 0.22 & 0.22 & 0.19 \\%
+        shoppinglist       & 0.77 & 0.78 & 0.84 & 0.77 & 0.77 & 0.77 \\%
+        periodical         & 0.85 & 0.86 & 0.87 & 0.84 & 0.83 & 0.85 \\%
+        drhoffmannsoftware & 0.44 & 0.48 & 0.45 & 0.44 & 0.46 & 0.48 \\%
+        activitydiary      & 0.74 & 0.82 & 0.82 & 0.73 & 0.74 & 0.75 \\%
+        bierverkostung     & 0.43 & 0.65 & 0.63 & 0.44 & 0.45 & 0.49 \\%
+        easy\_xkcd         & 0.62 & 0.73 & 0.68 & 0.57 & 0.59 & 0.60 \\%
+        markor             & 0.62 & 0.68 & 0.65 & 0.61 & 0.60 & 0.60 \\%
+        redreader          & 0.55 & 0.61 & 0.50 & 0.52 & 0.54 & 0.52 \\%
+        rentalcalc         & 0.67 & 0.77 & 0.76 & 0.68 & 0.68 & 0.71 \\%
+        \midrule%
+        Mean               & 0.59 & 0.68 & 0.66 & 0.58 & 0.59 & 0.60 \\%
+        \bottomrule%
+   \end{tabular}%
+   \label{tab:sample_label}%
+\end{table*}
 This subsection gives an overview of the SGM and related work on its optimization. Furthermore, the used optimizers in our experiments are explained.
@@
 \centering
-\includegraphics[scale=0.3]{images/window_ex}
+\includegraphics[scale=0.18]{images/window_ex}
 \caption{Example of a center word and it's context words}
@@
 \\&= log(\frac{1}{1+e^{-v_c \tilde{v_w } }}) + \sum_{k\in K} log(\frac{1}{1+e^{v_c \tilde{v_k} }})
-\\&= log(\sigma(v_c \tilde{v_w } ) + \sum_{k\in K} \sigma(log(-v_c \tilde{v_k} )) &&\text{where, $\sigma(x) = \frac{1}{1+e^{-x}}$}
+\\&= log(\sigma(v_c \tilde{v_w } ) + \sum_{k\in K} \sigma(log(-v_c \tilde{v_k} ))
 \end{align*}\label{eq:obj_neg_samples}
 
-Where $v_c$ and $\tilde{v_w }$, can be interpreted as before in Equation \ref{eq:softmax}.
+Where $v_c$ and $\tilde{v_w }$, can be interpreted as before in Equation \ref{eq:softmax} and $\sigma(x) = \frac{1}{1+e^{-x}}$.
 The goal is to maximize this objective function.
@@
 \centering
-\includegraphics[scale=0.45]{images/frequency_ex}
+\includegraphics[scale=0.30]{images/frequency_ex}
 \caption{Probabililty of a word, of the text8 dataset (sampled), to be chosen at random according to its frequency and the power to which the unigram distribution is raised}

### 2019-05-01 icdm/submission/includes/results.tex
splitted results into evalutation and results
updated adadelta table

@@
-\section{Results}\label{chap:results}
+\section{Evalutation}\label{chap:results}
 
@@
 
-\subsection{Results by optimizer}
+\subsection{Results}
 We ran multiple experiments for each optimizer.
@@
 \hline
-Adadelta Model & Convergence Time & Word similarity \\ \hline
-Without Shuffling & 20 & 0.59 \\ \hline
-With Shuffling & 3 & 0.59 \\ \hline
+
 \end{tabular}
+\caption{}
+\label{}
+\end{table}
+
+\begin{table}[tb]
     \caption{Convergence Time and Quality with Adadelta}
-\label{table:results_adadelta}
+    %\scriptsize
+    \begin{tabular}{l r r }%
+        \toprule
+Adadelta Model & Convergence Time & Word similarity \\ 
+        \midrule%
+        Without Shuffling & 20 & 0.59 \\ 
+With Shuffling & 3 & 0.59 \\
+        \midrule%
+   \end{tabular}%
+   \label{table:results_adadelta}%
 \end{table}

### 2019-05-01 icdm/submission/includes/evaluation.tex
new file for evalutation and changes to tables

--- 2019-05-01 icdm/submission/includes/results.tex
@@
 
-\begin{table}[]
-\centering
-\begin{tabular}{|l|l|l|l|l|}
-\hline
-Length of Sentences & 10 & 20 & 40 & 1 Document \\ \hline
-Training Time for one batch &8min & 10min & 11min & 18min \\ \hline
-Convergence Time &4 epochs & 3 epochs & 3 epochs & 3 epochs \\ \hline
+\begin{table}[tb]
+    \caption{Training and Convergence time according to choice of the length of sentences in text8 datasetCaption}
+    %\scriptsize
+    \begin{tabular}{l r r r r}%
+        \toprule
+          Length of Sentences & 10 & 20 & 40 & 1 Document \\
+
+        \midrule%
+Training Time for one batch &8min & 10min & 11min & 18min \\ 
+Convergence Time in epochs &4  & 3  & 3  & 3  \\ 
 Word Similarity& 0.65 & 0.66 & 0.66 & 0.66 \\ \hline
-\end{tabular}
-\caption{Training and Convergence time according to choice of the length of sentences in text8 dataset}
-\label{table:with_20}
+        \midrule%
+   \end{tabular}%
+   \label{table:with_20}%
 \end{table}
@@
 \hline
-Sampling Treshhold & 0 & $ 10^{-1}$&$ 10^{-2}$& $10^{-3} $ &$10^{-4} $ \\ \hline
-Number of words in Dataset & 16 mio & 15mio & 11 mio & 8mio & 4 mio \\ \hline
 \end{tabular}
-\caption{Size of the preprocessed text8 dataset according to sampling treshold}
+
 \label{table:treshold}
@@
 
-\begin{table}[]
-\begin{tabular}{|l|l|}
-\hline
+\begin{table}[tb]
+\caption{Size of the preprocessed text8 dataset according to sampling treshold}
+    %\scriptsize
+    \begin{tabular}{l r r r r r}%
+        \toprule
+Sampling Treshhold & 0 & $ 10^{-1}$&$ 10^{-2}$& $10^{-3} $ &$10^{-4} $ \\
+
+        \midrule%
+Words in Dataset  & 16mio & 15mio & 11mio & 8mio & 4mio \\ 
+        \midrule%
+   \end{tabular}%
+   \label{table:with_20}%
+\end{table}
+
+\begin{table}[tb]
+\caption{Example of a sentence with different sampling tresholds}
+    %\scriptsize
+    \begin{tabular}{l l}%
+        \toprule
+Sampling Treshhold & 0 \\
+
+        \midrule%
 Sampling Treshold & First sentence of Dataset \\ \hline
-$10^{-1}$ & \begin{tabular}[c]{@{}l@{}}Anarchism originated as a term of abuse first used against early\\ working class radicals including the diggers of the english\end{tabular} \\ \hline
-$10^{-2}$ & \begin{tabular}[c]{@{}l@{}}Anarchism originated as a term of abuse first used against early \\ working class radicals including diggers of english\end{tabular} \\ \hline
-$10^{-3}$ & \begin{tabular}[c]{@{}l@{}}Anarchism originated a term abuse first used against early \\ working class radicals including diggers the english\end{tabular} \\ \hline
-$10^{-4}$ & \begin{tabular}[c]{@{}l@{}}Anarchism originated abuse used against working class radicals\\ diggers english\end{tabular} \\ \hline
+$10^{-1}$ &Anarchism originated as a term of abuse first\tabularnewline used against early working class radicals \tabularnewline including the diggers of the english \\ \hline
+$10^{-2}$ & Anarchism originated as a term of abuse first used against early \tabularnewline working class radicals including diggers of english \\ \hline
+$10^{-3}$ & Anarchism originated a term abuse first used against early \tabularnewline working class radicals including diggers the english\\ \hline
+$10^{-4}$ & Anarchism originated abuse used against working class radicals\tabularnewline diggers english \\ \hline
 $10^{-5}$ & against radicals diggers \\ \hline        
-\end{tabular}
-\caption{Example of a sentence with different sampling tresholds}\label{table:treshold_examples}
+        \midrule%
+   \end{tabular}%
+\label{table:treshold_examples}
 \end{table}
@@
 We also stopped computation, if it took more then 20 epochs to converge.
-
-\subsection{Results}
-We ran multiple experiments for each optimizer.
-This subsection will only give an overview of the achieved results.
-Each subsection will give an explanation over the achieved result with a specific optimizer.
-
-\subsubsection{SGD}
-The first challenge for each optimizer was to find a correct learning rate.
-As SGD is the optimizer used in Gensim \cite{gensim} we first tried the same learning rate as Gensim \cite{gensim} and then performed a random search to find a better one.
-As expected a bell curve shape resulted, a learning rate that is too high leads to diversion and a learning rate that is too low leads to a training time that is too slow.
-The best value that we found for the learning rate is $0.0075$.
-With this setting SGD converged in 11 epochs.
-The second experiment was to add input shuffling.
-As seen in Figure \ref{fig:results_sgd}, for almost every learning rate the convergence time decreased.
-Our model, with the best setting, now converges in only 7 epochs.
-Another interesting fact to point out from Figure \ref{fig:results_sgd} is that with input shuffling we achieved better results with higher learning rates.
-As for learning rates of $0.01$ and $0.025$ we did converge in 11 epochs with input shuffling but did not converge in 20 epochs without it.
-
-\begin{figure}[h]
-\centering
-\includegraphics[scale=0.45]{images/results_sgd_shuffle}
-\caption{Training time Stochastic Gradient Descent with input Shuffling}
-\label{fig:results_sgd}
-\end{figure}
-\subsubsection{Momentum and Nesterov}
-Momentum and NAG \cite{nag} both have an additional hyperparameter $\gamma$, that, as described in Section \ref{optimizers}, defines the percentage of the previous gradient that will be added to the current gradients.
-We set $\gamma = 0.9$ as this is the typical value and did not alter it during our experiments.
-Momentum and Nesterov alone respectively only slightly decrease or increase the convergence time.
-Momentum optimally converges in 9 epochs and Nesterov in 13.
-If we combine these optimizers with input shuffling, interestingly the same phenomena as with plain SGD appear.
-The convergence time gets better, 8 epochs for Momentum and 3 epochs for NAG.
-The phenomena that a higher learning rat yields better results also happens with both of the optimizers.
-As Momentum does not converge in 20 epochs with a learning rate of 0.002 but does in 8 with input shuffling.
-\begin{figure}[h]
-\centering
-\begin{minipage}{.5\textwidth}
-\centering
-\includegraphics[scale=0.3]{images/results_mom_shuffle}
-\caption{Training time Momentum with input Shuffling}
-\label{fig:results_mom}
-\label{fig:test1}
-\end{minipage}%
-\begin{minipage}{.5\textwidth}
-\centering
-\includegraphics[scale=0.3]{images/results_nag_shuffle}
-\caption{Training time Nesterov with input Shuffling}
-\label{fig:results_nag}
-\end{minipage}
-\end{figure}
-\subsubsection{Adagrad}
-Adagrad \cite{adagrad} is a very interesting tool for learning word embeddings as it decreases the learning rate for very frequent occurring features, and vice versa (this is explained in detail in Section \ref{ssec:adagrad}).
-Because words that appear very frequently often do not have a semantic gain that is as important as words that appear less frequently to their context words, it's good to have a lower learning rate.
-So, in theory, Adagrad is particularly well suited for our task.
-This was confirmed empirically as our model converged in 4 epochs.
-When combined with shuffling Adagrad only took 3 epochs to converge.
-This shows the tendency of the skip gram model to converge faster with input shuffling and the big impact of having different learning rate for each feature.
-Here it's interesting to notice that a higher learning rate combined with input shuffling did not yield better results than without shuffling.
-Both of our best results happened with a learning rate of $0.1$, as shown in Figure \ref{fig:results_adagrad_shuffle}.
-\begin{figure}[h]
-\centering
-\includegraphics[scale=0.45]{images/results_adagrad_shuffle}
-\caption{Training time Adagrad with input Shuffling}
-\label{fig:results_adagrad_shuffle}
-\end{figure}
-\subsubsection{Adadelta}
-In theory Adadelta \cite{adadelta} should outperform Adagrad as it's an extension of the former.
-Because it didn't have any learning rate to tune, we only did 2 experiments, with and without input shuffling.
-We are aware of the fact that there are additional hyperparameters but did, for simplicity reason, and because their effect is not as high as the learning rate in other optimizers, decide not to tune it.
-We left it to its default value $\rho = 0.9$.
-This parameter defines the percentage taken when calculating the exponentially decaying average of past gradients, as explained in \ref{ssec:adadelta} Adadelta did not manage to achieve a word similarity of 0.66.
-It only converged to a similarity of 0.59.
-It did this in 20 epochs without input shuffling and in 3 with input shuffling, as can be seen in Table \ref{table:results_adadelta}
-\begin{table}[h]
-\centering
-\begin{tabular}{|l|l|l|}
-\hline
-
-\end{tabular}
-\caption{}
-\label{}
-\end{table}
-
-\begin{table}[tb]
-    \caption{Convergence Time and Quality with Adadelta}
-    %\scriptsize
-    \begin{tabular}{l r r }%
-        \toprule
-Adadelta Model & Convergence Time & Word similarity \\ 
-        \midrule%
-        Without Shuffling & 20 & 0.59 \\ 
-With Shuffling & 3 & 0.59 \\
-        \midrule%
-   \end{tabular}%
-   \label{table:results_adadelta}%
-\end{table}
-
-\subsubsection{Adam}
-Adam is the most advanced of all the optimizers used in our experiments and did yield the best results as seen in Figure \ref{fig:results_adam_shuffle}.
-Adam converged in 3 epochs without shuffling and 2 with.
-This is the best result that we got with any optimizer.
-It's also interesting to note that as same as with Adagrad it did not react to input shuffling the same way as SGD did.
-In fact, it worked in the opposite direction, as we achieved our best result with input shuffling while having a lower learning rate $0.001$ then we used to achieve the best result without input shuffling $0.05$.
-\begin{figure}[h]
-    \centering
-            \includegraphics[scale=0.45]{images/results_adam_shuffle} 
-    \caption{Training time Adam with input Shuffling}
-    \label{fig:results_adam_shuffle}
-\end{figure}

### 2019-05-01 icdm/submission/includes/evalutation.tex
new file for evalutation and changes to tables

--- 2019-05-01 icdm/submission/includes/results.tex
@@
 Furthermore, we applied a technique called subsampling to reduce the data set size, which is explained in the following subsection.
-
-\begin{table}[]
-\centering
-\begin{tabular}{|l|l|l|l|l|}
-\hline
-Length of Sentences & 10 & 20 & 40 & 1 Document \\ \hline
-Training Time for one batch &8min & 10min & 11min & 18min \\ \hline
-Convergence Time &4 epochs & 3 epochs & 3 epochs & 3 epochs \\ \hline
-Word Similarity& 0.65 & 0.66 & 0.66 & 0.66 \\ \hline
-\end{tabular}
-\caption{Training and Convergence time according to choice of the length of sentences in text8 dataset}
-\label{table:with_20}
-\end{table}
 

### 2019-05-01 icdm/submission/includes/results.tex
new file for evalutation and changes to tables

@@
-\section{Evalutation}\label{chap:results}
-
-%Describe the experimental setup, the used datasets/parameters and the experimental results achieved
-
-This subsection gives an overview of the used datasets, the used metric to evaluate our models, the configuration of our model and finally, the experimental results achieved.
-
-\subsection{Dataset}\label{sec:dataset}
-In this implementation we only used the text8 \footnote{http://mattmahoney.net/dc/enwik8.zip} dataset.
-We chose this dataset for two reasons.
-First of all, it's a very small dataset, that allowed us to do a lot of computations.
-Secondly, this dataset was used in related work \cite{intel} hence giving us a very good benchmark.
-The text8 dataset consists of 1702 lines of 1000 words, with a vocabulary of roughly 63000 words.
-Conveniently, there is no punctuation in the dataset.
-Therefore, we had to choose between building arbitrary sentences and keeping the dataset as it is.
-We chose the first option because it gives us a faster computation time, and did not show any significant loss in quality empirically, as shown in Table \ref{table:with_20}.
-We chose a length of sentences of 20.
-Furthermore, we applied a technique called subsampling to reduce the data set size, which is explained in the following subsection.
-
-\begin{table}[]
-\centering
-\begin{tabular}{|l|l|l|l|l|}
-\hline
-Length of Sentences & 10 & 20 & 40 & 1 Document \\ \hline
-Training Time for one batch &8min & 10min & 11min & 18min \\ \hline
-Convergence Time &4 epochs & 3 epochs & 3 epochs & 3 epochs \\ \hline
-Word Similarity& 0.65 & 0.66 & 0.66 & 0.66 \\ \hline
-\end{tabular}
-\caption{Training and Convergence time according to choice of the length of sentences in text8 dataset}
-\label{table:with_20}
-\end{table}
-
-\subsubsection{Subsampling}
-Subsampling is a technique introduced by Mikolov et al.
-\cite{mikolov} to reduce the dataset size while at the same time increasing the quality of the dataset, i.e getting better word embeddings with it.
-The idea behind subsampling is the removal of very frequent words such as: "the, as, it, of".
-These words do not give an intrinsic value to the words that appear in its context.
-Therefore, the goal of subsampling is to delete such words from the dataset.
-This will decrease the computation time, as it will reduce the number of training samples, and should, in theory, increase the accuracy of the model.
-The increase in accuracy can also be explained by the fact that words that would not have appeared in the context of each other, may now do because words between have been deleted.
-To choose which word to delete, Mikolov et al. \cite{mikolov2} chose the following equation to compute the deletion of a word $w$ in the data set:
-\begin{equation} \label{eq:sampling}
-P(w) = 1- \sqrt{{\frac{t}{f(w)}}}
-\end{equation}
-
-where $f(w)$ is the frequency of w, and $t$ is a threshold set empirically.
-As Equation \ref{eq:sampling} is a probability, subsampling is not a deterministic procedure, words that may have been deleted with a threshold of $10^-2$ may stay in the dataset with a lower threshold, as shown in Table \ref{table:treshold_examples}.
-Mikolov et al.
-recommend a value between $0$ and $10^{-5}$, depending on the size of the dataset.
-We experimented with different values and $10^{-4}$ seemed the most suited.
-We did this by looking at a random set of sentences and judging the results manually.
-An example of the first sentence with different sampling thresholds can be found in Table \ref{table:treshold_examples}.
-The table shows the first 20 words of our dataset, without the words that were subsampled according to a threshold sample.
-Stats about subsampling can be found in Table \ref{table:treshold}.
-\begin{table}[h]
-\centering
-\begin{tabular}{|l|l|l|l|l|l|}
-\hline
-Sampling Treshhold & 0 & $ 10^{-1}$&$ 10^{-2}$& $10^{-3} $ &$10^{-4} $ \\ \hline
-Number of words in Dataset & 16 mio & 15mio & 11 mio & 8mio & 4 mio \\ \hline
-\end{tabular}
-\caption{Size of the preprocessed text8 dataset according to sampling treshold}
-\label{table:treshold}
-\end{table}
-
-\begin{table}[]
-\begin{tabular}{|l|l|}
-\hline
-Sampling Treshold & First sentence of Dataset \\ \hline
-$10^{-1}$ & \begin{tabular}[c]{@{}l@{}}Anarchism originated as a term of abuse first used against early\\ working class radicals including the diggers of the english\end{tabular} \\ \hline
-$10^{-2}$ & \begin{tabular}[c]{@{}l@{}}Anarchism originated as a term of abuse first used against early \\ working class radicals including diggers of english\end{tabular} \\ \hline
-$10^{-3}$ & \begin{tabular}[c]{@{}l@{}}Anarchism originated a term abuse first used against early \\ working class radicals including diggers the english\end{tabular} \\ \hline
-$10^{-4}$ & \begin{tabular}[c]{@{}l@{}}Anarchism originated abuse used against working class radicals\\ diggers english\end{tabular} \\ \hline
-$10^{-5}$ & against radicals diggers \\ \hline
-\end{tabular}
-\caption{Example of a sentence with different sampling tresholds}\label{table:treshold_examples}
-\end{table}
-
-\subsubsection{Min count}
-We also deleted every word that did not appear more than 5 times in our dataset.
-We got this technique from Gensim \cite{gensim}, that introduced this parameter into their training.
-This is a good technique because of three reasons: First certain words of our data sets do not appear in a common lexicon (twigleg, melqu), or come from a foreign language (Volksvereinigung), or are names and acronyms.
-Secondly, each document often has spelling mistakes, those (as long as the same spelling mistake does not appear too often, what should be avoided in practice) would be deleted by sampling too, as the words do not have any meaning.
-Lastly, a word that only appears one time in our dataset will be very dependent on its original initialization.
-This is the case because it will only be updated with its context pairs once, which is only a dozen of times in practice and then won't be updated anymore.
-For all the above reasons, we applied this technique.
-Similar to subsampling, it should in theory improve the quality of the word embeddings and will decrease the computation time.
-
-\subsection{Evaluating word embedings}
-Evaluating word embedding is not an easy task, such as evaluating the accuracy of a common classifier.
-We cannot split our data set into train and test set.
-As the task that the network is learning is of no interest to us.
-Therefore, we need to verify that our embedding is of quality with other techniques.
-To define quality we first need to define a measure of similarity between two vectors.
-This requires knowledge of the Cosine distance, which is introduced in the following subsection.
-\subsubsection{Cosine distance}
-The cosine similarity, this is not the cosine distance, of vectors $v$ and $w$ is the cosine of the angle between the two vectors It can be calculated by taking the dot product of $v$ and $w$ and dividing it by the magnitude of $v$ and $w$ multiplied with each other.
-We get:
-\begin{equation}
-cos\_sim(v,w) = \frac{v \cdot w}{|v| |w|}
-\end{equation}
-The cosine of 0\textdegree ~is 1, it's 0 for two vectors that are orthogonal to each other and vectors that point in the opposite direction will have a cosine of their angle of -1.
-This is not a good distance measure as -1 is smaller than 0, and therefore two vectors pointing away from each other would be closer than two orthogonal vectors, but by subtracting 1 from the cosine of the angle we can create a good distance measure between the two vectors.
-This distance does not take into account any order of magnitude.
-Hence, for our tasks, two vectors will be considered equal if they are of different magnitude but point in the same direction.
-Apart from measuring the quality of word embedding well, this technique has another advantage.
-By normalizing the vectors the calculation of the cosine angle becomes the dot product of the two vectors.
-Which can be computed very fast on modern GPU's.
-Knowing that we have a measure to compute the similarity of two vectors let us introduce a way to rate the quality of our embeddings.
-
-\subsubsection{Word similarity and wordsim353}
-To measure the quality of our word embedding we will need a dataset to compare our results too.
-We chose wordsim353\footnote{http://www.cs.technion.ac.il/~gabr/resources/data/wordsim353/wordsim353.zip} for this task, as it's the most used in the related literature.
-The data set consists of 353 pairs of words rated by humans on their similarity.
-The similarity score is in the range of 1 and 10, with 10 being the best score.
-An example for two of such pairs can be found in Figure \ref{fig:ws353_ex}.
-We will rank our embeddings on the Pearson correlation coefficient between the cosine distance and the scores attributed by humans, as this is a common procedure.
-
-\begin{table}[]
-\centering
-\begin{tabular}{|l|l|l|}
-\hline
-Word1 & Word2 & Score \\ \hline
-\textquote{FBI} & \textquote{Investigation} & \textquote{8.31} \\ \hline
-\textquote{Mars} & \textquote{scientist} & \textquote{5.63} \\ \hline
-\end{tabular}
-\caption{Example of pairs and their rating in wordsim353}
-\label{fig:ws353_ex}
-\end{table}
-
-\subsection{Configuration of the network}
-The skip gram model has a lot of possible parameters, that can be tuned.
-We experimented with different models and finally decided for one that we tried to optimize.
-This subsection will give a short overview of each parameter, where we will explain the process in which we chose the value of the given parameter.
-The explanation of the parameters will be structured as follows:
-\texttt{Parameter} - Description and tuning - \textit{Value}
-\begin{itemize}
-\item \texttt{Negative Samples:} Here we have to find a trade-off between, setting the parameter too high which will result in increased accuracy but a longer computation time.
-For smaller data sets a higher number of negative samples is often needed.
-In their original paper Mikolov et al.
-\cite{mikolov2} recommend a value of 5-20.
-We tested a few values in the range of 5 to 15, as 10 yielded state of the art results we chose this value.
-- \textit{10}
-\item \texttt{Context Window:} The bigger the window the more training examples the network will have, but if the window is too big the semantic meaning of the window will be erased.
-Mikolov et al.
-\cite{mikolov} proposed a setting between 2-10, as all our sentences are of size 20 we chose 5.
-- \textit{5}
-\item\texttt{ Dimension of the embedding}: Here the choice is less obvious, as the dimension needs to be high enough to capture the meaning, but cannot be too high as this leads to a decrease in performance as shown by Yin and Shen \cite{dimension_size}.
-We, therefore, used Gensim to find the best embedding possible.
-- \textit{100}
-\item \texttt{Batch size}: As described in subsection \ref{ssec:b_SGM}, there is a tradeoff to find between quality and training time.
-We first used a batch size of 5000, but then decide after non conclusive results (see \ref{ssec:bs_lf}) that 2000 would be better - \textit{2000}
-\item \texttt{Alpha}: learning rate, this hyperparameter was tuned in every optimizer therefore only the range will be indicated - \textit{(1e-5,1)}
-\end{itemize}
-
-\iffalse
-\begin{table}[h]
-\centering
-\begin{tabular}{|l|l|}
-\hline
-Embedding Size & Word Similarity on Gensim \\ \hline
-50 & 0.65 \\ \hline
-100 & 0.67 \\ \hline
-200 & 0.65 \\ \hline
-300 & 0.63 \\ \hline
-\end{tabular}
-\caption{Word similarity in relation to the size of the embedding}
-\label{table:gensim_emb_size}
-\end{table}
-\fi
-\subsection{Input Shuffling}
-We used input shuffling as a technique to optimize the skip gram model.
-We will first describe input shuffling in a general way and then explain why we suppose that input shuffling could work well on the skip gram model.
-\\
-Let $X = {x_1...x_n}$ be our input data set. Input Shuffling describes the process of taking a random permutation of the dataset as input at each epoch.
-The idea behind this technique is to present our optimizer with different loss surfaces so that it's able to find the best optimum.
-Therefore, it's easier for the neural network to escape a local minimum.
-As for example if a network had converged to a local minimum after one epoch it could not escape it as all the parameters are the same.
-But if we change the shape of the loss function, by input shuffling, then there would be a greater probability for the network to escape the local minimum.
-\\
-There are two reasons why we think that input shuffling is particularly well suited for the skip gram model.
-The first one has to do with the fact that when we read our words sequentially that words that only appear very early will not benefit from the context words being already updated from others.
-The second idea is that we used the special batch technique described in Section \ref{ssec:b_SGM}.
-When using this technique and not using shuffling we will always have words that appear next to each other in a batch and will, therefore, update the same words at the same time.
-We, therefore, lose some quality.
-But if we would use input shuffling instead, then the words in one batch would likely not be similar and therefore only taking the average of a small part of pairs with the same words will be less likely.
-
-\subsection{Convergence time}
-To optimize convergence time we have to define it first.
-Therefore, we used the already available implementation of Gensim \cite{gensim}.
-Gensim is an open source software that proposes an implementation of the SGNS in Python.
-It is also written in Cython, therefore it has a fast computation time, but can be used inside a python implementation.
-Together with the knowledge from Ji et al.\cite{intel} that a score of $0.66$ in the task of word similarity, with the text8 dataset, is the state of the art, we tested Gensim (more on this process in subsection \ref{chap:discussion} and found out that it took 4 epochs to converge.
-Therefore, we defined the following criteria for convergence: \\
-$\rho - \rho_{prev} < 0.009 \vee \rho = 0.66$ \\
-where $\rho$ is the Pearson coefficient on the wordsim353 task.
-We also stopped computation, if it took more then 20 epochs to converge.
-
 \subsection{Results}

### 2019-05-01 icdm/submission/includes/evaluation.tex
table and figure form

@@
 
-\begin{table}[tb]
+\begin{table*}[tb]\centering
 \caption{Example of a sentence with different sampling tresholds}
@@
         \toprule
-Sampling Treshhold & 0 \\
-
+Sampling Treshold & First sentence of Dataset \\ 
         \midrule%
-Sampling Treshold & First sentence of Dataset \\ \hline
-$10^{-1}$ &Anarchism originated as a term of abuse first\tabularnewline used against early working class radicals \tabularnewline including the diggers of the english \\ \hline
-$10^{-2}$ & Anarchism originated as a term of abuse first used against early \tabularnewline working class radicals including diggers of english \\ \hline
-$10^{-3}$ & Anarchism originated a term abuse first used against early \tabularnewline working class radicals including diggers the english\\ \hline
-$10^{-4}$ & Anarchism originated abuse used against working class radicals\tabularnewline diggers english \\ \hline
+$10^{-1}$ & \makecell[l]{Anarchism originated as a term of abuse first used against early working class radicals including the diggers\\ of the english} \\ \hline
+$10^{-2}$  & \makecell[l]{ Anarchism originated as a term of abuse first used against early \\ working class radicals including diggers of english} \\ \hline
+$10^{-3}$ & \makecell[l]{Anarchism originated a term abuse first used against early \\ working class radicals including diggers the english}\\ \hline
+$10^{-4}$ &\makecell[l]{ Anarchism originated abuse used against working class radicals\\ diggers english} \\ \hline
 $10^{-5}$ & against radicals diggers \\ \hline        
@@
 \label{table:treshold_examples}
-\end{table}
+\end{table*}
 
@@
 
-\begin{table}[]
-\centering
-\begin{tabular}{|l|l|l|}
-\hline
+\begin{table}[tb]\centering
+\caption{Example of pairs and their rating in wordsim353}
+\begin{tabular}{l l l }
+        \toprule
 Word1 & Word2 & Score \\ \hline
-\textquote{FBI} & \textquote{Investigation} & \textquote{8.31} \\ \hline
-\textquote{Mars} & \textquote{scientist} & \textquote{5.63} \\ \hline
+        \midrule%
+
+\textquote{FBI} & \textquote{Investigation} & 8.31 \\ \hline
+\textquote{Mars} & \textquote{scientist} & 5.63 \\ \hline
+        \midrule%
 \end{tabular}
-\caption{Example of pairs and their rating in wordsim353}
 \label{fig:ws353_ex}

### 2019-05-01 icdm/submission/includes/discussion.tex
added discussion without subsection challenges

@@
 \centering
-\includegraphics[scale=0.45]{images/gensim_vs_adam}
+\includegraphics[scale=0.3]{images/gensim_vs_adam}
 \caption{Convergence time of SGD and Adam compared to Gensim}
 \label{fig:gensim_vs_adam}
-\end{figure}\subsection{Challenges faced}
-During our implementation we did encounter some problems, this subsection has the purpose of preventing the scientific community from making the same mistakes.
-\subsubsection{Using the wrong embeddings}
-To start with, we did not choose the same initialization value as Gensim in our word embeddings.
-We initialized the projection layer with a normal distribution between (-1,1), as opposed to Gensim which initializes all the weights to 0.
-The reason for this, is that at the beginning of our experiments we did not achieve good results with the same initialization as Gensim.
-Retrospectively we accord this to a learning rate that was too low.
-After a few simulations, we saw that we did not perform as good as Gensim.
-And as we changed the initialization of the projection layer back to 0, and already had adjusted the learning rate, we achieved good results.
-Therefore, this is a recommendation to future work to not set the initialization to (-1,1).
-
-\subsubsection{Batch size and loss function adjustements}\label{ssec:bs_lf}
-During our experiments, we faced a moment where we needed to us a very high learning rate (50) to achieve good results.
-As this is the complete opposite of what is standard we made a few changes to remedy the problem.\\
-First, we needed to find a good batch size. During the above-explained raise of the learning rate, we used a batch size of 5000. We then decided to take a batch size of 2000.\\
-Secondly, at the same time, we did experiment with different loss functions.
-As our batched model does not have exactly the same loss function as introduced by Mikolov et al.
-\cite{mikolov}, we needed to find one that suited our goals the best.
-At the beginning of our experiment, we took the average of all the scores stored in our final matrix, as explained in \ref{ssec:b_SGM}.
-But then chose to take the sum as the training did not seem optimal.\\
-With these two changes, we did increase convergence time and word similarity, while at the same time having a usual learning rate. \\
-Know the question arises why this happened?
-Is this because of the shape of the loss function better suits our advanced optimizers?
-Or because the loss function is closer to its original form?
-Our hypothesis looks the following way:
-When taking the average, our loss function is represented as follows:
-\begin{equation}
-\frac{-1}{b }* \sum_{x\in X} loss(x)
-\end{equation}
-where $X$ is our batch and $b = |X|$.
-When taking the partial derivative of each parameter, the constant $ \frac{-1}{b }$ will remain the same.
-Therefore, it's the same as taking the sum as the loss function and multiplying each gradient with $\frac{-1}{b }$.
-As our batch size was really high, i.e 2000 and 5000, this could highly influence the learning rate.
-This is also the conclusion of Goyal et al.
-\cite{fb}, that worked with very high batch sizes, i.e 8000: "When the minibatch size is multiplied by k, multiply the learning rate by k".
+\end{figure}
 

### 2019-05-01 icdm/submission/includes/related_work.tex
added example table and rescaled images
deleted context sensitive word embedings

@@
 This subsection will first give an overview of the optimization of the throughput and then present one paper that focused on context-sensitive word embeddings.
-
-\subsection{Optimization of the throughput}
 
@@
 
-\paragraph{Parallization in shared and Distributed Memory}
+\subsection{Parallization in shared and Distributed Memory}
 The first parallelized solution which was proposed by Ji et al.
@@
 
-\paragraph{Accelleration of word2vec by Using GPU's}
+\subsection{Accelleration of word2vec by Using GPU's}
 Seulki and Youngmin \cite{gpu} focused on getting a better throughput on the SGM when using GPU's.
@@
 %%%%%%%%%%%
-\paragraph{Gensim}\label{ssec:gensim}
+\subsection{Gensim}\label{ssec:gensim}
 Gensim \cite{gensim} is a pure Python library that holds a state of the art implementation of the SGM.
@@
 
-\subsubsection{Context sensitive word embedding}
-A word has different meanings dependent on its context.
-This is a problem that is not addressed by the SGM.
-Some new models, that have taken this issue into consideration, were proposed.
-A lot of work has been done in this direction, Liu et al.\cite{topicalWE} and Bartunov et al.\cite{breaking} for example, but the one reporting the best results is Liu et al.
-\cite{contextWithTensor}.
-The concept of this approach is to change the way we compute the objective function and variables we use in our conditional probability.
-The idea is to look if a word given a certain context word matches to a topic.
-\textit{Bank} would match to \textit{finance} given the context word \textit{money}.
-\textit{Bank} would also match to \textit{nature} if \textit{river} was the given context word.
-But \textit{Bank} would not match to \textit{nature} with the context word \textit{money}.
-Now one could ask how to achieve such a context sensitive word embedding?
-First, we have to introduce new variables, therefore let's look at the objective function used:
-\begin{equation}
-J(\Omega) = \sum_{(w,t,c)\in D} \sum_{(w,\tilde{t},\tilde{c} \in{\tilde{D}})} max(0,1- g(w,t,c) + g(w,\tilde{t},\tilde{c})) \lambda||\Omega||_{2}^2
-\end{equation}
-
-This approach uses the same negative sample technique as described in the previous subsections, $D$ is the corpus data and $\tilde{D}$ is the set of negative samples and $\lambda$is the hyper parameter used for the standard $L_2$ standardization.
-What is interesting here is the function $g(w,c,t)$, where $w$ is a word, $c$ the context word, and $t$ the context in which the word appears.
-$g$ is defined as follows:
-\begin{equation}
-g(w,c,t) = u^T \sigma(w^TM^{[1:k]}t+V_c^T(w \oplus t) + b_c)
-\end{equation}
-where, $u, V_c, b_c$ are standard parameters for a neural network.
-$\oplus$ is the vector concatenation.
-The most important parameter is $M^{[1:k]}$, which is a tensor layer, the tensor layer is used because of its ability to model multiple interactions in the data, as this will be useful for multiple contexts.
-They used SGD for the optimization of this objective function.
-They achieved interesting results as shown in \ref{fig:multipleContext}.\\
-\begin{figure}[ht]
-\centering
-\includegraphics[scale=0.7]{images/multipleContext.png}
-\caption{"Nearest neighbor words model and Skip-
-Gram. The first line in each block is the results of Skip-Gram;
-and the rest lines are the results of our model" \cite{contextWithTensor}}
-\label{fig:multipleContext}
-\end{figure}
-
 This will conclude our overview of the related work. We will now give the reader an outline of the different Gradient Descent Optimizer used in our experiments.

### 2019-05-01 icdm/submission/includes/background.tex
new file for evalutation and changes to tables
layers of sgm better explained

@@
 \section{Background}\label{chap:background}
-
+\iffalse
 \begin{table*}[tb]\centering
@@
 \end{table*}
+
+\fi
 This subsection gives an overview of the SGM and related work on its optimization. Furthermore, the used optimizers in our experiments are explained.
@@
 The SGM is a very simple model used to learn WE.
-The idea is to train the model, a neural network, on a fake task and then use some of the weights of the neural network as embeddings.
+The idea is to train a neural network, on a fake task and then use some of the weights of the network as embeddings.
 To understand this fake task center and context words must be defined.
@@
 See Figure \ref{fig:window_ex} as an example, where the context words are highlighted.
-An important thing to notice, is that instead of having a fixed window size $m$, each word will randomly have a window size between $1$ and $m$.
+An important aspect to notice, is that instead of having a fixed window size $m$, each word will randomly have a window size between $1$ and $m$.
 The idea behind having different window sizes is that words that are further away in the sentence, have less semantic correlation to the center word.
@@
 %TODO think about example matrix
-With these definitions the fake task can be defined: given a pair of center and context words, which will be our training samples, the goal of the network is to predict the probability of each word to appear in the context of the center word.
-One may ask how such a network is built.
-The network consists of two matrices, a projection and an output layer.
-Both of these matrices store word embeddings.
-The first one, the projection layer, will store the WE for one specific word in one specific row.
-It will have dimension $T\times d$, with $T$ being the size of our vocabulary and $d$ the dimension of our WE.
-The second layer, the output layer, will store one WE in each column.
-The idea behind these 2 layers is that the input layer will represent our words as center word and the output layer as context words.
+With these definitions the fake task can be defined: given a pair consisting of a center and a context words, which will be our training samples, the goal of the network is to predict the probability of each word to appear in the context of the center word.
+To achieve this goal, Mikolov et al.
+introduced the following model.
+The network consists of three layers.
+See fig x for an illustration.
+First the ntw has an input layer, a vector of size $V$ ( $V$ being the size of the vocabulary), represint words as an one hot encoding.
+Secondly, the projection layer wight Matrix $M_{in}$ soters $V$ $D$-dimensional vectors for the corresponding word in the vocabulary  in it's rows.
+The projection layer serves as a look up tavle for our WE (if a one hot vector is multiplied with $M_{in}$ the corresponding WE will result).
+Finally, the output layer wight matrix $M_{out}$ also stores $V$ $D$-dimensional vector representation of words.The idea behind these 2 Matrices is that the input layer will represent our words as center word and the output layer as context words.
 Since the task of the network being to predict the probability of each word appearing in the context of a given center word, the following probability should be maximized:\\
@@
 
-Here $\tilde{v}_{w_t}$ and $ v_{w_t}$ are the vector representations, stored respectively in the projection Layer (first Matrix) and in the output layer (second Matrix).
+Here $\tilde{v}_{w_t}$ and $ v_{w_t}$ are the vector representations, stored respectively in the projection Layer ($M_{in}$) and in the output layer ($M_{out}$).
 There is a problem with the classical softmax.

### 2019-05-01 icdm/submission/includes/background.tex
paramater changes in equation

@@
 \begin{equation} \label{eq:basicSkip}
-\prod_{t=1}^T \prod_{-m<j<m} p(w_{t+j}|w_t)
+\prod_{i=1}^V \prod_{-m<j<m} p(w_{i+j}|w_i)
 \end{equation}
-Where T is the number of words in the corpus data, $w_t$ the $t^{th}$ word in the corpus data and $m$ is the context window.
+Where $V$ is the number of words in the corpus data, $w_i$ the $i^{th}$ word in the corpus data and $m$ is the context window.
 This means that the $m$ nearest words to $w$ are considered as context words.
@@
 \begin{equation}
-\sum _{t=1}^T \sum_{-m<j<m} log( p(w_{t+j}|w_t) )
+\sum _{t=1}^V \sum_{-m<j<m} log( p(w_{i+j}|w_i) )
 \end{equation}
@@
 
-The basic Skip-Gram Model uses a classical Softmax to calculate the conditional probability $p(w_{t+j}|w_t)$:
+The basic Skip-Gram Model uses a classical Softmax to calculate the conditional probability $p(w_{i+j}|w_i)$:
 \begin{equation} \label{eq:softmax}
-p(w_{t+j}|w_t)= \frac{exp( \tilde{v}_{w_{t+j}}^Tv_{w_t})}{\sum_{w=1}^v exp(\tilde{v}_w^Tv_{ w_t})}
+p(w_{i+j}|w_i)= \frac{exp( \tilde{v}_{w_{i+j}}^Tv_{w_i})}{\sum_{w=1}^v exp(\tilde{v}_w^Tv_{ w_i)}}
 \end{equation}
 
-Here $\tilde{v}_{w_t}$ and $ v_{w_t}$ are the vector representations, stored respectively in the projection Layer ($M_{in}$) and in the output layer ($M_{out}$).
+Here $\tilde{v}_{w_i}$ and $ v_{w_i}$ are the vector representations, stored respectively in the projection Layer ($M_{in}$) and in the output layer ($M_{out}$).
 There is a problem with the classical softmax.
 As a matter of fact, it is unsuitable to compute the softmax.
-For the computation of $\sum_{w=1}^v exp(\tilde{v_w}^T w_t)$, the denominator in Equation \ref{eq:softmax}, one has to go over the whole corpus data.
+For the computation of $\sum_{w=1}^V exp(\tilde{v_w}^T w_t)$, the denominator in Equation \ref{eq:softmax}, one has to go over the whole corpus data.
 As very big data sets are needed to train the model, this is not a solution.

### 2019-05-01 icdm/submission/includes/background.tex
added fig config of ntw plus finished subsecction sgm

@@
 The network consists of three layers.
-See fig x for an illustration.
+See Figure \ref{fig:ntw_architecture} for an illustration.
 First the ntw has an input layer, a vector of size $V$ ( $V$ being the size of the vocabulary), represint words as an one hot encoding.
@@
 Finally, the output layer wight matrix $M_{out}$ also stores $V$ $D$-dimensional vector representation of words.The idea behind these 2 Matrices is that the input layer will represent our words as center word and the output layer as context words.
+
+\begin{figure}[ht]
+\centering
+\includegraphics[scale=0.625]{images/ntw_architecture}
+\caption{Structure of the network of the Skip Gram Model}
+\label{fig:ntw_architecture}
+\end{figure}
 Since the task of the network being to predict the probability of each word appearing in the context of a given center word, the following probability should be maximized:\\
@@
 There is a problem with the classical softmax.
-As a matter of fact, it is unsuitable to compute the softmax.
 For the computation of $\sum_{w=1}^V exp(\tilde{v_w}^T w_t)$, the denominator in Equation \ref{eq:softmax}, one has to go over the whole corpus data.
-As very big data sets are needed to train the model, this is not a solution.
+As very big data sets are needed to train the model, this is approach is unsuitable.
 Mikolov et al.

### 2019-05-02 icdm/submission/includes/background.tex
negative sampling reworked

@@
 \fi
-This subsection gives an overview of the SGM and related work on its optimization. Furthermore, the used optimizers in our experiments are explained.
+This section gives in its first subsection a broad overview over the functioning of the Skip-Gram Model, and its second section describes the Technique called Negative Sampling introduced by Mikolov et al.
+\cite{mikolov2} used to make the Skip-Gram model practicable.
 \subsection{The Skip-Gram Model}
@@
 \end{figure}
-%TODO think about example matrix
 With these definitions the fake task can be defined: given a pair consisting of a center and a context words, which will be our training samples, the goal of the network is to predict the probability of each word to appear in the context of the center word.
@@
 \cite{mikolov2} simplified NCE and called it Negative Sampling, which will be explained next.\\
-The main goal of negative sampling is to only update the output nodes of certain words.
+The main goal of negative sampling is to only update the output nodes of certain words, 5-20 in practice.
 This will obviously save an enormous amount of computation time.
-The idea is that given a pair $(c,w) \in D$, where $c$ is a word in the context window of $w$ we select $K$ random words $k_i$ from the corpus data $D$.
+The idea is that given a pair $(c,w) \in D$, wh$w$ ere $c$ is a word in the context window of we select $K$ random words $k_i$ from the corpus data $D$.
 We assume those words do not appear in the context of $w$.
-We denote the score that the $(c,w)$ wasn't drawn at random the following way: $p(y=1|c,w)$, and if $(k,w) $ is chosen at random this way: $p(y=0|k,w)$.
-Now we use logistic regression to update the weights of the $k$ selected context words and $c$.
 By doing so we will only have to update $k+1$ nodes.
+Mikolov et al.
+\cite{mikolov2} introduced the following objective function $J$, that should be maximized:
 
-Let's look at how we construct our objective function for a given word $w$ and one of its context words $c$:
-
-\begin{align*}
-p(c|w) &= p(y=1|c,w) + \prod_{k\in K} p(y=0|k,c)
-\\&= p(y=1|c,w) + \prod_{k\in K} 1- p(y=1|k,c)
-\\&= log((p(y=1|c,w)) + \sum_{k\in K} log(1- p(y=1|k,c))
-\\&= log(\frac{1}{1+e^{-v_c \tilde{v_w }}}) + \sum_{k\in K} log(1-\frac{1}{1+e^{-v_c \tilde{v_k}}})
-\\&= log(\frac{1}{1+e^{-v_c \tilde{v_w } }}) + \sum_{k\in K} log(\frac{1}{1+e^{v_c \tilde{v_k} }})
-\\&= log(\sigma(v_c \tilde{v_w } ) + \sum_{k\in K} \sigma(log(-v_c \tilde{v_k} ))
-\end{align*}\label{eq:obj_neg_samples}
+\begin{equation}
+J(c,w)= log(\sigma(v_c \tilde{v_w } ) + \sum_{k\in K} \sigma(log(-v_c \tilde{v_k} ))
+\end{equation}\label{eq:obj_neg_samples}
 
 Where $v_c$ and $\tilde{v_w }$, can be interpreted as before in Equation \ref{eq:softmax} and $\sigma(x) = \frac{1}{1+e^{-x}}$.
-The goal is to maximize this objective function.
-Another way to look at this function, apart from logistic regression, and understand why it's working so well, is to assume that two vectors are similar if their dot product is high.
-Therefore, we maximize the dot product of similar words $(w,c)$ ($c$ appears in the context of $w$) and minimize it for dissimilar words $(w,k_i)$ (those were selected randomly).
+To  understand why this loss function is working so well, is to assume that two vectors are similar if their dot product is high.
+Therefore, we maximize the dot product of similar words $(w,c)$ ($c$ appears in the context of $w$) and minimize it for dissimilar words $(w,k_i)$ (those were selected randomly, we therefore assume that their are not similar).
 We see that to compute our objective function we will only have to compute the sum over the number of negative samples $K$, which is very small in practice (2-20).
 To put things in perspective lets imagine our data set consists of 100000 words, we set $K=2$.
-Assume each output neuron has a weight vector $v$ with $|v| = 300$.
+Assume each output neuron has a weight vector of dimension $d = 300$.
 In consequence, when updating our weights we would only update $0.2*10^{-2}$ of the 300 million weights in the output layer.
 
-One question remains: how do we choose our random words? Mikolov et al. \cite{mikolov2} used the following unigram distribution:
+One question remains: how are the negative samples selected? Mikolov et al. \cite{mikolov2} used the following unigram distribution:
 
@@
 The value of $\frac{3}{4}$ is set empirically.
-Raising the unigram distribution to power of $\frac{3}{4}$ makes it less likely for a word to be drawn if it often appears in the dataset in comparison to the basic unigram distribution.
+Raising the unigram distribution to the power of $\frac{3}{4}$ makes it less likely for a word to be drawn if it often appears in the dataset in comparison to the basic unigram distribution.
 See figure \ref{fig:frequency_ex} for an example.

### 2019-05-02 icdm/submission/includes/related_work.tex
parallelization by the use of caching updated

@@
 \section{Related Work}
-Due to the popularity of the SGM, a lot of research went into optimizing it.
-This research can be divided into two categories, optimization of the throughput and the optimization of the algorithm's accuracy.
-The latter was achieved by allowing words to have multiple meanings, also called context-sensitive word embedings.
-For our work, the optimization of the throughput is of big interest while the semantic optimization is aimed at giving the reader a more holistic comprehension of the possible research directions.
-This subsection will first give an overview of the optimization of the throughput and then present one paper that focused on context-sensitive word embeddings.
-
 In the original model, the optimization is done with Stochastic Gradient Descent (SGD), which is a sequential algorithm.
 This process does not favor parallelization.
-To deal with this specific problem Mikolov et al.\cite{mikolov2} used a Hogwild tree proposed by Recht et al.\cite{hogwild}.
-The approach is to allow multiple threads to access a shared memory, in this case, the single model.
+To deal with this specific problem Mikolov et al.\cite{mikolov2} used a Hogwild!
+approach proposed by Recht et al.\cite{hogwild}.
+The idea is to allow multiple threads to access a shared memory, in this case, the single model.
 In the original SGM, the threads are constructed as follows: at the beginning of training, the dataset is split into $N$ evenly sized chunks and each of these chunks will be processed by a one thread.
@@
 But in the case of NLP, the problem seems to be a bit more significant, and especially for word embedding, as many words share the same context words.
-There were several attempts at solving this issue, and we are going to cover a few of them in the following subsubsections.
+There were several attempts at solving this issue, and optimize the throughput of the SGM.
 
-\paragraph{Parallelization by the use of caching}
-This idea was proposed by Vuurens et al.
-\cite{efficient}.
-The architecture used here is the basic skip gram model with a hierarchical softmax.
-The general idea is to cache the most frequently used nodes of the binary tree used to memorize the probability distribution and update them on the shared single model after a certain number of seen words (the paper used the number 10).
+\subsection{Parellelization by the use of caching}
+This idea was proposed by Vuurens et al. \cite{efficient}. The architecture used here is the SGM with a hierarchical softmax.
+
+The general idea is to cache the most frequently used weight vectors, and update the model after a certain number of seen training samples, while taking in consideration the changes that happend during these number of seen training samples (the paper used the number 10).
+The unfrequent weight vectors are handled as in the original paper, i.e updated in shared memory after each training sample.
+This technique  allows it to avoid collosision for the most frequent words, and the author therefore hoped to be able to increase the number of concurrent threads.The general idea is to cache the most frequently used nodes of the binary tree used to memorize the probability distribution and update them on the shared single model after a certain number of seen words.
 The paper produced interesting results as they managed to decrease execution time by increasing the number of cores used for the calculation.

### 2019-05-02 icdm/submission/includes/related_work.tex
intel update

@@
 \subsection{Parallization in shared and Distributed Memory}
-The first parallelized solution which was proposed by Ji et al.
-\cite{intel}, is to try to reduce the cost of our vector multiplication.
-The main idea in this paper is to convert the level 1-BLAS (Basic linear subprogram) vector to vector operations to a level 3-BLAS matrix multiplication operation, which are efficiently implemented into hardware and consequently faster.
+In contrast to the above approach, this one, introduced by Ji et al.
+\cite{intel} focused on the Skip-Gram Model with negative sampling.
+The focus of this work wasto reduce the cost of our vector to  vector multiplication in our loss funcion, see Equation \ref{eq:obj_neg_samples}, where the dot product of the word vector and the negative sampled as well as the context word is computed.
+The main idea in this paper is to convert the level 1-BLAS (Basic Linear Subprogram) vector to vector operations to a level 3-BLAS matrix multiplication operation, which are efficiently implemented into hardware and consequently faster.
 This is achieved, by using the same negative samples for each context word of a given word $w$.

### 2019-05-02 icdm/submission/includes/related_work.tex
update gpu acceleration

@@
 Consequently, it's very hard to parallelize at this level.
-To solve this problem, Seulki and Youngmin \cite{gpu} proposed the idea to parallelize the update of each dimension of the word embedding, since those are completely independent of each other.
+Since the dimensions of the word embeddings are completly independent of each other  Seulki and Youngmin \cite{gpu} tried to parallelize their update.
 They achieved this by mapping each dimension to a CUDA thread while mapping each sentence to a CUDA block.
 As each CUDA block runs independently, the training of the sentences is parallelized, and the fact that sentences have different length is of no problem.
-If the execution time of the GPU kernel is greater than time used to read the sentences, it could be a smart choice to use multiple GPU's.
+If the execution time of the GPU kernel is greater than the time used to read the sentences, it could be a smart choice to use multiple GPU's.
 According to Seulki and Youngmin \cite{gpu}, if multiple GPU's are used, there is a need for synchronizing the model, which will hinder run time performance.
@@
 The accomplished results were very good as they managed a 20x speedup compared to a single threaded CPU execution, which is a 3x increase in comparison to the original C code from Mikolov et al.
-\cite{mikolov2}, with no loss in accuracy.
-The problem with this and all the above optimization is that the code is not easily available.
+\cite{mikolov2}.
+Al this speedup was achieved, while maintaining the accuracy of the model.
+The problem with this and all the above optimization is that the code is not easily available or of use.
 Therefore, we need an optimized implementation of the SGM that is easily available.

### 2019-05-06 icdm/submission/includes/contribution.tex
intro contribution

@@
 This section will give an overview of our implementation.
-First, it will give a broad overview of the implementation idea and process.
-Then it will go into detail, explaining the forward process of our model and the construction of our dataset, by the use of the \texttt{Dataset} class.
 The proposed implementation is a slighlty altered version of the original SGNS.
-The idea behind this altering, is to compute the loss for multiple words and context pairs at the same time.
+The idea behind this altering, is to compute the loss for multiple words and context pairs at the same time, this is inspired from mini-batch training, a model often used in other machine learning problem.
+Instead of computing the loss function of each pair, we computed the sum over a certain number of pairs (i.e 2000) as a loss.
 The exact process will be described in the following paragraph.

### 2019-05-06 icdm/submission/paper.tex
added example table and rescaled images
new file for evalutation and changes to tables
table and figure form
added discussion without subsection challenges
intro contribution

@@
 \usepackage{csquotes}
+\usepackage{booktabs} % For formal tables`
+\usepackage{makecell} % For line breaks in tables
+
 \begin{document}
@@
 \begin{abstract}
-This document is a model and instructions for \LaTeX.
-This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
-or Math in Paper Title or Abstract.
+%This document is a model and instructions for \LaTeX.
+%This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
+%or Math in Paper Title or Abstract.
+The Skip-gram Model with negative sampling (SGNS) is an effective algorithm to create word embeddings.
+SGNS uses Stochastic Gradient Descent (SGD) as it's learning algorithm.
+While a lot of effort has gone into increasing the throughput of words of the SGNS, not much work has gone into optimizing the convergence time.
+Therefore, our work focuses on the latter.
+In this work we used two techniques to achieve a better convergence time, namely advanced optimizers and input shuffling.
+We used the Text8 dataset to train our model and measured the quality of our word embeddings with the wordsim353 dataset, which measures the quality of word embeddings by judging the similarity of different words.
+We trained our model with multiple advanced optimizers: momentum, Nesterov accelerated gradient, Adagrad, Adadelta, and Adam.
+We also applied input shuffling to the above optimizers.
+Adam combined with input shuffling outperformed every optimizer.
+Adam with shuffling also outperformed the current state of the art implementation Gensim.
+Adam converged to a similarity value of 0.66 (state of the art) in 2 epochs, while Gensim took 4 epochs.
+Hence, this work shows that advanced optimizers combined with input shuffling do decrease the convergence time of the SGNS.
+These results must be confirmed with other datasets, but optimizing our model in terms of throughput has the potential of further reducing the overall run time.
+
 \end{abstract}
@@
 \input{includes/contribution}
+\input{includes/evaluation}
 \input{includes/results}
+\input{includes/discussion}
 \input{includes/conclusion}

### 2019-05-06 icdm/submission/includes/contribution.tex
contribution done

@@
 The exact process will be described in the following paragraph.
+
 \subsection{Forwarding}
 As this represents the challenging part of our Implementation the forwarding method is explained step by step. Each time step is illustrated to make the explanation clearer.\\
-Let $X = {(v_1,c_1),(v_2,c_2),(v_3,c_3)}$, be the training batch, where $(v_i,c_i)$ is a training sample constituted by a word and on of its context words. \\
+Let $X = {(v_1,c_1),(v_2,c_2),(v_3,c_3)}$, be the training batch, where $(v_i,c_i)$ is a training sample constituted by a word and one of its context words. \\
 \textit{Input:}\\
@@
 For this task some definitions are necessary: let $A_j$ be the $j^th$ row of the matrix A, let $E_c(i,j)$ be the $d$ dimensional embedding of the word stored in $\tilde{A}(i,j)$.
-The dot product is computed with a so-called batch multiplication\footnote{Documentation: https://pytorch.org/docs/stable/torch.html\#torch.bmm} which will result in a matrix $S$ where $S(i,j) = E_c(i,j) \cdot A_j$.
+The dot product is computed with a so-called batch multiplication\footnote{Documentation: https://pytorch.org/docs/stable/torch.html\#torch.bmm} which will result in a matrix $S$ where $S(i,j) = E_c(i,j) \cdot {E_v}_j$.
 This will result in a $n\times k+1$ Matrix $S$.

### 2019-05-07 icdm/submission/includes/contribution.tex
conclusion contribution

@@
 
-This algorithm provides a way to compute the loss, to make the implementation successful, one still needs to access the context-pairs. The process of doing so is described in the next paragraph.
+This finishes the description of our batched approach. The next section will describe our experimental setting and the possibilities that exists to evaluate the quality of word embeddings.

### 2019-05-07 icdm/submission/includes/evaluation.tex
conclusion contribution
sampling plus min count

@@
-\section{Evalutation}\label{chap:results}
+\section{Evaluation}\label{chap:results}
 
@@
 
-This subsection gives an overview of the used datasets, the used metric to evaluate our models, the configuration of our model and finally, the experimental results achieved.
+This section gives an overview of the used datasets, the used metric to evaluate our models, the configuration of our model, and finally an explanation about inputa shuffling
 
@@
 We chose this dataset for two reasons.
-First of all, it's a very small dataset, that allowed us to do a lot of computations.
+First of all, it's a small dataset, that allowed us to do a lot of computations.
 Secondly, this dataset was used in related work \cite{intel} hence giving us a very good benchmark.
@@
 Therefore, the goal of subsampling is to delete such words from the dataset.
-This will decrease the computation time, as it will reduce the number of training samples, and should, in theory, increase the accuracy of the model.
+Subsampling will decrease the computation time, as it will reduce the number of training samples, and should, in theory, increase the accuracy of the model.
 The increase in accuracy can also be explained by the fact that words that would not have appeared in the context of each other, may now do because words between have been deleted.
-To choose which word to delete, Mikolov et al. \cite{mikolov2} chose the following equation to compute the deletion of a word $w$ in the data set:
+To choose which word to delete, Mikolov et al. \cite{mikolov2} chose the following equation to compute the probability of a word $w$  to be deleted from the data set:
 \begin{equation} \label{eq:sampling}
@@
 where $f(w)$ is the frequency of w, and $t$ is a threshold set empirically.
-As Equation \ref{eq:sampling} is a probability, subsampling is not a deterministic procedure, words that may have been deleted with a threshold of $10^-2$ may stay in the dataset with a lower threshold, as shown in Table \ref{table:treshold_examples}.
 Mikolov et al.
@@
 We did this by looking at a random set of sentences and judging the results manually.
+As Equation \ref{eq:sampling} is a probability, subsampling is not a deterministic procedure, words that may have been deleted with a threshold of $10^-2$ may stay in the dataset with a lower threshold.
 An example of the first sentence with different sampling thresholds can be found in Table \ref{table:treshold_examples}.
@@
 Sampling Treshhold & 0 & $ 10^{-1}$&$ 10^{-2}$& $10^{-3} $ &$10^{-4} $ \\ 
-
         \midrule%
 Words in Dataset  & 16mio & 15mio & 11mio & 8mio & 4mio \\ 
-        \midrule%
+        \bottomrule%
    \end{tabular}%
@@
 We also deleted every word that did not appear more than 5 times in our dataset.
-We got this technique from Gensim \cite{gensim}, that introduced this parameter into their training.
+This idea was also introduced by Mikolove et al.
+\cite{mikolov2}.
 This is a good technique because of three reasons: First certain words of our data sets do not appear in a common lexicon (twigleg, melqu), or come from a foreign language (Volksvereinigung), or are names and acronyms.

### 2019-05-07 icdm/submission/includes/evaluation.tex
input shuffling

@@
 \subsection{Evaluating word embedings}
-Evaluating word embedding is not an easy task, such as evaluating the accuracy of a common classifier.
+Evaluating word embedding is not an easy task, such as evaluating the accuracy of a common classifier for example.
 We cannot split our data set into train and test set.
 As the task that the network is learning is of no interest to us.
-Therefore, we need to verify that our embedding is of quality with other techniques.
-To define quality we first need to define a measure of similarity between two vectors.
+Therefore, to ensure the quality of word embeddings other techniques have to be used.To define quality we first need to define a measure of similarity between two vectors.
 This requires knowledge of the Cosine distance, which is introduced in the following subsection.
@@
 \subsubsection{Word similarity and wordsim353}
-To measure the quality of our word embedding we will need a dataset to compare our results too.
+To measure the quality of our word embedding we  use a dataset to compare our results too.
 We chose wordsim353\footnote{http://www.cs.technion.ac.il/~gabr/resources/data/wordsim353/wordsim353.zip} for this task, as it's the most used in the related literature.
@@
 The similarity score is in the range of 1 and 10, with 10 being the best score.
-An example for two of such pairs can be found in Figure \ref{fig:ws353_ex}.
-We will rank our embeddings on the Pearson correlation coefficient between the cosine distance and the scores attributed by humans, as this is a common procedure.
+An example for two of such pairs can be found in Table \ref{fig:ws353_ex}.
+We will rank our embeddings on the Pearson correlation coefficient between the cosine distance and the scores attributed by humans, as this is the common procedure.
 
@@
 \begin{itemize}
-\item \texttt{Negative Samples:} Here we have to find a trade-off between, setting the parameter too high which will result in increased accuracy but a longer computation time.
+\item \texttt{Negative Samples:} Here we have to find a trade-off between, a parameter that is too high (better accuracy) and low ( faster computation time).
 For smaller data sets a higher number of negative samples is often needed.
@@
 - \textit{100}
-\item \texttt{Batch size}: As described in subsection \ref{ssec:b_SGM}, there is a tradeoff to find between quality and training time.
-We first used a batch size of 5000, but then decide after non conclusive results (see \ref{ssec:bs_lf}) that 2000 would be better - \textit{2000}
+\item \texttt{Batch size}: As described in Section \ref{ssec:b_SGM}, there is a trade off to find between quality and training time.
+We first used a batch size of 5000, but then decide after non conclusive results that 2000 would be better - \textit{2000}
 \item \texttt{Alpha}: learning rate, this hyperparameter was tuned in every optimizer therefore only the range will be indicated - \textit{(1e-5,1)}
 \end{itemize}
+
+This conlcudes our overview of the possible parameters of our Skip-Gram Model, in the next section the reader will be given a description about the process of input shuffling.
 
@@
 \subsection{Input Shuffling}
-We used input shuffling as a technique to optimize the skip gram model.
-We will first describe input shuffling in a general way and then explain why we suppose that input shuffling could work well on the skip gram model.
+We used input shuffling as a technique to optimize the SGM.
+We will first describe input shuffling in a general way and then explain why we supposed that input shuffling could work well on the SGM.
 \\
-Let $X = {x_1...x_n}$ be our input data set. Input Shuffling describes the process of taking a random permutation of the dataset as input at each epoch.
-The idea behind this technique is to present our optimizer with different loss surfaces so that it's able to find the best optimum.
+Let $X = {x_1...x_n}$ be our input data set. Input Shuffling describes the process of taking a random permutation of the dataset as an input at each epoch.
+The idea behind this technique is to present our optimizer with different  surfaces of our loss function, in other to find the best optimum.
 Therefore, it's easier for the neural network to escape a local minimum.
-As for example if a network had converged to a local minimum after one epoch it could not escape it as all the parameters are the same.
+As for example if a network had converged to a local minimum after one epoch it probably cannot escape it.
 But if we change the shape of the loss function, by input shuffling, then there would be a greater probability for the network to escape the local minimum.
@@
 There are two reasons why we think that input shuffling is particularly well suited for the skip gram model.
-The first one has to do with the fact that when we read our words sequentially that words that only appear very early will not benefit from the context words being already updated from others.
+Firstly, we read our words sequentially, therefore,  words that only appear very early in the dataset will not benefit from the context words being already updated from others.
 The second idea is that we used the special batch technique described in Section \ref{ssec:b_SGM}.
-When using this technique and not using shuffling we will always have words that appear next to each other in a batch and will, therefore, update the same words at the same time.
-We, therefore, lose some quality.
-But if we would use input shuffling instead, then the words in one batch would likely not be similar and therefore only taking the average of a small part of pairs with the same words will be less likely.
+When we do not use input shuffling the same word will appear with all of its context words as training samples in a batch, therefore the updae will be an average of all of these training samples.
+But if we would use input shuffling instead, then the probability of a word appearing multiple time in one batch would be smaller.
+In consequence the accuracy of the model should be optimized leading to a better convergence time.
+A definition of convergence time is given in the next section.
 

### 2019-05-07 icdm/submission/includes/evaluation.tex
summary and transition evaluation

@@
 It is also written in Cython, therefore it has a fast computation time, but can be used inside a python implementation.
-Together with the knowledge from Ji et al.\cite{intel} that a score of $0.66$ in the task of word similarity, with the text8 dataset, is the state of the art, we tested Gensim (more on this process in subsection \ref{chap:discussion} and found out that it took 4 epochs to converge.
+Together with the knowledge from Ji et al.\cite{intel} that a score of $0.66$ in the task of word similarity, with the text8 dataset, is the state of the art, we tested Gensim (more on this process in subsection \ref{sec:discussion} and found out that it took 4 epochs to converge.
 Therefore, we defined the following criteria for convergence: \\
-$\rho - \rho_{prev} < 0.009 \vee \rho = 0.66$ \\
+\centerline{$\rho - \rho_{prev} < 0.009 \vee \rho = 0.66$} \\
 where $\rho$ is the Pearson coefficient on the wordsim353 task.
 We also stopped computation, if it took more then 20 epochs to converge.
+
+To summarize, we deleted words that appear very often (subsampling) and very infrequently (min count) from our dataset, i.e text8 dataset.
+To evaluate our word embeddings we defined the cosine distance and will compare our embeddings to a human judged dataset of pairs of words based on their similarity.
+The reader was also given an explanation about the configuration of our network and input shuffling.
+With all this knowledge we can know present the achieved experimental results in the next section.

### 2019-05-07 icdm/submission/includes/evaluation.tex
summary better
s

@@
 To summarize, we deleted words that appear very often (subsampling) and very infrequently (min count) from our dataset, i.e text8 dataset.
-To evaluate our word embeddings we defined the cosine distance and will compare our embeddings to a human judged dataset of pairs of words based on their similarity.
-The reader was also given an explanation about the configuration of our network and input shuffling.
+To evaluate our word embeddings we defined the cosine distance and will measur the quality our embeddings by comparing theim  to a human judged dataset of word pairs, based on their similarity.
+The reader was also given an explanation about the configuration of our network and introduced to the concept of input shuffling.
 With all this knowledge we can know present the achieved experimental results in the next section.

### 2019-05-07 icdm/submission/includes/introduction.tex
wip introduction

@@
-\section{Introduction}\label{chap:introduction}
+\section{Introduction}\label{sec:introduction}
 
 Representing words as vectors, i.e word embeddings (WE) is a fundamental aspect of Natural Language Processing (NLP).
-There are two ways to create such WE, either arbitrarily or with the purpose of capturing the semantics of the words, i.e.
+There are two ways to create such WE, either arbitrarily or with the purpose of capturing the semantic meaning of the words, i.e.
 vector representations of words that are syntactically or semantically similar will be near to each other.
 By capturing semantic or syntactic meaning WE have shown to facilitate a lot of subsequent NLP tasks, such as entity recognition, machine translation or sentence classification.
+%TODO CITE WORK
 The first attempt to create WE with neural networks was made by Bengio et al.
 \cite{bengio} but more recently Mikolov et al.
-\cite{mikolov} introduced a software package called w2vec that uses a simpler network and with this approach Mikolov et al.
-produced state of the art results.
+\cite{mikolov} introduced a software package called w2vec, which uses a simpler network produces state of the art results.
 One of the proposed algorithms in this software package is the Skip-Gram Model (SGM).
@@
 The weights of this network are then used as WE.
-
-The SGM gained a lot of attention, as it achieved very good results for a very simplistic model.
+Mikolov et al.
+\cite{mikolov2} then introduced an alteration to the SGM called negative sampling.
+The goal of the Skip-Gram Model with negative sampling (SGNS) is to distinguish for a given word $w$, context words $c_i$ (words that appear close to $w$ in a sentence) from words dram randomly (i.e negative samples) $k_i$.
+This is achieved by maximizing the dot product of $w$ and $c_i$ and minimizing the dot product of $w$ and $k_i$.
+The SGM, espacially SGNS, gained a lot of attention, as it achieved very good results for a very simplistic model.
 As a consequence, a lot of effort went into optimizing it.
@@
 Both of these approaches and most of the literature are focused on improving the throughput of the model, but not the convergence time.
-Therefore one could ask if the convergence time of the SGM can be optimized while at the same time maintaining its accuracy.
+Therefore one could ask if the convergence time of the SGM can be optimized while at the same time maintaining its accuracy.\\
  This work proposes an approach that uses advanced optimizers and input shuffling to optimize the convergence time of the SGM.
 In combination both of these techniques allowed us to decrease the convergence time of the SGM.
-If combined with an optimized throughput these result could lead to an overall decrease in runtime.
-
-This work is structured as follows: subsection \ref{chap:background}  describes the SGM and its optimizations.
+If combined with an optimized throughput these result could lead to an overall decrease in runtime.\\
+ The rest of the paper  is structured as follows: section \ref{sec:background}  describes the SGM and its optimizations.
 Furthermore, we will give an extended explanation of the gradient descent optimizers used in this work.
-The reader is introduced to our Implementation in subsection \ref{chap:implementation}.
-Results are presented in  subsection \ref{chap:results} , where we will also describe used datasets, the measure applied to compare the quality of word embeddings, and finally the empirical results.
-The last part will focus on the discussion of our results, and possible future work in subsection \ref{chap:discussion} followed by a conclusion in subsection \ref{chap:conclusion}.
+The reader is introduced to our Implementation in subsection \ref{sec:implementation}.
+Results are presented in  subsection \ref{sec:results} , where we will also describe used datasets, the measure applied to compare the quality of word embeddings, and finally the empirical results.
+The last part will focus on the discussion of our results, and possible future work in Section \ref{sec:discussion} followed by a conclusion in Section \ref{sec:conclusion}.

### 2019-05-07 icdm/submission/includes/evaluation.tex
referencing sections

@@
-\section{Evaluation}\label{chap:results}
+\section{Evaluation}\label{sec:evaluation}
 
@@
 - \textit{100}
-\item \texttt{Batch size}: As described in Section \ref{ssec:b_SGM}, there is a trade off to find between quality and training time.
+\item \texttt{Batch size}: As described in Section \ref{sec:contribution}, there is a trade off to find between quality and training time.
 We first used a batch size of 5000, but then decide after non conclusive results that 2000 would be better - \textit{2000}
@@
 Firstly, we read our words sequentially, therefore,  words that only appear very early in the dataset will not benefit from the context words being already updated from others.
-The second idea is that we used the special batch technique described in Section \ref{ssec:b_SGM}.
+The second idea is that we used the special batch technique described in Section \ref{sec:contribution}.
 When we do not use input shuffling the same word will appear with all of its context words as training samples in a batch, therefore the updae will be an average of all of these training samples.

### 2019-05-07 icdm/submission/includes/introduction.tex
referencing sections

@@
 The goal of the Skip-Gram Model with negative sampling (SGNS) is to distinguish for a given word $w$, context words $c_i$ (words that appear close to $w$ in a sentence) from words dram randomly (i.e negative samples) $k_i$.
-This is achieved by maximizing the dot product of $w$ and $c_i$ and minimizing the dot product of $w$ and $k_i$.
-The SGM, espacially SGNS, gained a lot of attention, as it achieved very good results for a very simplistic model.
+This is achieved by maximizing the dot product of $w$ and $c_i$ and minimizing the dot product of $w$ and $k_i$.\\
+The SGM, espacially the SGNS, gained a lot of attention, as it achieved very good results for a very simplistic model.
 As a consequence, a lot of effort went into optimizing it.
@@
 Furthermore, we will give an extended explanation of the gradient descent optimizers used in this work.
-The reader is introduced to our Implementation in subsection \ref{sec:implementation}.
-Results are presented in  subsection \ref{sec:results} , where we will also describe used datasets, the measure applied to compare the quality of word embeddings, and finally the empirical results.
+The reader is introduced to our Implementation in Section \ref{sec:contribution}.
+Results are presented in  Section \ref{sec:results} , where we will also describe used datasets, the measure applied to compare the quality of word embeddings, and finally the empirical results.
 The last part will focus on the discussion of our results, and possible future work in Section \ref{sec:discussion} followed by a conclusion in Section \ref{sec:conclusion}.

### 2019-05-07 icdm/submission/includes/related_work.tex
background overview finished
referencing sections

@@
-\section{Related Work}
+\section{Related Work}\label{sec:related_work}
 In the original model, the optimization is done with Stochastic Gradient Descent (SGD), which is a sequential algorithm.
@@
 
-This will conclude our overview of the related work. We will now give the reader an outline of the different Gradient Descent Optimizer used in our experiments.
+This will conclude our overview of the related work. The next Section will cover our implementation of the Skip-Gram Model with negative sampling.

### 2019-05-07 icdm/submission/includes/evaluation.tex
finished intro + modified summary evaluation

@@
 To summarize, we deleted words that appear very often (subsampling) and very infrequently (min count) from our dataset, i.e text8 dataset.
-To evaluate our word embeddings we defined the cosine distance and will measur the quality our embeddings by comparing theim  to a human judged dataset of word pairs, based on their similarity.
+To evaluate our word embeddings we defined the cosine distance and will measure the quality of our embeddings by comparing theim  to a human judged dataset of word pairs, based on their similarity, i.e wordsim353.
 The reader was also given an explanation about the configuration of our network and introduced to the concept of input shuffling.

### 2019-05-07 icdm/submission/includes/introduction.tex
finished intro + modified summary evaluation

@@
 Therefore one could ask if the convergence time of the SGM can be optimized while at the same time maintaining its accuracy.\\
- This work proposes an approach that uses advanced optimizers and input shuffling to optimize the convergence time of the SGM.
-In combination both of these techniques allowed us to decrease the convergence time of the SGM.
-If combined with an optimized throughput these result could lead to an overall decrease in runtime.\\
- The rest of the paper  is structured as follows: section \ref{sec:background}  describes the SGM and its optimizations.
-Furthermore, we will give an extended explanation of the gradient descent optimizers used in this work.
+This work focused on improving the convergence time of the SGNS.
+To facilitate our calculations we propose a slightly altered version of the original SNGS, in which we compute the loss for a very large amount, i.e 2000,  of training sample by computing the sum over all training samples in a batch.
+We combined this batched approach with advanced optimizers and input shuffling to optimize the convergence time of the SGNS.
+In combination advanced optimizers and input shuffling allowed us to decrease the convergence time of the SGNS.
+If combined with an optimized throughput these results could lead to an overall decrease in runtime.\\
+ The rest of the paper is structured as follows: section \ref{sec:background}  describes the SGM and the SGNS followed by its optimzations in Section \ref{sec:related_work}.
 The reader is introduced to our Implementation in Section \ref{sec:contribution}.
-Results are presented in  Section \ref{sec:results} , where we will also describe used datasets, the measure applied to compare the quality of word embeddings, and finally the empirical results.
+Furthermore  we will describe the used datasets, the measure applied to compare the quality of word embeddings and input shuffling in Section \ref{sec:evaluation}.
+Results are presented in  Section \ref{sec:results}.
 The last part will focus on the discussion of our results, and possible future work in Section \ref{sec:discussion} followed by a conclusion in Section \ref{sec:conclusion}.

### 2019-05-07 icdm/submission/includes/results.tex
table and figure form
added bib file
added discussion without subsection challenges
referencing sections
results improvement

@@
-\subsection{Results}
+\section{Results}\label{sec:results}
 We ran multiple experiments for each optimizer.
-This subsection will only give an overview of the achieved results.
+This Section will  give an overview of the achieved results.
 Each subsection will give an explanation over the achieved result with a specific optimizer.
@@
 The first challenge for each optimizer was to find a correct learning rate.
-As SGD is the optimizer used in Gensim \cite{gensim} we first tried the same learning rate as Gensim \cite{gensim} and then performed a random search to find a better one.
+As SGD is the optimizer used in Gensim \cite{gensim} we first tried the same learning rate as the default value in Gensim \cite{gensim}, i.e 0.01,  and then performed a random search to find a better one.
 As expected a bell curve shape resulted, a learning rate that is too high leads to diversion and a learning rate that is too low leads to a training time that is too slow.
@@
 \centering
-\includegraphics[scale=0.45]{images/results_sgd_shuffle}
+\includegraphics[scale=0.3]{images/results_sgd_shuffle}
 \caption{Training time Stochastic Gradient Descent with input Shuffling}
@@
 \subsubsection{Momentum and Nesterov}
-Momentum and NAG \cite{nag} both have an additional hyperparameter $\gamma$, that, as described in Section \ref{optimizers}, defines the percentage of the previous gradient that will be added to the current gradients.
-We set $\gamma = 0.9$ as this is the typical value and did not alter it during our experiments.
+Momentum and NAG \cite{nag} both have an additional hyperparameter $\gamma$, that, defines the percentage of the previous gradient that will be added to the current gradients.
+We set $\gamma = 0.9$ as this is a typical value and did not alter it during our experiments.
 Momentum and Nesterov alone respectively only slightly decrease or increase the convergence time.
@@
 \centering
-\begin{minipage}{.5\textwidth}
-\centering
+
 \includegraphics[scale=0.3]{images/results_mom_shuffle}
@@
 \label{fig:results_mom}
-\label{fig:test1}
-\end{minipage}%
-\begin{minipage}{.5\textwidth}
+\end{figure}
+
+\begin{figure}[h]
 \centering
@@
 \label{fig:results_nag}
-\end{minipage}
 \end{figure}
 \subsubsection{Adagrad}
-Adagrad \cite{adagrad} is a very interesting tool for learning word embeddings as it decreases the learning rate for very frequent occurring features, and vice versa (this is explained in detail in Section \ref{ssec:adagrad}).
-Because words that appear very frequently often do not have a semantic gain that is as important as words that appear less frequently to their context words, it's good to have a lower learning rate.
-So, in theory, Adagrad is particularly well suited for our task.
+Adagrad \cite{adagrad} is a very interesting tool for learning word embeddings as it decreases the learning rate for very frequent occurring features, and vice versa).
+Because words that appear very frequently often do not have a semantic gain that is as important as words that appear less frequently to their context words, it's good to have a lower learning rate for such frequent words.
+So, in theory, Adagrad is particularly well suited for our task, as for example Pennington et al.
+used Adagrad in the training of Glove \cite{glove}, another system used to create word embeddings.
 This was confirmed empirically as our model converged in 4 epochs.
@@
 \centering
-\includegraphics[scale=0.45]{images/results_adagrad_shuffle}
+\includegraphics[scale=0.3]{images/results_adagrad_shuffle}
 \caption{Training time Adagrad with input Shuffling}
@@
 Because it didn't have any learning rate to tune, we only did 2 experiments, with and without input shuffling.
-We are aware of the fact that there are additional hyperparameters but did, for simplicity reason, and because their effect is not as high as the learning rate in other optimizers, decide not to tune it.
-We left it to its default value $\rho = 0.9$.
-This parameter defines the percentage taken when calculating the exponentially decaying average of past gradients, as explained in \ref{ssec:adadelta} Adadelta did not manage to achieve a word similarity of 0.66.
+We are aware of the fact that there are additional hyper parameters to Adadelta.
+We decide not to tune theim for to reasons: first for simplicity reasons and second because their effect is not as high as the learning rate.
+The parameter that defines the percentage taken when calculating the exponentially decaying average of past gradients was set to $\rho = 0.9$.
+Adadelta did not manage to achieve a word similarity of 0.66.
 It only converged to a similarity of 0.59.
 It did this in 20 epochs without input shuffling and in 3 with input shuffling, as can be seen in Table \ref{table:results_adadelta}
-\begin{table}[h]
-\centering
-\begin{tabular}{|l|l|l|}
-\hline
-
-\end{tabular}
-\caption{}
-\label{}
-\end{table}
 
@@
     \centering
-            \includegraphics[scale=0.45]{images/results_adam_shuffle} 
+            \includegraphics[scale=0.3]{images/results_adam_shuffle} 
     \caption{Training time Adam with input Shuffling}

### 2019-05-07 icdm/submission/includes/conclusion.tex
referencing sections
conclusion improvement

@@
-\section{Conclusion}\label{chap:conclusion}
+\section{Conclusion}\label{sec:conclusion}
 
 This work provides an overview of the Skip Gram Model with negative Sampling (SGNS) and the numerous successful attempts of optimizing the throughput of the model.
-As this is the case, no effort went into optimizng the convergence time of the SGNS, therefore this work focused on this point.
-We decided to use advanced optimizers and input shuffling as optimizing techniques.
-After giving a short overview over Gradient Descent algorithms this work proposes a slighlty altered version of the SGNS, where the idea is to compute the loss over the sum of a high number of training samples, i.e 2000,  instead of computing it for each individually.
+As this is the case, no effort went into optimizng the convergence time of the SGNS, therefore this work focused on this task.
+We decided to use advanced optimizers and input shuffling to achieve a better convergence time.This work proposes a slighlty altered version of the SGNS, where the idea is to compute the loss over the sum of a high number of training samples, i.e 2000,  instead of computing it for each individually.
 We did this as it allowed us to compute more models and analyze the convergence time faster.
-We used the text8 dataset and used  word similarity as a quality measure for the word embeddings (WE).
-We used the State of the art implementation Gensim to compare our self.
-We did achieve a better convergence time than gensim with Adam as an optimizer and the use of input shuffling.
+We used the text8 dataset to train our model and  word similarity as a quality measure for the word embeddings (WE).
+We used the sate of the art implementation Gensim to compare our self.
+We did achieve a better convergence time than Gensim with Adam as an optimizer and the use of input shuffling.
 Gensim convereged in 4 epochs to a word similarity of 0.66 and our model only took 2 epochs to achieve the same quality.

### 2019-05-07 icdm/submission/paper.tex
abstract improved

@@
 %or Math in Paper Title or Abstract.
-The Skip-gram Model with negative sampling (SGNS) is an effective algorithm to create word embeddings.
+The Skip-Gram Model with negative sampling (SGNS) is an effective algorithm to create word embeddings.
 SGNS uses Stochastic Gradient Descent (SGD) as it's learning algorithm.
@@
 We used the Text8 dataset to train our model and measured the quality of our word embeddings with the wordsim353 dataset, which measures the quality of word embeddings by judging the similarity of different words.
-We trained our model with multiple advanced optimizers: momentum, Nesterov accelerated gradient, Adagrad, Adadelta, and Adam.
+We trained our model with multiple advanced optimizers: momentum, Nesterov Accelerated Gradient, Adagrad, Adadelta, and Adam.
 We also applied input shuffling to the above optimizers.
@@
 Hence, this work shows that advanced optimizers combined with input shuffling do decrease the convergence time of the SGNS.
-These results must be confirmed with other datasets, but optimizing our model in terms of throughput has the potential of further reducing the overall run time.
+These results must be confirmed with other datasets, but optimizing our model in terms of throughput has the potential of further reducing the overall run time of the SGNS.
 
@@
 \begin{IEEEkeywords}
-component, formatting, style, styling, insert
+NLP, Word Embeddings, Skip-Gram Model, Negative Sampling, Input Shuffling, Advanced Optimizers
 \end{IEEEkeywords}

### 2019-05-07 icdm/submission/includes/discussion.tex
summary and transition evaluation
discussion improved

@@
-\section{Discussion}\label{chap:discussion}
+\section{Discussion}\label{sec:discussion}
 
 %Discuss the results. What is the outcome of your experimetns?
-This subsection shortly discusses the empirical results and then extensively compares the findings of this work to the existing literature while trying to explain some differences.
-It is followed by a subsection describing the limitations and possible extensions of this work.
-This subsection is concluded by the problems we encountered during the implementation so others can avoid them in the future.
+This subsection shortly discusses the empirical results and then extensively compares the findings of this work to the existing literature while trying to explain some of the differences.
+It is concluded by a subsection describing the limitations and possible extensions of this work.
 
 \subsection{Our work}
-This subsection discusses our findings, first we will analyze how input shuffling could influence the learning rate and conclude with the discussion of unexpected results.
+This Section discusses our findings.
+First it will analyze the possible reasons behind the influence of input shuffling to the learning rate.
+Second it will conclude with the discussion of unexpected results.
 
 \subsubsection{Shuffling and learning rate with SGD}
-When the model used SGD as an optimizer it was able to use higher learning rates when input shuffling was used, as shown in figures \ref{fig:results_sgd}, \ref{fig:results_mom} and \ref{fig:results_nag} Therefore arises the questions why these phenomena happen, especially as it did not happen with advanced optimizers.
-One possibility is the use of the batched version of the SGNS.
-In consequence, when the input is not shuffled the same word will often appear in one batch, hence the optimizer is presented with an average value of the gradient, which can be imprecise.
+When the model used SGD as an optimizer it was able to use higher learning rates when input shuffling was used, as shown in figures \ref{fig:results_sgd}, \ref{fig:results_mom} and \ref{fig:results_nag}.
+Therefore arises the questions why these phenomena happen, especially as it did not happen with advanced optimizers.
+We attribute this phenomena to the use of the batched version of the SGNS.
+In consequence to this batched version, when input shuffling is not used a lot of words will appear multiple times in the same batch, as explained in \ref{ssec:shuffling}.
+Therefore, the gradients will be an average of all the training samples.
+When the input is shuffled, less words will appear multiple time which will make the gradients more precise.
 This is counter-attacked by advanced optimizers as they have adaptive learning rates, i.e a frequent appearing feature will have a low learning rate.
+Since high frequency words will appear more often multiple times in a batch, and therefore their average will not influence the results as much.
+The two above arguments could explain why a higher learning rate combined with SGD achieved better results with input shuffling but not when advanced optimizers are used.
 
@@
 \subsection{Comparison to Gensim}
-This subsection will compare our finding extensively to Gensim \cite{gensim}.
+This Section will compare our finding extensively to Gensim \cite{gensim}.
 As explained in Section \ref{ssec:gensim}, Gensim is optimized to have a very high throughput, this made it possible to achieve a lot of computations.
 Furthermore, Gensim provides access to the loss and the resulting word embeddings, which facilitated the comparison process.
+This Section is Structured as follows:First it will describe the used configuration of Gensim, second it will compare Gensim to our implementation of the SGNS with the use of SGD and finally compare our model with Adam to Gensim.
+A graphic showing the comparison of our models to Gensim can be found in Figure \ref{fig:gensim_vs_adam}.
 
@@
 \subsubsection{Configuration of Gensim}
-The training with Gensim has a lot of possible parameters an extended list can be found in the appendix.
-This subsection will only describe the parameters we changed from the default setting.
-The description of each parameter has the following form: \\
-\texttt{name} (type) -- \textit{Description} -- Value\\
-Parameters:
-\begin{itemize}
-\item \texttt{sentences} (iterable of iterables) -- \textit{Dataset} -- text8 document splitted into sentences of 20 words
-\item \texttt{size}(int) – \textit{Dimensionality of the word vectors } -- 100
-\item \texttt{window} (int) --\textit {Maximum distance between the current and predicted word within a sentence }-- 5
-\item \texttt{min\_count }(int) --\textit{ Ignores all words with total frequency lower than this }-- 5
-\item \texttt{workers} (int) -- \textit{ Number of threads used to train the model} -- 4
-\item \texttt{sg} ({0, 1}) --\textit{ Training algorithm: 1 for skip--gram}--1
-\item \texttt{negative} (int) --\textit{Number of negative samples}-- 10
-\item \texttt{ ns\_exponent} (float) --\textit{ Exponent in the unigram distribution, when choosing random samples, as shown in Equation \ref{eq:unigram} }-- 0.75
-\item \texttt{alpha} (float) --\textit{ The initial learning rate. }-- 0.025
-\item \texttt{min\_alpha} (float) --\textit{ Learning rate will linearly drop to min\_alpha as training progresses }--0.0001
-\item \texttt{sample} (float) --\textit{ Treshold for subsampling as described in Equation \ref{eq:sampling}} -- 1e--4
-\item \texttt{iter} (int) --\textit{ Number of iterations (epochs) over the corpus. }-- 10
-\
-\item \texttt{compute\_loss} (bool) --\textit{If True, loss is stored at the end of each batch}-- True
-\item \texttt{callbacks} (iterable of CallbackAny2Vec) --\textit{ Set of functions executed in order to follow the loss and the progress of the model in word similarity }-- see Appendix
-\end{itemize}
+To compare our-self in a correct  manner we used the same dataset, with the same preprocessing parameters, i.e subsampling and min count.
+The hyper parameters of the network (window size, embedding size, number of negative sample, exponent to which the unigram distribution is raised to chose neg samples) are equivalent to ours, described in Section \ref{ssec:config}.
+The only difference with our model is the learning rate.
+Gensim has a starting learning rate of $0.025$ and linearly decrease it at every epoch to $0.0001$.
 
@@
 The fact that we only used one dataset as well that it's a small one is problematic.
-By using a very small dataset we do not use the model in the condition it is most needed for, as the dataset used in practice usually consists of more than 1 billion words.
+\\
+~~\textit{Problem with a small dataset:} \\ By using a very small dataset we do not use the model in the condition it is most needed for, as the dataset used in practice usually consists of more than 1 billion words.
 There is a small argument that can be made for machine translation as the use of small parallel corpus is not unusual in this field.
-But the main issue with using only one data set it that it has been shown that some optimizers perform better with specific shapes of loss functions.
-To make a compelling argument it's, therefore, necessary to show that our model with the use of Input shuffling and Adam as its optimizer also outperform Gensim with other data sets, so that the claim can hold consistently, it needs to be confirmed with other datasets as well.
-Furthermore, our implementation did not outperform Gensim in run time, as this was not the goal of our work.
+\\
+~~\textit{Problem with using only one dataset:}\\
+It has been shown that some optimizers perform better with specific shapes of loss functions.
+To make a compelling argument it's, necessary to show that our model with the use of Input shuffling and Adam as its optimizer also outperform Gensim with other data sets, so that the claim can hold consistently.\\
+Finally, our implementation did not outperform Gensim in run time, as this was not the goal of our work.
 Therefore, one could improve an already existing, optimized version, with input shuffling and advanced optimizers and should achieve a better run time than Gensim.

### 2019-05-07 icdm/submission/includes/evaluation.tex
discussion improved

@@
 
-\subsection{Configuration of the network}
+\subsection{Configuration of the network}\label{ssec:config}
 The skip gram model has a lot of possible parameters, that can be tuned.
@@
 \fi
-\subsection{Input Shuffling}
+\subsection{Input Shuffling}\label{ssec:shuffling}
 We used input shuffling as a technique to optimize the SGM.

### 2019-05-09 icdm/submission/paper.tex
wip abstract

@@
 %or Math in Paper Title or Abstract.
-The Skip-Gram Model with negative sampling (SGNS) is an effective algorithm to create word embeddings.
+The Skip-Gram Model with negative sampling (SGNS) is a simple and effective algorithm to create word embeddings.
 SGNS uses Stochastic Gradient Descent (SGD) as it's learning algorithm.
@@
 Therefore, our work focuses on the latter.
-In this work we used two techniques to achieve a better convergence time, namely advanced optimizers and input shuffling.
-We used the Text8 dataset to train our model and measured the quality of our word embeddings with the wordsim353 dataset, which measures the quality of word embeddings by judging the similarity of different words.
-We trained our model with multiple advanced optimizers: momentum, Nesterov Accelerated Gradient, Adagrad, Adadelta, and Adam.
+To achieve a better convergence time we use advanced optimizers and input shuffling as optimization techniques.
+Here are the components used to evaluate our work: to train our model we used the text8 dataset - to measure the quality of our word embeddings we used the wordsim353 dataset, which assess that similar words have vector represenations that are close to each other.
+We trained our model with multiple advanced optimizers: Momentum, Nesterov Accelerated Gradient, Adagrad, Adadelta, and Adam.
 We also applied input shuffling to the above optimizers.

### 2019-05-09 icdm/submission/includes/introduction.tex
introduction reworked

@@
 There are two ways to create such WE, either arbitrarily or with the purpose of capturing the semantic meaning of the words, i.e.
-vector representations of words that are syntactically or semantically similar will be near to each other.
+vector representations of words that are syntactically or semantically similar will be close to each other in the vector space.
 By capturing semantic or syntactic meaning WE have shown to facilitate a lot of subsequent NLP tasks, such as entity recognition, machine translation or sentence classification.
@@
 \cite{bengio} but more recently Mikolov et al.
-\cite{mikolov} introduced a software package called w2vec, which uses a simpler network produces state of the art results.
+\cite{mikolov} introduced a software package called w2vec, which uses a simpler network that produces state of the art results.
 One of the proposed algorithms in this software package is the Skip-Gram Model (SGM).
@@
 \cite{mikolov2} then introduced an alteration to the SGM called negative sampling.
-The goal of the Skip-Gram Model with negative sampling (SGNS) is to distinguish for a given word $w$, context words $c_i$ (words that appear close to $w$ in a sentence) from words dram randomly (i.e negative samples) $k_i$.
+The goal of the Skip-Gram Model with negative sampling (SGNS) is to distinguish for a given word $w$, context words $c_i$ (words that appear close to $w$ in a sentence) from words drawn randomly (i.e negative samples) $k_i$.
 This is achieved by maximizing the dot product of $w$ and $c_i$ and minimizing the dot product of $w$ and $k_i$.\\
@@
 To remedy this problem Mikolov et al.
-used Hogwild \cite{hogwild}, where different threads can access a shared model and update it.
+used Hogwild!
+\cite{hogwild}, proposed by Recht et al., where different threads can access a shared model and update it.
 As this is not an optimal solution Yi et al.
@@
 This work focused on improving the convergence time of the SGNS.
-To facilitate our calculations we propose a slightly altered version of the original SNGS, in which we compute the loss for a very large amount, i.e 2000,  of training sample by computing the sum over all training samples in a batch.
-We combined this batched approach with advanced optimizers and input shuffling to optimize the convergence time of the SGNS.
+To facilitate our calculations we propose a slightly altered version of the original SNGS, in which we compute the loss for a very large amount of training samples, i.e 2000.
+The loss of such a batch was computed by taking the sum over the loss all individual training samples in the specific batch We combined this batched approach with advanced optimizers and input shuffling to decrease he convergence time of the SGNS.
 In combination advanced optimizers and input shuffling allowed us to decrease the convergence time of the SGNS.

### 2019-05-09 icdm/submission/includes/background.tex
referencing sections
background improved

@@
-\section{Background}\label{chap:background}
+\section{Background}\label{sec:background}
 \iffalse
@@
 This section gives in its first subsection a broad overview over the functioning of the Skip-Gram Model, and its second section describes the Technique called Negative Sampling introduced by Mikolov et al.
-\cite{mikolov2} used to make the Skip-Gram model practicable.
+\cite{mikolov2} used to make the Skip-Gram model more efficient
 \subsection{The Skip-Gram Model}
@@
 See Figure \ref{fig:ntw_architecture} for an illustration.
-First the ntw has an input layer, a vector of size $V$ ( $V$ being the size of the vocabulary), represint words as an one hot encoding.
-Secondly, the projection layer wight Matrix $M_{in}$ soters $V$ $D$-dimensional vectors for the corresponding word in the vocabulary  in it's rows.
-The projection layer serves as a look up tavle for our WE (if a one hot vector is multiplied with $M_{in}$ the corresponding WE will result).
-Finally, the output layer wight matrix $M_{out}$ also stores $V$ $D$-dimensional vector representation of words.The idea behind these 2 Matrices is that the input layer will represent our words as center word and the output layer as context words.
+First the network has an input layer, a vector of size $V$ ( $V$ being the size of the vocabulary), represent words as an one hot encoding.
+Secondly, the projection layer weight Matrix $M_{in}$ stores $V$ $D$-dimensional vectors for the corresponding word in the vocabulary  in it's rows.
+The projection layer serves as a look up table for our WE (if a one hot vector is multiplied with $M_{in}$ the corresponding WE will result).
+Finally, the output layer wight matrix $M_{out}$ also stores $V$ $D$-dimensional vector representation of words.The idea behind these 2 Matrices is that the projection layer will represent our words as center word and the output layer as context words.
 
@@
 \begin{equation} \label{eq:softmax}
-p(w_{i+j}|w_i)= \frac{exp( \tilde{v}_{w_{i+j}}^Tv_{w_i})}{\sum_{w=1}^v exp(\tilde{v}_w^Tv_{ w_i)}}
+p(w_{i+j}|w_i)= \frac{exp( \tilde{v}_{w_{i+j}}^Tv_{w_i})}{\sum_{w=1}^V exp(\tilde{v}_w^Tv_{ w_i)}}
 \end{equation}
 
-Here $\tilde{v}_{w_i}$ and $ v_{w_i}$ are the vector representations, stored respectively in the projection Layer ($M_{in}$) and in the output layer ($M_{out}$).
+Here $ v_{w_i}$ and $\tilde{v}_{w_i}$ are the vector representations, stored respectively in the projection Layer ($M_{in}$) and in the output layer ($M_{out}$).
 There is a problem with the classical softmax.
@@
 Where $v_c$ and $\tilde{v_w }$, can be interpreted as before in Equation \ref{eq:softmax} and $\sigma(x) = \frac{1}{1+e^{-x}}$.
-To  understand why this loss function is working so well, is to assume that two vectors are similar if their dot product is high.
+To  understand why this loss function is working so well, one needs to assume that two vectors are similar if their dot product is high.
 Therefore, we maximize the dot product of similar words $(w,c)$ ($c$ appears in the context of $w$) and minimize it for dissimilar words $(w,k_i)$ (those were selected randomly, we therefore assume that their are not similar).
@@
 
-One question remains: how are the negative samples selected? Mikolov et al. \cite{mikolov2} used the following unigram distribution:
+One question remains: how are the negative samples selected?
+Mikolov et al.
+\cite{mikolov2} used the following unigram distribution, to define the probability $P(w)$ of a word $w$ being drawn as a negative sample:
 
@@
 Raising the unigram distribution to the power of $\frac{3}{4}$ makes it less likely for a word to be drawn if it often appears in the dataset in comparison to the basic unigram distribution.
-See figure \ref{fig:frequency_ex} for an example.
+See figure \ref{fig:frequency_ex} for an example.\\
 \begin{figure}[ht]
@@
 \end{figure}
-It's quite easily observable that this approach will outperform the classical softmax in computation time, as it only compute the sum over $K$ output nodes.
-Now the question arises if the accuracy is good enough but according to Mikolov et al.
+With all of the above knowledge it's understandable that the SGNS will  will outperform the classical softmax in computation time.
+Now the question arises if the accuracy is maintained but according to Mikolov et al.
 \cite{mikolov2} the negative sampling method "is an extremely simple training method that learns accurate representations".
@@
 Therefore, it is a good solution to the problem raised by the classical softmax.
-We now have enough background knowledge about the SGM to look at how it can be optimized. The next subsection introduces earlier approaches to optimize the SGM.
+We now have enough background knowledge about the SGM to look at how it can be optimized. The next section introduces earlier approaches to optimize the SGM.

### 2019-05-09 icdm/submission/includes/related_work.tex
intel paper reworked good

@@
 \cite{intel} focused on the Skip-Gram Model with negative sampling.
-The focus of this work wasto reduce the cost of our vector to  vector multiplication in our loss funcion, see Equation \ref{eq:obj_neg_samples}, where the dot product of the word vector and the negative sampled as well as the context word is computed.
+The focus of this work was to reduce the cost of our vector to  vector multiplication in our loss funcion, see Equation \ref{eq:obj_neg_samples}, where the dot product of the word vector and the negative sampled as well as the context word is computed.
 The main idea in this paper is to convert the level 1-BLAS (Basic Linear Subprogram) vector to vector operations to a level 3-BLAS matrix multiplication operation, which are efficiently implemented into hardware and consequently faster.
-This is achieved, by using the same negative samples for each context word of a given word $w$.
-Instead of using for each context word a vector to vector multiplication we can transform this into a matrix multiplication, under the assumption that we will not lose accuracy by sharing the same negative samples.
+To achieve such a transformation, Ji et al.
+decided to use the same negative samples for given amount of training samples.
+The training samples that will have the same negative samples are chosen as follows: given a word $w_c$ one can construct all the training samples where $w_c$ is a context word.
+This will result in the following training batch $X = \{({w_{in}}^i, w_c) | w_c~appears~in~the~context~of~{w_{in}}^i\}$.
 The matrix multiplication can be represented in the following way.
@@
 \begin{bmatrix}
-w \\
+w_c \\
 w_{n_1} \\
@@
 \begin{bmatrix}
-w_{c_1}\\
+w_{in}^1\\
 \vdots\\
-w_{c_{2m}}\\
+{w_{in}}^{2m}\\
 \end{bmatrix}
 \]
-
-where $w$ is our given word, $w_{n_1}...w_{n_k}$ are the shared negative samples, with $k \in [5,20]$, and $w_{c_1}...w_{c_2m}$ are the words inside of the context window $m$ of $w$, with $m \in [10,20]$, also called a batch of input context words.
-After each batch the model updates the weights of the used vectors.
-This model achieves a 3.6 fold increase in throughput, by only losing 1\% of accuracy.
-An aspect that is not as useful to us is that the experiments were done on CPU, as modern GPU's are often used in many machine learning libraries, as with CUDA for example, there still need work to be done to optimize it with GPU's.
-This was done by Seulki and Youngmin \cite{gpu}, which is described in the next subsection.
+where, $w_{n_1}...w_{n_k}$ are the shared negative samples.
+The above Matrix multiplication will compute the loss for each training sample and the weights of the used vectors will be updated after each batch.
+This method achieves a 3.6 fold increase in throughput, by only losing 1\% of accuracy.
+The presented results were achieved on CPU's, as many modern machine learning libraries make use of the parallelizing capabilities of GPU's for matrix multiplications, the next subsection covers the work done by Seulki and Youngmin \cite{gpu} that parellelized the SGNS on GPU's.
 

### 2019-05-09 icdm/submission/includes/related_work.tex
caching plus start intel describption

@@
 This idea was proposed by Vuurens et al. \cite{efficient}. The architecture used here is the SGM with a hierarchical softmax. 
-
 The general idea is to cache the most frequently used weight vectors, and update the model after a certain number of seen training samples, while taking in consideration the changes that happend during these number of seen training samples (the paper used the number 10).
 The unfrequent weight vectors are handled as in the original paper, i.e updated in shared memory after each training sample.
-This technique  allows it to avoid collosision for the most frequent words, and the author therefore hoped to be able to increase the number of concurrent threads.The general idea is to cache the most frequently used nodes of the binary tree used to memorize the probability distribution and update them on the shared single model after a certain number of seen words.
+This technique  allows it to avoid collosision for the most frequent words, and the author therefore hoped to be able to increase the number of concurrent threads.
 The paper produced interesting results as they managed to decrease execution time by increasing the number of cores used for the calculation.
@@
 This work proposes a very good way to parallelize the SGM, as in particular, it allows using more cores during the computation.
-As this approached focused on the Hierarchical softmax, in contrast to our work which used negative sampling, the next subsection covers optimizations of the Skip-Gram Model with negative sampling(SGNS).
+As this approached focused on the Hierarchical softmax, in contrast to our work which uses negative sampling, the next subsection covers optimizations of the Skip-Gram Model with negative sampling(SGNS).
 
@@
 In contrast to the above approach, this one, introduced by Ji et al.
-\cite{intel} focused on the Skip-Gram Model with negative sampling.
-The focus of this work was to reduce the cost of our vector to  vector multiplication in our loss funcion, see Equation \ref{eq:obj_neg_samples}, where the dot product of the word vector and the negative sampled as well as the context word is computed.
-The main idea in this paper is to convert the level 1-BLAS (Basic Linear Subprogram) vector to vector operations to a level 3-BLAS matrix multiplication operation, which are efficiently implemented into hardware and consequently faster.
+\cite{intel} focused on the SGNS.
+The focus of this work was to speed up the computation of the loss function of the SGNS described in Equation \ref{eq:obj_neg_samples}.
+Where the dot product of the center word  with negative samples as well as the context word is computed.
+The dot product is a level 1-BLAS operation and is not as efficiently implemented into hardware, as matrix multiplication which are a level 3-BLAS operation.
 To achieve such a transformation, Ji et al.

### 2019-05-09 icdm/submission/includes/related_work.tex
shenanigans intel

@@
 The training samples that will have the same negative samples are chosen as follows: given a word $w_c$ one can construct all the training samples where $w_c$ is a context word.
-This will result in the following training batch $X = \{({w_{in}}^i, w_c) | w_c~appears~in~the~context~of~{w_{in}}^i\}$.
-The matrix multiplication can be represented in the following way.
+The following training batch $X$ will result: $X = \{({w_{in}}^i, w_c) | w_c~appears~in~the~context~of~{w_{in}}^i\}$.
+The matrix multiplication can be represented in the following way:
 \[
@@
 The above Matrix multiplication will compute the loss for each training sample and the weights of the used vectors will be updated after each batch.
-This method achieves a 3.6 fold increase in throughput, by only losing 1\% of accuracy.
+Ji et al.'s method achieves a 3.6 fold increase in throughput, by only losing 1\% of accuracy.
 The presented results were achieved on CPU's, as many modern machine learning libraries make use of the parallelizing capabilities of GPU's for matrix multiplications, the next subsection covers the work done by Seulki and Youngmin \cite{gpu} that parellelized the SGNS on GPU's.

### 2019-05-09 icdm/submission/includes/related_work.tex
gpu updated

@@
 The above Matrix multiplication will compute the loss for each training sample and the weights of the used vectors will be updated after each batch.
-Ji et al.'s method achieves a 3.6 fold increase in throughput, by only losing 1\% of accuracy.
+Ji et al.'s method achieves a 3.6 fold increase in throughput, by only losing 1\ % of accuracy.
 The presented results were achieved on CPU's, as many modern machine learning libraries make use of the parallelizing capabilities of GPU's for matrix multiplications, the next subsection covers the work done by Seulki and Youngmin \cite{gpu} that parellelized the SGNS on GPU's.
 
-\subsection{Accelleration of word2vec by Using GPU's}
-Seulki and Youngmin \cite{gpu} focused on getting a better throughput on the SGM when using GPU's.
-As the SGM is a sequential algorithm, is is not easy to parallelize it, especially if one wants to parallelize the training of individual training samples.
+\subsection{Accelleration of word2vec by using GPU's}
+Seulki and Youngmin \cite{gpu} focused on getting a better throughput on the SGM with the use of GPU's.
+As the SGM is a sequential algorithm, it is not easy to parallelize it, especially if one wants to parallelize the training of individual training samples.
 As the algorithm goes sequentially over a sentence, the samples next to each other, in order of execution, will almost every time have the same input word.
@@
 They achieved their best results with 2 concurrent GPU'S.
-The accomplished results were very good as they managed a 20x speedup compared to a single threaded CPU execution, which is a 3x increase in comparison to the original C code from Mikolov et al.
+The acomplished results were very good as they managed a 20x speedup compared to a single threaded CPU execution, which is a 3x increase in comparison to the original C code from Mikolov et al.
 \cite{mikolov2}.
-Al this speedup was achieved, while maintaining the accuracy of the model.
+The speedup was achieved, while maintaining the accuracy of the model.
 The problem with this and all the above optimization is that the code is not easily available or of use.

### 2019-05-09 icdm/submission/includes/related_work.tex
gensim updated

@@
 Furthermore, it made use of BLAS's and precomputed sigmoid tables, while also trying to parallelize the training of different sentences.
-This finally yielded in a 4x speedup in run time.
-Gensim is an important tool as it allows us, as a python library, to compare our data rather easily.
-It was also used in related work \cite{intel} and is therefore of value, as it allows to us compare our work in a simpler way.
-This concludes our overview of the optimizations of the throughput of the SGM.
-In the next subsection, we give a quick outlook of what has been done in the field of context-sensitive word embeddings.
-
-This will conclude our overview of the related work. The next Section will cover our implementation of the Skip-Gram Model with negative sampling.
+This finally yielded in a 4x speedup in run time compared to the orignal C code from Mikolov et al.
+Gensim as a python library, and because it was already used in related work \cite{intel} made the comparison easy and of value.
+\\ This concludes our overview of  related work.
+The next Section will cover our implementation of the Skip-Gram Model with negative sampling.

### 2019-05-09 icdm/submission/includes/contribution.tex
referencing sections
contribution updated

@@
-\section{Batched SkipGramModel} \label{ssec:b_SGM}
+\section{Batched SkipGramModel} \label{sec:contribution}
 This section will give an overview of our implementation.
-The proposed implementation is a slighlty altered version of the original SGNS.
+The proposed implementation is a slighlty altered version of the origina Skip-Gram Model with negative sampling (SGNS).
 The idea behind this altering, is to compute the loss for multiple words and context pairs at the same time, this is inspired from mini-batch training, a model often used in other machine learning problem.
@@
 The exact process will be described in the following paragraph.
-
 \subsection{Forwarding}
-As this represents the challenging part of our Implementation the forwarding method is explained step by step. Each time step is illustrated to make the explanation clearer.\\
-Let $X = {(v_1,c_1),(v_2,c_2),(v_3,c_3)}$, be the training batch, where $(v_i,c_i)$ is a training sample constituted by a word and one of its context words. \\
-\textit{Input:}\\
+As the computation of the new loss function represents the challenging part of our implementation the forwarding method is explained step by step.
+Each time step is illustrated to make the explanation clearer.\\
+Let $X = {(v_1,c_1),(v_2,c_2),(v_3,c_3)}$, be the training batch, where $(v_i,c_i)$ is a training sample constituted by a word and one of its context words. 
+\subsubsection{Input}
 The forwarding method will accept two vectors $v$ and $c$, and a matrix $A$ as an input.
@@
 
-\textit{Concatenation of samples:}\\
+\subsubsection{Concatenation of samples}
 The concatenation of the vector $c$ and the Matrix $A$ will result in a Matrix $\tilde{A}$, with:\\
@@
 c_3 & k_{1,3} & k_{2,3}& k_{3,3}\\
-\end{bmatrix}$\\
+\end{bmatrix}$
+\bigskip
 
-\textit{Embeddings:}\\
+\subsubsection{Embeddings}
 Now it's necessary to access the word embeddings.
@@
 \\
+
 $E_v = \begin{bmatrix}
@@
 
-\textit{Batch multiplication and negation of samples:}\\
+\subsubsection{Batch multiplication and negation of samples}
 Now we need to compute the dot product of each word vector with its pair and the negative samples, exactly as done as in the original loss function of Mikolov et al.
@@
 This will result in a $n\times k+1$ Matrix $S$.
-Now we only have to negate the last $k$ rows with minus one.
+Now we only have to negate the last $k$ rows.
 The sum of each row represents the loss computed in Equation \ref{eq:obj_neg_samples}, for each word context pair.
 Since computation time is too long with this approach the loss function is altered\\
+
 $S = \begin{bmatrix}
@@
 \tilde{v_3} \cdot \tilde{c_3} &-\tilde{v_3} \cdot c_3 \tilde{k_{1,3}} & -\tilde{v_3} \cdot c_3 \tilde{k_{2,3}}&-\tilde{v_3} \cdot \tilde{k_{3,3}}\\
-\end{bmatrix}$\\
+\end{bmatrix}$\bigskip
 
-\textit{Loss function:}\\
+\subsubsection{Loss function}
 Summing the matrix $S$ and multiplying it with $-1$ (to make the problem a minimizing problem) results in the loss for our entire batch.

### 2019-05-09 icdm/submission/includes/evaluation.tex
evalutation done

--- 2019-05-07 icdm/submission/includes/evaluation.tex
@@
 
-This section gives an overview of the used datasets, the used metric to evaluate our models, the configuration of our model, and finally an explanation about inputa shuffling
+This section gives an overview of the used datasets, the used metric to evaluate our models, the configuration of our model, and finally an explanation about input shuffling
 
@@
 Subsampling is a technique introduced by Mikolov et al.
-\cite{mikolov} to reduce the dataset size while at the same time increasing the quality of the dataset, i.e getting better word embeddings with it.
+\cite{mikolov} to reduce the size of the dataset while at the same time increasing the quality of the dataset, i.e increasing the quality of the word embeddings (WE).
 The idea behind subsampling is the removal of very frequent words such as: "the, as, it, of".
@@
 recommend a value between $0$ and $10^{-5}$, depending on the size of the dataset.
-We experimented with different values and $10^{-4}$ seemed the most suited.
+We experimented with different values and $10^{-4}$ seemed the most suited on the text8 dataset.
 We did this by looking at a random set of sentences and judging the results manually.
@@
 The table shows the first 20 words of our dataset, without the words that were subsampled according to a threshold sample.
-Stats about subsampling can be found in Table \ref{table:treshold}.
-\begin{table}[h]
-\centering
-\begin{tabular}{|l|l|l|l|l|l|}
-\hline
-\end{tabular}
-
-\label{table:treshold}
-\end{table}
+Stats about subsampling can be found in Table \ref{table:threshold}.
 
@@
    \end{tabular}%
-   \label{table:with_20}%
+   \label{table:threshold}%
 \end{table}
@@
 Lastly, a word that only appears one time in our dataset will be very dependent on its original initialization.
-This is the case because it will only be updated with its context pairs once, which is only a dozen of times in practice and then won't be updated anymore.
+The weights of the network that represents the word, will only be updated when the word itself is used as input.
+As the word appears very infrequently the word representation will not differ much from the initialization, hence being very dependent from it.
 For all the above reasons, we applied this technique.
-Similar to subsampling, it should in theory improve the quality of the word embeddings and will decrease the computation time.
+Similar to subsampling, it will decrease computation time and  should in theory improve the quality of the word embeddings.
 
@@
 \subsubsection{Cosine distance}
-The cosine similarity, this is not the cosine distance, of vectors $v$ and $w$ is the cosine of the angle between the two vectors It can be calculated by taking the dot product of $v$ and $w$ and dividing it by the magnitude of $v$ and $w$ multiplied with each other.
+The cosine similarity, this is not the cosine distance, of vectors $v$ and $w$ is the cosine of the angle between the two vectors.
+It can be calculated by taking the dot product of $v$ and $w$ and dividing it by the magnitude of $v$ and $w$ multiplied with each other.
 We get:
@@
 The cosine of 0\textdegree ~is 1, it's 0 for two vectors that are orthogonal to each other and vectors that point in the opposite direction will have a cosine of their angle of -1.
-This is not a good distance measure as -1 is smaller than 0, and therefore two vectors pointing away from each other would be closer than two orthogonal vectors, but by subtracting 1 from the cosine of the angle we can create a good distance measure between the two vectors.
+This is not a good distance measure as -1 is smaller than 0, and therefore two vectors pointing away from each other would be closer than two orthogonal vectors.
+But by subtracting 1 from the cosine of the angle we can create a good distance measure between the two vectors.
 This distance does not take into account any order of magnitude.
@@
 By normalizing the vectors the calculation of the cosine angle becomes the dot product of the two vectors.
-Which can be computed very fast on modern GPU's.
+Which can be computed very fast on modern GPU's, when using matrix multiplication.
 Knowing that we have a measure to compute the similarity of two vectors let us introduce a way to rate the quality of our embeddings.
@@
 \subsubsection{Word similarity and wordsim353}
-To measure the quality of our word embedding we  use a dataset to compare our results too.
+Too measure the quality of our word embedding, in question of word similarity we must use a dataset to compare our self too.
  We chose wordsim353\footnote{http://www.cs.technion.ac.il/~gabr/resources/data/wordsim353/wordsim353.zip} for this task, as it's the most used in the related literature.
@@
 We will rank our embeddings on the Pearson correlation coefficient between the cosine distance and the scores attributed by humans, as this is the common procedure.
+This concludes our description of the evaluation methods of word embeddings.
+The next section will present the configuration of our network used to optimize the SGNS.
 
@@
 
-\subsection{Configuration of the network}
-The skip gram model has a lot of possible parameters, that can be tuned.
+\subsection{Configuration of the network}\label{ssec:config}
+The SGNS has a lot of possible parameters, that can be tuned.
 We experimented with different models and finally decided for one that we tried to optimize.
@@
 The explanation of the parameters will be structured as follows:
-\texttt{Parameter} - Description and tuning - \textit{Value}
+\texttt{Parameter} - Description and Tuning - \textit{Value}
 \begin{itemize}
-\item \texttt{Negative Samples:} Here we have to find a trade-off between, a parameter that is too high (better accuracy) and low ( faster computation time).
+\item \texttt{Negative Samples:} For this parameter we have to find a trade-off between, a parameter that is too high (better accuracy slower computation time) and too low ( worse accuracy faster computation time).
 For smaller data sets a higher number of negative samples is often needed.
@@
 - \textit{10}
-\item \texttt{Context Window:} The bigger the window the more training examples the network will have, but if the window is too big the semantic meaning of the window will be erased.
+\item \texttt{Context Window:} The bigger the window size the more training examples the network will have, but if the window size is too big the semantic meaning of the window size will be erased.
 Mikolov et al.
-\cite{mikolov} proposed a setting between 2-10, as all our sentences are of size 20 we chose 5.
-- \textit{5}
-\item\texttt{ Dimension of the embedding}: Here the choice is less obvious, as the dimension needs to be high enough to capture the meaning, but cannot be too high as this leads to a decrease in performance as shown by Yin and Shen \cite{dimension_size}.
+\cite{mikolov} proposed a value between 2-10  - \textit{5}
+\item\texttt{ Dimension of the embedding}: The choice of the dimension is less obvious, as the dimension needs to be high enough to capture the meaning, but cannot be too high as this leads to a decrease in performance as shown by Yin and Shen \cite{dimension_size}.
 We, therefore, used Gensim to find the best embedding possible.
@@
 
-This conlcudes our overview of the possible parameters of our Skip-Gram Model, in the next section the reader will be given a description about the process of input shuffling.
+This conlcudes our overview of the possible parameters of our SGNS, in the next section the reader will be given a description about the process of input shuffling.
 
@@
 \fi
-\subsection{Input Shuffling}
-We used input shuffling as a technique to optimize the SGM.
-We will first describe input shuffling in a general way and then explain why we supposed that input shuffling could work well on the SGM.
+\subsection{Input Shuffling}\label{ssec:shuffling}
+We used input shuffling as a technique to optimize the SGNS.
+We will first describe input shuffling in a general way and then explain why we supposed that input shuffling could work well on the SGNS.
 \\
-Let $X = {x_1...x_n}$ be our input data set. Input Shuffling describes the process of taking a random permutation of the dataset as an input at each epoch.
-The idea behind this technique is to present our optimizer with different  surfaces of our loss function, in other to find the best optimum.
-Therefore, it's easier for the neural network to escape a local minimum.
-As for example if a network had converged to a local minimum after one epoch it probably cannot escape it.
+
+Let $X = {x_1...x_n}$ be our input data set.
+Input Shuffling describes the process of taking a random permutation of the dataset as an input at each epoch.
+The idea behind this technique is to present our optimizer with different  surfaces of our loss function, in order to find the best optimum.
+Furthermore, it's easier for the neural network to escape a local minimum.
+As for example if a network, without using shuffling,  had converged to a local minimum after one epoch it probably cannot escape it.
 But if we change the shape of the loss function, by input shuffling, then there would be a greater probability for the network to escape the local minimum.
 \\
-There are two reasons why we think that input shuffling is particularly well suited for the skip gram model.
+There are two reasons why we think that input shuffling is particularly well suited for the SGNS.
 Firstly, we read our words sequentially, therefore,  words that only appear very early in the dataset will not benefit from the context words being already updated from others.
@@
 But if we would use input shuffling instead, then the probability of a word appearing multiple time in one batch would be smaller.
-In consequence the accuracy of the model should be optimized leading to a better convergence time.
+In consequence the accuracy of the model should be optimized hence, leading to a better convergence time.
 A definition of convergence time is given in the next section.
@@
 Therefore, we used the already available implementation of Gensim \cite{gensim}.
-Gensim is an open source software that proposes an implementation of the SGNS in Python.
-It is also written in Cython, therefore it has a fast computation time, but can be used inside a python implementation.
-Together with the knowledge from Ji et al.\cite{intel} that a score of $0.66$ in the task of word similarity, with the text8 dataset, is the state of the art, we tested Gensim (more on this process in subsection \ref{sec:discussion} and found out that it took 4 epochs to converge.
+Ji.
+et al.
+defined a score of $0.66$ in the task of word similarity, with the text8 dataset, as the state of the art.
+We therefore could then use Gensim to find out the convergence time of the model, namely 4 epochs.
 Therefore, we defined the following criteria for convergence: \\

### 2019-05-13 icdm/submission/includes/evaluation.tex
summary evaluation reworked

@@
 
-To summarize, we deleted words that appear very often (subsampling) and very infrequently (min count) from our dataset, i.e text8 dataset.
-To evaluate our word embeddings we defined the cosine distance and will measure the quality of our embeddings by comparing theim  to a human judged dataset of word pairs, based on their similarity, i.e wordsim353.
+To summarize the evaluation, we deleted words that appear very often (subsampling) and very infrequently (min count) from our dataset, i.e text8 dataset.
+To evaluate the resulting word embeddings we defined the cosine distance, which is 1 minuse the cosine of the angle of two vectors.
+We will measure the quality of the embeddings by analyzing the correlation of the cosine distance of pairs of vectors to humanly judged similarity score of the corresponding words, and used the wordsim353 to collect these scores.
 The reader was also given an explanation about the configuration of our network and introduced to the concept of input shuffling.

### 2019-05-13 icdm/submission/includes/results.tex
intro sgd mom nag results reworked

@@
 \section{Results}\label{sec:results}
-We ran multiple experiments for each optimizer.
-This Section will  give an overview of the achieved results.
+As stated in the previous section we ran multiple experiments for each optimizer with the above configuration and datasets.
+.
+This Section will  give an overview of the experimental results.
 Each subsection will give an explanation over the achieved result with a specific optimizer.
@@
 As seen in Figure \ref{fig:results_sgd}, for almost every learning rate the convergence time decreased.
-Our model, with the best setting, now converges in only 7 epochs.
+Our model, with the best setting, know converges in only 7 epochs.
 Another interesting fact to point out from Figure \ref{fig:results_sgd} is that with input shuffling we achieved better results with higher learning rates.
@@
 \subsubsection{Momentum and Nesterov}
-Momentum and NAG \cite{nag} both have an additional hyperparameter $\gamma$, that, defines the percentage of the previous gradient that will be added to the current gradients.
+Momentum and  Nesterov Accelerated Gradient (NAG) \cite{nag} both have an additional hyperparameter $\gamma$, that defines the percentage of the previous gradient that will be added to the current gradients.
 We set $\gamma = 0.9$ as this is a typical value and did not alter it during our experiments.
-Momentum and Nesterov alone respectively only slightly decrease or increase the convergence time.
-Momentum optimally converges in 9 epochs and Nesterov in 13.
+Momentum and NAG alone respectively only slightly decrease or increase the convergence time.
+Momentum optimally converges in 9 epochs and NAG in 13.
 If we combine these optimizers with input shuffling, interestingly the same phenomena as with plain SGD appear.

### 2019-05-13 icdm/submission/includes/discussion.tex
discussion our work reworked

@@
 %Discuss the results. What is the outcome of your experimetns?
-This subsection shortly discusses the empirical results and then extensively compares the findings of this work to the existing literature while trying to explain some of the differences.
+This section shortly discusses the empirical results and then extensively compares the findings of this work to the existing literature while trying to explain some of the differences.
 It is concluded by a subsection describing the limitations and possible extensions of this work.
@@
 \subsection{Our work}
-This Section discusses our findings.
+This subsection discusses our findings.
 First it will analyze the possible reasons behind the influence of input shuffling to the learning rate.
@@
 \subsubsection{Shuffling and learning rate with SGD}
-When the model used SGD as an optimizer it was able to use higher learning rates when input shuffling was used, as shown in figures \ref{fig:results_sgd}, \ref{fig:results_mom} and \ref{fig:results_nag}.
-Therefore arises the questions why these phenomena happen, especially as it did not happen with advanced optimizers.
-We attribute this phenomena to the use of the batched version of the SGNS.
-In consequence to this batched version, when input shuffling is not used a lot of words will appear multiple times in the same batch, as explained in \ref{ssec:shuffling}.
+In this subsubsection we will use the term SGD for SGD, Momentum and NAG, and advanced optimzers for Adagrad and Adam.
+The models was able to use a higher learning rates when SGD optimizers where combined with input shuffling in comparison to as when not, as shown in Figures \ref{fig:results_sgd}, \ref{fig:results_mom} and \ref{fig:results_nag}.
+Therefore, arises the questions why these phenomena happen, especially as it did not happen with advanced optimizers.
+\\
+First, we suspect our batched version of the Skip-Gram model with negative sampling (SGNS), to be at the origin of this phenomena.
+In consequence to this batched version, when input shuffling is not used a lot of words will appear multiple times in the same batch, as explained in Section \ref{ssec:shuffling}.
 Therefore, the gradients will be an average of all the training samples.
-When the input is shuffled, less words will appear multiple time which will make the gradients more precise.
-This is counter-attacked by advanced optimizers as they have adaptive learning rates, i.e a frequent appearing feature will have a low learning rate.
-Since high frequency words will appear more often multiple times in a batch, and therefore their average will not influence the results as much.
-The two above arguments could explain why a higher learning rate combined with SGD achieved better results with input shuffling but not when advanced optimizers are used.
+When the input is shuffled, less words will appear multiple time which will make the gradients more precise.\\
+Second, advanced optimizers have a way to counter-attack this issue, namely adaptive learning rates, explained in Section \ref{ssec:results_adagrad}.
+Adaptive learning rates counter attack the issue in the following way: high frequency words  will appear more often multiple times in a batch, they will also have lower learning rates.
+Therefore the impact of appearing in the same batch will be reduced.
+\\  The two above arguments could explain why a higher learning rate combined with SGD achieved better results with input shuffling but not when advanced optimizers are used.
 

### 2019-05-13 icdm/submission/includes/results.tex
discussion our work reworked

@@
 \end{figure}
-\subsubsection{Adagrad}
-Adagrad \cite{adagrad} is a very interesting tool for learning word embeddings as it decreases the learning rate for very frequent occurring features, and vice versa).
-Because words that appear very frequently often do not have a semantic gain that is as important as words that appear less frequently to their context words, it's good to have a lower learning rate for such frequent words.
+\subsubsection{Adagrad}\label{ssec:results_adagrad}
+Adagrad \cite{adagrad} is a very interesting optimizing for creating word embeddings as it decreases the learning rate for very frequent occurring features, and vice versa.
+Words (features in our case) that appear very frequently (low learning rate) often do not have a semantic gain  that is as important as words that appear less frequently (high learning rate) to their context words.
 So, in theory, Adagrad is particularly well suited for our task, as for example Pennington et al.
-used Adagrad in the training of Glove \cite{glove}, another system used to create word embeddings.
-This was confirmed empirically as our model converged in 4 epochs.
+used Adagrad in the training of Glove \cite{glove}, another system which word embeddings.
+Our results confirmed empirically that Adagrad can work better for creating word embeddings.
+The  model converged in 4 epochs.
 When combined with shuffling Adagrad only took 3 epochs to converge.
-This shows the tendency of the skip gram model to converge faster with input shuffling and the big impact of having different learning rate for each feature.
-Here it's interesting to notice that a higher learning rate combined with input shuffling did not yield better results than without shuffling.
-Both of our best results happened with a learning rate of $0.1$, as shown in Figure \ref{fig:results_adagrad_shuffle}.
+These results confirms two hypothesis: first  the tendency of the Skip-Gram Model to converge faster with input shuffling and second  the big impact of having different learning rate for each feature.
+Another interesting fact to notice is that a higher learning rate combined with input shuffling did not yield better results than without shuffling, in comparison to the earlier described optimzers.
+Adagrad performed best with a learning rate of $0.1$, as shown in Figure \ref{fig:results_adagrad_shuffle}.
 \begin{figure}[h]
@@
 In theory Adadelta \cite{adadelta} should outperform Adagrad as it's an extension of the former.
-Because it didn't have any learning rate to tune, we only did 2 experiments, with and without input shuffling.
+Adadelta doesn't  have any learning rate to tune, therefore we only did 2 experiments, with and without input shuffling.
 We are aware of the fact that there are additional hyper parameters to Adadelta.
-We decide not to tune theim for to reasons: first for simplicity reasons and second because their effect is not as high as the learning rate.
-The parameter that defines the percentage taken when calculating the exponentially decaying average of past gradients was set to $\rho = 0.9$.
-Adadelta did not manage to achieve a word similarity of 0.66.
+We decide not to tune them for two reasons: first for simplicity reasons and second because their effect is not as high as the learning rate.
+The hyper parameter, $\rho$ that defines the percentage taken when calculating the exponentially decaying average of past gradients was set to $\rho = 0.9$.
+Adadelta did not manage to achieve a word similarity of 0.66, in comparison to all the others optimzers.
 It only converged to a similarity of 0.59.
@@
 Adam is the most advanced of all the optimizers used in our experiments and did yield the best results as seen in Figure \ref{fig:results_adam_shuffle}.
-Adam converged in 3 epochs without shuffling and 2 with.
-This is the best result that we got with any optimizer.
-It's also interesting to note that as same as with Adagrad it did not react to input shuffling the same way as SGD did.
-In fact, it worked in the opposite direction, as we achieved our best result with input shuffling while having a lower learning rate $0.001$ then we used to achieve the best result without input shuffling $0.05$.
+Adam converged in 3 epochs without shuffling and a learning rate of $0.05$ and 2 with shuffling and a learning rate of $0.001$, as shown in Figure \ref{fig:results_adam_shuffle}.
+Those are the best results achieved in any of our experiment.
+Exacctly as Adagrad, Adamd did not react to input shuffling the same way as SGD did.
 \begin{figure}[h]

### 2019-05-13 icdm/submission/includes/discussion.tex
discussion finished reworked

@@
 \subsubsection{Large differences with NAG and SGD when using shuffling}
-SGD and NAG both have very different values when using shuffling in comparison to unshuffled input, as shown in figures \ref{fig:results_sgd} and \ref{fig:results_nag}, plain SGD.
+SGD and NAG both have very different values when using shuffling in comparison to unshuffled input, as shown in figures \ref{fig:results_sgd} and \ref{fig:results_nag}.
 We do not only attribute those results to input shuffling but partially also to a good random initialization guess.
-Due to a lack of time these results were not replicate more than once.
+Due to a lack of time these experiments were not replicate more than once.
 
 \subsection{Comparison to Gensim}
-This Section will compare our finding extensively to Gensim \cite{gensim}.
+This subsection will compare our finding extensively to Gensim \cite{gensim}.
 As explained in Section \ref{ssec:gensim}, Gensim is optimized to have a very high throughput, this made it possible to achieve a lot of computations.
 Furthermore, Gensim provides access to the loss and the resulting word embeddings, which facilitated the comparison process.
-This Section is Structured as follows:First it will describe the used configuration of Gensim, second it will compare Gensim to our implementation of the SGNS with the use of SGD and finally compare our model with Adam to Gensim.
+This subsection is structured as follows: first it will describe the used configuration of Gensim, second it will compare Gensim to our implementation of the SGNS with the use of SGD and finally compare our model with Adam to Gensim.
 A graphic showing the comparison of our models to Gensim can be found in Figure \ref{fig:gensim_vs_adam}.
-
-\iffalse
-As Mikolov et al.
-published the original paper which introduced the SGNS, it is, of course, relevant to compare ourselves to their work.
-The first important thing to take under account is that Mikolov et al.
-only trained their model on a very large google news dataset incorporating more than 3 billion words.
-This makes the comparison of our work more difficult.
-But we will make some assumptions, as they could be of value if true.
-In their original paper, Mikolov et al.
-reported results from computations that took 1 and 3 epochs.
-We accord these good results, which clearly outperformed our SGD model and Gensim, to the very large dataset and furthermore, as a matter of fact, their results are better with 3 than with 1 epoch.
-We do not have any information about the convergence time or criterion.
-Hence it would be interesting to use their dataset for comparison.
-\\
-One thing we can compare is the quality of our word embeddings.
-Mikolov et al.
-did not report any results on their model with the text8 dataset, but they, therefore, published their code.
-Which was then used by Ji et al \cite{intel} and they tested the model on the text8 dataset.
-They reported a word similarity of 0.63 on the word similarity task.
-This is obviously outperformed by all our models.
-We did not find any explanation on why those results differ as much.
-\\
-The final assumption is that an advanced optimizer could maybe outperform SGD in terms of quality on a large data set. This will be discussed in further work.
-
-\fi
 
@@
 To compare our-self in a correct  manner we used the same dataset, with the same preprocessing parameters, i.e subsampling and min count.
-The hyper parameters of the network (window size, embedding size, number of negative sample, exponent to which the unigram distribution is raised to chose neg samples) are equivalent to ours, described in Section \ref{ssec:config}.
+The hyper parameters of the network (window size, embedding size, number of negative sample, exponent to which the unigram distribution, which decides how a negative sample is used, is raised) are equivalent to our parameters described in Section \ref{ssec:config}.
 The only difference with our model is the learning rate.
@@
 As shown in Figure \ref{fig:gensim_vs_adam} the convergence time was not the same between our implementation and Gensim.
-There are different possible reasons why this could be the case:\\ First, our batched approach could hinder performance in terms of convergence since our loss function is not exactly the same.
-When a word appears more than once in our batch, the gradient will be an average over the gradients of each pair alone, as it is done by Gensim.\\ Secondly, a difference to our implementation is the fact that Gensim checks whether negative samples are not equal to the context word.
+There are different possible reasons why this could be the case:\\ First, our batched approach could hinder performance in terms of convergence time since our loss function is not exactly the same.
+We compute the loss for multiple training samples by taking the sum over each score which is individually used as a loss by Gensim.\\ Secondly, a difference to our implementation is the fact that Gensim checks whether negative samples are not equal to the context word.
 If that is the case Gensim selects a new random sample.
@@
 \\ The first hypothesis, the fact that we used a batched approach, may be confirmed by the fact that when combined with input shuffling SGD does perform closer to Gensim, going from 11 to 7 epochs to converge, as input shuffling reduces the number of co-occurrence of the same word in a batch.
-Now the question arises if the 3 epochs, that Gensim is better, can be explained by the selection of better negative samples and the learning rate decay?
+Know the question arises if the 3 epochs, that Gensim is better, can be explained by the selection of better negative samples and the learning rate decay.
 
@@
 It has been shown that some optimizers perform better with specific shapes of loss functions.
-To make a compelling argument it's, necessary to show that our model with the use of Input shuffling and Adam as its optimizer also outperform Gensim with other data sets, so that the claim can hold consistently.\\
-Finally, our implementation did not outperform Gensim in run time, as this was not the goal of our work.
-Therefore, one could improve an already existing, optimized version, with input shuffling and advanced optimizers and should achieve a better run time than Gensim.
+To make a compelling argument it's necessary to show that our model with the use of input shuffling and Adam as its optimizer also outperforms Gensim with other data sets.\\
+Finally, as it wasn't the goal of our implementation to outperform Gensim in run time, one could improve an already existing, optimized version, with input shuffling and advanced optimizers and should achieve a better run time than Gensim.

### 2019-06-25 documents/preliminary_results.tex
created results document

%% Article Template
 \documentclass{scrartcl}

 %% UTF8 Encoding
 \usepackage[utf8]{inputenc}
 \usepackage[nenglish]{}
 \usepackage[T1]{fontenc}
 \usepackage{lmodern}
 \usepackage{subcaption}
 \usepackage{setspace}
\usepackage{booktabs} % For formal tables`
\usepackage{graphicx}

\title{Preliminary Results}
\author{Cédric Milinaire}
\date{Summer period 2019}
\begin{document}

\begin{titlepage}
\maketitle
\end{titlepage} 
 \begin{abstract}
     This document is used to summarize current results. 
 \end{abstract}
\listoffigures
\listoftables
\newpage
 \begin{abstract}
     This document is used to summarize current results. 
 \end{abstract}
\section{Sampling}
For Table \ref{tab:sampling_stats} the text8 dataset was used, all words that occured less than 5 times were deleted from the dataset before each sampling technique.

\begin{table*}[h]\centering
    \caption{Statistics about dataset with different sampling techniques}
    %\scriptsize
    \begin{tabular}{l r r r r}%
        \toprule
      &    \textbf{w/o sampling} & \textbf{Online Sampling} & \textbf{Mikolov} & \textbf{w/o outliers}  \\%
        \midrule%
        Min     &   34 & 34 & 34 & \\%
       Max    & 9Mio & 0.36 Mio. & 0.36 Mio & \\%
      	QTR1      & 79 & 79 & 79\\%
         Median &166 & 166 & 166 & \\%
         Mean & 2227 & 1125 & 1125 & \\%
       QTR3     & 533 & 533 & 534 & \\%
       QTR3 + 1.5IQR   & 1214 & 1214 & 1217  \\%
        \bottomrule%
   \end{tabular}%
   \label{tab:stats_pairs}%
\end{table*}
\begin{table*}[h]\centering
    \caption{Statistics about dataset with different sampling techniques}
    %\scriptsize
    \begin{tabular}{l r r r r}%
        \toprule
      &    \textbf{w/o sampling} & \textbf{Online Sampling} & \textbf{Mikolov} & \textbf{w/o outliers}  \\%
        \midrule%
        Size of Ds   (in words)     &  17 Mio.& 17 Mio. & 8 Mio.& \\%
        Number of Pairs       & 141 Mio. & 71 Mio. & 71. Mio & \\%
        Number of Sentences         & 0.8 Mio & 0.8 Mio & 0.4 Mio  \\%
        \bottomrule%
   \end{tabular}%
   \label{tab:sampling_stats}%
\end{table*}
\newpage

\section{ Boxplots}
\begin{figure}[h!]
\caption{Distribution of the number of pairs per context word without sampling}
\includegraphics[scale=0.4]{no_sampling_boxplot}
\centering
\end{figure}

\begin{figure}[h!]
\caption{Distribution of the number of pairs per context word with online sampling}
\includegraphics[scale=0.4]{online_sampling_boxplot}
\centering
\end{figure}

\begin{figure}[h!]
\caption{Distribution of the number of pairs per context word with preprocessing sampling}
\includegraphics[scale=0.45]{preprocessing_sampling_boxplot}
\centering
\end{figure}

\newpage
\section{Batch\_size}
This section shows a plot of the batch size per batch.
\begin{figure}[h!]
\caption{Batch size per batch}
\includegraphics[scale=0.65]{batch_sizes}
\centering
\end{figure}
\section{Results}
Figure shows the results of the training of the text8 dataset without outliers. 
\begin{figure}[h!]
\caption{Results text8 without outliers}
\includegraphics[scale=0.65]{text8_without_outliers}
\centering
\end{figure}

\end{document}

### 2019-06-25 documents/preliminary_results.tex
results added plus stats about dataset without outliers

@@
         \midrule%
-        Min     &   34 & 34 & 34 & \\%
-       Max    & 9Mio & 0.36 Mio. & 0.36 Mio & \\%
-      	QTR1      & 79 & 79 & 79\\%
-         Median &166 & 166 & 166 & \\%
-         Mean & 2227 & 1125 & 1125 & \\%
-       QTR3     & 533 & 533 & 534 & \\%
-       QTR3 + 1.5IQR   & 1214 & 1214 & 1217  \\%
+        Min     &   34 & 34 & 34 & 7 \\%
+       Max    & 9 Mio. & 0.36 Mio. & 0.36 Mio & 590 \\%
+      	QTR1      & 79 & 79 & 79&30\\%
+         Median &166 & 166 & 166 & 54 \\%
+       QTR3     & 533 & 533 & 534 & 123\\%
+       QTR3 + 1.5IQR   & 1214 & 1214 & 1217 &227  \\%
                \bottomrule%
+       Mean & 2227 & 1125 & 1125 & 100\\%
    \end{tabular}%
@@
         \midrule%
-        Size of Ds   (in words)     &  17 Mio.& 17 Mio. & 8 Mio.& \\%
-        Number of Pairs       & 141 Mio. & 71 Mio. & 71. Mio & \\%
-        Number of Sentences         & 0.8 Mio & 0.8 Mio & 0.4 Mio  \\%
+        Size of Ds   (in words)     &  17 Mio.& 17 Mio. & 8 Mio.& 2.8 Mio.\\%
+        Number of Pairs       & 141 Mio. & 71 Mio. & 71. Mio & 6 Mio. \\%
+        Number of Sentences         & 0.8 Mio & 0.8 Mio & 0.4 Mio  & 0.1 Mio.\\%
+        Vocabulary Size & 0.25 Mio & 0.25 Mio & 0.25 Mio & 0.06 Mio. \\%
         \bottomrule%
@@
 \newpage
+
+\begin{figure}
+\centering
+\begin{minipage}{.3\textwidth}
+  \centering
+  \includegraphics[scale=0.69]{online_sampling_example}
+  \captionof{figure}{Online Sampling}
+  \label{fig:test1}
+\end{minipage}%
+\begin{minipage}{.3\textwidth}
+  \centering
+  \includegraphics[scale=0.69]{preprocessing_sampling_example}
+  \captionof{figure}{Preprocessing S.}
+  \label{fig:test2}
+\end{minipage}
+\begin{minipage}{.3\textwidth}
+  \centering
+  \includegraphics[scale=0.69]{online_sampling_example}
+  \captionof{figure}{No Sampling}
+  \label{fig:test2}
+\end{minipage}
+\end{figure}
 
@@
 \section{Results}
-Figure shows the results of the training of the text8 dataset without outliers. 
+Figure \ref{fig:results_w/o_outliers} shows the results of the training of the text8 dataset without outliers. Those results are to be taken very lightly for two reasons: 
+\begin{enumerate}
+\item The data set is very small, i.e 60k vocabulary
+\item To assess word similarity only 12 words from the data set were taken. As those were the only one that are in the data set without outliers and in the wordsim dataset. 
+\end{enumerate}
+
+Figure shows that the deletion of the specific word ``the`` does not hinder performance.
+
+\begin{figure}[h!]
+\caption{Results text8 without the word ``the``}
+\includegraphics[scale=0.65]{results_without_the}
+   \label{fig:results_w/o_outliers}%
+
+\centering
+\end{figure}
+
 \begin{figure}[h!]
@@
 \includegraphics[scale=0.65]{text8_without_outliers}
+   \label{fig:results_w/o_outliers}%
+
 \centering

### 2019-08-27 presentation/slides/results.tex
presentation final

\begin{frame}\frametitle{Results}
\begin{Large}
Results
\end{Large}
\begin{itemize}
\item Rating our work
\begin{itemize}
\item Word similarity 
\item Convergence time
\end{itemize}
\item Results
\item Discussion
\begin{itemize}
\item Comparison to Gensim and other related work
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Word similarity}
\begin{Large}
What is word similarity? 
\end{Large}
\medskip
\begin{itemize}
\item Two word embeddings are close to each to other if their cosine distance is small. 
\item Pairs of word rated between 1 and 10 on their similarity, 
\item \includegraphics[scale=0.5]{images/wordsim353_example}
\item We are going to rank our model on the corelation between the distance of the word pairs and the human score.
\end{itemize}
\centerline{ \includegraphics[scale=0.4]{images/cosine}}
\end{frame}

\begin{frame}
\frametitle{Convergence time}
\begin{itemize}
\item Defined convergence time based on word similarity
\item Early Stoppage if: $ \rho -  \rho_{prev} < 0.009 \vee \rho > 0.66$ 
\item No more than 20 epochs. 
\end{itemize}
\includegraphics[scale=0.3]{images/wordsim_convergence}
\end{frame}

### 2019-08-27 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/background.tex
acm sigconf template

--- 2019-05-09 icdm/submission/includes/background.tex
@@
 This section gives in its first subsection a broad overview over the functioning of the Skip-Gram Model, and its second section describes the Technique called Negative Sampling introduced by Mikolov et al.
-\cite{mikolov2} used to make the Skip-Gram model more efficient
+\citep{mikolov2} used to make the Skip-Gram model more efficient
 \subsection{The Skip-Gram Model}
@@
 Mikolov et al.
-\cite{mikolov2} proposed different solutions.
-The first solution is to use a Hierarchical softmax introduced by Morin and Bengio \cite{hsoftmax}.
+\citep{mikolov2} proposed different solutions.
+The first solution is to use a Hierarchical softmax introduced by Morin and Bengio \citep{hsoftmax}.
 In this model, the probability distribution of the output nodes is saved in a binary tree which gives one a logarithmic computation time for each of these probabilities and makes it suitable to compute the softmax.
@@
 A second alternative solution to the use of a classical softmax is negative sampling.
-Negative Sampling is a simplification of Noise Contrastive Estimation (NCE) which was introduced by Gutmann and Hyv{\"a}rinen \cite{nce-original}, and first applied to NLP by Mnih and Teh \cite{mnih}.
+Negative Sampling is a simplification of Noise Contrastive Estimation (NCE) which was introduced by Gutmann and Hyv{\"a}rinen \citep{nce-original}, and first applied to NLP by Mnih and Teh \citep{mnih}.
 This subsection will shortly describe NCE and then how Mikolov et al.
-\cite{mikolov2} simplified it to create the technique called negative sampling.
+\citep{mikolov2} simplified it to create the technique called negative sampling.
 \\ The idea behind NCE is to distinguish data from noise.
@@
 Mikolov et al.
-\cite{mikolov2} simplified NCE and called it Negative Sampling, which will be explained next.\\
+\citep{mikolov2} simplified NCE and called it Negative Sampling, which will be explained next.\\
 The main goal of negative sampling is to only update the output nodes of certain words, 5-20 in practice.
@@
 Mikolov et al.
-\cite{mikolov2} introduced the following objective function $J$, that should be maximized:
+\citep{mikolov2} introduced the following objective function $J$, that should be maximized:
 
@@
 Mikolov et al.
-\cite{mikolov2} used the following unigram distribution, to define the probability $P(w)$ of a word $w$ being drawn as a negative sample:
+\citep{mikolov2} used the following unigram distribution, to define the probability $P(w)$ of a word $w$ being drawn as a negative sample:
 
@@
 Now the question arises if the accuracy is maintained but according to Mikolov et al.
-\cite{mikolov2} the negative sampling method "is an extremely simple training method that learns accurate representations".
+\citep{mikolov2} the negative sampling method "is an extremely simple training method that learns accurate representations".
 As a matter of fact, Mikolov et al.
-\cite{mikolov2} reported a 6\% accuracy improvement in comparison to a Hierarchical Softmax model.
+\citep{mikolov2} reported a 6\% accuracy improvement in comparison to a Hierarchical Softmax model.
 Therefore, it is a good solution to the problem raised by the classical softmax.

### 2019-08-27 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/evaluation.tex
acm sigconf template

--- 2019-05-13 icdm/submission/includes/evaluation.tex
@@
 First of all, it's a small dataset, that allowed us to do a lot of computations.
-Secondly, this dataset was used in related work \cite{intel} hence giving us a very good benchmark.
+Secondly, this dataset was used in related work \citep{intel} hence giving us a very good benchmark.
 The text8 dataset consists of 1702 lines of 1000 words, with a vocabulary of roughly 63000 words.
@@
 Subsampling is a technique introduced by Mikolov et al.
-\cite{mikolov} to reduce the size of the dataset while at the same time increasing the quality of the dataset, i.e increasing the quality of the word embeddings (WE).
+\citep{mikolov} to reduce the size of the dataset while at the same time increasing the quality of the dataset, i.e increasing the quality of the word embeddings (WE).
 The idea behind subsampling is the removal of very frequent words such as: "the, as, it, of".
@@
 The increase in accuracy can also be explained by the fact that words that would not have appeared in the context of each other, may now do because words between have been deleted.
-To choose which word to delete, Mikolov et al. \cite{mikolov2} chose the following equation to compute the probability of a word $w$  to be deleted from the data set:
+To choose which word to delete, Mikolov et al. \citep{mikolov2} chose the following equation to compute the probability of a word $w$  to be deleted from the data set:
 \begin{equation} \label{eq:sampling}
@@
 This idea was also introduced by Mikolove et al.
-\cite{mikolov2}.
+\citep{mikolov2}.
 This is a good technique because of three reasons: First certain words of our data sets do not appear in a common lexicon (twigleg, melqu), or come from a foreign language (Volksvereinigung), or are names and acronyms.
@@
 In their original paper Mikolov et al.
-\cite{mikolov2} recommend a value of 5-20.
+\citep{mikolov2} recommend a value of 5-20.
 We tested a few values in the range of 5 to 15, as 10 yielded state of the art results we chose this value.
@@
 Mikolov et al.
-\cite{mikolov} proposed a value between 2-10  - \textit{5}
-\item\texttt{ Dimension of the embedding}: The choice of the dimension is less obvious, as the dimension needs to be high enough to capture the meaning, but cannot be too high as this leads to a decrease in performance as shown by Yin and Shen \cite{dimension_size}.
+\citep{mikolov} proposed a value between 2-10  - \textit{5}
+\item\texttt{ Dimension of the embedding}: The choice of the dimension is less obvious, as the dimension needs to be high enough to capture the meaning, but cannot be too high as this leads to a decrease in performance as shown by Yin and Shen \citep{dimension_size}.
 We, therefore, used Gensim to find the best embedding possible.
@@
 To optimize convergence time we have to define it first.
-Therefore, we used the already available implementation of Gensim \cite{gensim}.
+Therefore, we used the already available implementation of Gensim \citep{gensim}.
 Ji.

### 2019-08-27 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/introduction.tex
acm sigconf template

--- 2019-05-09 icdm/submission/includes/introduction.tex
@@
 The first attempt to create WE with neural networks was made by Bengio et al.
-\cite{bengio} but more recently Mikolov et al.
-\cite{mikolov} introduced a software package called w2vec, which uses a simpler network that produces state of the art results.
+\citep{bengio} but more recently Mikolov et al.
+\citep{mikolov} introduced a software package called w2vec, which uses a simpler network that produces state of the art results.
 One of the proposed algorithms in this software package is the Skip-Gram Model (SGM).
@@
 Mikolov et al.
-\cite{mikolov2} then introduced an alteration to the SGM called negative sampling.
+\citep{mikolov2} then introduced an alteration to the SGM called negative sampling.
 The goal of the Skip-Gram Model with negative sampling (SGNS) is to distinguish for a given word $w$, context words $c_i$ (words that appear close to $w$ in a sentence) from words drawn randomly (i.e negative samples) $k_i$.
@@
 used Hogwild!
-\cite{hogwild}, proposed by Recht et al., where different threads can access a shared model and update it.
+\citep{hogwild}, proposed by Recht et al., where different threads can access a shared model and update it.
 As this is not an optimal solution Yi et al.
-\cite{intel} tried to optimize it, by using a mini-batch like approach and converting vector to vector multiplications into matrix to matrix multiplications.
+\citep{intel} tried to optimize it, by using a mini-batch like approach and converting vector to vector multiplications into matrix to matrix multiplications.
 This yielded two consequences: First the model is updated less frequently leading to less overwriting and offering the possibility to parallelize more.
 Secondly, it transformed level-1 BLAS operation into level-3 BLAS operations, and the algorithm could therefore effectively use computational resources.
-Another attempt at optimizing the throughput was made by Seulki and Youngmin \cite{gpu}.
+Another attempt at optimizing the throughput was made by Seulki and Youngmin \citep{gpu}.
 Their goal was to parallelize the algorithm on GPU's.

### 2019-08-27 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/related_work.tex
acm sigconf template

--- 2019-05-09 icdm/submission/includes/related_work.tex
@@
 This process does not favor parallelization.
-To deal with this specific problem Mikolov et al.\cite{mikolov2} used a Hogwild!
-approach proposed by Recht et al.\cite{hogwild}.
+To deal with this specific problem Mikolov et al.\citep{mikolov2} used a Hogwild!
+approach proposed by Recht et al.\citep{hogwild}.
 The idea is to allow multiple threads to access a shared memory, in this case, the single model.
@@
 Therefore, overwriting errors are bound to happen.
-According to Recht et al.\cite{hogwild} the overwriting errors won't lead to a significant accuracy loss if the data is sparse enough.
+According to Recht et al.\citep{hogwild} the overwriting errors won't lead to a significant accuracy loss if the data is sparse enough.
 But in the case of NLP, the problem seems to be a bit more significant, and especially for word embedding, as many words share the same context words.
@@
 \subsection{Parellelization by the use of caching}
-This idea was proposed by Vuurens et al. \cite{efficient}. The architecture used here is the SGM with a hierarchical softmax. 
+This idea was proposed by Vuurens et al. \citep{efficient}. The architecture used here is the SGM with a hierarchical softmax. 
 The general idea is to cache the most frequently used weight vectors, and update the model after a certain number of seen training samples, while taking in consideration the changes that happend during these number of seen training samples (the paper used the number 10).
@@
 It seems to indicate that too much overwriting was happening, as the number of concurrent threads surpasses a certain threshold.
-This can be seen in Figure \ref{fig:efficient}, where c31 is the model proposed by Vuurens et al.\cite{efficient}.
+This can be seen in Figure \ref{fig:efficient}, where c31 is the model proposed by Vuurens et al.\citep{efficient}.
 The model did not suffer any accuracy loss in comparison to the original SGM model.
@@
 \includegraphics[scale=0.3]{images/cachingEfficiency.png}
-\caption{Comparasion of the execution time in relation to the number of used cores \cite{efficient}}
+\caption{Comparasion of the execution time in relation to the number of used cores \citep{efficient}}
 \label{fig:efficient}
@@
 In contrast to the above approach, this one, introduced by Ji et al.
-\cite{intel} focused on the SGNS.
+\citep{intel} focused on the SGNS.
 The focus of this work was to speed up the computation of the loss function of the SGNS described in Equation \ref{eq:obj_neg_samples}.
@@
 Ji et al.'s method achieves a 3.6 fold increase in throughput, by only losing 1\ % of accuracy.
-The presented results were achieved on CPU's, as many modern machine learning libraries make use of the parallelizing capabilities of GPU's for matrix multiplications, the next subsection covers the work done by Seulki and Youngmin \cite{gpu} that parellelized the SGNS on GPU's.
+The presented results were achieved on CPU's, as many modern machine learning libraries make use of the parallelizing capabilities of GPU's for matrix multiplications, the next subsection covers the work done by Seulki and Youngmin \citep{gpu} that parellelized the SGNS on GPU's.
 
 \subsection{Accelleration of word2vec by using GPU's}
-Seulki and Youngmin \cite{gpu} focused on getting a better throughput on the SGM with the use of GPU's.
+Seulki and Youngmin \citep{gpu} focused on getting a better throughput on the SGM with the use of GPU's.
 As the SGM is a sequential algorithm, it is not easy to parallelize it, especially if one wants to parallelize the training of individual training samples.
@@
 Consequently, it's very hard to parallelize at this level.
-Since the dimensions of the word embeddings are completly independent of each other  Seulki and Youngmin \cite{gpu} tried to parallelize their update.
+Since the dimensions of the word embeddings are completly independent of each other  Seulki and Youngmin \citep{gpu} tried to parallelize their update.
 They achieved this by mapping each dimension to a CUDA thread while mapping each sentence to a CUDA block.
@@
 If the execution time of the GPU kernel is greater than the time used to read the sentences, it could be a smart choice to use multiple GPU's.
-According to Seulki and Youngmin \cite{gpu}, if multiple GPU's are used, there is a need for synchronizing the model, which will hinder run time performance.
+According to Seulki and Youngmin \citep{gpu}, if multiple GPU's are used, there is a need for synchronizing the model, which will hinder run time performance.
 They achieved their best results with 2 concurrent GPU'S.
 The acomplished results were very good as they managed a 20x speedup compared to a single threaded CPU execution, which is a 3x increase in comparison to the original C code from Mikolov et al.
-\cite{mikolov2}.
+\citep{mikolov2}.
 The speedup was achieved, while maintaining the accuracy of the model.
@@
 Therefore, we need an optimized implementation of the SGM that is easily available.
-This is provided by Gensim \cite{gensim}, which will be outlined in the next subsection.
+This is provided by Gensim \citep{gensim}, which will be outlined in the next subsection.
 %%%%%%%%%%%
 \subsection{Gensim}\label{ssec:gensim}
-Gensim \cite{gensim} is a pure Python library that holds a state of the art implementation of the SGM.
+Gensim \citep{gensim} is a pure Python library that holds a state of the art implementation of the SGM.
 Gensim is written in Cython, which first allowed Gensim to have the same runtime as the original C code.
@@
 This finally yielded in a 4x speedup in run time compared to the orignal C code from Mikolov et al.
-Gensim as a python library, and because it was already used in related work \cite{intel} made the comparison easy and of value.
+Gensim as a python library, and because it was already used in related work \citep{intel} made the comparison easy and of value.
 \\ This concludes our overview of  related work.

### 2019-08-27 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/results.tex
acm sigconf template

--- 2019-05-13 icdm/submission/includes/results.tex
@@
 The first challenge for each optimizer was to find a correct learning rate.
-As SGD is the optimizer used in Gensim \cite{gensim} we first tried the same learning rate as the default value in Gensim \cite{gensim}, i.e 0.01,  and then performed a random search to find a better one.
+As SGD is the optimizer used in Gensim \citep{gensim} we first tried the same learning rate as the default value in Gensim \citep{gensim}, i.e 0.01,  and then performed a random search to find a better one.
 As expected a bell curve shape resulted, a learning rate that is too high leads to diversion and a learning rate that is too low leads to a training time that is too slow.
@@
 \subsubsection{Momentum and Nesterov}
-Momentum and  Nesterov Accelerated Gradient (NAG) \cite{nag} both have an additional hyperparameter $\gamma$, that defines the percentage of the previous gradient that will be added to the current gradients.
+Momentum and  Nesterov Accelerated Gradient (NAG) \citep{nag} both have an additional hyperparameter $\gamma$, that defines the percentage of the previous gradient that will be added to the current gradients.
 We set $\gamma = 0.9$ as this is a typical value and did not alter it during our experiments.
@@
 \subsubsection{Adagrad}\label{ssec:results_adagrad}
-Adagrad \cite{adagrad} is a very interesting optimizing for creating word embeddings as it decreases the learning rate for very frequent occurring features, and vice versa.
+Adagrad \citep{adagrad} is a very interesting optimizing for creating word embeddings as it decreases the learning rate for very frequent occurring features, and vice versa.
 Words (features in our case) that appear very frequently (low learning rate) often do not have a semantic gain  that is as important as words that appear less frequently (high learning rate) to their context words.
 So, in theory, Adagrad is particularly well suited for our task, as for example Pennington et al.
-used Adagrad in the training of Glove \cite{glove}, another system which word embeddings.
+used Adagrad in the training of Glove \citep{glove}, another system which word embeddings.
 Our results confirmed empirically that Adagrad can work better for creating word embeddings.
@@
 \subsubsection{Adadelta}
-In theory Adadelta \cite{adadelta} should outperform Adagrad as it's an extension of the former.
+In theory Adadelta \citep{adadelta} should outperform Adagrad as it's an extension of the former.
 Adadelta doesn't  have any learning rate to tune, therefore we only did 2 experiments, with and without input shuffling.

### 2019-08-27 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/sample-sigconf.tex
acm sigconf template

\documentclass[sigconf]{acmart}

\usepackage{booktabs} % For formal tables
\usepackage{csquotes}
\usepackage{makecell} % For line breaks in tables
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}

% Copyright
%\setcopyright{none}
\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\acmDOI{xx.xxx/xxx_x}

% ISBN
\acmISBN{978-1-4503-6866-7/20/03}

%Conference
\acmConference[SAC'20]{ACM SAC Conference}{March 30-April 3, 2020}{Brno, Czech Republic} 
\acmYear{2020}
\copyrightyear{2020}

\acmArticle{4}
\acmPrice{15.00}

% These commands are optional
%\acmBooktitle{Transactions of the ACM Woodstock conference}
%\editor{Jennifer B. Sartor}
%\editor{Theo D'Hondt}
%\editor{Wolfgang De Meuter}

\begin{document}
\title{SIG Proceedings Paper in LaTeX Format}
\titlenote{Produces the permission block, and
  copyright information}
\subtitle{Extended Abstract}
\subtitlenote{The full version of the author's guide is available as
  \texttt{acmart.pdf} document}

\author{Ben Trovato}
\authornote{Dr.~Trovato insisted his name be first.}
\orcid{1234-5678-9012}
\affiliation{%
  \institution{Institute for Clarity in Documentation}
  \streetaddress{P.O. Box 1212}
  \city{Dublin} 
  \state{Ohio} 
  \postcode{43017-6221}
}
\email{trovato@corporation.com}

\author{G.K.M. Tobin}
\authornote{The secretary disavows any knowledge of this author's actions.}
\affiliation{%
  \institution{Institute for Clarity in Documentation}
  \streetaddress{P.O. Box 1212}
  \city{Dublin} 
  \state{Ohio} 
  \postcode{43017-6221}
}
\email{webmaster@marysville-ohio.com}

\author{Lars Th{\o}rv{\"a}ld}
\authornote{This author is the
  one who did all the really hard work.}
\affiliation{%
  \institution{The Th{\o}rv{\"a}ld Group}
  \streetaddress{1 Th{\o}rv{\"a}ld Circle}
  \city{Hekla} 
  \country{Iceland}}
\email{larst@affiliation.org}

\author{Valerie B\'eranger}
\affiliation{%
  \institution{Inria Paris-Rocquencourt}
  \city{Rocquencourt}
  \country{France}
}
\author{Aparna Patel} 
\affiliation{%
 \institution{Rajiv Gandhi University}
 \streetaddress{Rono-Hills}
 \city{Doimukh} 
 \state{Arunachal Pradesh}
 \country{India}}
\author{Huifen Chan}
\affiliation{%
  \institution{Tsinghua University}
  \streetaddress{30 Shuangqing Rd}
  \city{Haidian Qu} 
  \state{Beijing Shi}
  \country{China}
}

\author{Charles Palmer}
\affiliation{%
  \institution{Palmer Research Laboratories}
  \streetaddress{8600 Datapoint Drive}
  \city{San Antonio}
  \state{Texas} 
  \postcode{78229}}
\email{cpalmer@prl.com}

\author{John Smith}
\affiliation{\institution{The Th{\o}rv{\"a}ld Group}}
\email{jsmith@affiliation.org}

\author{Julius P.~Kumquat}
\affiliation{\institution{The Kumquat Consortium}}
\email{jpkumquat@consortium.net}

% The default list of authors is too long for headers}
\renewcommand{\shortauthors}{B. Trovato et al.}

\begin{abstract}
The Skip-Gram Model with negative sampling (SGNS) is a simple and effective algorithm to create word embeddings. SGNS uses Stochastic Gradient Descent (SGD) as it's learning algorithm. While a lot of effort has gone into increasing the throughput of words of the SGNS, not much work has gone into optimizing the convergence time. Therefore, our work focuses on the latter. To achieve a better convergence time we use advanced optimizers and input shuffling as optimization techniques. Here are the components used to evaluate our work: to train our model we used the text8 dataset - to measure the quality of our word embeddings we used the wordsim353 dataset, which assess that similar words have vector represenations that are close to each other. We trained our model with multiple advanced optimizers: Momentum, Nesterov Accelerated Gradient, Adagrad, Adadelta, and Adam. We also applied input shuffling to the above optimizers. Adam combined with input shuffling outperformed every optimizer. Adam with shuffling also outperformed the current state of the art implementation Gensim. Adam converged to a similarity value of 0.66 (state of the art) in 2 epochs, while Gensim took 4 epochs. Hence, this work shows that advanced optimizers combined with input shuffling do decrease the convergence time of the SGNS. These results must be confirmed with other datasets, but optimizing our model in terms of throughput has the potential of further reducing the overall run time of the SGNS.

\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>  
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Embedded systems}
\ccsdesc[300]{Computer systems organization~Redundancy}
\ccsdesc{Computer systems organization~Robotics}
\ccsdesc[100]{Networks~Network reliability}

\keywords{ACM proceedings, \LaTeX, text tagging}

\maketitle

\input{includes/introduction}
\input{includes/background}
\input{includes/related_work}
\input{includes/contribution}
\input{includes/evaluation}
\input{includes/results}
\input{includes/discussion}
\input{includes/conclusion}

\bibliographystyle{ACM-Reference-Format}
\bibliography{sample-bibliography}

\end{document}

### 2019-08-27 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/background.tex
icdm work to acm te mplate

@@
 Raising the unigram distribution to the power of $\frac{3}{4}$ makes it less likely for a word to be drawn if it often appears in the dataset in comparison to the basic unigram distribution.
-See figure \ref{fig:frequency_ex} for an example.\\
+\iffalse See figure \ref{fig:frequency_ex} for an example.\\
 \begin{figure}[ht]
@@
 \end{figure}
+\fi
 With all of the above knowledge it's understandable that the SGNS will  will outperform the classical softmax in computation time.

### 2019-08-27 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/evaluation.tex
icdm work to acm te mplate

@@
 Stats about subsampling can be found in Table \ref{table:threshold}.
-
+\iffalse
 \begin{table}[tb]
@@
 \end{table*}
+\fi
 
@@
 Therefore, to ensure the quality of word embeddings other techniques have to be used.To define quality we first need to define a measure of similarity between two vectors.
-This requires knowledge of the Cosine distance, which is introduced in the following subsection.
-\subsubsection{Cosine distance}
-The cosine similarity, this is not the cosine distance, of vectors $v$ and $w$ is the cosine of the angle between the two vectors.
-It can be calculated by taking the dot product of $v$ and $w$ and dividing it by the magnitude of $v$ and $w$ multiplied with each other.
-We get:
-\begin{equation}
-cos\_sim(v,w) = \frac{v \cdot w}{|v| |w|}
-\end{equation}
-The cosine of 0\textdegree ~is 1, it's 0 for two vectors that are orthogonal to each other and vectors that point in the opposite direction will have a cosine of their angle of -1.
-This is not a good distance measure as -1 is smaller than 0, and therefore two vectors pointing away from each other would be closer than two orthogonal vectors.
-But by subtracting 1 from the cosine of the angle we can create a good distance measure between the two vectors.
-This distance does not take into account any order of magnitude.
-Hence, for our tasks, two vectors will be considered equal if they are of different magnitude but point in the same direction.
-Apart from measuring the quality of word embedding well, this technique has another advantage.
-By normalizing the vectors the calculation of the cosine angle becomes the dot product of the two vectors.
-Which can be computed very fast on modern GPU's, when using matrix multiplication.
-Knowing that we have a measure to compute the similarity of two vectors let us introduce a way to rate the quality of our embeddings.
+For this task we will use the cosine distance.
+Two vectors will be judged similiar if they point in the same direction.
 
@@
 
-\subsection{Convergence time}
-To optimize convergence time we have to define it first.
-Therefore, we used the already available implementation of Gensim \citep{gensim}.
-Ji.
-et al.
-defined a score of $0.66$ in the task of word similarity, with the text8 dataset, as the state of the art.
-We therefore could then use Gensim to find out the convergence time of the model, namely 4 epochs.
-Therefore, we defined the following criteria for convergence: \\
-\centerline{$\rho - \rho_{prev} < 0.009 \vee \rho = 0.66$} \\
-where $\rho$ is the Pearson coefficient on the wordsim353 task.
-We also stopped computation, if it took more then 20 epochs to converge.
-
 To summarize the evaluation, we deleted words that appear very often (subsampling) and very infrequently (min count) from our dataset, i.e text8 dataset.

### 2019-08-27 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/evaluation.tex
big cutting

@@
 Therefore, we had to choose between building arbitrary sentences and keeping the dataset as it is.
-We chose the first option because it gives us a faster computation time, and did not show any significant loss in quality empirically, as shown in Table \ref{table:with_20}.
+We chose the first option because it gives us a faster computation time, and did not show any significant loss in quality empirically.
 We chose a length of sentences of 20.
 Furthermore, we applied a technique called subsampling to reduce the data set size, which is explained in the following subsection.
-
+\iffalse
 \begin{table}[tb]
@@
 \end{table}
+\fi
 
@@
 Subsampling is a technique introduced by Mikolov et al.
-\citep{mikolov} to reduce the size of the dataset while at the same time increasing the quality of the dataset, i.e increasing the quality of the word embeddings (WE).
+\citep{mikolov} to reduce the size of the dataset while at the same increasing the quality of the word embeddings (WE).
 The idea behind subsampling is the removal of very frequent words such as: "the, as, it, of".
@@
 The increase in accuracy can also be explained by the fact that words that would not have appeared in the context of each other, may now do because words between have been deleted.
+Details about the probability of the deletion of a word can be found in the orignial paper from Mikolov \citep{mikolov2}.
+\iffalse
 To choose which word to delete, Mikolov et al. \citep{mikolov2} chose the following equation to compute the probability of a word $w$  to be deleted from the data set:
@@
 Stats about subsampling can be found in Table \ref{table:threshold}.
-\iffalse
 \begin{table}[tb]
@@
 This is a good technique because of three reasons: First certain words of our data sets do not appear in a common lexicon (twigleg, melqu), or come from a foreign language (Volksvereinigung), or are names and acronyms.
-Secondly, each document often has spelling mistakes, those (as long as the same spelling mistake does not appear too often, what should be avoided in practice) would be deleted by sampling too, as the words do not have any meaning.
+Secondly, each document often has spelling mistakes, those would be deleted by sampling too, as the words do not have any meaning.
 Lastly, a word that only appears one time in our dataset will be very dependent on its original initialization.
-The weights of the network that represents the word, will only be updated when the word itself is used as input.
-As the word appears very infrequently the word representation will not differ much from the initialization, hence being very dependent from it.
 For all the above reasons, we applied this technique.
@@
 The next section will present the configuration of our network used to optimize the SGNS.
-
+\iffalse
 \begin{table}[tb]\centering
@@
 We experimented with different models and finally decided for one that we tried to optimize.
-This subsection will give a short overview of each parameter, where we will explain the process in which we chose the value of the given parameter.
-The explanation of the parameters will be structured as follows:
-\texttt{Parameter} - Description and Tuning - \textit{Value}
+The network was configured the following way:
 \begin{itemize}
-\item \texttt{Negative Samples:} For this parameter we have to find a trade-off between, a parameter that is too high (better accuracy slower computation time) and too low ( worse accuracy faster computation time).
-For smaller data sets a higher number of negative samples is often needed.
-In their original paper Mikolov et al.
-\citep{mikolov2} recommend a value of 5-20.
-We tested a few values in the range of 5 to 15, as 10 yielded state of the art results we chose this value.
-- \textit{10}
-\item \texttt{Context Window:} The bigger the window size the more training examples the network will have, but if the window size is too big the semantic meaning of the window size will be erased.
-Mikolov et al.
-\citep{mikolov} proposed a value between 2-10  - \textit{5}
-\item\texttt{ Dimension of the embedding}: The choice of the dimension is less obvious, as the dimension needs to be high enough to capture the meaning, but cannot be too high as this leads to a decrease in performance as shown by Yin and Shen \citep{dimension_size}.
-We, therefore, used Gensim to find the best embedding possible.
-- \textit{100}
-\item \texttt{Batch size}: As described in Section \ref{sec:contribution}, there is a trade off to find between quality and training time.
-We first used a batch size of 5000, but then decide after non conclusive results that 2000 would be better - \textit{2000}
-\item \texttt{Alpha}: learning rate, this hyperparameter was tuned in every optimizer therefore only the range will be indicated - \textit{(1e-5,1)}
+\item \texttt{Negative Samples:} - \textit{10}
+\item \texttt{Context Window:}  - \textit{5}
+\item\texttt{Dimension of the embedding}: - \textit{100}
+\item \texttt{Batch size}: - \textit{2000}
+\item \texttt{Alpha}:  - \textit{(1e-5,1)}
 \end{itemize}
-
-This conlcudes our overview of the possible parameters of our SGNS, in the next section the reader will be given a description about the process of input shuffling.
+This concludes our overview of the possible parameters of our SGNS, in the next section the reader will be given a description about the process of input shuffling.
 

### 2019-08-27 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/related_work.tex
big cutting

@@
 It seems to indicate that too much overwriting was happening, as the number of concurrent threads surpasses a certain threshold.
-This can be seen in Figure \ref{fig:efficient}, where c31 is the model proposed by Vuurens et al.\citep{efficient}.
 The model did not suffer any accuracy loss in comparison to the original SGM model.
+\iffalse
 \begin{figure}[ht]
@@
 \end{figure}
+\fi
+
 This work proposes a very good way to parallelize the SGM, as in particular, it allows using more cores during the computation.

### 2019-08-27 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/contribution.tex
start word freq distribution

--- 2019-05-09 icdm/submission/includes/contribution.tex
@@
 As some words may appear more than once in the batch this will more be an average of it than as with Mikolov et al.
-\cite{mikolov2} the exact update per pair.
+\citep{mikolov2} the exact update per pair.
 \\
@@
 This finishes the description of our batched approach. The next section will describe our experimental setting and the possibilities that exists to evaluate the quality of word embeddings.
+
+\subsection{Distribution of word frequencies}
+To tune the batch size and create batches of better quality we chose to analyze the distribution of word frequencies.
+A box plot of the distribution, on the text8 dataset, can be found in Figure \ref{fig:boxplot_freq}.
+\begin{figure}[h]
+\centering
+\includegraphics[scale=0.3]{images/boxplot_freq}
+\caption{Distribution of word frequencies in the text8 datasedt}
+\label{fig:boxplot_freq}
+\end{figure} 
+This shows that very few words are responsible for the majority of the context pairs.
+For example the word \textit{the} creates 50\% of the pairs in the text8 dataset.
+Therefore if we create batches, the word the will appear multiple time and the learning will be hindered.

### 2019-08-28 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/contribution.tex
started including word distribution

@@
-\section{Batched SkipGramModel} \label{sec:contribution}
+\section{Batched Skip-Gram Model} \label{sec:contribution}
 This section will give an overview of our implementation.
@@
 \subsection{Distribution of word frequencies}
+\subsubsection{Boxplot}
 To tune the batch size and create batches of better quality we chose to analyze the distribution of word frequencies.
@@
 \centering
-\includegraphics[scale=0.3]{images/boxplot_freq}
+\includegraphics[scale=0.3]{images/no_sampling_boxplot}
 \caption{Distribution of word frequencies in the text8 datasedt}
@@
 For example the word \textit{the} creates 50\% of the pairs in the text8 dataset.
-Therefore if we create batches, the word the will appear multiple time and the learning will be hindered.
+Therefore if we create arbitrary batches, most words  will appear multiple times in a batch and the learning will be hindered.
+Therefore we propose a solution where each word will only appear once in a batch.
+\subsubsection{Batches}
+To solve this problem we allowed each batch to have a different size.
+Each batch will contain the most possible pairs where each center word does not appear twice.
+The problem with this approach is that the batch size decreases very rapidly.
+This is due to the distribution of words.
+In consequence the average batch size is small and training time is increased, to remedy this problem we tried to reduce the number of batches, while maintaining the quality of learning.
+\subsubsection{Deletion of outliers}
+The idea is to delete every word which is an outlier, according to the whiskers plot.
+This means every word which frequency is bigger than the frequency of the word that marks the boundary of the 3/4 percentile +1.5 IQR.
+This technique drastically reduces the dataset size,  and is for some datasets, too drastic.
+We therefore experimented with different frequencies, that could be a good treshhold.

### 2019-08-28 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/contribution.tex
contribution reworked

--- 2019-05-09 icdm/submission/includes/contribution.tex
@@
 
-This finishes the description of our batched approach. The next section will describe our experimental setting and the possibilities that exists to evaluate the quality of word embeddings.
+This finishes the description of our batched approach.
+
+\subsection{Distribution of word frequencies}
+\subsubsection{Boxplot}
+The quality of a batch is dependent on  the number of multiple appearances of words in the batch.
+The more frequently a word appears in a batch, the training quality of the word decreases.
+For this matter we chose to analyze the distribution of word frequencies.
+A box plot of the distribution, on the text8 dataset, can be found in Figure \ref{fig:boxplot_freq}.
+\begin{figure}[h]
+\centering
+\includegraphics[scale=0.3]{images/no_sampling_boxplot}
+\caption{Distribution of word frequencies in the text8 datasedt}
+\label{fig:boxplot_freq}
+\end{figure} 
+This shows that very few words appear very frequently, and are therefore responsible for the building of the most context pairs.
+For example the word \textit{the} creates 50\% of the pairs in the text8 dataset.
+In consequence if we create arbitrary batches, most words  will appear multiple times in a batch and the learning will be hindered.
+A solution would be to only allow words to appear once in a batch, in the next section we present such a solution with a variable batch size.
+\subsubsection{Batches}
+To solve this problem we only chose to block multiple appearance in a batch on the center words.
+The context words can be the same for two different center words.
+Each batch will contain the highest possible number of pairs where each center word does not appear twice.
+The problem with this approach is that the batch size decreases very rapidly.
+This is due to the distribution of words.
+In consequence the average batch size is small and training time is increased, to remedy this problem we tried to reduce the number of very small batches, while maintaining the quality of learning.
+\subsubsection{Deletion of outliers}
+To solve this matter, we looked again at the distribution of word frequencies (Figure \ref{fig:boxplot_freq}).
+We experimented with different settings (mean, median, average, outlier) and found that deleting all words that had a frequency greater than the 99\% was best.
+This technique drastically reduces the dataset size.
+
+The next section will describe our experimental setting and the possibilities that exists to evaluate the quality of word embeddings.

### 2019-08-28 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/results.tex
big cutting
started including word distribution
results first draft

@@
 
+\subsection{Simple Batch Version}
 \subsubsection{SGD}
@@
 \end{figure}
+\iffalse 
 \subsubsection{Momentum and Nesterov}
@@
 \end{figure}
+
 \subsubsection{Adagrad}\label{ssec:results_adagrad}
@@
 \end{figure}
+
+\subsection{Advanced Batches Version}
+The advanced batched version is largely dependent on the definition of outlier.
+As shown in Table if the value for outliers is too small then the model convergence into score that is too small.
+If the threshold is higher training time is increased.
+We found that 99\% is a good reference value.
+With this setting our model converged in 1 epoch for the text8 dataset, which is our best result, and 7 for the enwik9 corpus.
+A recap of all our results with the text8 dataset can be found in Table.
+
+\begin{table}[tb]
+    \caption{Results on the text8 dataset}
+    \begin{tabular}{l r r }
+        \toprule
+Model    & Convergence Time & Word Similarity \\ \hline
+SGD & {11}              & 0.65            \\ 
+SGD w/shuffling & {7}              & 0.66            \\ 
+Adam & {3}              & 0.66            \\ 
+Adam w/ shuffling & {2}              & 0.66      \\ 
+Gensim   & 4          & 0.66            \\ \hline
+        %
+\textbf{Adam w/o outliers }& \textbf{1} & 0.66 \\  \hline
+\bottomrule
+\end{tabular}
+\end{table}

### 2019-08-28 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/introduction.tex
introcution w/ word freq

@@
 Representing words as vectors, i.e word embeddings (WE) is a fundamental aspect of Natural Language Processing (NLP).
-There are two ways to create such WE, either arbitrarily or with the purpose of capturing the semantic meaning of the words, i.e.
+There are two ways to create such WE, either arbitrarily or with the purpose of capturing the semantic  of the words, i.e.
 vector representations of words that are syntactically or semantically similar will be close to each other in the vector space.
@@
 Most of this effort was trying to improve the throughput of the model, i.e the number of words that are processed per second.
-The SGM uses Stochastic Gradient Descent as an optimization algorithm and is therefore inherently sequential.
-To remedy this problem Mikolov et al.
-used Hogwild!
-\citep{hogwild}, proposed by Recht et al., where different threads can access a shared model and update it.
-As this is not an optimal solution Yi et al.
-\citep{intel} tried to optimize it, by using a mini-batch like approach and converting vector to vector multiplications into matrix to matrix multiplications.
-This yielded two consequences: First the model is updated less frequently leading to less overwriting and offering the possibility to parallelize more.
-Secondly, it transformed level-1 BLAS operation into level-3 BLAS operations, and the algorithm could therefore effectively use computational resources.
-Another attempt at optimizing the throughput was made by Seulki and Youngmin \citep{gpu}.
-Their goal was to parallelize the algorithm on GPU's.
-They, therefore, chose to parallelize the update of the dimensions of each word representation.
+As remarkable progress was made \citep{intel} , \citep{gpu}.
 Both of these approaches and most of the literature are focused on improving the throughput of the model, but not the convergence time.
 Therefore one could ask if the convergence time of the SGM can be optimized while at the same time maintaining its accuracy.\\
-This work focused on improving the convergence time of the SGNS.
+Our work focused on improving the convergence time of the SGNS.
 To facilitate our calculations we propose a slightly altered version of the original SNGS, in which we compute the loss for a very large amount of training samples, i.e 2000.
-The loss of such a batch was computed by taking the sum over the loss all individual training samples in the specific batch We combined this batched approach with advanced optimizers and input shuffling to decrease he convergence time of the SGNS.
-In combination advanced optimizers and input shuffling allowed us to decrease the convergence time of the SGNS.
+The loss of such a batch was computed by taking the sum over the loss of all individual training samples in the specific batch.
+We combined this batched approach with advanced optimizers and input shuffling to decrease the convergence time of the SGNS.
+Furthermore we analyzed the distribution of word frequencies of different datasets to create batches of better quality.
+As a matter of fact a word that appear multiple times in a batch will hinder the quality of the batch.
+Therefore we focused our work on creating batches that only contained words once.
+In combination advanced optimizers, input shuffling and our batched approach allowed us to decrease the convergence time of the SGNS.
 If combined with an optimized throughput these results could lead to an overall decrease in runtime.\\

### 2019-08-28 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/discussion.tex
acm sigconf template
big cutting
discussion added improvements word freq

--- 2019-05-13 icdm/submission/includes/discussion.tex
@@
 
+\iffalse
 \subsubsection{Large differences with NAG and SGD when using shuffling}
@@
 Due to a lack of time these experiments were not replicate more than once.
+\fi
 
 \subsection{Comparison to Gensim}
-This subsection will compare our finding extensively to Gensim \cite{gensim}.
+This subsection will compare our finding extensively to Gensim \citep{gensim}.
 As explained in Section \ref{ssec:gensim}, Gensim is optimized to have a very high throughput, this made it possible to achieve a lot of computations.
@@
 
+\subsection{The quality of batches can be optimized}
+As shown in Table \ref{table:results_text8}, the creation of batches where each word only appears once in a batch outperforms the simple batch version.
+This shows empiracly that our suspission that words that appear multiple time in a batch hinder performance.
+Furthermore the deletion of outliers, speeds up training without hindering performance.
+A reason could be that words that appear very often have mostly the same context words and therefore will have a very similar representation as they learn from the same words.
+
 \subsection{Future Work}
-This work shows that the convergence time of the SGNS could be improved by using input shuffling and advanced optimizers.
+This work shows that the convergence time of the SGNS could be improved by using input shuffling, advanced optimizers and a batched version with specific rules to the creation of the batches.
 As with every work, there still exists possible extensions.

### 2019-08-28 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/conclusion.tex
conclusion w/ freq words

--- 2019-05-07 icdm/submission/includes/conclusion.tex
@@
 This work provides an overview of the Skip Gram Model with negative Sampling (SGNS) and the numerous successful attempts of optimizing the throughput of the model.
-As this is the case, no effort went into optimizng the convergence time of the SGNS, therefore this work focused on this task.
-We decided to use advanced optimizers and input shuffling to achieve a better convergence time.This work proposes a slighlty altered version of the SGNS, where the idea is to compute the loss over the sum of a high number of training samples, i.e 2000,  instead of computing it for each individually.
-We did this as it allowed us to compute more models and analyze the convergence time faster.
+As this is the case, no effort went into optimizng the convergence time of the SGNS, therefore this work focused on the convergence time.
+We decided to use advanced optimizers, input shuffling and a batched approach with specific rules for the batch creation to achieve a better convergence time.
+We propose a batched version of the SGNS, where the idea is to compute the loss over the sum of a high number of training samples,  instead of computing it for each individually.
+The quality of each individual batch is hindered by words appearing multiple time in a batch, we therefore decided to create batches where this would not be the case.
+Furthermore we  decided to cut the dataset size by deleting each word that appears more often than the 99 percentile, this technique maintained the quality of the dataset, while drastically improving the computation time.
 We used the text8 dataset to train our model and  word similarity as a quality measure for the word embeddings (WE).
-We used the sate of the art implementation Gensim to compare our self.
+To compare our work, Gensim was used, which holds a state of the art implementation of the SGNS.
 We did achieve a better convergence time than Gensim with Adam as an optimizer and the use of input shuffling.
 Gensim convereged in 4 epochs to a word similarity of 0.66 and our model only took 2 epochs to achieve the same quality.
+The latter model was improved by deleting each  word that appears more often than the 99 percentile and using batches where words only appear once.
 Those results still need to be confirmed with more datasets.

### 2019-08-28 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/sample-sigconf.tex
absatract w/ word freq

@@
 \subtitle{Extended Abstract}
-\subtitlenote{The full version of the author's guide is available as
-  \texttt{acmart.pdf} document}
 
 \author{Ben Trovato}
-\authornote{Dr.~Trovato insisted his name be first.}
 \orcid{1234-5678-9012}
@@
 \author{G.K.M. Tobin}
-\authornote{The secretary disavows any knowledge of this author's actions.}
+
 \affiliation{%
@@
 \author{Lars Th{\o}rv{\"a}ld}
-\authornote{This author is the
-  one who did all the really hard work.}
 \affiliation{%
@@
 \email{larst@affiliation.org}
-
-\author{Valerie B\'eranger}
-\affiliation{%
-  \institution{Inria Paris-Rocquencourt}
-  \city{Rocquencourt}
-  \country{France}
-}
-\author{Aparna Patel} 
-\affiliation{%
- \institution{Rajiv Gandhi University}
- \streetaddress{Rono-Hills}
- \city{Doimukh} 
- \state{Arunachal Pradesh}
- \country{India}}
-\author{Huifen Chan}
-\affiliation{%
-  \institution{Tsinghua University}
-  \streetaddress{30 Shuangqing Rd}
-  \city{Haidian Qu} 
-  \state{Beijing Shi}
-  \country{China}
-}
-
-\author{Charles Palmer}
-\affiliation{%
-  \institution{Palmer Research Laboratories}
-  \streetaddress{8600 Datapoint Drive}
-  \city{San Antonio}
-  \state{Texas} 
-  \postcode{78229}}
-\email{cpalmer@prl.com}
-
-\author{John Smith}
-\affiliation{\institution{The Th{\o}rv{\"a}ld Group}}
-\email{jsmith@affiliation.org}
-
-\author{Julius P.~Kumquat}
-\affiliation{\institution{The Kumquat Consortium}}
-\email{jpkumquat@consortium.net}
 
@@
 Therefore, our work focuses on the latter.
-To achieve a better convergence time we use advanced optimizers and input shuffling as optimization techniques.
+We propose a new batched version of the SGNS.
+.
 Here are the components used to evaluate our work: to train our model we used the text8 dataset - to measure the quality of our word embeddings we used the wordsim353 dataset, which assess that similar words have vector represenations that are close to each other.
-We trained our model with multiple advanced optimizers: Momentum, Nesterov Accelerated Gradient, Adagrad, Adadelta, and Adam.
-We also applied input shuffling to the above optimizers.
-Adam combined with input shuffling outperformed every optimizer.
-Adam with shuffling also outperformed the current state of the art implementation Gensim.
-Adam converged to a similarity value of 0.66 (state of the art) in 2 epochs, while Gensim took 4 epochs.
-Hence, this work shows that advanced optimizers combined with input shuffling do decrease the convergence time of the SGNS.
+To improve the quality of our batched version, we only allowed each word to appear once in a batch.
+Furthermore we deleted every word that appear more frequently than the 99 percentile.
+The advanced optimizer Adam combined with our batched approach outperformed the current state of the art implementation Gensim.
+Adam converged to a similarity value of 0.66 (state of the art) in 1 epochs, while Gensim took 4 epochs.
+Hence, this work shows that our batched version with advanced optimizers, and the deletion of outliers do decrease the convergence time of the SGNS.
 These results must be confirmed with other datasets, but optimizing our model in terms of throughput has the potential of further reducing the overall run time of the SGNS.

### 2019-08-28 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/sample-sigconf.tex
abstract reworked

@@
 The Skip-Gram Model with negative sampling (SGNS) is a simple and effective algorithm to create word embeddings.
-SGNS uses Stochastic Gradient Descent (SGD) as it's learning algorithm.
+SGNS uses Stochastic Gradient Descent (SGD) as it's optimizer.
 While a lot of effort has gone into increasing the throughput of words of the SGNS, not much work has gone into optimizing the convergence time.
@@
 We propose a new batched version of the SGNS.
-.
+To improve the quality of our batched version, we only allowed each word to appear once in a batch.
+Furthermore we deleted every word that appear more frequently than the 99 percentile in our datasets.
+Finally we also used advanced optimizers and input shuffling to decrease the convergence time.
 Here are the components used to evaluate our work: to train our model we used the text8 dataset - to measure the quality of our word embeddings we used the wordsim353 dataset, which assess that similar words have vector represenations that are close to each other.
-To improve the quality of our batched version, we only allowed each word to appear once in a batch.
-Furthermore we deleted every word that appear more frequently than the 99 percentile.
 The advanced optimizer Adam combined with our batched approach outperformed the current state of the art implementation Gensim.
-Adam converged to a similarity value of 0.66 (state of the art) in 1 epochs, while Gensim took 4 epochs.
+Adam converged to a similarity value of 0.66 (state of the art) in 1 epoch, while Gensim took 4 epochs.
 Hence, this work shows that our batched version with advanced optimizers, and the deletion of outliers do decrease the convergence time of the SGNS.

### 2019-08-28 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/contribution.tex
contribution reworked

@@
 
-\subsection{Distribution of word frequencies}
-\subsubsection{Boxplot}
-The quality of a batch is dependent on  the number of multiple appearances of words in the batch.
-The more frequently a word appears in a batch, the training quality of the word decreases.
-For this matter we chose to analyze the distribution of word frequencies.
-A box plot of the distribution, on the text8 dataset, can be found in Figure \ref{fig:boxplot_freq}.
+\subsection{Optimization of the quality of batches}
+There are two main ways to create a batch.
+Either arbitrary or with specific rules.
+Both of these approaches have benefits, the first one do not need any preprocessing  and the second should in theory provide better learning.
+In our case the quality of a batch is dependent on  the number of multiple appearances of words in the batch.
+The more frequently a word appears in a batch, the more the quality of the word embedding diminishes.
+Therefore the next sections will go into details on how we constructed our batches.
+
+\subsubsection{Analyzing word frequencies}
+. For this matter we chose to analyze the distribution of word frequencies. A box plot of the distribution, on the text8 dataset, can be found in Figure \ref{fig:boxplot_freq}. 
 \begin{figure}[h]
@@
 In consequence if we create arbitrary batches, most words  will appear multiple times in a batch and the learning will be hindered.
-A solution would be to only allow words to appear once in a batch, in the next section we present such a solution with a variable batch size.
-\subsubsection{Batches}
-To solve this problem we only chose to block multiple appearance in a batch on the center words.
+A solution would be to only allow words to appear once in a batch, in the next subsection we present such a solution with a variable batch size.
+\subsubsection{Criteria for batches}
+To solve the above problem we only chose to block multiple appearance in a batch on the center words.
 The context words can be the same for two different center words.
 Each batch will contain the highest possible number of pairs where each center word does not appear twice.
-The problem with this approach is that the batch size decreases very rapidly.
-This is due to the distribution of words.
-In consequence the average batch size is small and training time is increased, to remedy this problem we tried to reduce the number of very small batches, while maintaining the quality of learning.
+As with all batches, each pair will only appear once in a batch.
+Therefore the batch size will be very high at the beginning of the training and will then decrease very rapidly.
+In consequence the average batch size is small and training time is increased, in comparison to arbitrary created batches.
+To remedy this problem we tried to reduce the number of very small batches by altering the dataset, while maintaining the quality of learning.
+We call it, deletion of outliers and explain it in the next subsection.
 \subsubsection{Deletion of outliers}
-To solve this matter, we looked again at the distribution of word frequencies (Figure \ref{fig:boxplot_freq}).
-We experimented with different settings (mean, median, average, outlier) and found that deleting all words that had a frequency greater than the 99\% was best.
-This technique drastically reduces the dataset size.
+To solve the above matter, we looked  at the distribution of word frequencies (Figure \ref{fig:boxplot_freq}).
+We experimented with different settings (mean, median, average, outlier) and found that deleting all words that had a frequency greater than the 99 percentile was best.
+This technique drastically reduces the data set size, it also deletes words that appear very frequently, it therefore reduces the number of very small batches.
 

### 2019-08-28 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/evaluation.tex
evaluation reworked

@@
 Therefore, the goal of subsampling is to delete such words from the dataset.
-Subsampling will decrease the computation time, as it will reduce the number of training samples, and should, in theory, increase the accuracy of the model.
-The increase in accuracy can also be explained by the fact that words that would not have appeared in the context of each other, may now do because words between have been deleted.
+Subsampling will decrease the computation time, as it will reduce the number of training samples, and increases the accuracy of the model.
 Details about the probability of the deletion of a word can be found in the orignial paper from Mikolov \citep{mikolov2}.
@@
 \subsubsection{Min count}
-We also deleted every word that did not appear more than 5 times in our dataset.
-This idea was also introduced by Mikolove et al.
-\citep{mikolov2}.
-This is a good technique because of three reasons: First certain words of our data sets do not appear in a common lexicon (twigleg, melqu), or come from a foreign language (Volksvereinigung), or are names and acronyms.
-Secondly, each document often has spelling mistakes, those would be deleted by sampling too, as the words do not have any meaning.
+Another idea introduced by Mikolove et al.
+\citep{mikolov2} is to delete every word that appear less than 5 times and is a good technique because of three reasons: First certain words of our data sets do not appear in a common lexicon (twigleg, melqu), or are names and acronyms.
+Secondly, each document often has spelling mistakes.
 Lastly, a word that only appears one time in our dataset will be very dependent on its original initialization.
@@
 As the task that the network is learning is of no interest to us.
-Therefore, to ensure the quality of word embeddings other techniques have to be used.To define quality we first need to define a measure of similarity between two vectors.
+Therefore, to ensure the quality of word embeddings other techniques have to be used.
+To define quality we first need to define a measure of similarity between two vectors.
 For this task we will use the cosine distance.
 Two vectors will be judged similiar if they point in the same direction.
+Too measure the quality of our word embedding, in question of word similarity we must use a dataset to compare our self too.
+The  next section describes the chosen dataset.
 
-\subsubsection{Word similarity and wordsim353}
-Too measure the quality of our word embedding, in question of word similarity we must use a dataset to compare our self too.
+\subsubsection{wordsim353}
  We chose wordsim353\footnote{http://www.cs.technion.ac.il/~gabr/resources/data/wordsim353/wordsim353.zip} for this task, as it's the most used in the related literature.
@@
 \end{itemize}
-This concludes our overview of the possible parameters of our SGNS, in the next section the reader will be given a description about the process of input shuffling.
-
+This concludes our overview of the possible parameters of our model of the SGNS. \\
 \iffalse
@@
 \fi
-\subsection{Input Shuffling}\label{ssec:shuffling}
-We used input shuffling as a technique to optimize the SGNS.
-We will first describe input shuffling in a general way and then explain why we supposed that input shuffling could work well on the SGNS.
-\\
-
-Let $X = {x_1...x_n}$ be our input data set.
-Input Shuffling describes the process of taking a random permutation of the dataset as an input at each epoch.
-The idea behind this technique is to present our optimizer with different  surfaces of our loss function, in order to find the best optimum.
-Furthermore, it's easier for the neural network to escape a local minimum.
-As for example if a network, without using shuffling,  had converged to a local minimum after one epoch it probably cannot escape it.
-But if we change the shape of the loss function, by input shuffling, then there would be a greater probability for the network to escape the local minimum.
-\\
-There are two reasons why we think that input shuffling is particularly well suited for the SGNS.
-Firstly, we read our words sequentially, therefore,  words that only appear very early in the dataset will not benefit from the context words being already updated from others.
-The second idea is that we used the special batch technique described in Section \ref{sec:contribution}.
-When we do not use input shuffling the same word will appear with all of its context words as training samples in a batch, therefore the updae will be an average of all of these training samples.
-But if we would use input shuffling instead, then the probability of a word appearing multiple time in one batch would be smaller.
-In consequence the accuracy of the model should be optimized hence, leading to a better convergence time.
-A definition of convergence time is given in the next section.
-
 To summarize the evaluation, we deleted words that appear very often (subsampling) and very infrequently (min count) from our dataset, i.e text8 dataset.

### 2019-08-30 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/contribution.tex
structure change w/ freq words

@@
 This shows that very few words appear very frequently, and are therefore responsible for the building of the most context pairs.
-For example the word \textit{the} creates 50\% of the pairs in the text8 dataset.
+For example the word \textit{the} creates 50\% of the pairs in the text8\footnote{See Section \ref{sec:evalutaion}} dataset.
 In consequence if we create arbitrary batches, most words  will appear multiple times in a batch and the learning will be hindered.
@@
 Therefore the batch size will be very high at the beginning of the training and will then decrease very rapidly.
-In consequence the average batch size is small and training time is increased, in comparison to arbitrary created batches.
+In consequence the average batch size is small and training time is increased, in comparison to arbitrary created batches where the batch size is fixed.
 To remedy this problem we tried to reduce the number of very small batches by altering the dataset, while maintaining the quality of learning.
@@
 We experimented with different settings (mean, median, average, outlier) and found that deleting all words that had a frequency greater than the 99 percentile was best.
-This technique drastically reduces the data set size, it also deletes words that appear very frequently, it therefore reduces the number of very small batches.
+This technique drastically reduces the data set size,by deleting words that appear very frequently, it therefore reduces the number of very small batches.
 

### 2019-08-30 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/results.tex
discussion added improvements word freq
structure change w/ freq words

@@
 \section{Results}\label{sec:results}
-As stated in the previous section we ran multiple experiments for each optimizer with the above configuration and datasets.
-.
-This Section will  give an overview of the experimental results.
-Each subsection will give an explanation over the achieved result with a specific optimizer.
+As stated in the previous section we ran multiple experiments for each optimizer with the above configuration and datasets . This Section will  give an overview of the experimental results. 
+The Section is structured as follows: The first part will describe the results achieved by our batched version with advanced optimizers and input shuffling the second part will focus on batches with improved quality and the deletion of outliers.
 
 \subsection{Simple Batch Version}
+This subsection will give an explanation over the achieved results with different optimizers, combined with input shuffling on our batched version of the Skip-Gram Model with negative sampling (SGNS).
+It will first explain the result achieved with Stochastic Gradient Descent (SGD) followed by the result achieved with adam as optimizers.
 \subsubsection{SGD}
@@
 
+This conlcudes the results achieved with arbitrary batches.
+The following section will cover the results achieved when creating batches that only contains center words once and how the deletion of outliers affect the results.
+
 \subsection{Advanced Batches Version}
@@
 \bottomrule
-\end{tabular}
+\end{tabular}\label{table:results_text8}
 \end{table}

### 2019-08-30 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/discussion.tex
final changes draft paper

@@
 When the input is shuffled, less words will appear multiple time which will make the gradients more precise.\\
-Second, advanced optimizers have a way to counter-attack this issue, namely adaptive learning rates, explained in Section \ref{ssec:results_adagrad}.
+Second, advanced optimizers have a way to counter-attack this issue, namely adaptive learning rates.
 Adaptive learning rates counter attack the issue in the following way: high frequency words  will appear more often multiple times in a batch, they will also have lower learning rates.
@@
 \iffalse
-\subsubsection{Large differences with NAG and SGD when using shuffling}
+\subsubsection{Large differences with SGD when using shuffling}
 SGD and NAG both have very different values when using shuffling in comparison to unshuffled input, as shown in figures \ref{fig:results_sgd} and \ref{fig:results_nag}.

### 2019-08-30 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/results.tex
final changes draft paper

@@
 \subsection{Advanced Batches Version}
-The advanced batched version is largely dependent on the definition of outlier.
-As shown in Table if the value for outliers is too small then the model convergence into score that is too small.
-If the threshold is higher training time is increased.
-We found that 99\% is a good reference value.
-With this setting our model converged in 1 epoch for the text8 dataset, which is our best result, and 7 for the enwik9 corpus.
-A recap of all our results with the text8 dataset can be found in Table.
+The advanced batched version in combination with the deletion of outliers and Adam as it's optimizer greatly outperformed the other versions.
+With this setting our model converged in 1 epoch for the text8 dataset, which is our best result.
+A recap of all our results with the text8 dataset can be found in Table \ref{table:results_text8}.
 

### 2019-09-27 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/conclusion.tex
changed back to paper without word freqs

--- 2019-05-07 icdm/submission/includes/conclusion.tex
@@
 As this is the case, no effort went into optimizng the convergence time of the SGNS, therefore this work focused on this task.
-We decided to use advanced optimizers and input shuffling to achieve a better convergence time.This work proposes a slighlty altered version of the SGNS, where the idea is to compute the loss over the sum of a high number of training samples, i.e 2000,  instead of computing it for each individually.
-We did this as it allowed us to compute more models and analyze the convergence time faster.
+We decided to use advanced optimizers and input shuffling to achieve a better convergence time.
+This work proposes a slighlty altered version of the SGNS, where the idea is to compute the loss over the sum of a high number of training samples, i.e 2000,  instead of computing it for each individually.
+The batched version allowed us to compute more models and analyze the convergence time faster.
 We used the text8 dataset to train our model and  word similarity as a quality measure for the word embeddings (WE).
-We used the sate of the art implementation Gensim to compare our self.
+To compare our work, Gensim was used, which holds a state of the art implementation of the SGNS.
 We did achieve a better convergence time than Gensim with Adam as an optimizer and the use of input shuffling.

### 2019-09-27 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/contribution.tex
changed back to paper without word freqs

--- 2019-08-28 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/contribution.tex
@@
-\section{Batched Skip-Gram Model} \label{sec:contribution}
+\section{Batched SkipGramModel} \label{sec:contribution}
 This section will give an overview of our implementation.
@@
 
-This finishes the description of our batched approach.
-
-\subsection{Distribution of word frequencies}
-\subsubsection{Boxplot}
-The quality of a batch is dependent on  the number of multiple appearances of words in the batch.
-The more frequently a word appears in a batch, the training quality of the word decreases.
-For this matter we chose to analyze the distribution of word frequencies.
-A box plot of the distribution, on the text8 dataset, can be found in Figure \ref{fig:boxplot_freq}.
-\begin{figure}[h]
-\centering
-\includegraphics[scale=0.3]{images/no_sampling_boxplot}
-\caption{Distribution of word frequencies in the text8 datasedt}
-\label{fig:boxplot_freq}
-\end{figure} 
-This shows that very few words appear very frequently, and are therefore responsible for the building of the most context pairs.
-For example the word \textit{the} creates 50\% of the pairs in the text8 dataset.
-In consequence if we create arbitrary batches, most words  will appear multiple times in a batch and the learning will be hindered.
-A solution would be to only allow words to appear once in a batch, in the next section we present such a solution with a variable batch size.
-\subsubsection{Batches}
-To solve this problem we only chose to block multiple appearance in a batch on the center words.
-The context words can be the same for two different center words.
-Each batch will contain the highest possible number of pairs where each center word does not appear twice.
-The problem with this approach is that the batch size decreases very rapidly.
-This is due to the distribution of words.
-In consequence the average batch size is small and training time is increased, to remedy this problem we tried to reduce the number of very small batches, while maintaining the quality of learning.
-\subsubsection{Deletion of outliers}
-To solve this matter, we looked again at the distribution of word frequencies (Figure \ref{fig:boxplot_freq}).
-We experimented with different settings (mean, median, average, outlier) and found that deleting all words that had a frequency greater than the 99\% was best.
-This technique drastically reduces the dataset size.
-
-The next section will describe our experimental setting and the possibilities that exists to evaluate the quality of word embeddings.
+This finishes the description of our batched approach. The next section will describe our experimental setting and the possibilities that exists to evaluate the quality of word embeddings.

### 2019-09-28 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/background.tex
bib modifiactions plus background and introduction review

@@
 \fi
-This section gives in its first subsection a broad overview over the functioning of the Skip-Gram Model, and its second section describes the Technique called Negative Sampling introduced by Mikolov et al.
-\citep{mikolov2} used to make the Skip-Gram model more efficient
+This section gives in its first subsection a broad overview over the functioning of the Skip-Gram Model, and its second section describes the technique called Negative Sampling introduced by Mikolov et al.
+\citep{mikolov2} used to make the Skip-Gram model more efficient.
 \subsection{The Skip-Gram Model}
@@
 This will obviously save an enormous amount of computation time.
-The idea is that given a pair $(c,w) \in D$, wh$w$ ere $c$ is a word in the context window of we select $K$ random words $k_i$ from the corpus data $D$.
+The idea is that given a pair $(c,w) \in D$, where $w$  $c$ is a word in the context window of $c$ we select $K$ random words $k_i$ from the corpus data $D$.
 We assume those words do not appear in the context of $w$.
@@
 Assume each output neuron has a weight vector of dimension $d = 300$.
-In consequence, when updating our weights we would only update $0.2*10^{-2}$ of the 300 million weights in the output layer.
+In consequence, when updating our weights we would only update $0.9*10^{3}$ of the 300 million weights in the output layer.
 
@@
 \fi
-With all of the above knowledge it's understandable that the SGNS will  will outperform the classical softmax in computation time.
+With all of the above knowledge it's understandable that the SGNS will  outperform the classical softmax in computation time.
 Now the question arises if the accuracy is maintained but according to Mikolov et al.

### 2019-09-28 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/results.tex
bib modifiactions plus background and introduction review

--- 2019-08-28 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/results.tex
@@
 
-\subsection{Simple Batch Version}
 \subsubsection{SGD}
@@
 The second experiment was to add input shuffling.
-As seen in Figure \ref{fig:results_sgd}, for almost every learning rate the convergence time decreased.
+For almost every learning rate the convergence time decreased.
 Our model, with the best setting, know converges in only 7 epochs.
-Another interesting fact to point out from Figure \ref{fig:results_sgd} is that with input shuffling we achieved better results with higher learning rates.
+Another interesting fact to point out is that with input shuffling we achieved better results with higher learning rates.
 As for learning rates of $0.01$ and $0.025$ we did converge in 11 epochs with input shuffling but did not converge in 20 epochs without it.
-
+\iffalse
 \begin{figure}[h]
@@
 \end{figure}
+\fi
 \iffalse 
@@
 
-\subsubsection{Adagrad}\label{ssec:results_adagrad}
-Adagrad \citep{adagrad} is a very interesting optimizing for creating word embeddings as it decreases the learning rate for very frequent occurring features, and vice versa.
-Words (features in our case) that appear very frequently (low learning rate) often do not have a semantic gain  that is as important as words that appear less frequently (high learning rate) to their context words.
-So, in theory, Adagrad is particularly well suited for our task, as for example Pennington et al.
-used Adagrad in the training of Glove \citep{glove}, another system which word embeddings.
-Our results confirmed empirically that Adagrad can work better for creating word embeddings.
-The  model converged in 4 epochs.
-When combined with shuffling Adagrad only took 3 epochs to converge.
-These results confirms two hypothesis: first  the tendency of the Skip-Gram Model to converge faster with input shuffling and second  the big impact of having different learning rate for each feature.
-Another interesting fact to notice is that a higher learning rate combined with input shuffling did not yield better results than without shuffling, in comparison to the earlier described optimzers.
-Adagrad performed best with a learning rate of $0.1$, as shown in Figure \ref{fig:results_adagrad_shuffle}.
-\begin{figure}[h]
-\centering
-\includegraphics[scale=0.3]{images/results_adagrad_shuffle}
-\caption{Training time Adagrad with input Shuffling}
-\label{fig:results_adagrad_shuffle}
-\end{figure}
 \subsubsection{Adadelta}
@@
 
+\subsubsection{Adagrad}\label{ssec:results_adagrad}
+Adagrad \citep{adagrad} is a very interesting optimizer for creating word embeddings as it decreases the learning rate for very frequent occurring features, and vice versa.
+Words (features in our case) that appear very frequently (low learning rate) often do not have a semantic gain  that is as important as words that appear less frequently (high learning rate) to their context words.
+So, in theory, Adagrad is particularly well suited for our task, as for example Pennington et al.
+used Adagrad in the training of Glove \citep{glove}, another system which word embeddings.
+Our results confirmed empirically that Adagrad can work better for creating word embeddings.
+The  model converged in 4 epochs.
+When combined with shuffling Adagrad only took 3 epochs to converge, as shown in Table \ref{table:results}.
+These results confirms two hypothesis: first  the tendency of the Skip-Gram Model to converge faster with input shuffling and second  the big impact of having different learning rate for each feature.
+Another interesting fact to notice is that a higher learning rate combined with input shuffling did not yield better results than without shuffling, in comparison to the earlier described optimzers.
+Adagrad performed best with a learning rate of $0.1$.
+\iffalse
+\begin{figure}[h]
+\centering
+\includegraphics[scale=0.3]{images/results_adagrad_shuffle}
+\caption{Training time Adagrad with input Shuffling}
+\label{fig:results_adagrad_shuffle}
+\end{figure}
+\fi
 \subsubsection{Adam}
-Adam is the most advanced of all the optimizers used in our experiments and did yield the best results as seen in Figure \ref{fig:results_adam_shuffle}.
-Adam converged in 3 epochs without shuffling and a learning rate of $0.05$ and 2 with shuffling and a learning rate of $0.001$, as shown in Figure \ref{fig:results_adam_shuffle}.
+Adam is the most advanced of all the optimizers used in our experiments and did yield the best results as seen in F.
+Adam converged in 3 epochs without shuffling and a learning rate of $0.05$ and 2 with shuffling and a learning rate of $0.001$, as shown in Table \ref{table:results}.
 Those are the best results achieved in any of our experiment.
 Exacctly as Adagrad, Adamd did not react to input shuffling the same way as SGD did.
+    \begin{table}[]
+\begin{tabular}{|l|l|l|}
+\hline
+Model    & Convergence Time & Word Similarity \\ \hline
+SGD & {11}              & 0.65            \\ \hline
+SGD w/shuffling & {7}              & 0.66            \\ \hline
+Adam & {3}              & 0.66            \\ \hline
+Adam w/ shuffling & \textbf{2}              & 0.66      \\ \hline
+\end{tabular}
+\label{table:results}
+\caption{Results on the text8 dataset with different optimizers and input shuffling}
+
+\end{table}
+\iffalse
 \begin{figure}[h]
@@
 \end{figure}
-
-\subsection{Advanced Batches Version}
-The advanced batched version is largely dependent on the definition of outlier.
-As shown in Table if the value for outliers is too small then the model convergence into score that is too small.
-If the threshold is higher training time is increased.
-We found that 99\% is a good reference value.
-With this setting our model converged in 1 epoch for the text8 dataset, which is our best result, and 7 for the enwik9 corpus.
-A recap of all our results with the text8 dataset can be found in Table.
-
-\begin{table}[tb]
-    \caption{Results on the text8 dataset}
-    \begin{tabular}{l r r }
-        \toprule
-Model    & Convergence Time & Word Similarity \\ \hline
-SGD & {11}              & 0.65            \\ 
-SGD w/shuffling & {7}              & 0.66            \\ 
-Adam & {3}              & 0.66            \\ 
-Adam w/ shuffling & {2}              & 0.66      \\ 
-Gensim   & 4          & 0.66            \\ \hline
-        %
-\textbf{Adam w/o outliers }& \textbf{1} & 0.66 \\  \hline
-\bottomrule
-\end{tabular}
-\end{table}
+\fi

### 2019-09-28 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/sample-sigconf.tex
structure change w/ freq words
changed back to paper without word freqs
bib modifiactions plus background and introduction review

--- 2019-08-27 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/sample-sigconf.tex
@@
 
-\author{Valerie B\'eranger}
-\affiliation{%
-  \institution{Inria Paris-Rocquencourt}
-  \city{Rocquencourt}
-  \country{France}
-}
-\author{Aparna Patel} 
-\affiliation{%
- \institution{Rajiv Gandhi University}
- \streetaddress{Rono-Hills}
- \city{Doimukh} 
- \state{Arunachal Pradesh}
- \country{India}}
-\author{Huifen Chan}
-\affiliation{%
-  \institution{Tsinghua University}
-  \streetaddress{30 Shuangqing Rd}
-  \city{Haidian Qu} 
-  \state{Beijing Shi}
-  \country{China}
-}
-
-\author{Charles Palmer}
-\affiliation{%
-  \institution{Palmer Research Laboratories}
-  \streetaddress{8600 Datapoint Drive}
-  \city{San Antonio}
-  \state{Texas} 
-  \postcode{78229}}
-\email{cpalmer@prl.com}
-
-\author{John Smith}
-\affiliation{\institution{The Th{\o}rv{\"a}ld Group}}
-\email{jsmith@affiliation.org}
-
-\author{Julius P.~Kumquat}
-\affiliation{\institution{The Kumquat Consortium}}
-\email{jpkumquat@consortium.net}
-
 % The default list of authors is too long for headers}
@@
 The Skip-Gram Model with negative sampling (SGNS) is a simple and effective algorithm to create word embeddings.
-SGNS uses Stochastic Gradient Descent (SGD) as it's learning algorithm.
+The SGNS uses Stochastic Gradient Descent (SGD) as it's learning algorithm.
 While a lot of effort has gone into increasing the throughput of words of the SGNS, not much work has gone into optimizing the convergence time.

### 2019-09-28 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/evaluation.tex
review

--- 2019-08-27 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/evaluation.tex
@@
 As the task that the network is learning is of no interest to us.
-Therefore, to ensure the quality of word embeddings other techniques have to be used.To define quality we first need to define a measure of similarity between two vectors.
+Therefore, to ensure the quality of word embeddings other techniques have to be used.
+To define quality we first need to define a measure of similarity between two vectors.
 For this task we will use the cosine distance.
@@
 \subsubsection{Word similarity and wordsim353}
-Too measure the quality of our word embedding, in question of word similarity we must use a dataset to compare our self too.
+To measure the quality of our word embedding, in question of word similarity we must use a dataset to compare our self too.
  We chose wordsim353\footnote{http://www.cs.technion.ac.il/~gabr/resources/data/wordsim353/wordsim353.zip} for this task, as it's the most used in the related literature.
@@
 The next section will present the configuration of our network used to optimize the SGNS.
-\iffalse
 \begin{table}[tb]\centering
@@
 \end{table}
-\fi
 
@@
 \end{itemize}
-This concludes our overview of the possible parameters of our SGNS, in the next section the reader will be given a description about the process of input shuffling.
+This concludes our overview of the possible parameters of our SGNS, in the next subsection the reader will be given a description about the process of input shuffling.
 
@@
 But if we would use input shuffling instead, then the probability of a word appearing multiple time in one batch would be smaller.
-In consequence the accuracy of the model should be optimized hence, leading to a better convergence time.
-A definition of convergence time is given in the next section.
+In consequence the accuracy of the model should be optimized, hence leading to a better convergence time.
 
 To summarize the evaluation, we deleted words that appear very often (subsampling) and very infrequently (min count) from our dataset, i.e text8 dataset.
-To evaluate the resulting word embeddings we defined the cosine distance, which is 1 minuse the cosine of the angle of two vectors.
+To evaluate the resulting word embeddings we defined the cosine distance, which is 1 minus the cosine of the angle of two vectors.
 We will measure the quality of the embeddings by analyzing the correlation of the cosine distance of pairs of vectors to humanly judged similarity score of the corresponding words, and used the wordsim353 to collect these scores.

### 2019-09-28 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/related_work.tex
review

@@
 The idea is to allow multiple threads to access a shared memory, in this case, the single model.
-In the original SGM, the threads are constructed as follows: at the beginning of training, the dataset is split into $N$ evenly sized chunks and each of these chunks will be processed by a one thread.
+In the original SGM, the threads are constructed as follows: at the beginning of training, the dataset is split into $N$ evenly sized chunks and each of these chunks will be processed by one thread.
 The threads run parallel and have access to the shared memory.
@@
 The focus of this work was to speed up the computation of the loss function of the SGNS described in Equation \ref{eq:obj_neg_samples}.
-Where the dot product of the center word  with negative samples as well as the context word is computed.
+In this Equation the dot product of the center word  with negative samples as well as the context word is computed.
 The dot product is a level 1-BLAS operation and is not as efficiently implemented into hardware, as matrix multiplication which are a level 3-BLAS operation.
@@
 The above Matrix multiplication will compute the loss for each training sample and the weights of the used vectors will be updated after each batch.
-Ji et al.'s method achieves a 3.6 fold increase in throughput, by only losing 1\ % of accuracy.
+Ji et al.'s method achieves a 3.6 fold increase in throughput, by only losing 1\% of accuracy.
 The presented results were achieved on CPU's, as many modern machine learning libraries make use of the parallelizing capabilities of GPU's for matrix multiplications, the next subsection covers the work done by Seulki and Youngmin \citep{gpu} that parellelized the SGNS on GPU's.
@@
 Since the dimensions of the word embeddings are completly independent of each other  Seulki and Youngmin \citep{gpu} tried to parallelize their update.
-They achieved this by mapping each dimension to a CUDA thread while mapping each sentence to a CUDA block.
+They achieved this by mapping each dimension of a weight vector represinting a word embedding(WE) to a CUDA thread while mapping each sentence to a CUDA block.
 As each CUDA block runs independently, the training of the sentences is parallelized, and the fact that sentences have different length is of no problem.

### 2019-09-28 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/discussion.tex
final tex work

--- 2019-05-13 icdm/submission/includes/discussion.tex
@@
 \subsubsection{Shuffling and learning rate with SGD}
-In this subsubsection we will use the term SGD for SGD, Momentum and NAG, and advanced optimzers for Adagrad and Adam.
-The models was able to use a higher learning rates when SGD optimizers where combined with input shuffling in comparison to as when not, as shown in Figures \ref{fig:results_sgd}, \ref{fig:results_mom} and \ref{fig:results_nag}.
+In this subsubsection we will use the term advanced optimzers for Adagrad and Adam.
+The models was able to use a higher learning rates when SGD was combined with input shuffling in comparison to as when not.
 Therefore, arises the questions why these phenomena happen, especially as it did not happen with advanced optimizers.
@@
 There are different possible reasons why this could be the case:\\ First, our batched approach could hinder performance in terms of convergence time since our loss function is not exactly the same.
-We compute the loss for multiple training samples by taking the sum over each score which is individually used as a loss by Gensim.\\ Secondly, a difference to our implementation is the fact that Gensim checks whether negative samples are not equal to the context word.
-If that is the case Gensim selects a new random sample.
-Therefore, the learning of the input and output context is optimized.
-\\Finally, another possibility is the decay of the learning rate used by Gensim.
+We compute the loss for multiple training samples by taking the sum over each score which is individually used as a loss by Gensim.\\ Finally, another possibility is the decay of the learning rate used by Gensim.
 In fact, decaying the learning rate has been proven in a lot of work to decrease the convergence time.
 Gensim linearly decreases the learning rate, as we did not use this technique, the decay of the learning rate could help explain the noted differences.
-\\ The first hypothesis, the fact that we used a batched approach, may be confirmed by the fact that when combined with input shuffling SGD does perform closer to Gensim, going from 11 to 7 epochs to converge, as input shuffling reduces the number of co-occurrence of the same word in a batch.
-Know the question arises if the 3 epochs, that Gensim is better, can be explained by the selection of better negative samples and the learning rate decay.
+\\ The first hypothesis, the fact that we used a batched approach, may be confirmed by the fact that when combined with input shuffling SGD does perform closer to Gensim, going from 7 to 5 epochs to converge, as input shuffling reduces the number of co-occurrence of the same word in a batch.
+Know the question arises if the 3 epochs, that Gensim is better can be explained by the batched approach and the learning rate decay.
 

### 2019-09-28 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/results.tex
final tex work

@@
 As stated in the previous section we ran multiple experiments for each optimizer with the above configuration and datasets.
-.
 This Section will  give an overview of the experimental results.
@@
 The first challenge for each optimizer was to find a correct learning rate.
-As SGD is the optimizer used in Gensim \citep{gensim} we first tried the same learning rate as the default value in Gensim \citep{gensim}, i.e 0.01,  and then performed a random search to find a better one.
+As SGD is the optimizer used in Gensim \citep{gensim} we first tried the same learning rate as the default value in Gensim \citep{gensim}, i.e 0.025,  and then performed a random search to find a better one.
 As expected a bell curve shape resulted, a learning rate that is too high leads to diversion and a learning rate that is too low leads to a training time that is too slow.
 The best value that we found for the learning rate is $0.0075$.
-With this setting SGD converged in 11 epochs.
+With this setting SGD converged in 7 epochs.
 The second experiment was to add input shuffling.
-For almost every learning rate the convergence time decreased.
-Our model, with the best setting, know converges in only 7 epochs.
-Another interesting fact to point out is that with input shuffling we achieved better results with higher learning rates.
-As for learning rates of $0.01$ and $0.025$ we did converge in 11 epochs with input shuffling but did not converge in 20 epochs without it.
+For almost every learning rate the convergence time decreased. Our model, with the best setting, know converges in only 5 epochs. 
 \iffalse
@@
 \subsubsection{Adam}
-Adam is the most advanced of all the optimizers used in our experiments and did yield the best results as seen in F.
+Adam is the most advanced of all the optimizers used in our experiments and did yield the best results as seen in Table \ref{table:results}.
 Adam converged in 3 epochs without shuffling and a learning rate of $0.05$ and 2 with shuffling and a learning rate of $0.001$, as shown in Table \ref{table:results}.

### 2019-09-29 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/introduction.tex
structure change w/ freq words
introduction plus abstract

--- 2019-08-27 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/introduction.tex
@@
 Therefore one could ask if the convergence time of the SGM can be optimized while at the same time maintaining its accuracy.\\
-This work focused on improving the convergence time of the SGNS.
-To facilitate our calculations we propose a slightly altered version of the original SNGS, in which we compute the loss for a very large amount of training samples, i.e 2000.
-The loss of such a batch was computed by taking the sum over the loss all individual training samples in the specific batch We combined this batched approach with advanced optimizers and input shuffling to decrease he convergence time of the SGNS.
-In combination advanced optimizers and input shuffling allowed us to decrease the convergence time of the SGNS.
+We propose a slightly altered version of the original SNGS, in which we compute the loss for a very large amount of training samples, i.e 2000.
+The loss of such a batch was computed by taking the sum over the loss all individual training samples in the specific batch.
+We combined this batched approach with advanced optimizers and input shuffling  This work focused on demonstrating that a batch approach, combined with advanced optimizers and input shuflling, does not hinder performance nor convergence time.
 If combined with an optimized throughput these results could lead to an overall decrease in runtime.\\

### 2019-09-29 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/results.tex
introduction plus abstract

@@
 Model    & Convergence Time & Word Similarity \\ \hline
-SGD & {11}              & 0.65            \\ \hline
-SGD w/shuffling & {7}              & 0.66            \\ \hline
-Adam & {3}              & 0.66            \\ \hline
-Adam w/ shuffling & \textbf{2}              & 0.66      \\ \hline
+SGD & {7}              & 0.66            \\ \hline
+SGD w/shuffling & {3}              & 0.65     \\ \hline
+Adam & {4}              & 0.67           \\ \hline
+Adam w/ shuffling & \textbf{4}              & 0.66      \\ \hline
 \end{tabular}

### 2019-09-29 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/sample-sigconf.tex
introduction plus abstract

@@
 The SGNS uses Stochastic Gradient Descent (SGD) as it's learning algorithm.
-While a lot of effort has gone into increasing the throughput of words of the SGNS, not much work has gone into optimizing the convergence time.
-Therefore, our work focuses on the latter.
-To achieve a better convergence time we use advanced optimizers and input shuffling as optimization techniques.
+A lot of effort has gone into increasing the throughput of words of the SGNS .
+Therefore we introduce a batched approach of the SGNS that should in theory reduce the runtime of the SGNS.
+In our batched approach the loss is computed for a high number of training samples (i.e.
+2000), which is then summed together.
+This sum is the used as loss for our model.
+This approach has the potential of further reducing the runtime of the SGNS.
+We combinded our batched approach with input shuffling and advanced optimizers.
 Here are the components used to evaluate our work: to train our model we used the text8 dataset - to measure the quality of our word embeddings we used the wordsim353 dataset, which assess that similar words have vector represenations that are close to each other.
@@
 Adam combined with input shuffling outperformed every optimizer.
-Adam with shuffling also outperformed the current state of the art implementation Gensim.
-Adam converged to a similarity value of 0.66 (state of the art) in 2 epochs, while Gensim took 4 epochs.
-Hence, this work shows that advanced optimizers combined with input shuffling do decrease the convergence time of the SGNS.
+Adam with shuffling converged to a similarity value of 0.67 epoch (1\% loss of accuracy) in 4 epochs.
+The current state of the art implementation Gensim converges in 4 epochs to a similarity value of 0.68.
+Hence this work shows that our batched approach does not hinder performance and has the potential of further reducing the runtime of the SGNS.
 These results must be confirmed with other datasets, but optimizing our model in terms of throughput has the potential of further reducing the overall run time of the SGNS.

### 2019-09-29 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/conclusion.tex
final changes

@@
 
-This work provides an overview of the Skip Gram Model with negative Sampling (SGNS) and the numerous successful attempts of optimizing the throughput of the model.
-As this is the case, no effort went into optimizng the convergence time of the SGNS, therefore this work focused on this task.
-We decided to use advanced optimizers and input shuffling to achieve a better convergence time.
-This work proposes a slighlty altered version of the SGNS, where the idea is to compute the loss over the sum of a high number of training samples, i.e 2000,  instead of computing it for each individually.
+This work provides an overview of the Skip Gram Model with negative Sampling (SGNS) and the numerous successful attempts of optimizing the throughput of the SGNS.
+This work also proposes a slighlty altered version of the SGNS, where the idea is to compute the loss over the sum of a high number of training samples, i.e 2000,  instead of computing it for each individually.
 The batched version allowed us to compute more models and analyze the convergence time faster.
+We showed that the convergence time and the quality of the created word embeddings are not hindered by our implementation.
 We used the text8 dataset to train our model and  word similarity as a quality measure for the word embeddings (WE).
 To compare our work, Gensim was used, which holds a state of the art implementation of the SGNS.
-We did achieve a better convergence time than Gensim with Adam as an optimizer and the use of input shuffling.
-Gensim convereged in 4 epochs to a word similarity of 0.66 and our model only took 2 epochs to achieve the same quality.
+We did achieve the same convergence time as Gensim with Adam as an optimizer and the use of input shuffling.
+Gensim convereged in 4 epochs to a word similarity of 0.68 and our model  took 4 epochs to achieve a score of 0.68.
 Those results still need to be confirmed with more datasets.

### 2019-09-29 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/discussion.tex
introduction plus abstract
final changes

@@
 Gensim linearly decreases the learning rate, as we did not use this technique, the decay of the learning rate could help explain the noted differences.
-\\ The first hypothesis, the fact that we used a batched approach, may be confirmed by the fact that when combined with input shuffling SGD does perform closer to Gensim, going from 7 to 5 epochs to converge, as input shuffling reduces the number of co-occurrence of the same word in a batch.
+\\ The first hypothesis, the fact that we used a batched approach, may be confirmed by the fact that when combined with input shuffling SGD does perform closer to Gensim, going from 7 to 3 epochs to converge, as input shuffling reduces the number of co-occurrence of the same word in a batch.
 Know the question arises if the 3 epochs, that Gensim is better can be explained by the batched approach and the learning rate decay.
@@
 \subsubsection{Gensim vs. Adam}
-The Adam optimizer did outperform the Gensim application in quality of word embeddings (only slightly: 0.01 correlation coefficient better) and convergence time.
-Adam converged in 2 epochs while Gensim in 4.
-To confirm the achieved results we ran each computation 40 times.
+The Adam optimizer did almost yield the same results as Gensim in accuracy  (only slightly: 0.01 correlation coefficient worse) and convergence time.
+Adam converged in the same number of  as Gensim, namely 4.
+To confirm the achieved results we ran each computation 10 times.
 The results can be seen in Figure \ref{fig:gensim_vs_adam}.
@@
 \centering
-\includegraphics[scale=0.3]{images/gensim_vs_adam}
+\includegraphics[scale=0.45]{images/comparison}
 \caption{Convergence time of SGD and Adam compared to Gensim}
@@
 To make a compelling argument it's necessary to show that our model with the use of input shuffling and Adam as its optimizer also outperforms Gensim with other data sets.\\
-Finally, as it wasn't the goal of our implementation to outperform Gensim in run time, one could improve an already existing, optimized version, with input shuffling and advanced optimizers and should achieve a better run time than Gensim.
+Finally, as it wasn't the goal of our implementation to outperform Gensim in run time, one could improve an already existing, optimized version, with input shuffling, advanced optimizers  and our batched approach and should achieve a better run time than Gensim.

### 2019-09-29 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/introduction.tex
final changes

@@
 By capturing semantic or syntactic meaning WE have shown to facilitate a lot of subsequent NLP tasks, such as entity recognition, machine translation or sentence classification.
-%TODO CITE WORK 
 The first attempt to create WE with neural networks was made by Bengio et al.
@@
 They, therefore, chose to parallelize the update of the dimensions of each word representation.
-Both of these approaches and most of the literature are focused on improving the throughput of the model, but not the convergence time.
-Therefore one could ask if the convergence time of the SGM can be optimized while at the same time maintaining its accuracy.\\
+Both of these approaches and most of the literature are focused on improving the throughput of the model, but not the convergence time.\\
 We propose a slightly altered version of the original SNGS, in which we compute the loss for a very large amount of training samples, i.e 2000.

### 2019-09-29 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/results.tex
final changes

@@
 As expected a bell curve shape resulted, a learning rate that is too high leads to diversion and a learning rate that is too low leads to a training time that is too slow.
-The best value that we found for the learning rate is $0.0075$.
+The best value that we found for the learning rate is $0.006$, as shown in Table \ref{test}.
 With this setting SGD converged in 7 epochs.
 The second experiment was to add input shuffling.
-For almost every learning rate the convergence time decreased. Our model, with the best setting, know converges in only 5 epochs. 
+For almost every learning rate the convergence time decreased.
+The best result was a learning rate of $0.0075$ combined with input shuffling where it converged to a similarity value of $0.65$ in 3 epochs, as shown in Table \ref{test}.
 \iffalse
@@
 Adagrad performed best with a learning rate of $0.1$.
+\fi
 \iffalse
@@
 \subsubsection{Adam}
-Adam is the most advanced of all the optimizers used in our experiments and did yield the best results as seen in Table \ref{table:results}.
-Adam converged in 3 epochs without shuffling and a learning rate of $0.05$ and 2 with shuffling and a learning rate of $0.001$, as shown in Table \ref{table:results}.
+Adam is the most advanced of all the optimizers used in our experiments and did yield the best results as seen in Table \ref{test}.
+Adam converged in 4 epochs without shuffling and a learning rate of $0.05$ to a similarity value of 0.66.
+With input shuffling and a learning rate of $0.001$ Adam converged to a similarity value of 0.67 in 4 epochs, as shown in Table \ref{test}.
 Those are the best results achieved in any of our experiment.
-Exacctly as Adagrad, Adamd did not react to input shuffling the same way as SGD did.
+
     \begin{table}[]
+    \caption{Results on the text8 dataset with different optimizers and input shuffling}
 \begin{tabular}{|l|l|l|}
@@
 \end{tabular}
-\label{table:results}
-\caption{Results on the text8 dataset with different optimizers and input shuffling}
+\label{test}
+\end{table}
 
-\end{table}
 \iffalse

### 2019-09-29 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/discussion.tex
final changes2

@@
 Gensim linearly decreases the learning rate, as we did not use this technique, the decay of the learning rate could help explain the noted differences.
-\\ The first hypothesis, the fact that we used a batched approach, may be confirmed by the fact that when combined with input shuffling SGD does perform closer to Gensim, going from 7 to 3 epochs to converge, as input shuffling reduces the number of co-occurrence of the same word in a batch.
+\\ The first hypothesis, the fact that we used a batched approach, may be confirmed by the fact that when combined with input shuffling SGD does perform closer to Gensim, going from 8 to 3 epochs to converge, as input shuffling reduces the number of co-occurrence of the same word in a batch.
 Know the question arises if the 3 epochs, that Gensim is better can be explained by the batched approach and the learning rate decay.

### 2019-09-29 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/results.tex
final changes2

@@
 The best value that we found for the learning rate is $0.006$, as shown in Table \ref{test}.
-With this setting SGD converged in 7 epochs.
+With this setting SGD converged in 8 epochs.
 The second experiment was to add input shuffling.
@@
 Adam converged in 4 epochs without shuffling and a learning rate of $0.05$ to a similarity value of 0.66.
-With input shuffling and a learning rate of $0.001$ Adam converged to a similarity value of 0.67 in 4 epochs, as shown in Table \ref{test}.
+With input shuffling and a learning rate of $0.003$ Adam converged to a similarity value of 0.67 in 4 epochs, as shown in Table \ref{test}.
 Those are the best results achieved in any of our experiment.
@@
 Model    & Convergence Time & Word Similarity \\ \hline
-SGD & {7}              & 0.66            \\ \hline
+SGD & {8}              & 0.66            \\ \hline
 SGD w/shuffling & {3}              & 0.65     \\ \hline
